### 2023-06-20 article .tex
started tex file for the article

\documentclass[
  journal=large,
  manuscript=article-type,
  year=2023,
  volume=?,
]{cup-journal}

\title{Resource Use and the Value-Productivity Tradeoff in Australian Winegrowing Regions}

\author{B, Polley}
\affiliation{Queensland University of Technology, Australian Wine Research Institute, Adelaide, 5047, South Australia, Australia}
\email{bryce.polley@awri.com.au}

\keywords{}

\begin{document}

\section{Introduction}
Viticulture differs from other agricultural pursuits as it is integrated within the wine industry, meaning that when strategies for sustainable viticulture are assessed, there is a trade off between trying to balance the amount of resources invested and the resultant quality and quantity of grapes produced (Montalvo-Falcón et al., 2023). These considerations have become increasingly important within the Australian wine industry due to recent economic, social and environmental pressures, such as bush fires (Wine Australia, 2020), tariffs (Wine Australia, 2022), and shifting attitudes to alcohol consumption (Toumbourou et al., 2018).

There is an extensive amount of research into how a variety of factors effect the outcome of a vineyard’s grape quality and yield, however individual effects are usually studied in isolation (Abbal et al., 2016). Due to a lack of long-term and in-depth data sources, there has been little ability to gain statistical insight at a large scale across multiple regions into vineyards’ resource use and its relationship to vineyards’ outputs (Keith Jones, 2002; Knight et al., 2019). Here we look at the broad effects of resource input and its relation to resultant quality and quantity; where we define quality by the financial value of the grapes per tonne. This analysis looks at these relationships through the use of linear models, using a significantly larger sample size compared to previous research. The analysis includes a multitude of vineyards located over a diverse range of locations and climates, allowing for a more comprehensive understanding of the relationships between resource use and vineyard outputs.

\section{Methods}
We created four linear models to explore the relationships between resource use and vineyard outputs (see table 1). These models look at the value and quantity of grapes produced by vineyards, as a total and as a scale of area harvested. We compare both values to observe how economies of scale effect the use of resources. The direct relationship between yield and average sale price was reviewed referring to the various models explanation of the determining factors.

Table 1: Summary of models, their predictors, covariates and variable interactions.

Response
Predictors
Covariates
Interactions
Model 1
Yield
Water Used, Scope 1 Emissions
Area Harvested, Year, GI Region
N/A
Model 2

Water Used, Scope 1 Emissions
Area Harvested, Year, GI Region
Area Harvested * Scope 1 Emissions,
Area Harvested * Water Use,
Year * Region
Model 3

Water Used, Scope 1 Emissions
Area Harvested, Year, GI Region
N/A
Model 4

Water Used, Scope 1 Emissions
Area Harvested, Year, GI Region
Area Harvested * Scope 1 Emissions,
Area Harvested * Water Use,
Year * Region
\subsection{Data}
The data used in this analysis was provided by Sustainable Winegrowing Australia (SWA), Australia’s national wine industry sustainability program, which aims to facilitate grape-growers and winemakers in demonstrating and improving their sustainability (SWA, 2022). The data recorded by SWA is entered manually by growers using a web based interface, with some fields being optional. Two subsets of this data were used for this analysis, both containing data for a vineyard’s region, harvest year, yield, area harvested, water used and fuel used (in litres for diesel, petrol, biodiesel and LPG). To enable comparisons, total fuel was converted to amount of carbon emissions in metric tons.

The first dataset, used for Model 1 and Model 2, contained 5298 samples spanning the period from 2012 to 2022, covering 57 GI Regions and 1432 separate vineyards.

The second dataset was a subset of the first, containing vineyards which recorded a value for their average sale price of grapes per tonne. This dataset was used for Model 3 and Model 4. This dataset contained 2878 samples spanning the period from 2015 to 2022, covering 51 GI Regions and 944 separate vineyards.

Data from the SWA was limited to samples that had recorded values for the variables used in the created models (see Table 1). The data was logarithmically transformed, centred and scaled by standard deviation before being used. Two values for average sale price were removed from the dataset, due to a recording of \$1. And, the unreported values for average prices per tonne were filled in using regional averages taken from Wine Australia’s yearly reports, where they were available (Wine Australia, 2022, 2021, 2020, 2019; Winemakers’ Federation of Australia, 2018, 2017, 2016, 2015, 2014, 2013, 2013).

\subsection{Total Emissions}
Emissions were calculated from the total diesel, petrol, bio-diesel and LPG used for irrigation and any activities within the vineyard. Taken from the Australian National Greenhouse Accounts Factors, Eq. 1 was used to convert the quantity of fuel in litres, Q, using a prescribed Energy Content, EC, and emission factors of scope one, EF1, and scope three, EF3, to tonnes of Carbon Dioxide equivalent, tCO2e (Department of Climate Change, Energy, the Environment and Water, 2022).
 									(1)

\subsection{Region}
The site of a vineyard predetermines several physical parameters such as climate, geology and soil, making location a widely considered key determinant of grape yield and quality (Abbal et al., 2016; Agosta et al., 2012; Fraga et al., 2017). A vineyards site and its significance was captured in this analysis through the use of including each vineyard’s Geographical Indicator Region (GI Region). GI Regions are the defined winegrowing regions of Australia, used to describe the rich and diverse landscape of Australian vineyards; with each GI Region having its own unique mixture of climatic and geophysical properties (Halliday, 2009; Oliver et al., 2013; SOAR et al., 2008).

The climatic properties of a GI Region are summarised by the SWA (SWA, 2021), where regions of similar climatic properties are amalgamated together into superset regions. This was used to illustrate similar trends and potentially explain the similarities and differences between sets of regions.

\subsection{Analysis}
General Linear Models were chosen as they offered the ability to produce statistical models that were explicit in the relationships between predictors and response variables. They also allowed the exploration interactions between predictors and easily comparable differences in influence and magnitude of relationships.

The Python programming language (G. van Rossum, 1995) was used for all data preprocessing such as logarithmic transformations prior to analysis and to connect to the SWA database. Linear models were created using the R statistical programming language (R Core Team, 2021). These models were created iteratively to explore a variety of variable interactions and approaches to modelling the data. Not all of the explored approaches yielded further improvements or accurate models, with different approaches attempted including the use of Splines, hierarchical regression and Generalised Linear Models. Due to the optional reporting of data many other variables were also explored but not used due to low reporting values such as fertiliser, tractor and contractor use. The use of only scope one emissions was due to the same reason where scope 2 sources were recorded sporadically at best.

\subsection{Model Validation}
K-fold cross validation was used to validate the models using the R Caret Package (Kuhn, 2008). K-fold cross validation works by removing a subset of data from the sample used to train models and then predicts those variables to determine how sensitive the model is to changes in the sample data. For this analysis each model was validated using 10 folds, repeated 100 times.

\section{Results}
\subsection{Exploratory Analysis}
The variables were initially reviewed for any simple linear relationships by using a Pearson’s Correlation Coefficient (see Tables 1, 2 and 3). This was undertaken for data on the original scale (see Table 1) and for data as a logarithmic transform (see Table 2). All P-values were found to be significant (< 2.200E-16), with the exception of the non-transformed values for water used (see Table 3). The logarithmic transforms performed the best due to a skew likely caused by a greater number of smaller vineyards within the dataset (see Table 4).

Table 1: Pearson correlation coefficients for each variable on the original scale.
Variable
Yield
Area
Water Used
Scope One Emissions
Yield/Area
Average Price Per Tonne
Average Price for Yield/Area
Yield
1.000E+00
7.440E-01
-4.309E-03
7.290E-01
3.500E-01
-2.262E-01
-1.644E-01
Area
7.440E-01
1.000E+00
-5.331E-03
8.921E-01
7.854E-02
-1.178E-01
-2.042E-01
Water Used
-4.309E-03
-5.331E-03
1.000E+00
-1.929E-03
-5.600E-03
-3.562E-02
-2.669E-02
Scope One Emissions
7.290E-01
8.921E-01
-1.929E-03
1.000E+00
9.357E-02
-9.422E-02
-1.933E-01
Yield/Area
3.500E-01
7.854E-02
-5.600E-03
9.357E-02
1.000E+00
-4.849E-01
-1.698E-01
Average Price Per Tonne
-2.262E-01
-1.178E-01
-3.562E-02
-9.422E-02
-4.849E-01
1.000E+00
4.732E-01
Average Price for Yield/Area
-1.644E-01
-2.042E-01
-2.669E-02
-1.933E-01
-1.698E-01
4.732E-01
1.000E+00

Table 2: Pearson correlation coefficients for each logarithmically transformed variable.

Variable
Yield
Area
Water Used
Scope One Emissions
Yield/Area
Average Price Per Tonne
Average Price for Yield/Area
Yield
1.000E+00
8.822E-01
8.245E-01
7.617E-01
9.353E-01
-4.591E-01
-8.918E-01
Area
8.822E-01
1.000E+00
7.750E-01
8.311E-01
6.742E-01
-1.911E-01
-8.474E-01
Water Used
8.245E-01
7.750E-01
1.000E+00
6.668E-01
7.292E-01
-4.881E-01
-8.300E-01
Scope One Emissions
7.617E-01
8.311E-01
6.668E-01
1.000E+00
6.086E-01
-1.559E-01
-7.063E-01
Yield/Area
9.353E-01
6.742E-01
7.292E-01
6.086E-01
1.000E+00
-5.625E-01
-8.076E-01
Average Price Per Tonne
-4.591E-01
-1.911E-01
-4.881E-01
-1.559E-01
-5.625E-01
1.000E+00
6.592E-01
Average Price for Yield/Area
-8.918E-01
-8.474E-01
-8.300E-01
-7.063E-01
-8.076E-01
6.592E-01
1.000E+00

Table 3: P-values for the non-transformed water used variable’s Pearson correlation coefficients.
Variable
Water Used
Yield
7.538E-01
Area
6.981E-01
Scope One Emissions
8.883E-01
Yield/Area
6.836E-01
Average Price Per Tonne
5.600E-02
Average Price for Yield/Area
1.522E-01

Table 4: Summary statistics for each variable on the original scale.
Variable
Mean
Std
Minimum
Median
Maximum
Water Used (ML)
8.06E+06
5.86E+08
1.00E+00
4.70E+01
4.27E+10
Scope One Emissions (tCO2e)
4.35E+04
8.72E+04
6.76E+00
1.49E+04
2.11E+06
Grapes Harvested (tonnes)
8.35E+02
2.28E+03
1.00E+00
2.19E+02
7.23E+04
Area Harvested (M2)
7.14E+04
1.40E+05
2.90E+02
2.72E+04
2.44e+06
Average Per Tonne ($)
1.45E+03
9.32E+02
1.60E+02
1.37E+03
2.60E+04
Grapes Harvested (tonnes/M2)
1.04E-02
8.27E-03
4.00E-05
7.53E-03
8.63E-02
Average Per Tonne ($/m2)
1.12E-01
2.30E-01
1.75E-04
4.08E-02
5.92E+00

\subsection{General Linear Models}
All variables were considered significant contributors to the models, with the exception of scope one emissions in Models 3 and 4 (see Tables 5, 6, 7, 8). The inclusion of scope one emissions was done for comparison and as an interaction with area harvested. Even though scope one emissions was not a significant contributor within models 3 and 4, it was found to be significantly correlated on its own with both model average price per tonne, especially when it was scaled by area (see table 2). However, noting that irrigation systems still use fuel and that the application of water is a significant variable in every model, scope one emissions’ lack of significance and contribution given its F-statistics (See Tables 7 and 8), indicated that it is possible other vineyard activities requiring fuel are not as determining factors for a vineyards grape quality.

The comparison of models performance shows that the average price per tonne of grapes describes a great deal of the relationship between predictor and response when comparing model 2 to model 4 (see Table 10). This relationship between yield and average price was also illustrated in the correlation values between them (see Table 2).

Table 5: Model 1 ANOVA summarising variable significance at the .5 level.
Variable
Df
Sum Sq
Mean Sq
F Value
Pr(>F)
Year
9
7.060E+01
7.800E+00
8.353E+01
< 2.20E-16 ***
GI Region
54
1.507E+03
2.790E+01
2.972E+02
< 2.20E-16 ***
Area Harvested
1
3.211E+03
3.211E+03
3.419E+04
< 2.20E-16 ***
Water Used
1
1.040E+01
1.040E+01
1.103E+02
< 2.20E-16 ***
Scope One Emissions
1
6.600E+00
6.600E+00
7.056E+01
< 2.20E-16 ***

Table 6: Model 2 ANOVA summarising variable significance at the .5 level.
Variable
Df
Sum Sq
Mean Sq
F Value
Pr(>F)
Area Harvested
1
2.407E+03
2.407E+03
1.080E+04
< 2.20E-16 ***
Scope One Emissions
1
3.989E+01
3.989E+01
1.789E+02
< 2.20E-16 ***
Water Used
1
5.500E+02
5.500E+02
2.467E+03
< 2.20E-16 ***
Area Harvested*Scope One Emissions
1
6.921E+01
6.921E+01
3.104E+02
< 2.20E-16 ***
Area Harvested * Water Used
1
1.040E+00
1.040E+00
4.686E+00
3.045E-02 **
Year * GI Region
424
1.144E+03
2.700E+00
1.210E+01
< 2.20E-16 ***

Table 7: Model 3 ANOVA summarising variable significance at the .5 level.
Variable
Df
Sum Sq
Mean Sq
F Value
Pr(>F)
Year
6
1.324E+01
2.210E+00
8.748E+01
< 2.20E-16 ***
GI Region
50
6.498E+02
1.300E+01
5.151E+02
< 2.20E-16 ***
Area Harvested
1
2.142E+03
2.142E+03
8.491E+04
< 2.20E-16 ***
Water Used
1
3.200E-01
3.200E-01
1.259E+01
3.947E-04 **
Scope One Emissions
1
4.000E-02
4.000E-02
1.492E+00
2.221E-01

Table 8: Model 4 ANOVA summarising variable significance at the .5 level.
Variable
Df
Sum Sq
Mean Sq
F Value
Pr(>F)
Area Harvested
1
2.066E+03
2.066E+03
5.700E+04
< 2.20E-16 ***
Scope One Emissions
1
6.000E-02
6.000E-02
1.569E+00
2.105E-01
Water Used
1
2.014E+02
2.014E+02
5.557E+03
< 2.20E-16 ***
Area Harvested*Scope One Emissions
1
5.246E+01
5.246E+01
1.448E+03
< 2.20E-16 ***
Area Harvested * Water Used
1
7.270E+00
7.270E+00
2.005E+02
< 2.20E-16 ***
Year * GI Region
243
4.546E+02
1.870E+00
5.162E+01
< 2.20E-16 ***

Table 9: Comparison of Model Residuals

Df
Sum Sq
Mean Sq
Model 1
5231
4.913E+02
1.000E-01
Model 2
4868
1.085E+03
2.200E-01
Model 3
2818
7.111E+01
3.000E-02
Model 4
2629
9.528E+01
4.000E-02

Table 10: Comparison of Model performance.

RSE
R2
Adjusted R2
F-statistic
P-Value
Model 1
3.065E-01
9.072E-01
9.061E-01
7.753E+02
< 2.2e-16
Model 2
4.722E-01
7.951E-01
7.770E-01
4.403E+01
< 2.2e-16
Model 3
1.589E-01
9.753E-01
9.748E-01
1.885E+03
< 2.2e-16
Model 4
1.904E-01
9.669E-01
9.638E-01
3.095E+02
< 2.2e-16

Models 1 and 2 suffered from overestimating yield, with Models 3 and 4 underestimating crop value (See Figures 1, 2, 3 and 4). Reviewing the data to uncover reasons for this included the use of binary variables such as the utilisation of renewable energy, contractors, and the occurrence of disease, fire and frost; however none of these variables were able to explain why some vineyards produced less, or why other vineyards sold at higher prices than predicted. A wide variety of these influences were likely already explained within the use of year and GI Region, or the interaction of both variables. The change between some regions was dramatic, with particularly warmer and drier regions producing much higher volumes of grapes at lower prices (See Figures 5 and 6). The use of other variables and methods, specifically splines, were able to create a more normally distributed set of residuals but at a drastically reduced accuracy when comparing R2 and RSE. The introduction of known average prices per tonne also helped increase R2 values a small amount; it is important to not that it is common practice for wineries to purchase grapes at a regional average rate, likely resulting in much less variance within a region.

Figure 1: Model 1 residuals vs fitted (left) and QQ plot (right).

Figure 2:  Model 2 residuals vs fitted (left) and QQ plot (right).

Figure 3:  Model 3 residuals vs fitted (left) and QQ plot (right).

Figure 4:  Model 4 residuals vs fitted (left) and QQ plot (right).

Figure 5: Average Yield per hectare for each GI Region

Figure 6: Average Price Per Tonne for each GI Region

Although the initial correlation between average sales price and yield was a negative trend (see table 2), these models were able to shed light on several contributing factors to yield and average sales price. Correlation values showed that water and emissions increased with yield but decreased with average sale price (see Table 4). In alternative attempts at models it was found that without the incorporation of GI Region or year the predictions greatly under performed. The possible reason behind this effect was that different strategies are likely employed between different regions, where some regions target the mass production of cheaper grapes over quality. This is most notable when grouping regions by climate, especially when considering GI Regions in the ‘Hot Very Dry’ climate (see Figure 7). The effect of climate when used in models was not more significant than the more granular use of GI regions. The interaction between year and GI Region likely accounted for localised events such as bushfires, which would be impactful, but only at a local level in both time and space.

Figure 7: Value of tonnes harvested per hectare against the yield per hectare. With the variable transformed, with means centred on zero and scaled by standard deviation.

\subsection{Model Validation}
To validate the performance of these models k-fold cross validation was used. This was done using 10 folds, k=10, repeated 100 times. The models performed similarly to their original counter parts (see Table 11).

Table 11: Model validation using k-fold cross validation, for 10 folds repeated 100 times.

RMSE
R2
MAE
Model 1
3.087E-01
9.045E-01
2.165E-01
Model 2
5.104E-01
7.409E-01
3.493E-01
Model 3
1.652E-01
9.723E-01
1.008E-01
Model 4
2.235E-01
9.500E-01
1.279E-01

\section{Discussion}
This study investigated the general relationships between input resources of a vineyard, including fuel and water, and the outputs including yield and value. Some regions appeared to produce many low quality grapes at scale compared to attempting to produce fewer higher quality grapes. This behaviour can be observed when reviewing Wine Australia’s annual reports, where it is apparent that warm inland regions such as the Riverland are known to only produce large amounts of lower graded grapes (Wine Australia, 2022; Winemakers’ Federation of Australia, 2017). Comparatively, regions such as Tasmania only produce A grade grapes but in much smaller quantities than the Riverland. Knowing that the difference in pricing per tonne can exceed a magnitude of 10 between grades E and A, the operations in regions that target different grades would have varied priorities. However, some regions such as the Yarra Valley produce a Variety of different grades of grapes, from C to A, highlighting that vineyard priorities, although may be somewhat present within regional classifications, are not necessarily aligned within a given region.

The opportunity to target different grades of grapes may not always be available, with some regions being more renowned than others, and likely to be sought after regardless (Halliday, 2009). The Barossa is an example of this, known for its quality could also lend itself to a bias in purchasers not considering other regions that may be capable of similar quality. This effect could stifle the potential for market opportunities within these lesser known regions. A further possibility is that there may be regional upper limits with the relationship between resource input and the value gained becoming no longer proportional due to diminishing returns. Climate was considered to be a large determinant of the ability to grow a larger quantity of grapes, as well as a determinant in grape quality (Agosta et al., 2012); however there were vineyards in similar regions that were able to produce exceptionally better results than others (See Figure 7).

The issue of model 1 and 2 over predicting yield, may have been due to preventative measures brought on by regional pressures such as fire, frost and disease. Where, more resources were required to prevent these issues from spreading within a region, thus disproportionately effecting some vineyards compared to others locally. This type of maintenance is not well captured especially when considering that some regions, those in warmer areas are not as prone to disease as cooler climates and could potentially have lower operating costs per hectare. This could create a discrepancy in vineyards that utilise preventative measures in wetter regions, as opposed to those who do not, and thus expend less fuel and energy but risk disease. When reviewing the differences between regions it is important to consider that vineyards in Hot Very Dry areas can be hundreds of times the size of those in other regions. It is interesting that while area, although significantly correlated to the ratio of yield to area, was still lower than water and about the same as emissions. This points to economies of scale playing a role but still being only one consideration alongside the potential resources that can be used. The negative trend between size and average sales price could also be a side effect of mass supply verse demand, especially when looking at the level of difference in production of some vineyards (see Table 4). The relationships between yield, value and area are not simply about efficiently producing the most grapes; sales price and by association grape quality, are integral to the profitability, and this is strongly linked to resource use and thus the longevity and sustainability of a vineyard.

Literature shows that there are many on-the-ground decisions that influence both quality and yield. Where these decisions are governed by complex physical and social forces such as international market demands, disease pressures and natural disasters (Abad et al., 2021; Cortez et al., 2009; Hall et al., 2011; I. Goodwin, et al., 2009; Kasimati et al., 2022; Oliver et al., 2013; Srivastava and Sadistap, 2018). Many of these occurrences being highlighted throughout the past decades vintage reports (Wine Australia, 2022, 2021, 2020, 2019; Winemakers’ Federation of Australia, 2018, 2017, 2016, 2015, 2014, 2013, 2012). It is also important to consider that these reports show that the warm inland regions have seen a decline in profit during this period, as they were often compared to other regions that focused more on quality than quantity. This is an important consideration, as the size of some of these vineyards when considering their ratio of value to area would only require a marginal increase to out compete other regions. There are also differences when comparing winegrowers to other agricultural industries as they are vertically integrated within the wine industry, tying them to secondary and tertiary industries, such as wine production, packaging, transport and sales. This results in unique issues and considerations for each vineyard, where these on-the-ground decisions may be influenced by other wine industry’s choices, such as the use of sustainable practices in vineyards as a requirement for sale in overseas markets; notably these interactions are further complicated by some winegrowers being totally integrated into wine companies, while others are not (Knight et al., 2019). Incorporating such decisions into the model could help describe the contributing factors to regional differences beyond resource consumption and regional differences.

Having more data for each region would also be an improvement, allowing greater comparison between regions. More variables may also help to discern vineyards that can produce larger volumes of grapes at higher prices. The use of semi transparent tools such as random forests and decision trees alongside more variables and data may help to uncover the reasons for values that were under or over estimated. These differences could be caused by the use of alternative sustainable practices in the field. While there is evidence to suggest that environmentally sustainable practices can reduce costs, increase efficiency, whilst improving the quality of grapes, more research is needed to link these benefits across different regions and climates (Baiano, 2021; Mariani and Vastola, 2015; Montalvo-Falcón et al., 2023).

\end{document}

References
Abad, J., Hermoso de Mendoza, I., Marín, D., Orcaray, L., Santesteban, L.G., 2021. Cover crops in viticulture. A systematic review (1): <br>Implications on soil characteristics and biodiversity in vineyard. OENO One 55, 295–312. https://doi.org/10.20870/oeno-one.2021.55.1.3599
Abbal, P., Sablayrolles, J.-M., Matzner-Lober, É., Boursiquot, J.-M., Baudrit, C., Carbonneau, A., 2016. Decision Support System for Vine Growers Based on a Bayesian Network. J. Agric. Biol. Environ. Stat. 21, 131–151. https://doi.org/10.1007/s13253-015-0233-2
Agosta, E., Canziani, P., Cavagnaro, M., 2012. Regional climate variability impacts on the annual grape yield in Mendoza, Argentina. J. Appl. Meteorol. Climatol. 51, 993–1009.
Baiano, A., 2021. An Overview on Sustainability in the Wine Production Chain. Beverages 7. https://doi.org/10.3390/beverages7010015
Cortez, P., Teixeira, J., Cerdeira, A., Almeida, F., Matos, T., Reis, J., 2009. Using data mining for wine quality assessment, in: Discovery Science: 12th International Conference, DS 2009, Porto, Portugal, October 3-5, 2009 12. Springer, pp. 66–79.
Department of Climate Change, Energy, the Environment and Water, 2022. Australian National Greenhouse Accounts Factors.
Fraga, H., Costa, R., Santos, J.A., 2017. Multivariate clustering of viticultural terroirs in the Douro winemaking region. Ciênc. Téc Vitiv 32, 142–153.
G. van Rossum, 1995. Python tutorial, Technical Report CS-R9526.
Hall, A., Lamb, D.W., Holzapfel, B.P., Louis, J.P., 2011. Within-season temporal variation in correlations between vineyard canopy and winegrape composition and yield. Precis. Agric. 12, 103–117.
Halliday, J.C. (James C.), 2009. Australian Wine Encyclopedia. Hardie Grant Books, VIC.
I. Goodwin, L. McClymont, D. Lanyon, A. Zerihun, J. Hornbuckle, M. Gibberd, D. Mowat, D. Smith, M. Barnes, R. Correll, 2009. Managing soil and water to target quality and reduce environmental impact.
Kasimati, A., Espejo-García, B., Darra, N., Fountas, S., 2022. Predicting Grape Sugar Content under Quality Attributes Using Normalized Difference Vegetation Index Data and Automated Machine Learning. Sensors 22. https://doi.org/10.3390/s22093249
Keith Jones, 2002. Australian Wine Industry Environment Strategy.
Knight, H., Megicks, P., Agarwal, S., Leenders, M.A.A.M., 2019. Firm resources and the development of environmental sustainability among small and medium-sized enterprises: Evidence from the Australian wine industry. Bus. Strategy Environ. 28, 25–39. https://doi.org/10.1002/bse.2178
Kuhn, M., 2008. Building Predictive Models in R Using the caret Package. J. Stat. Softw. Artic. 28, 1–26. https://doi.org/10.18637/jss.v028.i05
Mariani, A., Vastola, A., 2015. Sustainable winegrowing: Current perspectives. Int. J. Wine Res. 7, 37–48.
Montalvo-Falcón, J.V., Sánchez-García, E., Marco-Lajara, B., Martínez-Falcó, J., 2023. Sustainability Research in the Wine Industry: A Bibliometric Approach. Agronomy 13. https://doi.org/10.3390/agronomy13030871
Oliver, D.P., Bramley, R.G.V., Riches, D., Porter, I., Edwards, J., 2013. Review: soil physical and chemical properties as indicators of soil quality in Australian viticulture. Aust. J. Grape Wine Res. 19, 129–139. https://doi.org/10.1111/ajgw.12016
R Core Team, 2021. R: A Language and Environment for Statistical Computing.
SOAR, C.J., SADRAS, V.O., PETRIE, P.R., 2008. Climate drivers of red wine quality in four contrasting Australian wine regions. Aust. J. Grape Wine Res. 14, 78–90. https://doi.org/10.1111/j.1755-0238.2008.00011.x
Srivastava, S., Sadistap, S., 2018. Non-destructive sensing methods for quality assessment of on-tree fruits: a review. J. Food Meas. Charact. 12, 497–526.
SWA, S.W.A., 2022. Sustainable Wingrowing Australia [WWW Document]. Case Stud. URL https://sustainablewinegrowing.com.au/case-studies/ (accessed 11.5.22).
SWA, S.W.A., 2021. Sustainable Winegrowing Australia User Manual.
Toumbourou, J.W., Rowland, B., Ghayour-Minaie, M., Sherker, S., Patton, G.C., Williams, J.W., 2018. Student survey trends in reported alcohol use and influencing factors in Australia. Drug Alcohol Rev. 37, S58–S66. https://doi.org/10.1111/dar.12645
Wine Australia, 2022. National Vintage Report 2022.
Wine Australia, 2021. National Vintage Report 2021.
Wine Australia, 2020. National Vintage Report 2020.
Wine Australia, 2019. National Vintage Report 2019.
Winemakers’ Federation of Australia, 2018. National Vintage Report 2018.
Winemakers’ Federation of Australia, 2017. National Vintage Report 2017.
Winemakers’ Federation of Australia, 2016. National Vintage Report 2016.
Winemakers’ Federation of Australia, 2015. National Vintage Report 2015.
Winemakers’ Federation of Australia, 2014. National Vintage Report 2014.
Winemakers’ Federation of Australia, 2013. National Vintage Report 2013.
Winemakers’ Federation of Australia, 2012. National Vintage Report 2012.

### 2023-06-22 article.tex
updating the tex file. This should be done at the end of the day when writing for future referencegit add article.tex What if your computer breaks dont risk itgit add article.tex

\documentclass[12pt,australian]{article}
\usepackage{lingmacros}
\usepackage{tree-dvips}
\usepackage[utf8]{inputenc}
\usepackage{csquotes,xpatch}% recommended
\usepackage[australian]{babel}
\usepackage{amsmath}
\usepackage[backend=biber,<options>]{biblatex}
\usepackage{graphicx}
\addbibresource[location=local]{references.bib}

\title{Resource Use and the Value-Productivity Tradeoff in Australian Winegrowing Regions}
\author{Bryce Polley}
\date{20/06/2023}

\begin{document}
\begin{sloppypar}

\section{Introduction}
When assessing strategies for viticulture there is a trade-off between the amount of resources invested and the resultant quality and quantity of grapes produced \autocite{montalvo-falconSustainabilityResearchWine2023}. Due to recent economic, social and environmental pressures, such as: bush fires, tariffs, and shifting attitudes to alcohol consumption; the chosen emphasis taken in balancing the value-productivity tradeoff has become more important \autocite{toumbourouStudentSurveyTrends2018,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2022}. When considering pressures such as those prior listed, the integration of viticulture within the Australian wine industry is a compounding effect.

In this analysis we observed relationships between quantity and quality through the use of linear models. For the purpose of this study; quality was defined by the financial value of grapes per tonne. An extensive amount of research into a variety of factor's effect on grape quality and yield exists. Due to the lack of long-term and in-depth data sources, individual effects are often studied in isolation \autocite{abbalDecisionSupportSystem2016}. The lack of consolidated datasets also restricts the ability to gain statistical insight at a large scale across multiple regions \autocite{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}. The dataset used for this analysis includes data collected for the past 10 years from a multitude of vineyards located over a diverse range of Australian winegrowing regions.

\section{Methods}
We created four linear models to explore the relationships between resource use and vineyard outputs (see table 1). The models looked at the value and quantity of grapes produced by vineyards, as a total and as a scale of area harvested. We compared values to observe how economies of scale affect the use of resources.

\begin{table}[]
    \begin{center}
      \caption{Summary of models; their predictors, covariates and variable interactions.}
      \label{tab:tab1}
      \begin{tabular}{ccccc}
        & Response & Predictors & Covariates & Interactions \\
        Model 1 & Yield & Water Used, Scope 1 Emissions & Area Harvested, Year, GI Region & N\textbackslash A \\
        Model 2 & ${\text{Yield}}\over{\text{Area Harvested}}$ & Water Used, Scope 1 Emissions & Area Harvested, Year, GI Region & Area Harvested * Scope 1 Emissions, Area Harvested * Water Use, Year * Region \\
        Model 3 & \small{$\text{Yield} {\times} \text{Average Sale Price}$} &  Water Used, Scope 1 Emissions& Area Harvested, Year, GI Region & Area Harvested * Scope 1 Emissions, Area Harvested * Water Use, Year * Region  \\
        Model 4 & ${\text{Yield} \times \text{Average Sale Price}}\over{\text{Area Harvested}}$ & Water Used, Scope 1 Emissions & Area Harvested, Year, GI Region  & N\textbackslash A
      \end{tabular}
    \end{center}
\end{table}

\subsection{Data}
The data used in this analysis was provided by Sustainable Winegrowing Australia, Australia's national wine industry sustainability program. Sustainable winegrowing Australia aims to facilitate grape-growers and winemakers in demonstrating and improving their sustainability \autocite{swaSustainableWingrowingAustralia2022}. Data recorded by Sustainable Winegrowing Australia is entered manually by winegrowers using a web based interface; with some fields being optional. Two subsets of this data were defined by vineyards that recorded values for average price of grape sales per tonne, and vineyards who did not. Both subsets contained: region, harvest year, yield, area harvested, water used and fuel used (in litres for diesel, petrol, biodiesel and LPG). To enable comparisons, total fuel was converted to amount of carbon emissions in metric tons.

The first subset of data was used for Model 1 and Model 2 (see Table 1). This subset contained 5298 samples spanning the period from 2012 to 2022, covering 57 GI Regions and 1432 separate vineyards.

The second subset of data, was limited to vineyards that recorded a value for their average sale price of grapes per tonne. This subset was used for Model 3 and Model 4 (see Table 1); and contained 2878 samples spanning the period from 2015 to 2022, covering 51 GI Regions and 944 separate vineyards.

Data was limited to samples that had recorded values for variables used (see Table 1). After reviewing correlation coefficients the data was logarithmically transformed, centred and scaled by standard deviation. Two values for average sale price were removed from the dataset, due to a recording of \$1. Unreported values for average prices per tonne were filled in using regional averages taken from Wine Australia's annual reports, where they were available \autocite{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2012,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.

Other variables including the use of renewable energy, contractors; and pressures such as frost, fire and disease were also explored. Variables that did not significantly contribute to the prediction of a response variable were excluded.

\subsection{Total Emissions}
Emissions were calculated from the total diesel, petrol, bio-diesel and LPG used for irrigation and activities within the vineyard. The equations given from the Australian National Greenhouse Accounts Factors, shown as

\begin{equation}
\label{(1)}
    tCO_{2}e={{Q \times EC \times EF1 + EF3}\over{1000}},
\end{equation}

was used to convert the quantity of fuel in litres, $Q$, using a prescribed Energy Content, $EC$, and emission factors of scope one, $EF1$, and scope three, $EF3$, to tonnes of Carbon Dioxide equivalent, $tCO2e$ \autocite{departmentofclimatechangeenergytheenvironmentandwaterAustralianNationalGreenhouse2022}.

The variables were reviewed for correlations by using a Pearson's Correlation Coefficient (see Tables 1, 2 and 3). This was undertaken for data on the original scale (see Table 1) and for data as a logarithmic transform (see Table 2). All P-values were found to be significant (< 2.200E-16), except the non-transformed values for water used (see Table 3). The logarithmic transforms performed the best due to a skew likely caused by a greater number of smaller vineyards within the dataset (see Table 4).

\subsection{Region}
The site of a vineyard predetermines several physical parameters such as climate, geology and soil; making location a widely considered key determinant of grape yield and quality \autocite{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}. Differences in vineyard locations were captured through the use of Geographical Indicator Regions (GI Regions). Each GI Region has its own unique mixture of climatic and geophsical properties that describes a unique winegrowing region within Australia; these regions were predefined by Wine Australia \autocite{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.

The climatic properties of a GI Region are summarised in the Sustainable Winegrowing Australia user manual \autocite{swaSustainableWinegrowingAustralia2021}. The user manual describes climates by rainfall and temperature, creating supersets of Regions of similar climatic properties. The climatic groups were used to illustrate similarities and differences occurring in areas larger than GI regions.

\subsection{Analysis}
General Linear Models were used as they offered the ability to produce statistical models that were explicit in the relationships between predictors and response variables. They also allowed the exploration of interactions between predictors and easily comparable differences in the influence and magnitude of relationships.

Data preprocessing, such as logarithmic transforms, was done using the Python programming language \autocite{g.vanrossumPythonTutorialTechnical1995}. Linear models were created using the R statistical programming language \autocite{rcoreteamLanguageEnvironmentStatistical2021}. These models were created iteratively to explore a variety of variable interactions and approaches to modelling the data. Not all explored approaches yielded improvements or accurate models. Alternate approaches included the use of Splines, hierarchical regression, Additive and Generalised Linear Models. Other variables were also explored but not used due to low reporting values such as fertiliser, tractor and contractor use. The use of only scope one emissions was due to the same reason where scope 2 sources were recorded sporadically at best.

\subsection{Model Validation}
Models were validated using K-fold cross validation calculated through the R Caret Package \autocite{kuhnBuildingPredictiveModels2008}. K-fold cross validation works by removing a subset of data from the sample used to train models and then predicts those variables to determine how sensitive the model is to changes in the sample data. For this analysis each model was validated using 10 folds, repeated 100 times. % A reference for this is absolutely needed.

\section{Results}
\subsection{Exploratory Analysis}
Simple linear relationships between variables were explored using Pearson Correlation Coefficients. This was undertaken for data on the original scale (see Table 2) and for data as a logarithmic transform (see Table 3). Strong relationships were found to be present, as all P-values were considered significant (< 2.200E-16, see Tables 2 and 3), except for the non-transformed values for water used (see Table 4). The logarithmic transforms showed the strongest correlations, this was likely due to a skew caused by a greater number of smaller vineyards within the dataset (see Table 5).

\begin{table}[]
  \caption{Summary of models, their predictors, covariates and variable interactions.}
  \label{tab:tab2}
  \begin{tabular}{cccccccc}
  Variable                             & Yield      & Area       & Water Used & Scope One Emissions & $\text{Yield}\over\text{Area}$ & Average Price Per Tonne & ${\text{Average Price per tonne}\over\text{Area}}$ \\
  Yield                                & 1.000E+00  & 7.440E-01  & -4.309E-03 & 7.290E-01           & 3.500E-01            & -2.262E-01              & -1.644E-01                           \\
  Area                                 & 7.440E-01  & 1.000E+00  & -5.331E-03 & 8.921E-01           & 7.854E-02            & -1.178E-01              & -2.042E-01                           \\
  Water Used                           & -4.309E-03 & -5.331E-03 & 1.000E+00  & -1.929E-03          & -5.600E-03           & -3.562E-02              & -2.669E-02                           \\
  Scope One Emissions                  & 7.290E-01  & 8.921E-01  & -1.929E-03 & 1.000E+00           & 9.357E-02            & -9.422E-02              & -1.933E-01                           \\
  $\text{Yield}\over\text{Area}$                 & 3.500E-01  & 7.854E-02  & -5.600E-03 & 9.357E-02           & 1.000E+00            & -4.849E-01              & -1.698E-01                           \\
  Average Price Per Tonne              & -2.262E-01 & -1.178E-01 & -3.562E-02 & -9.422E-02          & -4.849E-01           & 1.000E+00               & 4.732E-01                            \\
  ${\text{Average Price per tonne}\over\text{Area}}$ & -1.644E-01 & -2.042E-01 & -2.669E-02 & -1.933E-01          & -1.698E-01           & 4.732E-01               & 1.000E+00                           
  \end{tabular}
  \end{table}

  \begin{table}[]
    \caption{Pearson correlation coefficients for each logarithmically transformed variable.}
    \label{tab:tab3}
    \begin{tabular}{clllllll}
    Variable                                           & \multicolumn{1}{c}{Yield} & \multicolumn{1}{c}{Area} & \multicolumn{1}{c}{Water Used} & \multicolumn{1}{c}{Scope One Emissions} & \multicolumn{1}{c}{$\text{Yield}\over\text{Area}$} & \multicolumn{1}{c}{Average Price Per Tonne} & \multicolumn{1}{c}{${\text{Average Price per tonne}\over\text{Area}}$} \\
    Yield                                              & 1.000E+00                 & 8.822E-01                & 8.245E-01                      & 7.617E-01                               & 9.353E-01                                          & -4.591E-01                                  & -8.918E-01                                                             \\
    Area                                               & 8.822E-01                 & 1.000E+00                & 7.750E-01                      & 8.311E-01                               & 6.742E-01                                          & -1.911E-01                                  & -8.474E-01                                                             \\
    Water Used                                         & 8.245E-01                 & 7.750E-01                & 1.000E+00                      & 6.668E-01                               & 7.292E-01                                          & -4.881E-01                                  & -8.300E-01                                                             \\
    Scope One Emissions                                & 7.617E-01                 & 8.311E-01                & 6.668E-01                      & 1.000E+00                               & 6.086E-01                                          & -1.559E-01                                  & -7.063E-01                                                             \\
    $\text{Yield}\over\text{Area}$                     & 9.353E-01                 & 6.742E-01                & 7.292E-01                      & 6.086E-01                               & 1.000E+00                                          & -5.625E-01                                  & -8.076E-01                                                             \\
    Average Price Per Tonne                            & -4.591E-01                & -1.911E-01               & -4.881E-01                     & -1.559E-01                              & -5.625E-01                                         & 1.000E+00                                   & 6.592E-01                                                              \\
    ${\text{Average Price per tonne}\over\text{Area}}$ & -8.918E-01                & -8.474E-01               & -8.300E-01                     & -7.063E-01                              & -8.076E-01                                         & 6.592E-01                                   & 1.000E+00                                                             
    \end{tabular}
    \end{table}

    \begin{table}[]
      \caption{P-values for the non-transformed water used variable's Pearson correlation coefficients.}
    \label{tab:tab4}
      \begin{tabular}{cl}
      Variable                                           & Water Used \\
      Yield                                              & 7.538E-01  \\
      Area                                               & 6.981E-01  \\
      Scope One Emissions                                & 8.883E-01  \\
      $\text{Yield}\over\text{Area}$                     & 6.836E-01  \\
      Average Price Per Tonne                            & 5.600E-02  \\
      ${\text{Average Price per tonne}\over\text{Area}}$ & 1.522E-01 
      \end{tabular}
      \end{table}

    \begin{table}[]
      \caption{Summary statistics for each variable on the original scale..}
      \label{tab:tab5}
      \begin{tabular}{clllllll}
      Variable &
        \multicolumn{1}{c}{Yield} &
        \multicolumn{1}{c}{Area} &
        \multicolumn{1}{c}{Water Used} &
        \multicolumn{1}{c}{Scope One Emissions} &
        \multicolumn{1}{c}{$\text{Yield}\over\text{Area}$} &
        \multicolumn{1}{c}{Average Price Per Tonne} &
        \multicolumn{1}{c}{${\text{Average Price per tonne}\over\text{Area}}$} \\
      Yield                                              & 1.000E+00  & 8.822E-01  & 8.245E-01  & 7.617E-01  & 9.353E-01  & -4.591E-01 & -8.918E-01 \\
      Area                                               & 8.822E-01  & 1.000E+00  & 7.750E-01  & 8.311E-01  & 6.742E-01  & -1.911E-01 & -8.474E-01 \\
      Water Used                                         & 8.245E-01  & 7.750E-01  & 1.000E+00  & 6.668E-01  & 7.292E-01  & -4.881E-01 & -8.300E-01 \\
      Scope One Emissions                                & 7.617E-01  & 8.311E-01  & 6.668E-01  & 1.000E+00  & 6.086E-01  & -1.559E-01 & -7.063E-01 \\
      $\text{Yield}\over\text{Area}$                     & 9.353E-01  & 6.742E-01  & 7.292E-01  & 6.086E-01  & 1.000E+00  & -5.625E-01 & -8.076E-01 \\
      Average Price Per Tonne                            & -4.591E-01 & -1.911E-01 & -4.881E-01 & -1.559E-01 & -5.625E-01 & 1.000E+00  & 6.592E-01  \\
      ${\text{Average Price per tonne}\over\text{Area}}$ & -8.918E-01 & -8.474E-01 & -8.300E-01 & -7.063E-01 & -8.076E-01 & 6.592E-01  & 1.000E+00 
      \end{tabular}
      \end{table}

\subsection{General Linear Models}
Models 1 and 2 showed significant relationships between each of the predictors and their response variable (see Tables 6 and 7). Variables in models 3 and 4 reported similar significance; except for scope 1 emissions (see Tables 8 and 9). Scope one emissions was included in all models to directly compare the response variables as ratios of vineyard size to raw values. Even though not significant within models 3 and 4, when using the Pearson Correlation Coefficients, scope one emissions was strongly correlated to every Model's response variable; this was especially so for Model 1 and 4 (Yeild and average price per tonne as a ratio to area harvested, respectively).

The comparison of models performance shows that the average price per tonne of grapes describes a great deal of the relationship between predictor and response when comparing model 2 to model 4 (see Table 10). This relationship between yield and average price was also illustrated in the correlation values between them (see Table 2).

\begin{table}[]
  \label{tab:tab6}
  \caption{Model 1 ANOVA summarising variable significance at the .5 level.}
  \begin{tabular}{llllll}
  Variable            & Df & Sum Sq    & Mean Sq   & F Value   & Pr(\textgreater{}F)    \\
  Year                & 9  & 7.060E+01 & 7.800E+00 & 8.353E+01 & \textless 2.20E-16 *** \\
  GI Region           & 54 & 1.507E+03 & 2.790E+01 & 2.972E+02 & \textless 2.20E-16 *** \\
  Area Harvested      & 1  & 3.211E+03 & 3.211E+03 & 3.419E+04 & \textless 2.20E-16 *** \\
  Water Used          & 1  & 1.040E+01 & 1.040E+01 & 1.103E+02 & \textless 2.20E-16 *** \\
  Scope One Emissions & 1  & 6.600E+00 & 6.600E+00 & 7.056E+01 & \textless 2.20E-16 ***
  \end{tabular}
\end{table}

\begin{table}[]
    \label{tab:tab7}
    \caption{Model 2 ANOVA summarising variable significance at the .5 level.}
    \begin{tabular}{llllll}
    Variable                    & Df  & Sum Sq    & Mean Sq   & F Value   & Pr(\textgreater{}F)    \\
    Area Harvested              & 1   & 2.407E+03 & 2.407E+03 & 1.080E+04 & \textless 2.20E-16 *** \\
    Scope One Emissions         & 1   & 3.989E+01 & 3.989E+01 & 1.789E+02 & \textless 2.20E-16 *** \\
    Water Used                  & 1   & 5.500E+02 & 5.500E+02 & 2.467E+03 & \textless 2.20E-16 *** \\
    Area Harvested*Scope One Emissions & 1 & 6.921E+01 & 6.921E+01 & 3.104E+02 & \textless 2.20E-16 *** \\
    Area Harvested * Water Used & 1   & 1.040E+00 & 1.040E+00 & 4.686E+00 & 3.045E-02 **           \\
    Year * GI Region            & 424 & 1.144E+03 & 2.700E+00 & 1.210E+01 & \textless 2.20E-16 ***
    \end{tabular}
\end{table}

\begin{table}[]
    \caption{Model 3 ANOVA summarising variable significance at the .5 level.}
    \label{tab:tab8}
    \begin{tabular}{llllll}
    Variable            & Df & Sum Sq    & Mean Sq   & F Value   & Pr(\textgreater{}F)    \\
    Year                & 6  & 1.324E+01 & 2.210E+00 & 8.748E+01 & \textless 2.20E-16 *** \\
    GI Region           & 50 & 6.498E+02 & 1.300E+01 & 5.151E+02 & \textless 2.20E-16 *** \\
    Area Harvested      & 1  & 2.142E+03 & 2.142E+03 & 8.491E+04 & \textless 2.20E-16 *** \\
    Water Used          & 1  & 3.200E-01 & 3.200E-01 & 1.259E+01 & 3.947E-04 **           \\
    Scope One Emissions & 1  & 4.000E-02 & 4.000E-02 & 1.492E+00 & 2.221E-01             
    \end{tabular}
\end{table}

\begin{table}[]
  \label{tab:tab9}
  \caption{Model 4 ANOVA summarising variable significance at the .5 level.}
  \begin{tabular}{llllll}
  Variable            & Df  & Sum Sq    & Mean Sq   & F Value   & Pr(\textgreater{}F)    \\
  Area Harvested      & 1   & 2.066E+03 & 2.066E+03 & 5.700E+04 & \textless 2.20E-16 *** \\
  Scope One Emissions & 1   & 6.000E-02 & 6.000E-02 & 1.569E+00 & 2.105E-01              \\
  Water Used          & 1   & 2.014E+02 & 2.014E+02 & 5.557E+03 & \textless 2.20E-16 *** \\
  Area Harvested*Scope One Emissions & 1 & 5.246E+01 & 5.246E+01 & 1.448E+03 & \textless 2.20E-16 *** \\
  Area Harvested * Water Used        & 1 & 7.270E+00 & 7.270E+00 & 2.005E+02 & \textless 2.20E-16 *** \\
  Year * GI Region    & 243 & 4.546E+02 & 1.870E+00 & 5.162E+01 & \textless 2.20E-16 ***
  \end{tabular}
\end{table}

\begin{table}[]
  \label{tab:tab 10}
  \caption{Comparison of Model Residuals}
  \begin{tabular}{llll}
        & Df   & Sum Sq    & Mean Sq   \\
  Model 1 & 5231 & 4.913E+02 & 1.000E-01 \\
  Model 2 & 4868 & 1.085E+03 & 2.200E-01 \\
  Model 3 & 2818 & 7.111E+01 & 3.000E-02 \\
  Model 4 & 2629 & 9.528E+01 & 4.000E-02
  \end{tabular}
\end{table}

\begin{table}[]
  \caption{Comparison of Model performance.}
  \label{tab:tab 11}
  \begin{tabular}{llllll}
          & RSE       & R2        & Adjusted 			R2 & F-statistic & P-Value           \\
  Model 1 & 3.065E-01 & 9.072E-01 & 9.061E-01      & 7.753E+02   & \textless 2.2e-16 \\
  Model 2 & 4.722E-01 & 7.951E-01 & 7.770E-01      & 4.403E+01   & \textless 2.2e-16 \\
  Model 3 & 1.589E-01 & 9.753E-01 & 9.748E-01      & 1.885E+03   & \textless 2.2e-16 \\
  Model 4 & 1.904E-01 & 9.669E-01 & 9.638E-01      & 3.095E+02   & \textless 2.2e-16
  \end{tabular}
\end{table}

Limitations included overestimating yield for models 1 and 2, (see Figures 1 and 2) and underestimating crop value in models 3 and 4 (see Figures 3 and 4).

 Reviewing the data to uncover reasons for this included the use of binary variables such as the utilisation of renewable energy, contractors, and the occurrence of disease, fire and frost; however none of these variables were able to explain why some vineyards produced less, or why other vineyards sold at higher prices than predicted. A wide variety of these influences were likely already explained within the use of year and GI Region, or the interaction of both variables. The change between some regions was dramatic, with particularly warmer and drier regions producing much higher volumes of grapes at lower prices (See Figures 5 and 6). The use of other variables and methods, specifically splines, were able to create a more normally distributed set of residuals but at a drastically reduced accuracy when comparing R2 and RSE. The introduction of known average prices per tonne also helped increase R2 values a small amount; it is important to not that it is common practice for wineries to purchase grapes at a regional average rate, likely resulting in much less variance within a region.

%\begin{figure}
%  \includegraphics[width=\linewidth]{figure1.jpg}
%  \caption{Model 1 residuals vs fitted (left) and QQ plot (right).}
%  \label{fig:fig1}
%\end{figure}

%\begin{figure}
%  \includegraphics[width=\linewidth]{figure2.jpg}
%  \caption{Model 2 residuals vs fitted (left) and QQ plot (right).}
%  \label{fig:fig2}
%\end{figure}

%\begin{figure}
%  \includegraphics[width=\linewidth]{figure3.jpg}
%  \caption{Model 3 residuals vs fitted (left) and QQ plot (right).}
%  \label{fig:fig3}
%\end{figure}

%\begin{figure}
%  \includegraphics[width=\linewidth]{figure4.jpg}
%  \caption{Model 4 residuals vs fitted (left) and QQ plot (right).}
%  \label{fig:fig4}
%\end{figure}

%\begin{figure}
%  \includegraphics[width=\linewidth]{figure5.jpg}
%  \caption{Average Yield per hectare for each GI Region.}
%  \label{fig:fig5}
%\end{figure}

%\begin{figure}
%  \includegraphics[width=\linewidth]{figure6.jpg}
%  \caption{Average Price Per Tonne for each GI Region.}
%  \label{fig:fig6}
%\end{figure}

% This belongs in the discussion
The relationship between scope one emissions and the response variables that included average sales price

It is possible that the relationships between scope one emissions and the response variables were closely tied to a vineyards area. This possibility could be explained through the emissions

Noting that irrigation systems use fuel
and that the application of water was a significant variable in each model
scope one emissions' lack of significance and contribution given its F-statistics (See Tables 7 and 8),
 indicated that it is possible other vineyard activities requiring fuel are not as determining factors for a vineyards grape quality.

\printbibliography{}
\end{sloppypar}
\end{document}

### 2023-06-22 article.tex
updated r analysis file, fixing continuity in variable use and fixing graph construction. updated reference list to include three articles on value vs quantity tradeoff in agri sector. added a requirements file for python libraries. updated the article with further edits and tables - a lot still to do.

@@
 \section{Introduction}
-When assessing strategies for viticulture there is a trade-off between the amount of resources invested and the resultant quality and quantity of grapes produced \autocite{montalvo-falconSustainabilityResearchWine2023}.
-Due to recent economic, social and environmental pressures, such as: bush fires, tariffs, and shifting attitudes to alcohol consumption; the chosen emphasis taken in balancing the value-productivity tradeoff has become more important \autocite{toumbourouStudentSurveyTrends2018,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2022}.
-When considering pressures such as those prior listed, the integration of viticulture within the Australian wine industry is a compounding effect.
 
-In this analysis we observed relationships between quantity and quality through the use of linear models.
-For the purpose of this study; quality was defined by the financial value of grapes per tonne.
+The global focus on sustainability in agronomic industries has changed the way in which these enterprises do business.
+When strategies for a sustainable industry are assessed, there is a trade-off between balancing the amount of resources invested and the resultant quality verses quantity of crops produced.
+This dilemma exists across agriculture, sharing fundemental considerations such as water and nitrogen input \autocite{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016,zhuEffectsNitrogenLevel2017}.
+Within viticulture (the cultivation of grapes for wine production) quality is driven by its integration within the wine industry; with a wine's potential quality being initially defined through the chemical makeup of the grapes used in its production.
+The consideration of sustainability is further complicated by environmental and socio-demographic pressures.
+In the Australian context, these include biosecurity, climate and international market demands.
+
+In this analysis we observe relationships between quantity and quality through the use of linear models.
+Quality can be defined in a variety of ways for grapes, such as # # #.
+For the purpose of this study quality was defined by the financial value of grapes per tonne; this was due to varying differences in defining quality between grape varietals.
+
 An extensive amount of research into a variety of factor's effect on grape quality and yield exists.
@@
 Limitations included overestimating yield for models 1 and 2, (see Figures 1 and 2) and underestimating crop value in models 3 and 4 (see Figures 3 and 4).
-
+% This section needs to be heavily revised. Why are there tails on the distributions? What does a heavy tail indicate?
+% What does it look like with respect to disaster events?
+% split the discussion sections out of the results!! Even the questions I am asking are answered in the discussion, however they need to be highlighted as aspects of the models here beforehand.
  Reviewing the data to uncover reasons for this included the use of binary variables such as the utilisation of renewable energy, contractors, and the occurrence of disease, fire and frost; however none of these variables were able to explain why some vineyards produced less, or why other vineyards sold at higher prices than predicted.
@@
 
+% It is odd to have a paragraph on its own without a label, it does look a bit lost in the article.
+The correlation between average sales price and yield was a negative trend (see table 2); the contributing factors to yield and average sales price was ???.
+Correlation values showed that water and emissions increased with yield but decreased with average sale price (see Table 4).
+In alternative attempts at models it was found that without the incorporation of GI Region or year the predictions greatly under performed.
+The possible reason behind this effect was that different strategies are likely employed between different regions, where some regions target the mass production of cheaper grapes over quality.
+% ALSO THAT CATASTROPHES HAPPEN EVEN ON SMALL SCALES THis needs to be tied up better.
+ This is most notable when grouping regions by climate, especially when considering GI Regions in the 'Hot Very Dry' climate (see Figure 7).
+The effect of climate in the models was not more significant than the more granular use of GI regions.
+The interaction between year and GI Region likely accounted for localised events such as bushfires, which would be impactful, but only at a local level in both time and space.
+
+%\begin{figure}
+%  \includegraphics[width=\linewidth]{figure7.jpg}
+%  \caption{A depiction of the relationship between Value of tonnes harvested per hectare and yield per hectare.Both variables have been transformed to be scaled by standard deviation with mean 0.}
+%  \label{fig:fig7}
+%\end{figure}
+
+\subsection{Model Validation}
+To validate the performance of these models k-fold cross validation was used.
+This was done using 10 folds, k=10, repeated 100 times.
+The models performed similarly to their original counter parts (see Table 11).
+
+% really this is it?
+
+\begin{table}[]
+  \label{tab:tab12}
+  \caption{Model validation using k-fold cross validation, for 10 folds repeated 100 times.}
+  \begin{tabular}{llll}
+          & RMSE      & R2        & MAE       \\
+  Model 1 & 3.087E-01 & 9.045E-01 & 2.165E-01 \\
+  Model 2 & 5.104E-01 & 7.409E-01 & 3.493E-01 \\
+  Model 3 & 1.652E-01 & 9.723E-01 & 1.008E-01 \\
+  Model 4 & 2.235E-01 & 9.500E-01 & 1.279E-01
+  \end{tabular}
+  \end{table}
+
+\section{Discussion}
+% this needs serious revision.
+This study investigated the general relationships between input resources of a vineyard, including fuel and water, and the outputs including yield and value.
+Some regions appeared to produce many low quality grapes at scale compared to attempting to produce fewer higher quality grapes.
+This behaviour can be observed when reviewing Wine Australia’s annual reports, where it is apparent that warm inland regions such as the Riverland are known to only produce large amounts of lower graded grapes \autocite{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
+Comparatively, regions such as Tasmania only produce A grade grapes but in much smaller quantities than the Riverland.
+Knowing that the difference in pricing per tonne can exceed a magnitude of 10 between grades E and A, the operations in regions that target different grades would have varied priorities.
+However, some regions such as the Yarra Valley produce a Variety of different grades of grapes, from C to A, highlighting that vineyard priorities, although may be somewhat present within regional classifications, are not necessarily aligned within a given region.
+
+The opportunity to target different grades of grapes may not always be available, with some regions being more renowned than others, and likely to be sought after regardless \autocite{hallidayAustralianWineEncyclopedia2009}.
+The Barossa is an example of this, known for its quality could also lend itself to a bias in purchasers not considering other regions that may be capable of similar quality.
+This effect could stifle the potential for market opportunities within these lesser known regions.
+A further possibility is that there may be regional upper limits with the relationship between resource input and the value gained becoming no longer proportional due to diminishing returns.
+Climate was considered to be a large determinant of the ability to grow a larger quantity of grapes, as well as a determinant in grape quality \autocite{agostaRegionalClimateVariability2012}; however there were vineyards in similar regions that were able to produce exceptionally better results than others (See Figure 7).
+
+The issue of model 1 and 2 over predicting yield, may have been due to preventative measures brought on by regional pressures such as fire, frost and disease.
+Where, more resources were required to prevent these issues from spreading within a region, thus disproportionately effecting some vineyards compared to others locally.
+This type of maintenance is not well captured especially when considering that some regions, those in warmer areas are not as prone to disease as cooler climates and could potentially have lower operating costs per hectare.
+This could create a discrepancy in vineyards that utilise preventative measures in wetter regions, as opposed to those who do not, and thus expend less fuel and energy but risk disease.
+When reviewing the differences between regions it is important to consider that vineyards in Hot Very Dry areas can be hundreds of times the size of those in other regions.
+It is interesting that while area, although significantly correlated to the ratio of yield to area, was still lower than water and about the same as emissions.
+This points to economies of scale playing a role but still being only one consideration alongside the potential resources that can be used.
+The negative trend between size and average sales price could also be a side effect of mass supply verse demand, especially when looking at the level of difference in production of some vineyards (see Table 4).
+The relationships between yield, value and area are not simply about efficiently producing the most grapes; sales price and by association grape quality, are integral to the profitability, and this is strongly linked to resource use and thus the longevity and sustainability of a vineyard.
+
+Literature shows that there are many on-the-ground decisions that influence both quality and yield.
+Where these decisions are governed by complex physical and social forces such as international market demands, disease pressures and natural disasters \autocite{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}.
+Many of these occurrences being highlighted throughout the past decades vintage reports \autocite{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
+%This definitely needs to be broken up, what were these occurences, which year was significant for what?
+It is also important to consider that these reports show that the warm inland regions have seen a decline in profit during this period, as they were often compared to other regions that focused more on quality than quantity.
+This is an important consideration, as the size of some of these vineyards when considering their ratio of value to area would only require a marginal increase to out compete other regions.
+There are also differences when comparing winegrowers to other agricultural industries as they are vertically integrated within the wine industry, tying them to secondary and tertiary industries, such as wine production, packaging, transport and sales.
+This results in unique issues and considerations for each vineyard, where these on-the-ground decisions may be influenced by other wine industry's choices, such as the use of sustainable practices in vineyards as a requirement for sale in overseas markets; notably these interactions are further complicated by some winegrowers being totally integrated into wine companies, while others are not (Knight et al., 2019).
+Incorporating such decisions into the model could help describe the contributing factors to regional differences beyond resource consumption and regional differences.
+
+Having more data for each region would also be an improvement, allowing greater comparison between regions.
+More variables may also help to discern vineyards that can produce larger volumes of grapes at higher prices.
+The use of semi transparent tools such as random forests and decision trees alongside more variables and data may help to uncover the reasons for values that were under or over estimated.
+These differences could be caused by the use of alternative sustainable practices in the field.
+While there is evidence to suggest that environmentally sustainable practices can reduce costs, increase efficiency, whilst improving the quality of grapes, more research is needed to link these benefits across different regions and climates \autocite{baianoOverviewSustainabilityWine2021,marianiSustainableWinegrowingCurrent2015,montalvo-falconSustainabilityResearchWine2023}.
+
 % This belongs in the discussion

### 2023-06-22 analysis.r

@@ -1,6 +1,8 @@
+library(ggplot2)
+
 # data_prep.py has to be executed in order for df.csv to exist
 set.seed(100)
-df <- read.csv("df.csv")
+df <- read.csv("dfnh.csv")
 
 model1 <- lm(tonnes_grapes_harvested ~ data_year_id+giregion+area_harvested+water_used+scope1,data=df)
 model2 <- lm(ha_tonnes_grapes_harvested ~ area_harvested*((scope1)+water_used)+data_year_id:giregion, data=df)
@@ -10,9 +12,13 @@ model4 <- lm(ha_value ~ area_harvested*((scope1)+water_used)+data_year_id:giregi
 # We create our first model
 
 # MODEL 1
-anova(lm(tonnes_grapes_harvested ~ data_year_id+giregion+area_harvested+water_used+scope1,data=df))
-summary(lm(tonnes_grapes_harvested ~ data_year_id+giregion+area_harvested+water_used+scope1,data=df))
-#Linear Regression 
+anova(model1)
+summary(model1)
+model1.residuals <- residuals(model1)
+model1.fitted <- fitted(model1)
+ggplot(data.frame(y=model1.residuals), aes(sample=y)) + stat_qq() + stat_qq_line()
+ggplot(data.frame(y=model1.residuals,x=model1.fitted), aes(x=x,y=y)) + geom_point() + geom_hline(yintercept=0)
+#Linear Regression
 #
 #6068 samples
 #   5 predictor
@@ -30,8 +36,8 @@ summary(lm(tonnes_grapes_harvested ~ data_year_id+giregion+area_harvested+water_
 
 # MODEL 2
 # Lets look at ratios
-anova(lm(df$ha_tonnes_grapes_harvested ~ df$area_harvested*((df$scope1)+df$water_used)+df$data_year_id:df$giregion))
-summary(lm(df$ha_tonnes_grapes_harvested ~ df$area_harvested*((df$scope1)+df$water_used)+df$data_year_id:df$giregion))
+anova(model2)
+summary(model2)
 #Linear Regression 
 #
 #6068 samples
@@ -49,8 +55,8 @@ summary(lm(df$ha_tonnes_grapes_harvested ~ df$area_harvested*((df$scope1)+df$wat
 
 # MODEL 3
 # We add Value to the model :P
-anova(lm(value ~ data_year_id+giregion+area_harvested+water_used+scope1,data=df))
[…]

### 2023-06-26 analysis.r
Added questionaire and update to analysis

@@ -2,7 +2,7 @@ library(ggplot2)
 
 # data_prep.py has to be executed in order for df.csv to exist
 set.seed(100)
-df <- read.csv("dfnh.csv")
+df <- read.csv("df.csv")
 
 model1 <- lm(tonnes_grapes_harvested ~ data_year_id+giregion+area_harvested+water_used+scope1,data=df)
 model2 <- lm(ha_tonnes_grapes_harvested ~ area_harvested*((scope1)+water_used)+data_year_id:giregion, data=df)

### 2023-06-29 article.tex
Fixed referencesgit add article.tex! It is all in the correct review style now.

@@
-\documentclass[12pt,australian]{article}
-\usepackage{lingmacros}
-\usepackage{tree-dvips}
-\usepackage[utf8]{inputenc}
-\usepackage{csquotes,xpatch}% recommended
-\usepackage[australian]{babel}
-\usepackage{amsmath}
-\usepackage[backend=biber,<options>]{biblatex}
-\usepackage{graphicx}
-\addbibresource[location=local]{references.bib}
+%% 
+%% Copyright 2007-2020 Elsevier Ltd
+%% 
+%% This file is part of the 'Elsarticle Bundle'.
+%% ---------------------------------------------
+%% 
+%% It may be distributed under the conditions of the LaTeX Project Public
+%% License, either version 1.2 of this license or (at your option) any
+%% later version.  The latest version of this license is in
+%%    http://www.latex-project.org/lppl.txt
+%% and version 1.2 or later is part of all distributions of LaTeX
+%% version 1999/12/01 or later.
+%% 
+%% The list of all files belonging to the 'Elsarticle Bundle' is
+%% given in the file `manifest.txt'.
+%% 
+%% Template article for Elsevier's document class `elsarticle'
+%% with harvard style bibliographic references
+
+\documentclass[review,12pt,authoryear]{elsarticle}
+
+%% Use the option review to obtain double line spacing
+%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
+
+%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
+%% for a journal layout:
+%% \documentclass[final,1p,times,authoryear]{elsarticle}
+%% \documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
+%% \documentclass[final,3p,times,authoryear]{elsarticle}
+%% \documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
+%% \documentclass[final,5p,times,authoryear]{elsarticle}
+%% \documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}
+
+%% For including figures, graphicx.sty has been loaded in
+%% elsarticle.cls. If you prefer to use the old commands
+%% please give \usepackage{epsfig}
+
+%% The amssymb package provides various useful mathematical symbols
+\usepackage{amssymb}
+%% The amsthm package provides extended theorem environments
+%% \usepackage{amsthm}
+
+%% The lineno packages adds line numbers. Start line numbering with
+%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
+%% for the whole article with \linenumbers.
+\usepackage{lineno}
+
+\journal{Geography and Sustainability}
+
+\begin{document}
+\begin{linenumbers}
+\begin{frontmatter}
+
+%% Title, authors and addresses
+
+%% use the tnoteref command within \title for footnotes;
+%% use the tnotetext command for theassociated footnote;
+%% use the fnref command within \author or \affiliation for footnotes;
+%% use the fntext command for theassociated footnote;
+%% use the corref command within \author for corresponding author footnotes;
+%% use the cortext command for theassociated footnote;
+%% use the ead command for the email address,
+%% and the form \ead[url] for the home page:
+%% \title{Title\tnoteref{label1}}
+%% \tnotetext[label1]{}
+%% \author{Name\corref{cor1}\fnref{label2}}
+%% \ead{email address}
+%% \ead[url]{home page}
+%% \fntext[label2]{}
+%% \cortext[cor1]{}
+%% \affiliation{organization={},
+%%            addressline={}, 
+%%            city={},
+%%            postcode={}, 
+%%            state={},
+%%            country={}}
+%% \fntext[label3]{}
 
     \title{Resource Use and the Value-Productivity Tradeoff in Australian Winegrowing Regions}
-\author{Bryce Polley}
+
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1,label2]{}
+%% \affiliation[label1]{organization={},
+%%             addressline={},
+%%             city={},
+%%             postcode={},
+%%             state={},
+%%             country={}}
+%%
+%% \affiliation[label2]{organization={},
+%%             addressline={},
+%%             city={},
+%%             postcode={},
+%%             state={},
+%%             country={}}
+
+    \affiliation[label1]{organization={QUT},
+                 addressline={},
+                 city={},
+                 postcode={},
+                 state={QLD},
+                 country={}}
+    \affiliation[label2]{organization={AWRI},
+                 addressline={},
+                 city={},
+                 postcode={},
+                 state={SA},
+                 country={}}
+    \affiliation[label3]{organization={Food Agility CRC},
+                 addressline={},
+                 city={},
+                 postcode={},
+                 state={Vic},
+                 country={}}
+\author[label1,label2,label3]{Bryce Polley}
 \date{20/06/2023}
 
-\begin{document}
-\begin{sloppypar}
+    \begin{abstract}
+      %% Text of abstract
+      Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+      \end{abstract}
+
+      %%Graphical abstract
+      \begin{graphicalabstract}
+        \includegraphics{graphical_abstract.jpeg}
+      \end{graphicalabstract}
+
+      \begin{keyword}
+      %% keywords here, in the form: keyword \sep keyword
+      keyword one \sep keyword two
+      %% PACS codes here, in the form: \PACS code \sep code
+      \PACS 0000 \sep 1111
+      %% MSC codes here, in the form: \MSC code \sep code
+      %% or \MSC[2008] code \sep code (2000 is the default)
+      \MSC 0000 \sep 1111
+      \end{keyword}
+
+      %%Research highlights
+      \begin{highlights}
+        \item Research highlight 1
+        \item Research highlight 2
+        \end{highlights}
+
+      \end{frontmatter}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                main text             %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \section{Introduction}
-
 The global focus on sustainability in agronomic industries has changed the way in which these enterprises do business.
 When strategies for a sustainable industry are assessed, there is a trade-off between balancing the amount of resources invested and the resultant quality verses quantity of crops produced.
-This dilemma exists across agriculture, sharing fundemental considerations such as water and nitrogen input \autocite{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016,zhuEffectsNitrogenLevel2017}.
+This dilemma exists across agriculture, sharing fundemental considerations such as water and nitrogen input
+\citep{abadCoverCropsViticulture2021}
+
+\cite{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016, zhuEffectsNitrogenLevel2017}. 
+\citet{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016, zhuEffectsNitrogenLevel2017}.
+
 Within viticulture (the cultivation of grapes for wine production) quality is driven by its integration within the wine industry; with a wine's potential quality being initially defined through the chemical makeup of the grapes used in its production.
@@
 In this analysis we observe relationships between quantity and quality through the use of linear models.
-Quality can be defined in a variety of ways for grapes, such as # # #.
+Quality can be defined in a variety of ways for grapes, such as *.
 For the purpose of this study quality was defined by the financial value of grapes per tonne; this was due to varying differences in defining quality between grape varietals.
@@
 An extensive amount of research into a variety of factor's effect on grape quality and yield exists.
-Due to the lack of long-term and in-depth data sources, individual effects are often studied in isolation \autocite{abbalDecisionSupportSystem2016}.
-The lack of consolidated datasets also restricts the ability to gain statistical insight at a large scale across multiple regions \autocite{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
+Due to the lack of long-term and in-depth data sources, individual effects are often studied in isolation \citep{abbalDecisionSupportSystem2016}.
+The lack of consolidated datasets also restricts the ability to gain statistical insight at a large scale across multiple regions \citep{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
 The dataset used for this analysis includes data collected for the past 10 years from a multitude of vineyards located over a diverse range of Australian winegrowing regions.
@@
         Model 1 & Yield & Water Used, Scope 1 Emissions & Area Harvested, Year, GI Region & N\textbackslash A \\
-        Model 2 & ${\text{Yield}}\over{\text{Area Harvested}}$ & Water Used, Scope 1 Emissions & Area Harvested, Year, GI Region & Area Harvested * Scope 1 Emissions, Area Harvested * Water Use, Year * Region \\
-        Model 3 & \small{$\text{Yield} {\times} \text{Average Sale Price}$} &  Water Used, Scope 1 Emissions& Area Harvested, Year, GI Region & Area Harvested * Scope 1 Emissions, Area Harvested * Water Use, Year * Region  \\
-        Model 4 & ${\text{Yield} \times \text{Average Sale Price}}\over{\text{Area Harvested}}$ & Water Used, Scope 1 Emissions & Area Harvested, Year, GI Region  & N\textbackslash A
+        Model 2 & ${ \textrm{Yield}}\over{ \textrm{Area Harvested}}$ & Water Used, Scope 1 Emissions & Area Harvested, Year, GI Region & Area Harvested * Scope 1 Emissions, Area Harvested * Water Use, Year * Region \\
+        Model 3 & \small{$\textrm{Yield} {\times} \textrm{Average Sale Price}$} &  Water Used, Scope 1 Emissions& Area Harvested, Year, GI Region & Area Harvested * Scope 1 Emissions, Area Harvested * Water Use, Year * Region  \\
+        Model 4 & ${\textrm{Yield} \times \textrm{Average Sale Price}}\over{\textrm{Area Harvested}}$ & Water Used, Scope 1 Emissions & Area Harvested, Year, GI Region  & N\textbackslash A
       \end{tabular}
@@
 The data used in this analysis was provided by Sustainable Winegrowing Australia, Australia's national wine industry sustainability program.
-Sustainable winegrowing Australia aims to facilitate grape-growers and winemakers in demonstrating and improving their sustainability \autocite{swaSustainableWingrowingAustralia2022}.
+Sustainable winegrowing Australia aims to facilitate grape-growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
 Data recorded by Sustainable Winegrowing Australia is entered manually by winegrowers using a web based interface; with some fields being optional.
@@
 Two values for average sale price were removed from the dataset, due to a recording of \$1.
-Unreported values for average prices per tonne were filled in using regional averages taken from Wine Australia's annual reports, where they were available \autocite{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2012,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
+Unreported values for average prices per tonne were filled in using regional averages taken from Wine Australia's annual reports, where they were available \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2012,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
 
@@
 
-was used to convert the quantity of fuel in litres, $Q$, using a prescribed Energy Content, $EC$, and emission factors of scope one, $EF1$, and scope three, $EF3$, to tonnes of Carbon Dioxide equivalent, $tCO2e$ \autocite{departmentofclimatechangeenergytheenvironmentandwaterAustralianNationalGreenhouse2022}.
+was used to convert the quantity of fuel in litres, $Q$, using a prescribed Energy Content, $EC$, and emission factors of scope one, $EF1$, and scope three, $EF3$, to tonnes of Carbon Dioxide equivalent, $tCO2e$ \citep{departmentofclimatechangeenergytheenvironmentandwaterAustralianNationalGreenhouse2022}.
 
@@
 \subsection{Region}
-The site of a vineyard predetermines several physical parameters such as climate, geology and soil; making location a widely considered key determinant of grape yield and quality \autocite{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
+The site of a vineyard predetermines several physical parameters such as climate, geology and soil; making location a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
 Differences in vineyard locations were captured through the use of Geographical Indicator Regions (GI Regions).
-Each GI Region has its own unique mixture of climatic and geophsical properties that describes a unique winegrowing region within Australia; these regions were predefined by Wine Australia \autocite{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
+Each GI Region has its own unique mixture of climatic and geophsical properties that describes a unique winegrowing region within Australia; these regions were predefined by Wine Australia \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
 
-The climatic properties of a GI Region are summarised in the Sustainable Winegrowing Australia user manual \autocite{swaSustainableWinegrowingAustralia2021}.
+The climatic properties of a GI Region are summarised in the Sustainable Winegrowing Australia user manual \citep{swaSustainableWinegrowingAustralia2021}.
 The user manual describes climates by rainfall and temperature, creating supersets of Regions of similar climatic properties.
@@
 
-Data preprocessing, such as logarithmic transforms, was done using the Python programming language \autocite{g.vanrossumPythonTutorialTechnical1995}.
-Linear models were created using the R statistical programming language \autocite{rcoreteamLanguageEnvironmentStatistical2021}.
+Data preprocessing, such as logarithmic transforms, was done using the Python programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
+Linear models were created using the R statistical programming language \citep{rcoreteamLanguageEnvironmentStatistical2021}.
 These models were created iteratively to explore a variety of variable interactions and approaches to modelling the data.
@@
 \subsection{Model Validation}
-Models were validated using K-fold cross validation calculated through the R Caret Package \autocite{kuhnBuildingPredictiveModels2008}.
+Models were validated using K-fold cross validation calculated through the R Caret Package \citep{kuhnBuildingPredictiveModels2008}.
 K-fold cross validation works by removing a subset of data from the sample used to train models and then predicts those variables to determine how sensitive the model is to changes in the sample data.
@@
   \begin{tabular}{cccccccc}
-  Variable                             & Yield      & Area       & Water Used & Scope One Emissions & $\text{Yield}\over\text{Area}$ & Average Price Per Tonne & ${\text{Average Price per tonne}\over\text{Area}}$ \\
+  Variable                             & Yield      & Area       & Water Used & Scope One Emissions & $\textrm{Yield}\over\textrm{Area}$ & Average Price Per Tonne & ${\textrm{Average Price per tonne}\over\textrm{Area}}$ \\
   Yield                                & 1.000E+00  & 7.440E-01  & -4.309E-03 & 7.290E-01           & 3.500E-01            & -2.262E-01              & -1.644E-01                           \\
@@
   Scope One Emissions                  & 7.290E-01  & 8.921E-01  & -1.929E-03 & 1.000E+00           & 9.357E-02            & -9.422E-02              & -1.933E-01                           \\
-  $\text{Yield}\over\text{Area}$                 & 3.500E-01  & 7.854E-02  & -5.600E-03 & 9.357E-02           & 1.000E+00            & -4.849E-01              & -1.698E-01                           \\
+  $\textrm{Yield}\over\textrm{Area}$                 & 3.500E-01  & 7.854E-02  & -5.600E-03 & 9.357E-02           & 1.000E+00            & -4.849E-01              & -1.698E-01                           \\
   Average Price Per Tonne              & -2.262E-01 & -1.178E-01 & -3.562E-02 & -9.422E-02          & -4.849E-01           & 1.000E+00               & 4.732E-01                            \\
-  ${\text{Average Price per tonne}\over\text{Area}}$ & -1.644E-01 & -2.042E-01 & -2.669E-02 & -1.933E-01          & -1.698E-01           & 4.732E-01               & 1.000E+00
+  ${\textrm{Average Price per tonne}\over\textrm{Area}}$ & -1.644E-01 & -2.042E-01 & -2.669E-02 & -1.933E-01          & -1.698E-01           & 4.732E-01               & 1.000E+00
   \end{tabular}
@@
     \begin{tabular}{clllllll}
-    Variable                                           & \multicolumn{1}{c}{Yield} & \multicolumn{1}{c}{Area} & \multicolumn{1}{c}{Water Used} & \multicolumn{1}{c}{Scope One Emissions} & \multicolumn{1}{c}{$\text{Yield}\over\text{Area}$} & \multicolumn{1}{c}{Average Price Per Tonne} & \multicolumn{1}{c}{${\text{Average Price per tonne}\over\text{Area}}$} \\
+    Variable                                           & \multicolumn{1}{c}{Yield} & \multicolumn{1}{c}{Area} & \multicolumn{1}{c}{Water Used} & \multicolumn{1}{c}{Scope One Emissions} & \multicolumn{1}{c}{$\textrm{Yield}\over\textrm{Area}$} & \multicolumn{1}{c}{Average Price Per Tonne} & \multicolumn{1}{c}{${\textrm{Average Price per tonne}\over\textrm{Area}}$} \\
     Yield                                              & 1.000E+00                 & 8.822E-01                & 8.245E-01                      & 7.617E-01                               & 9.353E-01                                          & -4.591E-01                                  & -8.918E-01                                                             \\
@@
     Scope One Emissions                                & 7.617E-01                 & 8.311E-01                & 6.668E-01                      & 1.000E+00                               & 6.086E-01                                          & -1.559E-01                                  & -7.063E-01                                                             \\
-    $\text{Yield}\over\text{Area}$                     & 9.353E-01                 & 6.742E-01                & 7.292E-01                      & 6.086E-01                               & 1.000E+00                                          & -5.625E-01                                  & -8.076E-01                                                             \\
+    $\textrm{Yield}\over\textrm{Area}$                     & 9.353E-01                 & 6.742E-01                & 7.292E-01                      & 6.086E-01                               & 1.000E+00                                          & -5.625E-01                                  & -8.076E-01                                                             \\
     Average Price Per Tonne                            & -4.591E-01                & -1.911E-01               & -4.881E-01                     & -1.559E-01                              & -5.625E-01                                         & 1.000E+00                                   & 6.592E-01                                                              \\
-    ${\text{Average Price per tonne}\over\text{Area}}$ & -8.918E-01                & -8.474E-01               & -8.300E-01                     & -7.063E-01                              & -8.076E-01                                         & 6.592E-01                                   & 1.000E+00
+    ${\textrm{Average Price per tonne}\over\textrm{Area}}$ & -8.918E-01                & -8.474E-01               & -8.300E-01                     & -7.063E-01                              & -8.076E-01                                         & 6.592E-01                                   & 1.000E+00
     \end{tabular}
@@
       Scope One Emissions                                & 8.883E-01  \\
-      $\text{Yield}\over\text{Area}$                     & 6.836E-01  \\
+      $\textrm{Yield}\over\textrm{Area}$                     & 6.836E-01  \\
       Average Price Per Tonne                            & 5.600E-02  \\
-      ${\text{Average Price per tonne}\over\text{Area}}$ & 1.522E-01 
+      ${\textrm{Average Price per tonne}\over\textrm{Area}}$ & 1.522E-01 
       \end{tabular}
@@
         \multicolumn{1}{c}{Scope One Emissions} &
-        \multicolumn{1}{c}{$\text{Yield}\over\text{Area}$} &
+        \multicolumn{1}{c}{$\textrm{Yield}\over\textrm{Area}$} &
         \multicolumn{1}{c}{Average Price Per Tonne} &
-        \multicolumn{1}{c}{${\text{Average Price per tonne}\over\text{Area}}$} \\
+        \multicolumn{1}{c}{${\textrm{Average Price per tonne}\over\textrm{Area}}$} \\
       Yield                                              & 1.000E+00  & 8.822E-01  & 8.245E-01  & 7.617E-01  & 9.353E-01  & -4.591E-01 & -8.918E-01 \\
@@
       Scope One Emissions                                & 7.617E-01  & 8.311E-01  & 6.668E-01  & 1.000E+00  & 6.086E-01  & -1.559E-01 & -7.063E-01 \\
-      $\text{Yield}\over\text{Area}$                     & 9.353E-01  & 6.742E-01  & 7.292E-01  & 6.086E-01  & 1.000E+00  & -5.625E-01 & -8.076E-01 \\
+      $\textrm{Yield}\over\textrm{Area}$                     & 9.353E-01  & 6.742E-01  & 7.292E-01  & 6.086E-01  & 1.000E+00  & -5.625E-01 & -8.076E-01 \\
       Average Price Per Tonne                            & -4.591E-01 & -1.911E-01 & -4.881E-01 & -1.559E-01 & -5.625E-01 & 1.000E+00  & 6.592E-01  \\
-      ${\text{Average Price per tonne}\over\text{Area}}$ & -8.918E-01 & -8.474E-01 & -8.300E-01 & -7.063E-01 & -8.076E-01 & 6.592E-01  & 1.000E+00 
+      ${\textrm{Average Price per tonne}\over\textrm{Area}}$ & -8.918E-01 & -8.474E-01 & -8.300E-01 & -7.063E-01 & -8.076E-01 & 6.592E-01  & 1.000E+00 
       \end{tabular}
@@
 Some regions appeared to produce many low quality grapes at scale compared to attempting to produce fewer higher quality grapes.
-This behaviour can be observed when reviewing Wine Australia’s annual reports, where it is apparent that warm inland regions such as the Riverland are known to only produce large amounts of lower graded grapes \autocite{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
+This behaviour can be observed when reviewing Wine Australia’s annual reports, where it is apparent that warm inland regions such as the Riverland are known to only produce large amounts of lower graded grapes \cite{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
 Comparatively, regions such as Tasmania only produce A grade grapes but in much smaller quantities than the Riverland.
@@
 
-The opportunity to target different grades of grapes may not always be available, with some regions being more renowned than others, and likely to be sought after regardless \autocite{hallidayAustralianWineEncyclopedia2009}.
+The opportunity to target different grades of grapes may not always be available, with some regions being more renowned than others, and likely to be sought after regardless \citep{hallidayAustralianWineEncyclopedia2009}.
 The Barossa is an example of this, known for its quality could also lend itself to a bias in purchasers not considering other regions that may be capable of similar quality.
@@
 A further possibility is that there may be regional upper limits with the relationship between resource input and the value gained becoming no longer proportional due to diminishing returns.
-Climate was considered to be a large determinant of the ability to grow a larger quantity of grapes, as well as a determinant in grape quality \autocite{agostaRegionalClimateVariability2012}; however there were vineyards in similar regions that were able to produce exceptionally better results than others (See Figure 7).
+Climate was considered to be a large determinant of the ability to grow a larger quantity of grapes, as well as a determinant in grape quality \citep{agostaRegionalClimateVariability2012}; however there were vineyards in similar regions that were able to produce exceptionally better results than others (See Figure 7).
 
@@
 Literature shows that there are many on-the-ground decisions that influence both quality and yield.
-Where these decisions are governed by complex physical and social forces such as international market demands, disease pressures and natural disasters \autocite{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}.
-Many of these occurrences being highlighted throughout the past decades vintage reports \autocite{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
+Where these decisions are governed by complex physical and social forces such as international market demands, disease pressures and natural disasters \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}.
+Many of these occurrences being highlighted throughout the past decades vintage reports \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
 %This definitely needs to be broken up, what were these occurences, which year was significant for what?
@@
 These differences could be caused by the use of alternative sustainable practices in the field.
-While there is evidence to suggest that environmentally sustainable practices can reduce costs, increase efficiency, whilst improving the quality of grapes, more research is needed to link these benefits across different regions and climates \autocite{baianoOverviewSustainabilityWine2021,marianiSustainableWinegrowingCurrent2015,montalvo-falconSustainabilityResearchWine2023}.
+While there is evidence to suggest that environmentally sustainable practices can reduce costs, increase efficiency, whilst improving the quality of grapes, more research is needed to link these benefits across different regions and climates \citep{baianoOverviewSustainabilityWine2021,marianiSustainableWinegrowingCurrent2015,montalvo-falconSustainabilityResearchWine2023}.
 
@@
 
-\printbibliography{}
-\end{sloppypar}
+\bibliography{references}
+\bibliographystyle{elsarticle-harv}
+
+%% The Appendices part is started with the command \appendix;
+%% appendix sections are then done as normal sections
+%% \appendix
+
+%% \section{}
+%% \label{}
+
+%% If you have bibdatabase file and want bibtex to generate the
+%% bibitems, please use
+%%
+
+%% else use the following coding to input the bibitems directly in the
+%% TeX file.
+
+\end{linenumbers}
   \end{document}
+
+  \endinput

### 2023-06-29 article.tex
Fixed introduction and start of methods.

@@
 The global focus on sustainability in agronomic industries has changed the way in which these enterprises do business.
-When strategies for a sustainable industry are assessed, there is a trade-off between balancing the amount of resources invested and the resultant quality verses quantity of crops produced.
-This dilemma exists across agriculture, sharing fundemental considerations such as water and nitrogen input
-\citep{abadCoverCropsViticulture2021}
-
-\cite{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016, zhuEffectsNitrogenLevel2017}. 
-\citet{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016, zhuEffectsNitrogenLevel2017}.
-
+When strategies for a sustainable winegrowing industry are assessed, there is a trade-off between balancing the amount of resources invested and the resultant yield verses quality produced.
+This dilemma exists across agriculture within shared fundamental considerations such as water and nitrogen inputs \citep{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016, zhuEffectsNitrogenLevel2017}.
 Within viticulture (the cultivation of grapes for wine production) quality is driven by its integration within the wine industry; with a wine's potential quality being initially defined through the chemical makeup of the grapes used in its production.
@@
 In the Australian context, these include biosecurity, climate and international market demands.
-
-In this analysis we observe relationships between quantity and quality through the use of linear models.
-Quality can be defined in a variety of ways for grapes, such as *.
-For the purpose of this study quality was defined by the financial value of grapes per tonne; this was due to varying differences in defining quality between grape varietals.
-
-An extensive amount of research into a variety of factor's effect on grape quality and yield exists.
-Due to the lack of long-term and in-depth data sources, individual effects are often studied in isolation \citep{abbalDecisionSupportSystem2016}.
-The lack of consolidated datasets also restricts the ability to gain statistical insight at a large scale across multiple regions \citep{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
+\newline
+In this analysis we observe relationships between yield and quality through the use of linear models.
+Quality can be defined in a variety of ways, for example analysing grapes' aroma, chemical composition and color.
+For the purpose of this study, quality was defined by winegrape crops' financial value per tonne.
+This definition assumes due diligence on the side of those that purchased the grapes; where market value of grapes heavily relies on grape quality \citep{yeggeInfluenceSensoryNonsensory2001}.
+Wine Australia also links grape quality to price per tonne, by explicitly defining grape quality within discrete price brackets.
+\newline
+An extensive amount of research into a variety of factors' effect on grape quality and yield exists.
+% there needs to be references if there is extensive research
+Due to the lack of long-term and in-depth data, individual effects are often studied in isolation \citep{abbalDecisionSupportSystem2016}.
+The lack of consolidated datasets also restricts the ability to gain statistical insights at large scales and across multiple regions \citep{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
 The dataset used for this analysis includes data collected for the past 10 years from a multitude of vineyards located over a diverse range of Australian winegrowing regions.
-
+\newline
+We aim to use this broad dataset to confirm the existence of a yield verse quality trade off within Australian winegrowing; one not prior confirmed explicitly across such extensive diversities.
+% you require a reference here
+In achieving this, the context of how resource use relates to yield and quality will also be described.
+We link these relations to the potential for improvement through decision-making processes, and highlight that the way moving forward will require the optimisation of this process.
+The practical addition of these aims is a baseline for comparison - given a vineyard within Australia, one could extrapolate their comparative efficiency with regard to the tradeoff between invested resources, yield and quality.
 \section{Methods}
-We created four linear models to explore the relationships between resource use and vineyard outputs (see table 1).
-The models looked at the value and quantity of grapes produced by vineyards, as a total and as a scale of area harvested.
-We compared values to observe how economies of scale affect the use of resources.
+We created four linear models to explore relationships between resources used and vineyard outputs (see Table 1).
+The response variables of the models were yield and quality, with yield being measured in tonnes and quality being the product of yield and the average sale price per tonne.
+Both response variables were examined as totals and as scales of area harvested.
+Values were compared in this manner to observe how economies of scale affect the use of resources.
 
@@
 \subsection{Data}
-The data used in this analysis was provided by Sustainable Winegrowing Australia, Australia's national wine industry sustainability program.
-Sustainable winegrowing Australia aims to facilitate grape-growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
-Data recorded by Sustainable Winegrowing Australia is entered manually by winegrowers using a web based interface; with some fields being optional.
-Two subsets of this data were defined by vineyards that recorded values for average price of grape sales per tonne, and vineyards who did not.
-Both subsets contained: region, harvest year, yield, area harvested, water used and fuel used (in litres for diesel, petrol, biodiesel and LPG).
-To enable comparisons, total fuel was converted to amount of carbon emissions in metric tons.
+Data used in this analysis was provided by Sustainable Winegrowing Australia, Australia's national wine industry sustainability program; which aims to facilitate grape-growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
+Data recorded by Sustainable Winegrowing Australia is entered manually by winegrowers using a web based interface, with some fields being optional.
+Two subsets of this data were defined through vineyards which did not record average price of sale per tonne, and vineyards which did or which were in a region of known average price.
+Vineyards which did not record a value for average price of sale per tonne but were within regions with a recorded average price of sale per tonne by Wine Australia were filled in using this regional average.
+Both subsets contained: region, harvest year, yield, area harvested, water used and fuel used (diesel, petrol, biodiesel and LPG).
+To enable comparisons, total fuel was converted to equivalent carbon emissions in metric tons.
 
@@
 Two values for average sale price were removed from the dataset, due to a recording of \$1.
-Unreported values for average prices per tonne were filled in using regional averages taken from Wine Australia's annual reports, where they were available \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2012,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
+% \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2012,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
 

### 2023-06-30 article.tex
coffee break edit update

@@
 \subsection{Data}
-Data used in this analysis was provided by Sustainable Winegrowing Australia, Australia's national wine industry sustainability program; which aims to facilitate grape-growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
-Data recorded by Sustainable Winegrowing Australia is entered manually by winegrowers using a web based interface, with some fields being optional.
-Two subsets of this data were defined through vineyards which did not record average price of sale per tonne, and vineyards which did or which were in a region of known average price.
-Vineyards which did not record a value for average price of sale per tonne but were within regions with a recorded average price of sale per tonne by Wine Australia were filled in using this regional average.
-Both subsets contained: region, harvest year, yield, area harvested, water used and fuel used (diesel, petrol, biodiesel and LPG).
-To enable comparisons, total fuel was converted to equivalent carbon emissions in metric tons.
-
+Data used in this analysis was sampled by Sustainable Winegrowing Australia and Wine Australia.
+Sustainable Winegrowing Australia is Australia's national wine industry sustainability program, which aims to facilitate grape-growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
+Wine Australia is an Australian Government statutory authority governed by the Wine Australia Act 2013 \citep{WineAustraliaAct2019}.
+\newline
+Data sampled by Wine Australia included: summary statistics such as yield and average price of sale per tonne, for contexts such as region and grape varietal; this data was collected via phone surveys.
+Data recorded by Sustainable Winegrowing Australia was entered manually by winegrowers using a web based interface with some fields being optional, variables included: region, harvest year, yield, area harvested, water used and fuel used (diesel, petrol, biodiesel and LPG).
+To enable comparisons, total fuel was converted to tonnes of Carbon Dioxide equivalent.
+Two subsets of Sustainable Winegrowing Australia's data were defined.
+One subset contained all vineyards.
+The other subset contained vineyards which either recorded a value for average price of sale per tonne or had a regional value recorded by Wine Australia.
+\newline
 The first subset of data was used for Model 1 and Model 2 (see Table 1). This subset contained 5298 samples spanning the period from 2012 to 2022, covering 57 GI Regions and 1432 separate vineyards.
-
+\newline
 The second subset of data, was limited to vineyards that recorded a value for their average sale price of grapes per tonne.
 This subset was used for Model 3 and Model 4 (see Table 1); and contained 2878 samples spanning the period from 2015 to 2022, covering 51 GI Regions and 944 separate vineyards.
-
-Data was limited to samples that had recorded values for variables used (see Table 1).
+1842 of the values for average price of sale per tonne were extracted from Wine Australia surveys with the remaining 1036 being from Sustainable Winegrowing Australia's dataset.
+\newline
+Data used in each model was limited to samples that had recorded values for the variables used (see Table 1).
 After reviewing correlation coefficients the data was logarithmically transformed, centred and scaled by standard deviation.
@@
 % \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2012,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
-
 Other variables including the use of renewable energy, contractors; and pressures such as frost, fire and disease were also explored.
@@
 \subsection{Total Emissions}
-Emissions were calculated from the total diesel, petrol, bio-diesel and LPG used for irrigation and activities within the vineyard.
-The equations given from the Australian National Greenhouse Accounts Factors, shown as
+The equation given from the Australian National Greenhouse Accounts Factors, shown as
 
@@
 was used to convert the quantity of fuel in litres, $Q$, using a prescribed Energy Content, $EC$, and emission factors of scope one, $EF1$, and scope three, $EF3$, to tonnes of Carbon Dioxide equivalent, $tCO2e$ \citep{departmentofclimatechangeenergytheenvironmentandwaterAustralianNationalGreenhouse2022}.
+Emissions were calculated for total diesel, petrol, bio-diesel and LPG used.
 
-The variables were reviewed for correlations by using a Pearson's Correlation Coefficient (see Tables 1, 2 and 3).
+% What the hell is the below paragraph and why is it here?
+%The variables were reviewed for correlations by using a Pearson's Correlation Coefficient (see Tables 1, 2 and 3).
 This was undertaken for data on the original scale (see Table 1) and for data as a logarithmic transform (see Table 2).
@@
 \subsection{Region}
-The site of a vineyard predetermines several physical parameters such as climate, geology and soil; making location a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
 Differences in vineyard locations were captured through the use of Geographical Indicator Regions (GI Regions).
@@
 
-The climatic properties of a GI Region are summarised in the Sustainable Winegrowing Australia user manual \citep{swaSustainableWinegrowingAustralia2021}.
+The site of a vineyard predetermines several physical parameters such as climate, geology and soil; making location a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
+The climatic properties of each GI Region were summarised by using predefined climatic classifications as per the \citet{swaSustainableWinegrowingAustralia2021} user manual.
 The user manual describes climates by rainfall and temperature, creating supersets of Regions of similar climatic properties.

### 2023-07-06 article.tex
aaahhhh why didnt you commit any of these changes this weekgit add --all!

@@
 
-    \title{Resource Use and the Value-Productivity Tradeoff in Australian Winegrowing Regions}
+    \title{Resource-use and the Value-Productivity Tradeoff in Australian Winegrowing Regions}
 
@@
 When strategies for a sustainable winegrowing industry are assessed, there is a trade-off between balancing the amount of resources invested and the resultant yield verses quality produced.
-This dilemma exists across agriculture within shared fundamental considerations such as water and nitrogen inputs \citep{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016, zhuEffectsNitrogenLevel2017}.
-Within viticulture (the cultivation of grapes for wine production) quality is driven by its integration within the wine industry; with a wine's potential quality being initially defined through the chemical makeup of the grapes used in its production.
-The consideration of sustainability is further complicated by environmental and socio-demographic pressures.
-In the Australian context, these include biosecurity, climate and international market demands.
+This dilemma exists across agriculture through shared fundamental considerations such as water use and nitrogen levels \citep{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016, zhuEffectsNitrogenLevel2017}.
+Quality in viticulture (the cultivation of grapes for wine production) is driven through its integration within the wine industry; with a wine's potential quality being initially defined through the chemical makeup of the grapes used in its production.
+The consideration of sustainability within viticulture is further complicated by environmental and socio-demographic pressures.
+In the Australian context, these include: biosecurity, climate and international market demands.
 \newline
 In this analysis we observe relationships between yield and quality through the use of linear models.
-Quality can be defined in a variety of ways, for example analysing grapes' aroma, chemical composition and color.
-For the purpose of this study, quality was defined by winegrape crops' financial value per tonne.
-This definition assumes due diligence on the side of those that purchased the grapes; where market value of grapes heavily relies on grape quality \citep{yeggeInfluenceSensoryNonsensory2001}.
-Wine Australia also links grape quality to price per tonne, by explicitly defining grape quality within discrete price brackets.
-\newline
-An extensive amount of research into a variety of factors' effect on grape quality and yield exists.
-% there needs to be references if there is extensive research
-Due to the lack of long-term and in-depth data, individual effects are often studied in isolation \citep{abbalDecisionSupportSystem2016}.
+Although an extensive amount of research into a variety of factors' effect on grape quality and yield exists; due to the lack of long-term and in-depth data, individual effects are often studied in isolation \citep{abbalDecisionSupportSystem2016}.
 The lack of consolidated datasets also restricts the ability to gain statistical insights at large scales and across multiple regions \citep{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
@@
 % you require a reference here
-In achieving this, the context of how resource use relates to yield and quality will also be described.
-We link these relations to the potential for improvement through decision-making processes, and highlight that the way moving forward will require the optimisation of this process.
+In achieving this, the context of how resource-use relates to yield and quality will also be described.
+We link these relations to the potential for improvement through decision-making processes, whilst highlighting that the way moving forward will require the optimisation of these processes.
 The practical addition of these aims is a baseline for comparison - given a vineyard within Australia, one could extrapolate their comparative efficiency with regard to the tradeoff between invested resources, yield and quality.
 \section{Methods}
-We created four linear models to explore relationships between resources used and vineyard outputs (see Table 1).
-The response variables of the models were yield and quality, with yield being measured in tonnes and quality being the product of yield and the average sale price per tonne.
-Both response variables were examined as totals and as scales of area harvested.
-Values were compared in this manner to observe how economies of scale affect the use of resources.
+We created four linear models to explore relationships between resource-use and vineyard outputs (see Table 1).
+The data was sourced from Sustainable Winegrowing Australia and Wine Australia.
+Variables used included: yield, average sale price, region, water use, emissions, area harvested and year.
+After fitting to the data, each model was validated using k-fold cross validation.
 
@@
 
+\subsection{Analysis}
+Before models were fit to the data, Pearson Correlation Coefficients were used to look at the existence of linear relationships between predictor variables.
+These relationships were summarised in correlation matrices to compare levels of interaction present between predictor variables.
+The relationships between the predictors and response variables were then modelled using General Linear Models.
+Both the Pearson Correlation Coefficients and General Linear Models were created using the R statistical programming language \citep{rcoreteamLanguageEnvironmentStatistical2021}.
+General Linear Models were chosen as they offer the ability to produce statistical models that are explicit in the relationships between predictors and response variables.
+%reference?
+General Linear Models also allow the exploration of interactions between predictors and easily comparable differences in the influence and magnitude of relationships.
+% reference?
+ A variety of alternate methods to General Linear Models was also explored, including: Splines, hierarchical regression, Additive and Generalised Linear Models.
+These alternative approaches were not used as final models due to offering no further improvements or to producing inaccurate models.
+\newline
+The response variables of the models were yield and quality.
+The definition of quality was an important consideration, as quality can be defined in a variety of ways, for example analysing grapes': aroma, chemical composition and color.
+For the purpose of this study, quality was defined by the financial value of winegrape crops' average sale price per tonne.
+This definition was used due to the market value of winegrapes being reliant on grape quality and because Wine Australia explicitly defines grape quality through the use of discrete price brackets in their annual reports
+% We need a reference to where they did this
+% Did they do this more prolifically at some point as well?
+; the generalisation made to reflect quality through using average price assumed a due diligence of those who purchased the grapes \citep{yeggeInfluenceSensoryNonsensory2001}.
+Yield was defined as the total tonnes of grapes harvested.
+Both response variables were examined as totals and as scales of area harvested.
+Values were compared in this manner to observe how economies of scale affect the use of resources.
 \subsection{Data}
@@
 \newline
-Data sampled by Wine Australia included: summary statistics such as yield and average price of sale per tonne, for contexts such as region and grape varietal; this data was collected via phone surveys.
+Data sampled by Wine Australia was collected via phone surveys and included: summary statistics such as yield and average price of sale per tonne; these values were summarised by region and grape varietal.
 Data recorded by Sustainable Winegrowing Australia was entered manually by winegrowers using a web based interface with some fields being optional, variables included: region, harvest year, yield, area harvested, water used and fuel used (diesel, petrol, biodiesel and LPG).
-To enable comparisons, total fuel was converted to tonnes of Carbon Dioxide equivalent.
-Two subsets of Sustainable Winegrowing Australia's data were defined.
-One subset contained all vineyards.
-The other subset contained vineyards which either recorded a value for average price of sale per tonne or had a regional value recorded by Wine Australia.
+To enable direct comparisons between fuels, they were converted to tonnes of Carbon Dioxide equivalent.
+\newline
+The inclusion of Wine Australia data was due to average sale price being an optional field in Sustainable Winegrowing Australia's dataset. 
+Regional average prices from Wine Australia were filled into values that were missing from the Sustainable Winegrowing Australia data; the common practice of purchasing grapes at regional prices was an important consideration in this decision.
+%{It would be lovely to have a reference for this but it is currently only anacdotal.}
+Two subsets of data were then created for the analysis.
+The first subset contained all vineyards and was used for Models 1 and 3.
+The second subset contained vineyards which either recorded a value for average price of sale per tonne through Sustainable Winegrowing Australia, or were within a region with an average price of sale recorded by Wine Australia; this subset was used for Models 2 and 4.
+These subsets meant that the data would be limited to samples which had recorded values for the response variables (see Table 1), where every sample had a recorded value for yield but not average price of sale per tonne.
 \newline
@@
 \newline
-Data used in each model was limited to samples that had recorded values for the variables used (see Table 1).
-After reviewing correlation coefficients the data was logarithmically transformed, centred and scaled by standard deviation.
-Two values for average sale price were removed from the dataset, due to a recording of \$1.
+Additional variables were considered for analysis but were excluded due to being either underreported or had insignificant contributions to model accuracies.
+Variables explored but not used due to low reporting values included: fertiliser, tractor and contractor use, and scope 2 emissions.
+Variables considered but ultimately removed due to a lack of significant contributions to models, included: the use of renewable energy, contractors, and pressures such as frost, fire and disease.
+\newline
+Data preprocessing was conducted prior to analysis using the Python programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
+Preprocessing included logarithmic transformations, centring and scaling by standard deviation.
+Variables such as scope 1, which required prior calculations were also computed using Python.
+% Two values for average sale price were removed from the dataset, due to a recording of \$1. 
 % \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2012,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
-Other variables including the use of renewable energy, contractors; and pressures such as frost, fire and disease were also explored.
-Variables that did not significantly contribute to the prediction of a response variable were excluded.
 
@@
 The equation given from the Australian National Greenhouse Accounts Factors, shown as
-
+\newline
 \begin{equation}
@@
 \end{equation}
-
+\newline
 was used to convert the quantity of fuel in litres, $Q$, using a prescribed Energy Content, $EC$, and emission factors of scope one, $EF1$, and scope three, $EF3$, to tonnes of Carbon Dioxide equivalent, $tCO2e$ \citep{departmentofclimatechangeenergytheenvironmentandwaterAustralianNationalGreenhouse2022}.
 Emissions were calculated for total diesel, petrol, bio-diesel and LPG used.
-
+%\newline
 % What the hell is the below paragraph and why is it here?
@@
 The logarithmic transforms performed the best due to a skew likely caused by a greater number of smaller vineyards within the dataset (see Table 4).
-
 \subsection{Region}
 Differences in vineyard locations were captured through the use of Geographical Indicator Regions (GI Regions).
-Each GI Region has its own unique mixture of climatic and geophsical properties that describes a unique winegrowing region within Australia; these regions were predefined by Wine Australia \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
-
+Each GI Region has its own unique mixture of climatic and geophysical properties that describes a unique winegrowing region within Australia; these regions were predefined by Wine Australia \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
+Both Wine Australia and Sustainable Winegrowing Australia used the same GI Region format to describe location.
+\newline
 The site of a vineyard predetermines several physical parameters such as climate, geology and soil; making location a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
-The climatic properties of each GI Region were summarised by using predefined climatic classifications as per the \citet{swaSustainableWinegrowingAustralia2021} user manual.
+The climatic properties of each GI Region were summarised by using predefined classifications as per the \citet{swaSustainableWinegrowingAustralia2021} user manual.
 The user manual describes climates by rainfall and temperature, creating supersets of Regions of similar climatic properties.
-The climatic groups were used to illustrate similarities and differences occurring in areas larger than GI regions.
-
-\subsection{Analysis}
-General Linear Models were used as they offered the ability to produce statistical models that were explicit in the relationships between predictors and response variables.
-They also allowed the exploration of interactions between predictors and easily comparable differences in the influence and magnitude of relationships.
-
-Data preprocessing, such as logarithmic transforms, was done using the Python programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
-Linear models were created using the R statistical programming language \citep{rcoreteamLanguageEnvironmentStatistical2021}.
-These models were created iteratively to explore a variety of variable interactions and approaches to modelling the data.
-Not all explored approaches yielded improvements or accurate models.
-Alternate approaches included the use of Splines, hierarchical regression, Additive and Generalised Linear Models.
-Other variables were also explored but not used due to low reporting values such as fertiliser, tractor and contractor use.
-The use of only scope one emissions was due to the same reason where scope 2 sources were recorded sporadically at best.
-
+The climatic groups were used to illustrate similarities and differences occurring in areas larger than GI Regions.
+\newline
 \subsection{Model Validation}
@@
 \subsection{Exploratory Analysis}
-Simple linear relationships between variables were explored using Pearson Correlation Coefficients.
+Linear relationships between variables were explored using Pearson Correlation Coefficients.
+Values for these coefficients range between -1 and 1, and are indicative of .
 This was undertaken for data on the original scale (see Table 2) and for data as a logarithmic transform (see Table 3).
@@
 \subsection{General Linear Models}
-Models 1 and 2 showed significant relationships between each of the predictors and their response variable (see Tables 6 and 7).
+Models 1 and 2 showed significant relationships between each of the predictors and their response variables (see Tables 6 and 7).
 Variables in models 3 and 4 reported similar significance; except for scope 1 emissions (see Tables 8 and 9).
@@
 The negative trend between size and average sales price could also be a side effect of mass supply verse demand, especially when looking at the level of difference in production of some vineyards (see Table 4).
-The relationships between yield, value and area are not simply about efficiently producing the most grapes; sales price and by association grape quality, are integral to the profitability, and this is strongly linked to resource use and thus the longevity and sustainability of a vineyard.
+The relationships between yield, value and area are not simply about efficiently producing the most grapes; sales price and by association grape quality, are integral to the profitability, and this is strongly linked to resource-use and thus the longevity and sustainability of a vineyard.
 

### 2023-07-06 article.tex
working with the results.

@@
 % A reference for this is absolutely needed.
-
 \section{Results}
@@
 Linear relationships between variables were explored using Pearson Correlation Coefficients.
-Values for these coefficients range between -1 and 1, and are indicative of .
+Values for these coefficients are a measure of the linear relation between two variables; this measure is scale between -1 and 1, indicating the strength of the relationship.
+Transforming the data prior to calculating these values changes several things: The logarithmic transform of the data changes the interpretation of the coefficients to percentage changes, such that a coefficient will be indicative of the ratio a variable will change in percentage compared to a one percent increase in the other; being scaled by standard deviation this increase will be relative to the standard deviation, making it a percent of that variables standard deviation.
+Scaling by standard deviation also makes the Pearson Correlation Coefficient equal to the covariance of the two variables.
+%
 This was undertaken for data on the original scale (see Table 2) and for data as a logarithmic transform (see Table 3).

### 2023-07-07 article.tex
Up to general linear model in result section.

@@
 Linear relationships between variables were explored using Pearson Correlation Coefficients.
-Values for these coefficients are a measure of the linear relation between two variables; this measure is scale between -1 and 1, indicating the strength of the relationship.
-Transforming the data prior to calculating these values changes several things: The logarithmic transform of the data changes the interpretation of the coefficients to percentage changes, such that a coefficient will be indicative of the ratio a variable will change in percentage compared to a one percent increase in the other; being scaled by standard deviation this increase will be relative to the standard deviation, making it a percent of that variables standard deviation.
+Values for these coefficients reflect the linear relation between two variables, on a scale between -1 and 1; the magnitude and sign of a coefficient indicates the strength of the relation, and whether the relation is positive or negative respectively.
+This was undertaken for data on the original scale (see Table 2) and for data as a logarithmic transform (see Table 3).
+The logarithmic transforms showed the strongest correlations, this was likely due to a skew caused by a greater number of smaller vineyards within the dataset (see Table 5).
+Transforming data prior to calculating the coefficients changes several things: The logarithmic transform of the data alters the interpretation of the coefficients to percentage change \- a coefficient will be indicative of the change in percentage of one variable compared to the other; scaling by standard deviation also changes this interpretation to be a percentage of that variables standard deviation.
 Scaling by standard deviation also makes the Pearson Correlation Coefficient equal to the covariance of the two variables.
-%
-This was undertaken for data on the original scale (see Table 2) and for data as a logarithmic transform (see Table 3).
+A correlation of 1 would indicate that given the change of one variable by one percentage of its standard deviation, the other variable would change by one percent of its standard deviation; the importance of this is the unitless nature of these relationships and that it can be translated directly to any case that has a well known distribution.
+\newline
+To determine if a coefficient was indicative of a strong relationship, P-values were used.
+P-values reflect the significance of a given correlation coefficient when considering its relation to sample size via its incorporation as an element of standard error.
 Strong relationships were found to be present, as all P-values were considered significant (< 2.200E-16, see Tables 2 and 3), except for the non-transformed values for water used (see Table 4).
-The logarithmic transforms showed the strongest correlations, this was likely due to a skew caused by a greater number of smaller vineyards within the dataset (see Table 5).
 
@@
 \subsection{General Linear Models}
+% What are these tables?
+% what are these models?
+% What are F values?
+% Why are some significant?
+% What are degrees of freedom
+% what are mean sq values?
+% What are sum sq values?
+% What are the ramifications of any of this???
+% Break this down into each model separately!! DO NOT GENERALISE
 Models 1 and 2 showed significant relationships between each of the predictors and their response variables (see Tables 6 and 7).
+% Why are they significant?
 Variables in models 3 and 4 reported similar significance; except for scope 1 emissions (see Tables 8 and 9).

### 2023-07-07 analysis.r
Added graphs and coef collection to each model

@@ -4,20 +4,48 @@ library(ggplot2)
 set.seed(100)
 df <- read.csv("df.csv")
 
-model1 <- lm(tonnes_grapes_harvested ~ data_year_id+giregion+area_harvested+water_used+scope1,data=df)
-model2 <- lm(ha_tonnes_grapes_harvested ~ area_harvested*((scope1)+water_used)+data_year_id:giregion, data=df)
-model3 <- lm(value ~ data_year_id+giregion+area_harvested+water_used+scope1,data=df)
-model4 <- lm(ha_value ~ area_harvested*((scope1)+water_used)+data_year_id:giregion,data=df)
+model1 <- lm(
+    tonnes_grapes_harvested ~
+    data_year_id + giregion + area_harvested + water_used + scope1
+    , data = df)
+model2 <- lm(
+    ha_tonnes_grapes_harvested ~
+    area_harvested * ((scope1) + water_used) + data_year_id:giregion
+    , data = df)
+model3 <- lm(
+    value ~
+    data_year_id + giregion + area_harvested + water_used + scope1
+    , data = df)
+model4 <- lm(
+    ha_value ~
+    area_harvested * ((scope1) + water_used) + data_year_id:giregion
+    , data = df)
 
-# We create our first model
+# We review our first model
 
 # MODEL 1
 anova(model1)
 summary(model1)
-model1.residuals <- residuals(model1)
-model1.fitted <- fitted(model1)
-ggplot(data.frame(y=model1.residuals), aes(sample=y)) + stat_qq() + stat_qq_line()
-ggplot(data.frame(y=model1.residuals,x=model1.fitted), aes(x=x,y=y)) + geom_point() + geom_hline(yintercept=0)
+coefs1 <- coefs(model1)
+coefs1_giregion <- coefs1[grep("gi*", row.names(ok)), ]
+coefs1_year <- coefs1[grep("data*", row.names(ok)), ]
+model1_residuals <- residuals(model1)
+model1_fitted <- fitted(model1)
+ggplot(
+    data.frame(
+        y = model1_residuals)
+        , aes(sample = y)
+       ) +
+       stat_qq() +
+       stat_qq_line()
+ggplot(
+    data.frame(
+        y = model1_residuals
+        , x = model1_fitted)
+    , aes(x = x, y = y)
+    ) +
+    geom_point() +
+    geom_hline(yintercept = 0)
 #Linear Regression
 #
 #6068 samples
[…]

### 2023-07-10 article.tex
Further edits. Title change. Work on results.

@@
 
-    \title{Resource-use and the Value-Productivity Tradeoff in Australian Winegrowing Regions}
+    \title{An exploratory analysis of the influence of resource use on the yield verse quality trade-off in Australian vineyards}
 
@@
 In achieving this, the context of how resource-use relates to yield and quality will also be described.
+% Describing baselines is more important - it needs to be emphasised more so than something that is known but not shown. This is the heart of what your are doing sell it more.
 We link these relations to the potential for improvement through decision-making processes, whilst highlighting that the way moving forward will require the optimisation of these processes.
@@
 % reference?
- A variety of alternate methods to General Linear Models was also explored, including: Splines, hierarchical regression, Additive and Generalised Linear Models.
-These alternative approaches were not used as final models due to offering no further improvements or to producing inaccurate models.
+ A variety of alternate methods were also explored, including: Splines, hierarchical regression, General Additive Models, and Generalised Linear Models.
+These alternative approaches were not used as final models due to offering no further insights or improvements in accuracy.
 \newline
 The response variables of the models were yield and quality.
+Yield was defined as the total tonnes of grapes harvested.
+For the purpose of this study, quality was defined by the financial value of winegrape crops' average sale price per tonne.
 The definition of quality was an important consideration, as quality can be defined in a variety of ways, for example analysing grapes': aroma, chemical composition and color.
-For the purpose of this study, quality was defined by the financial value of winegrape crops' average sale price per tonne.
-This definition was used due to the market value of winegrapes being reliant on grape quality and because Wine Australia explicitly defines grape quality through the use of discrete price brackets in their annual reports
+% References to were this was done.
+Using sale price as a defining trait of quality was due to the market value of winegrapes being reliant on grape quality and because Wine Australia explicitly defines grape quality through the use of discrete price brackets in their annual reports
 % We need a reference to where they did this
@@
 ; the generalisation made to reflect quality through using average price assumed a due diligence of those who purchased the grapes \citep{yeggeInfluenceSensoryNonsensory2001}.
-Yield was defined as the total tonnes of grapes harvested.
 Both response variables were examined as totals and as scales of area harvested.
@@
 Additional variables were considered for analysis but were excluded due to being either underreported or had insignificant contributions to model accuracies.
-Variables explored but not used due to low reporting values included: fertiliser, tractor and contractor use, and scope 2 emissions.
-Variables considered but ultimately removed due to a lack of significant contributions to models, included: the use of renewable energy, contractors, and pressures such as frost, fire and disease.
+Variables explored but not used due to low reporting values included: fertiliser, and scope 2 emissions.
+Variables considered but ultimately removed due to a lack of significant contributions to models, included: the use of renewable energy, contractor use, and pressures such as frost, fire and disease.
 \newline
@@
 The site of a vineyard predetermines several physical parameters such as climate, geology and soil; making location a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
-The climatic properties of each GI Region were summarised by using predefined classifications as per the \citet{swaSustainableWinegrowingAustralia2021} user manual.
+The climatic properties of each GI Region were summarised by using predefined classifications as per the \citet{sustainablewinegrowingaustraliaSustainableWinegrowingAustralia2021} user manual.
 The user manual describes climates by rainfall and temperature, creating supersets of Regions of similar climatic properties.
 The climatic groups were used to illustrate similarities and differences occurring in areas larger than GI Regions.
-\newline
 \subsection{Model Validation}
@@
 This was undertaken for data on the original scale (see Table 2) and for data as a logarithmic transform (see Table 3).
-The logarithmic transforms showed the strongest correlations, this was likely due to a skew caused by a greater number of smaller vineyards within the dataset (see Table 5).
-Transforming data prior to calculating the coefficients changes several things: The logarithmic transform of the data alters the interpretation of the coefficients to percentage change \- a coefficient will be indicative of the change in percentage of one variable compared to the other; scaling by standard deviation also changes this interpretation to be a percentage of that variables standard deviation.
+The logarithmic transforms showed the strongest correlations, likely due to a skew caused by a greater number of smaller vineyards within the dataset (see Table 5).
+Transforming data prior to calculating the coefficients changes several things: The logarithmic transform of the data alters the interpretation of the coefficients to percentage change - a coefficient will be indicative of the change in percentage of one variable compared to the other; scaling by standard deviation also changes this interpretation to be a percentage of that variables standard deviation.
 Scaling by standard deviation also makes the Pearson Correlation Coefficient equal to the covariance of the two variables.
-A correlation of 1 would indicate that given the change of one variable by one percentage of its standard deviation, the other variable would change by one percent of its standard deviation; the importance of this is the unitless nature of these relationships and that it can be translated directly to any case that has a well known distribution.
+With all this in mind, when considering the logarithmically transformed variables, a coefficient of 1 would indicate that: given the change of one variable by one percentage of its standard deviation, the other variable would change by one percent of its own standard deviation.
+The importance of this is the unitless nature of these relationships and that it can be translated directly to any vineyard's case that has a well known distribution.
 \newline
-To determine if a coefficient was indicative of a strong relationship, P-values were used.
-P-values reflect the significance of a given correlation coefficient when considering its relation to sample size via its incorporation as an element of standard error.
+To determine if a coefficient was indicative of a strong relationship, confidence intervals were used.
+P-values reflected the significance of a given correlation coefficient when considering its relation to sample size via its incorporation as an element of standard error.
 Strong relationships were found to be present, as all P-values were considered significant (< 2.200E-16, see Tables 2 and 3), except for the non-transformed values for water used (see Table 4).
@@
 \subsection{General Linear Models}
+% The tables need to be in the appropriate sections!!
+%
+%Where are the coefficients???? You cant use the models without them!!
+% You need to actually be able to use this research and look at what these
+% models are showing.
+%
 % What are these tables?
@@
 % Break this down into each model separately!! DO NOT GENERALISE
-Models 1 and 2 showed significant relationships between each of the predictors and their response variables (see Tables 6 and 7).
-% Why are they significant?
-Variables in models 3 and 4 reported similar significance; except for scope 1 emissions (see Tables 8 and 9).
+% You can draw similarities between the model, as this is not a generalisation.
+%
+%
+%
+% What is a general linear model
+% what does this mean
+To describe the response variables relative to the predictors' values, General Linear Models were used.
+Log transformed variables were used as inputs to these models as they resulted in higher $R^2$ values and described the relationships proportionally; reflecting coefficient values as percentages of a variable's standard deviation.
+\newline
+% F-tests
+To determine if individually each model's predictors were significantly linearly related to the response variables, F-tests were conducted.
+Models 1 and 2 showed significant linear relationships between each of the predictors and their response variables (see Tables 6 and 7).
+Predictors in models 3 and 4 reported similar significance; except for scope 1 emissions (see Tables 8 and 9).
 Scope one emissions was included in all models to directly compare the response variables as ratios of vineyard size to raw values.
-Even though not significant within models 3 and 4, when using the Pearson Correlation Coefficients, scope one emissions was strongly correlated to every Model's response variable; this was especially so for Model 1 and 4 (Yeild and average price per tonne as a ratio to area harvested, respectively).
-
-The comparison of models performance shows that the average price per tonne of grapes describes a great deal of the relationship between predictor and response when comparing model 2 to model 4 (see Table 10).
-This relationship between yield and average price was also illustrated in the correlation values between them (see Table 2).
-
+Even though not significant within models 3 and 4, when using the Pearson Correlation Coefficients, scope one emissions was strongly correlated to every Model's response variable; this was especially so for Model 1 and 4 (Yield and average price per tonne as a ratio to area harvested, respectively).
+\newline
+% t-tests
+Predictors were also reviewed using t-tests to determine if a predictor significantly contributed to the model when accounting for other variables.
+We can consider this the marginal significance of each variable.
+Using t-tests allowed a more granular examination of each factor (or categorical variable) and interaction, showing that some years and areas contributed marginally significantly and some did not (see appendix).
+With regard to continuous variables, Model 1 and 2 showed water use, scope 1 emissions and area harvested were marginally significant.
+For Model 3 all continuous variables except scope 1 emissions were significant.Model 4 showed scope 1 emissions and water use to only be significant when considered as an interaction with area harvested but not as individual variables (see Table ?).
+% should these tables all go into the appendix solely or should they be summarised as well?
+%\begin{table}[]
+%  \begin{tabular}{lllll}
+%                    & Coef      & Std. Error & t-value   & Pr(\textgreater{}|t|) \\
+%  Area Harvested    & 7.418E-01 & 1.003E-02  & 7.390E+01 & 2.000E-16             \\
+%  Water Used        & 8.660E-02 & 8.885E-03  & 9.747E+00 & 2.000E-16             \\
+%  Scope 1 Emissions & 6.731E-02 & 8.013E-03  & 8.400E+00 & 2.000E-16            
+%  \end{tabular}
+%  \end{table}
+%
 \begin{table}[]
@@
   Variable       & Df & Sum Sq    & Mean Sq   & F Value   & Pr(\textgreater{}F) \\
-  Year                & 9  & 7.060E+01 & 7.800E+00 & 8.353E+01 & \textless 2.20E-16 *** \\
-  GI Region           & 54 & 1.507E+03 & 2.790E+01 & 2.972E+02 & \textless 2.20E-16 *** \\
-  Area Harvested      & 1  & 3.211E+03 & 3.211E+03 & 3.419E+04 & \textless 2.20E-16 *** \\
-  Water Used          & 1  & 1.040E+01 & 1.040E+01 & 1.103E+02 & \textless 2.20E-16 *** \\
-  Scope One Emissions & 1  & 6.600E+00 & 6.600E+00 & 7.056E+01 & \textless 2.20E-16 ***
+  Year                & 9  & 7.060E+01 & 7.800E+00 & 8.353E+01 & \textless 2.20E-16 \\
+  GI Region           & 54 & 1.507E+03 & 2.790E+01 & 2.972E+02 & \textless 2.20E-16 \\
+  Area Harvested      & 1  & 3.211E+03 & 3.211E+03 & 3.419E+04 & \textless 2.20E-16 \\
+  Water Used          & 1  & 1.040E+01 & 1.040E+01 & 1.103E+02 & \textless 2.20E-16 \\
+  Scope One Emissions & 1  & 6.600E+00 & 6.600E+00 & 7.056E+01 & \textless 2.20E-16
   \end{tabular}
 \end{table}
-
 \begin{table}[]
@@
 \end{table}
-
+%
+The comparison of models performance shows that the average price per tonne of grapes describes a great deal of the relationship between predictor and response when comparing model 2 to model 4 (see Table 10).
+This relationship between yield and average price was also illustrated in the correlation values between them (see Table 2).
+%
 \begin{table}[]

### 2023-07-13 article.tex
Please do end of day commitsgit add analysis.r  you have no idea what you change dotherwise :|

@@
 
+% for adjusting table width automatically
+\usepackage{adjustbox}
+\usepackage{tabulary, ragged2e}
+\usepackage{booktabs}
+
 \journal{Geography and Sustainability}
+%1.Full Length Article A full-length article should be a substantial and in-depth research study regarding a particular state of issue through several techniques or approaches. 
+% The main text should be approximately 6,000 words in length, but it should not exceed 8,000 words (excluding abstract, references, tables, figures, and appendices).
+%A maximum of 250 words abstract and up to 10 displayed items (figures and tables) is allowed. A full-length article should include an Introduction,
+%Materials and methods, Results, Discussion, Conclusions, and References, which can be accompanied by Supplementary material.
 
@@
 After fitting to the data, each model was validated using k-fold cross validation.
-
 \begin{table}[]
-    \begin{center}
+  %\centering
+    %\begin{center}
       \caption{Summary of models; their predictors, covariates and variable interactions.}
       \label{tab:tab1}
-      \begin{tabular}{ccccc}
-        & Response & Predictors & Covariates & Interactions \\
-        Model 1 & Yield & Water Used, Scope 1 Emissions & Area Harvested, Year, GI Region & N\textbackslash A \\
-        Model 2 & ${ \textrm{Yield}}\over{ \textrm{Area Harvested}}$ & Water Used, Scope 1 Emissions & Area Harvested, Year, GI Region & Area Harvested * Scope 1 Emissions, Area Harvested * Water Use, Year * Region \\
-        Model 3 & \small{$\textrm{Yield} {\times} \textrm{Average Sale Price}$} &  Water Used, Scope 1 Emissions& Area Harvested, Year, GI Region & Area Harvested * Scope 1 Emissions, Area Harvested * Water Use, Year * Region  \\
-        Model 4 & ${\textrm{Yield} \times \textrm{Average Sale Price}}\over{\textrm{Area Harvested}}$ & Water Used, Scope 1 Emissions & Area Harvested, Year, GI Region  & N\textbackslash A
-      \end{tabular}
-    \end{center}
+        \resizebox{\textwidth}{!}{\begin{tabular}{@{}ccccc@{}}
+          \toprule
+           & Response & Predictors & Covariates & Interactions \\ \midrule
+          Model 1 & Yield & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope 1 Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & N/A \\
+          \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
+          Model 2 & ${ \textrm{Yield}}\over{ \textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope 1 Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * Scope 1 Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular} \\
+          \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
+          Model 3 & {$\textrm{Yield} {\times} \textrm{Average Sale Price}$} & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope 1 Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & N/A \\
+          \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
+          Model 4 & ${\textrm{Yield} \times \textrm{Average Sale Price}}\over{\textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope 1 Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * Scope 1 Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular}
+          \end{tabular}}
+    %\end{adjustbox}
+    %\end{center}
 \end{table}
-
 \subsection{Analysis}
@@
 % \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2012,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
-
 \subsection{Total Emissions}
@@
 Strong relationships were found to be present, as all P-values were considered significant (< 2.200E-16, see Tables 2 and 3), except for the non-transformed values for water used (see Table 4).
-
-\begin{table}[]
-  \caption{Summary of models, their predictors, covariates and variable interactions.}
-  \label{tab:tab2}
-  \begin{tabular}{cccccccc}
-  Variable                             & Yield      & Area       & Water Used & Scope One Emissions & $\textrm{Yield}\over\textrm{Area}$ & Average Price Per Tonne & ${\textrm{Average Price per tonne}\over\textrm{Area}}$ \\
-  Yield                                & 1.000E+00  & 7.440E-01  & -4.309E-03 & 7.290E-01           & 3.500E-01            & -2.262E-01              & -1.644E-01                           \\
-  Area                                 & 7.440E-01  & 1.000E+00  & -5.331E-03 & 8.921E-01           & 7.854E-02            & -1.178E-01              & -2.042E-01                           \\
-  Water Used                           & -4.309E-03 & -5.331E-03 & 1.000E+00  & -1.929E-03          & -5.600E-03           & -3.562E-02              & -2.669E-02                           \\
-  Scope One Emissions                  & 7.290E-01  & 8.921E-01  & -1.929E-03 & 1.000E+00           & 9.357E-02            & -9.422E-02              & -1.933E-01                           \\
-  $\textrm{Yield}\over\textrm{Area}$                 & 3.500E-01  & 7.854E-02  & -5.600E-03 & 9.357E-02           & 1.000E+00            & -4.849E-01              & -1.698E-01                           \\
-  Average Price Per Tonne              & -2.262E-01 & -1.178E-01 & -3.562E-02 & -9.422E-02          & -4.849E-01           & 1.000E+00               & 4.732E-01                            \\
-  ${\textrm{Average Price per tonne}\over\textrm{Area}}$ & -1.644E-01 & -2.042E-01 & -2.669E-02 & -1.933E-01          & -1.698E-01           & 4.732E-01               & 1.000E+00
-  \end{tabular}
-  \end{table}
-
-  \begin{table}[]
-    \caption{Pearson correlation coefficients for each logarithmically transformed variable.}
-    \label{tab:tab3}
-    \begin{tabular}{clllllll}
-    Variable                                           & \multicolumn{1}{c}{Yield} & \multicolumn{1}{c}{Area} & \multicolumn{1}{c}{Water Used} & \multicolumn{1}{c}{Scope One Emissions} & \multicolumn{1}{c}{$\textrm{Yield}\over\textrm{Area}$} & \multicolumn{1}{c}{Average Price Per Tonne} & \multicolumn{1}{c}{${\textrm{Average Price per tonne}\over\textrm{Area}}$} \\
-    Yield                                              & 1.000E+00                 & 8.822E-01                & 8.245E-01                      & 7.617E-01                               & 9.353E-01                                          & -4.591E-01                                  & -8.918E-01                                                             \\
-    Area                                               & 8.822E-01                 & 1.000E+00                & 7.750E-01                      & 8.311E-01                               & 6.742E-01                                          & -1.911E-01                                  & -8.474E-01                                                             \\
-    Water Used                                         & 8.245E-01                 & 7.750E-01                & 1.000E+00                      & 6.668E-01                               & 7.292E-01                                          & -4.881E-01                                  & -8.300E-01                                                             \\
-    Scope One Emissions                                & 7.617E-01                 & 8.311E-01                & 6.668E-01                      & 1.000E+00                               & 6.086E-01                                          & -1.559E-01                                  & -7.063E-01                                                             \\
-    $\textrm{Yield}\over\textrm{Area}$                     & 9.353E-01                 & 6.742E-01                & 7.292E-01                      & 6.086E-01                               & 1.000E+00                                          & -5.625E-01                                  & -8.076E-01                                                             \\
-    Average Price Per Tonne                            & -4.591E-01                & -1.911E-01               & -4.881E-01                     & -1.559E-01                              & -5.625E-01                                         & 1.000E+00                                   & 6.592E-01                                                              \\
-    ${\textrm{Average Price per tonne}\over\textrm{Area}}$ & -8.918E-01                & -8.474E-01               & -8.300E-01                     & -7.063E-01                              & -8.076E-01                                         & 6.592E-01                                   & 1.000E+00
-    \end{tabular}
-    \end{table}
-
-    \begin{table}[]
-      \caption{P-values for the non-transformed water used variable's Pearson correlation coefficients.}
-    \label{tab:tab4}
-      \begin{tabular}{cl}
-      Variable                                           & Water Used \\
-      Yield                                              & 7.538E-01  \\
-      Area                                               & 6.981E-01  \\
-      Scope One Emissions                                & 8.883E-01  \\
-      $\textrm{Yield}\over\textrm{Area}$                     & 6.836E-01  \\
-      Average Price Per Tonne                            & 5.600E-02  \\
-      ${\textrm{Average Price per tonne}\over\textrm{Area}}$ & 1.522E-01 
-      \end{tabular}
-      \end{table}
-
-    \begin{table}[]
-      \caption{Summary statistics for each variable on the original scale..}
-      \label{tab:tab5}
-      \begin{tabular}{clllllll}
-      Variable &
-        \multicolumn{1}{c}{Yield} &
-        \multicolumn{1}{c}{Area} &
-        \multicolumn{1}{c}{Water Used} &
-        \multicolumn{1}{c}{Scope One Emissions} &
-        \multicolumn{1}{c}{$\textrm{Yield}\over\textrm{Area}$} &
-        \multicolumn{1}{c}{Average Price Per Tonne} &
-        \multicolumn{1}{c}{${\textrm{Average Price per tonne}\over\textrm{Area}}$} \\
-      Yield                                              & 1.000E+00  & 8.822E-01  & 8.245E-01  & 7.617E-01  & 9.353E-01  & -4.591E-01 & -8.918E-01 \\
-      Area                                               & 8.822E-01  & 1.000E+00  & 7.750E-01  & 8.311E-01  & 6.742E-01  & -1.911E-01 & -8.474E-01 \\
-      Water Used                                         & 8.245E-01  & 7.750E-01  & 1.000E+00  & 6.668E-01  & 7.292E-01  & -4.881E-01 & -8.300E-01 \\
-      Scope One Emissions                                & 7.617E-01  & 8.311E-01  & 6.668E-01  & 1.000E+00  & 6.086E-01  & -1.559E-01 & -7.063E-01 \\
-      $\textrm{Yield}\over\textrm{Area}$                     & 9.353E-01  & 6.742E-01  & 7.292E-01  & 6.086E-01  & 1.000E+00  & -5.625E-01 & -8.076E-01 \\
-      Average Price Per Tonne                            & -4.591E-01 & -1.911E-01 & -4.881E-01 & -1.559E-01 & -5.625E-01 & 1.000E+00  & 6.592E-01  \\
-      ${\textrm{Average Price per tonne}\over\textrm{Area}}$ & -8.918E-01 & -8.474E-01 & -8.300E-01 & -7.063E-01 & -8.076E-01 & 6.592E-01  & 1.000E+00 
-      \end{tabular}
-      \end{table}
-
 \subsection{General Linear Models}
-% The tables need to be in the appropriate sections!!
-%
-%Where are the coefficients???? You cant use the models without them!!
-% You need to actually be able to use this research and look at what these
-% models are showing.
-%
-% What are these tables?
-% what are these models?
-% What are F values?
-% Why are some significant?
-% What are degrees of freedom
-% what are mean sq values?
-% What are sum sq values?
-% What are the ramifications of any of this???
-% Break this down into each model separately!! DO NOT GENERALISE
-% You can draw similarities between the model, as this is not a generalisation.
-%
-%
-%
-% What is a general linear model
-% what does this mean
 To describe the response variables relative to the predictors' values, General Linear Models were used.
@@
 \newline
-% F-tests
-To determine if individually each model's predictors were significantly linearly related to the response variables, F-tests were conducted.
-Models 1 and 2 showed significant linear relationships between each of the predictors and their response variables (see Tables 6 and 7).
-Predictors in models 3 and 4 reported similar significance; except for scope 1 emissions (see Tables 8 and 9).
-Scope one emissions was included in all models to directly compare the response variables as ratios of vineyard size to raw values.
+Each model showed a strong relationship between the predictors and the response (see Table \ref{tab:modelsummary}).
+The strength of the models was measured in $R^2$, as this allowed an easy comparison between their performances and their validation.
+\begin{table}[]
+  \caption{Summary of models; their performance, F-statistics and Residual error.}
+      \label{tab:modelsummary}
+      \resizebox{\textwidth}{!}{\begin{tabular}{@{}cccccccc@{}}
+        \toprule
+         & $R^2$ & Adjusted $R^2$ & F-statistic & P-Value & \begin{tabular}[c]{@{}c@{}}Residual\\ Standard Error\end{tabular} & \begin{tabular}[c]{@{}c@{}}Residual Sum\\ of Squares\end{tabular} & \begin{tabular}[c]{@{}c@{}}Residual Mean\\ of Squares\end{tabular} \\ \midrule
+        Model 1 & 9.072E-01 & 9.061E-01 & 7.753E+02 & 2.200e-16 & 3.065E-01 & 4.913E+02 & 1.000E-01 \\
+        Model 2 & 7.951E-01 & 7.770E-01 & 4.403E+01 & 2.200e-16 & 4.722E-01 & 1.085E+03 & 2.200E-01 \\
+        Model 3 & 9.753E-01 & 9.748E-01 & 1.885E+03 & 2.200e-16 & 1.589E-01 & 7.111E+01 & 3.000E-02 \\
+        Model 4 & 9.669E-01 & 9.638E-01 & 3.095E+02 & 2.200e-16 & 1.904E-01 & 9.528E+01 & 4.000E-02 \\ \bottomrule
+        \end{tabular}}
+  \end{table}
+\subsubsection{F-tests}
+To determine if predictors significantly related to a Model's response variable, F-tests were conducted.
+Models 1 and 2 showed significant linear relationships present between all predictors and their response variables (see Tables 6 and 7).
+Predictors in Models 3 and 4 reported the same significance for each variable except for scope 1 emissions (see Tables 8 and 9).
+Scope 1 emissions was included in all models to directly compare the response variables as ratios of vineyard size to raw values.
 Even though not significant within models 3 and 4, when using the Pearson Correlation Coefficients, scope one emissions was strongly correlated to every Model's response variable; this was especially so for Model 1 and 4 (Yield and average price per tonne as a ratio to area harvested, respectively).
-\newline
-% t-tests
+\subsubsection{t-tests}
 Predictors were also reviewed using t-tests to determine if a predictor significantly contributed to the model when accounting for other variables.
-We can consider this the marginal significance of each variable.
-Using t-tests allowed a more granular examination of each factor (or categorical variable) and interaction, showing that some years and areas contributed marginally significantly and some did not (see appendix).
-With regard to continuous variables, Model 1 and 2 showed water use, scope 1 emissions and area harvested were marginally significant.
-For Model 3 all continuous variables except scope 1 emissions were significant.Model 4 showed scope 1 emissions and water use to only be significant when considered as an interaction with area harvested but not as individual variables (see Table ?).
-% should these tables all go into the appendix solely or should they be summarised as well?
-%\begin{table}[]
-%  \begin{tabular}{lllll}
-%                    & Coef      & Std. Error & t-value   & Pr(\textgreater{}|t|) \\
-%  Area Harvested    & 7.418E-01 & 1.003E-02  & 7.390E+01 & 2.000E-16             \\
-%  Water Used        & 8.660E-02 & 8.885E-03  & 9.747E+00 & 2.000E-16             \\
-%  Scope 1 Emissions & 6.731E-02 & 8.013E-03  & 8.400E+00 & 2.000E-16            
-%  \end{tabular}
-%  \end{table}
-%
-\begin{table}[]
-  \label{tab:tab6}
-  \caption{Model 1 ANOVA summarising variable significance at the .5 level.}
-  \begin{tabular}{llllll}
-  Variable       & Df & Sum Sq    & Mean Sq   & F Value   & Pr(\textgreater{}F) \\
-  Year                & 9  & 7.060E+01 & 7.800E+00 & 8.353E+01 & \textless 2.20E-16 \\
-  GI Region           & 54 & 1.507E+03 & 2.790E+01 & 2.972E+02 & \textless 2.20E-16 \\
-  Area Harvested      & 1  & 3.211E+03 & 3.211E+03 & 3.419E+04 & \textless 2.20E-16 \\
-  Water Used          & 1  & 1.040E+01 & 1.040E+01 & 1.103E+02 & \textless 2.20E-16 \\
-  Scope One Emissions & 1  & 6.600E+00 & 6.600E+00 & 7.056E+01 & \textless 2.20E-16
-  \end{tabular}
-\end{table}
-\begin{table}[]
-  \label{tab:tab7}
-  \caption{Model 2 ANOVA summarising variable significance at the .5 level.}
-  \begin{tabular}{llllll}
-  Variable                    & Df  & Sum Sq    & Mean Sq   & F Value   & Pr(\textgreater{}F)    \\
-  Area Harvested              & 1   & 2.407E+03 & 2.407E+03 & 1.080E+04 & \textless 2.20E-16 *** \\
-  Scope One Emissions         & 1   & 3.989E+01 & 3.989E+01 & 1.789E+02 & \textless 2.20E-16 *** \\
-  Water Used                  & 1   & 5.500E+02 & 5.500E+02 & 2.467E+03 & \textless 2.20E-16 *** \\
-  Area Harvested*Scope One Emissions & 1 & 6.921E+01 & 6.921E+01 & 3.104E+02 & \textless 2.20E-16 *** \\
-  Area Harvested * Water Used & 1   & 1.040E+00 & 1.040E+00 & 4.686E+00 & 3.045E-02 **           \\
-  Year * GI Region            & 424 & 1.144E+03 & 2.700E+00 & 1.210E+01 & \textless 2.20E-16 ***
-  \end{tabular}
-\end{table}
-%
-The comparison of models performance shows that the average price per tonne of grapes describes a great deal of the relationship between predictor and response when comparing model 2 to model 4 (see Table 10).
+Using t-tests allowed a more granular examination of each factor (or categorical variable) and interaction, showing which specific years and areas contributed significantly and which did not (see appendix).
+With regard to continuous variables, Model 1 and 2 showed water use, scope 1 emissions and area harvested were significant when accounting for other variables.
+T-tests for Model 3 showed all continuous variables except scope 1 emissions were significant.
+Model 4 showed scope 1 emissions and water use to only be significant when considered as an interaction with area harvested but not when considered on their own (see Table ?).
+%\subsubsection{Model Comparison}
+% wHY ARE WE COMPARING MODELS?
+%Models were compared using $R^2$ values.
+The comparison of models performance showed that the average price per tonne of grapes describes a great deal of the relationship between predictor and response when comparing model 2 to model 4 (see Table 10).
 This relationship between yield and average price was also illustrated in the correlation values between them (see Table 2).
-%
-\begin{table}[]
-    \caption{Model 3 ANOVA summarising variable significance at the .5 level.}
-    \label{tab:tab8}
-    \begin{tabular}{llllll}
-    Variable            & Df & Sum Sq    & Mean Sq   & F Value   & Pr(\textgreater{}F)    \\
-    Year                & 6  & 1.324E+01 & 2.210E+00 & 8.748E+01 & \textless 2.20E-16 *** \\
-    GI Region           & 50 & 6.498E+02 & 1.300E+01 & 5.151E+02 & \textless 2.20E-16 *** \\
-    Area Harvested      & 1  & 2.142E+03 & 2.142E+03 & 8.491E+04 & \textless 2.20E-16 *** \\
-    Water Used          & 1  & 3.200E-01 & 3.200E-01 & 1.259E+01 & 3.947E-04 **           \\
-    Scope One Emissions & 1  & 4.000E-02 & 4.000E-02 & 1.492E+00 & 2.221E-01             
-    \end{tabular}
-\end{table}
 
-\begin{table}[]
-  \label{tab:tab9}
-  \caption{Model 4 ANOVA summarising variable significance at the .5 level.}
-  \begin{tabular}{llllll}
-  Variable            & Df  & Sum Sq    & Mean Sq   & F Value   & Pr(\textgreater{}F)    \\
-  Area Harvested      & 1   & 2.066E+03 & 2.066E+03 & 5.700E+04 & \textless 2.20E-16 *** \\
-  Scope One Emissions & 1   & 6.000E-02 & 6.000E-02 & 1.569E+00 & 2.105E-01              \\
-  Water Used          & 1   & 2.014E+02 & 2.014E+02 & 5.557E+03 & \textless 2.20E-16 *** \\
-  Area Harvested*Scope One Emissions & 1 & 5.246E+01 & 5.246E+01 & 1.448E+03 & \textless 2.20E-16 *** \\
-  Area Harvested * Water Used        & 1 & 7.270E+00 & 7.270E+00 & 2.005E+02 & \textless 2.20E-16 *** \\
-  Year * GI Region    & 243 & 4.546E+02 & 1.870E+00 & 5.162E+01 & \textless 2.20E-16 ***
-  \end{tabular}
-\end{table}
-
-\begin{table}[]
-  \label{tab:tab 10}
-  \caption{Comparison of Model Residuals}
-  \begin{tabular}{llll}
-        & Df   & Sum Sq    & Mean Sq   \\
-  Model 1 & 5231 & 4.913E+02 & 1.000E-01 \\
-  Model 2 & 4868 & 1.085E+03 & 2.200E-01 \\
-  Model 3 & 2818 & 7.111E+01 & 3.000E-02 \\
-  Model 4 & 2629 & 9.528E+01 & 4.000E-02
-  \end{tabular}
-\end{table}
-
-\begin{table}[]
-  \caption{Comparison of Model performance.}
-  \label{tab:tab 11}
-  \begin{tabular}{llllll}
-          & RSE       & R2        & Adjusted 			R2 & F-statistic & P-Value           \\
-  Model 1 & 3.065E-01 & 9.072E-01 & 9.061E-01      & 7.753E+02   & \textless 2.2e-16 \\
-  Model 2 & 4.722E-01 & 7.951E-01 & 7.770E-01      & 4.403E+01   & \textless 2.2e-16 \\
-  Model 3 & 1.589E-01 & 9.753E-01 & 9.748E-01      & 1.885E+03   & \textless 2.2e-16 \\
-  Model 4 & 1.904E-01 & 9.669E-01 & 9.638E-01      & 3.095E+02   & \textless 2.2e-16
-  \end{tabular}
-\end{table}
-
-Limitations included overestimating yield for models 1 and 2, (see Figures 1 and 2) and underestimating crop value in models 3 and 4 (see Figures 3 and 4).
 % This section needs to be heavily revised. Why are there tails on the distributions? What does a heavy tail indicate?
@@
   \caption{Model validation using k-fold cross validation, for 10 folds repeated 100 times.}
-  \begin{tabular}{llll}
+  \begin{tabular}{cccc}
           & RMSE      & R2        & MAE       \\
@@
 \section{Discussion}
+\subsection{Limitations}
+Limitations included overestimating yield for models 1 and 2, (see Figures 1 and 2) and underestimating crop value in models 3 and 4 (see Figures 3 and 4).
 % this needs serious revision.
@@
 %% appendix sections are then done as normal sections
-%% \appendix
+ \appendix
 
+ \begin{table}[]
+  \centering
+  \caption{Summary of models, their predictors, covariates and variable interactions.}
+  \label{tab:tab2}
+  \begin{tabulary}{\textwidth}{CCCCCCCC}
+  Variable                             & Yield      & Area       & Water Used & Scope One Emissions & $\textrm{Yield}\over\textrm{Area}$ & Average Price Per Tonne & ${\textrm{Average Price per tonne}\over\textrm{Area}}$ \\
+  Yield                                & 1.000E+00  & 7.440E-01  & -4.309E-03 & 7.290E-01           & 3.500E-01            & -2.262E-01              & -1.644E-01                           \\
+  Area                                 & 7.440E-01  & 1.000E+00  & -5.331E-03 & 8.921E-01           & 7.854E-02            & -1.178E-01              & -2.042E-01                           \\
+  Water Used                           & -4.309E-03 & -5.331E-03 & 1.000E+00  & -1.929E-03          & -5.600E-03           & -3.562E-02              & -2.669E-02                           \\
+  Scope One Emissions                  & 7.290E-01  & 8.921E-01  & -1.929E-03 & 1.000E+00           & 9.357E-02            & -9.422E-02              & -1.933E-01                           \\
+  $\textrm{Yield}\over\textrm{Area}$                 & 3.500E-01  & 7.854E-02  & -5.600E-03 & 9.357E-02           & 1.000E+00            & -4.849E-01              & -1.698E-01                           \\
+  Average Price Per Tonne              & -2.262E-01 & -1.178E-01 & -3.562E-02 & -9.422E-02          & -4.849E-01           & 1.000E+00               & 4.732E-01                            \\
+  ${\textrm{Average Price per tonne}\over\textrm{Area}}$ & -1.644E-01 & -2.042E-01 & -2.669E-02 & -1.933E-01          & -1.698E-01           & 4.732E-01               & 1.000E+00
+  \end{tabulary}
+  \end{table}
+
+  \begin{table}[]
+    \caption{Pearson correlation coefficients for each logarithmically transformed variable.}
+    \label{tab:tab3}
+    \begin{tabular}{clllllll}
+    Variable                                           & \multicolumn{1}{c}{Yield} & \multicolumn{1}{c}{Area} & \multicolumn{1}{c}{Water Used} & \multicolumn{1}{c}{Scope One Emissions} & \multicolumn{1}{c}{$\textrm{Yield}\over\textrm{Area}$} & \multicolumn{1}{c}{Average Price Per Tonne} & \multicolumn{1}{c}{${\textrm{Average Price per tonne}\over\textrm{Area}}$} \\
+    Yield                                              & 1.000E+00                 & 8.822E-01                & 8.245E-01                      & 7.617E-01                               & 9.353E-01                                          & -4.591E-01                                  & -8.918E-01                                                             \\
+    Area                                               & 8.822E-01                 & 1.000E+00                & 7.750E-01                      & 8.311E-01                               & 6.742E-01                                          & -1.911E-01                                  & -8.474E-01                                                             \\
+    Water Used                                         & 8.245E-01                 & 7.750E-01                & 1.000E+00                      & 6.668E-01                               & 7.292E-01                                          & -4.881E-01                                  & -8.300E-01                                                             \\
+    Scope One Emissions                                & 7.617E-01                 & 8.311E-01                & 6.668E-01                      & 1.000E+00                               & 6.086E-01                                          & -1.559E-01                                  & -7.063E-01                                                             \\
+    $\textrm{Yield}\over\textrm{Area}$                     & 9.353E-01                 & 6.742E-01                & 7.292E-01                      & 6.086E-01                               & 1.000E+00                                          & -5.625E-01                                  & -8.076E-01                                                             \\
+    Average Price Per Tonne                            & -4.591E-01                & -1.911E-01               & -4.881E-01                     & -1.559E-01                              & -5.625E-01                                         & 1.000E+00                                   & 6.592E-01                                                              \\
+    ${\textrm{Average Price per tonne}\over\textrm{Area}}$ & -8.918E-01                & -8.474E-01               & -8.300E-01                     & -7.063E-01                              & -8.076E-01                                         & 6.592E-01                                   & 1.000E+00
+    \end{tabular}
+    \end{table}
+
+    \begin{table}[]
+      \caption{P-values for the non-transformed water used variable's Pearson correlation coefficients.}
+    \label{tab:tab4}
+      \begin{tabular}{cl}
+      Variable                                           & Water Used \\
+      Yield                                              & 7.538E-01  \\
+      Area                                               & 6.981E-01  \\
+      Scope One Emissions                                & 8.883E-01  \\
+      $\textrm{Yield}\over\textrm{Area}$                     & 6.836E-01  \\
+      Average Price Per Tonne                            & 5.600E-02  \\
+      ${\textrm{Average Price per tonne}\over\textrm{Area}}$ & 1.522E-01 
+      \end{tabular}
+      \end{table}
+
+    \begin{table}[]
+      \caption{Summary statistics for each variable on the original scale..}
+      \label{tab:tab5}
+      \begin{tabular}{clllllll}
+      Variable &
+        \multicolumn{1}{c}{Yield} &
+        \multicolumn{1}{c}{Area} &
+        \multicolumn{1}{c}{Water Used} &
+        \multicolumn{1}{c}{Scope One Emissions} &
+        \multicolumn{1}{c}{$\textrm{Yield}\over\textrm{Area}$} &
+        \multicolumn{1}{c}{Average Price Per Tonne} &
+        \multicolumn{1}{c}{${\textrm{Average Price per tonne}\over\textrm{Area}}$} \\
+      Yield                                              & 1.000E+00  & 8.822E-01  & 8.245E-01  & 7.617E-01  & 9.353E-01  & -4.591E-01 & -8.918E-01 \\
+      Area                                               & 8.822E-01  & 1.000E+00  & 7.750E-01  & 8.311E-01  & 6.742E-01  & -1.911E-01 & -8.474E-01 \\
+      Water Used                                         & 8.245E-01  & 7.750E-01  & 1.000E+00  & 6.668E-01  & 7.292E-01  & -4.881E-01 & -8.300E-01 \\
+      Scope One Emissions                                & 7.617E-01  & 8.311E-01  & 6.668E-01  & 1.000E+00  & 6.086E-01  & -1.559E-01 & -7.063E-01 \\
+      $\textrm{Yield}\over\textrm{Area}$                     & 9.353E-01  & 6.742E-01  & 7.292E-01  & 6.086E-01  & 1.000E+00  & -5.625E-01 & -8.076E-01 \\
+      Average Price Per Tonne                            & -4.591E-01 & -1.911E-01 & -4.881E-01 & -1.559E-01 & -5.625E-01 & 1.000E+00  & 6.592E-01  \\
+      ${\textrm{Average Price per tonne}\over\textrm{Area}}$ & -8.918E-01 & -8.474E-01 & -8.300E-01 & -7.063E-01 & -8.076E-01 & 6.592E-01  & 1.000E+00 
+      \end{tabular}
+      \end{table}
+% should these tables all go into the appendix solely or should they be summarised as well?
+%\begin{table}[]
+%  \begin{tabular}{lllll}
+%                    & Coef      & Std. Error & t-value   & Pr(\textgreater{}|t|) \\
+%  Area Harvested    & 7.418E-01 & 1.003E-02  & 7.390E+01 & 2.000E-16             \\
+%  Water Used        & 8.660E-02 & 8.885E-03  & 9.747E+00 & 2.000E-16             \\
+%  Scope 1 Emissions & 6.731E-02 & 8.013E-03  & 8.400E+00 & 2.000E-16            
+%  \end{tabular}
+%  \end{table}
+%
+\begin{table}[]
+  \label{tab:tab6}
+  \caption{Model 1 ANOVA summarising variable significance at the .5 level.}
+  \begin{tabular}{llllll}
+  Variable       & Df & Sum Sq    & Mean Sq   & F Value   & Pr(\textgreater{}F) \\
+  Year                & 9  & 7.060E+01 & 7.800E+00 & 8.353E+01 & \textless 2.20E-16 \\
+  GI Region           & 54 & 1.507E+03 & 2.790E+01 & 2.972E+02 & \textless 2.20E-16 \\
+  Area Harvested      & 1  & 3.211E+03 & 3.211E+03 & 3.419E+04 & \textless 2.20E-16 \\
+  Water Used          & 1  & 1.040E+01 & 1.040E+01 & 1.103E+02 & \textless 2.20E-16 \\
+  Scope One Emissions & 1  & 6.600E+00 & 6.600E+00 & 7.056E+01 & \textless 2.20E-16
+  \end{tabular}
+\end{table}
+\begin{table}[]
+  \label{tab:tab7}
+  \caption{Model 2 ANOVA summarising variable significance at the .5 level.}
+  \begin{tabular}{llllll}
+  Variable                    & Df  & Sum Sq    & Mean Sq   & F Value   & Pr(\textgreater{}F)    \\
+  Area Harvested              & 1   & 2.407E+03 & 2.407E+03 & 1.080E+04 & \textless 2.20E-16 *** \\
+  Scope One Emissions         & 1   & 3.989E+01 & 3.989E+01 & 1.789E+02 & \textless 2.20E-16 *** \\
+  Water Used                  & 1   & 5.500E+02 & 5.500E+02 & 2.467E+03 & \textless 2.20E-16 *** \\
+  Area Harvested*Scope One Emissions & 1 & 6.921E+01 & 6.921E+01 & 3.104E+02 & \textless 2.20E-16 *** \\
+  Area Harvested * Water Used & 1   & 1.040E+00 & 1.040E+00 & 4.686E+00 & 3.045E-02 **           \\
+  Year * GI Region            & 424 & 1.144E+03 & 2.700E+00 & 1.210E+01 & \textless 2.20E-16 ***
+  \end{tabular}
+\end{table}
+%
+\begin{table}[]
+  \caption{Model 3 ANOVA summarising variable significance at the .5 level.}
+  \label{tab:tab8}
+  \begin{tabular}{llllll}
+  Variable            & Df & Sum Sq    & Mean Sq   & F Value   & Pr(\textgreater{}F)    \\
+  Year                & 6  & 1.324E+01 & 2.210E+00 & 8.748E+01 & \textless 2.20E-16 *** \\
+  GI Region           & 50 & 6.498E+02 & 1.300E+01 & 5.151E+02 & \textless 2.20E-16 *** \\
+  Area Harvested      & 1  & 2.142E+03 & 2.142E+03 & 8.491E+04 & \textless 2.20E-16 *** \\
+  Water Used          & 1  & 3.200E-01 & 3.200E-01 & 1.259E+01 & 3.947E-04 **           \\
+  Scope One Emissions & 1  & 4.000E-02 & 4.000E-02 & 1.492E+00 & 2.221E-01             
+  \end{tabular}
+\end{table}
+
+\begin{table}[]
+\label{tab:tab9}
+\caption{Model 4 ANOVA summarising variable significance at the .5 level.}
+\begin{tabular}{llllll}
+Variable            & Df  & Sum Sq    & Mean Sq   & F Value   & Pr(\textgreater{}F)    \\
+Area Harvested      & 1   & 2.066E+03 & 2.066E+03 & 5.700E+04 & \textless 2.20E-16 *** \\
+Scope One Emissions & 1   & 6.000E-02 & 6.000E-02 & 1.569E+00 & 2.105E-01              \\
+Water Used          & 1   & 2.014E+02 & 2.014E+02 & 5.557E+03 & \textless 2.20E-16 *** \\
+Area Harvested*Scope One Emissions & 1 & 5.246E+01 & 5.246E+01 & 1.448E+03 & \textless 2.20E-16 *** \\
+Area Harvested * Water Used        & 1 & 7.270E+00 & 7.270E+00 & 2.005E+02 & \textless 2.20E-16 *** \\
+Year * GI Region    & 243 & 4.546E+02 & 1.870E+00 & 5.162E+01 & \textless 2.20E-16 ***
+\end{tabular}
+\end{table}
+
+\begin{table}[]
+\label{tab:tab 10}
+\caption{Comparison of Model Residuals}
+\begin{tabular}{llll}
+      & Df   & Sum Sq    & Mean Sq   \\
+Model 1 & 5231 & 4.913E+02 & 1.000E-01 \\
+Model 2 & 4868 & 1.085E+03 & 2.200E-01 \\
+Model 3 & 2818 & 7.111E+01 & 3.000E-02 \\
+Model 4 & 2629 & 9.528E+01 & 4.000E-02
+\end{tabular}
+\end{table}
+
+\begin{table}[]
+\caption{Comparison of Model performance.}
+\label{tab:tab 11}
+\begin{tabular}{llllll}
+        & RSE       & R2        & Adjusted 			R2 & F-statistic & P-Value           \\
+Model 1 & 3.065E-01 & 9.072E-01 & 9.061E-01      & 7.753E+02   & \textless 2.2e-16 \\
+Model 2 & 4.722E-01 & 7.951E-01 & 7.770E-01      & 4.403E+01   & \textless 2.2e-16 \\
+Model 3 & 1.589E-01 & 9.753E-01 & 9.748E-01      & 1.885E+03   & \textless 2.2e-16 \\
+Model 4 & 1.904E-01 & 9.669E-01 & 9.638E-01      & 3.095E+02   & \textless 2.2e-16
+\end{tabular}
+\end{table}
+
+%
 %% \section{}

### 2023-07-13 analysis.r

@@ -245,23 +245,4 @@ sbar(1000, lonlat=TRUE, label="1000km")
 north(xy="bottomleft", type=2)
 
 # below is just a way to grab the coefs from a linear model.
-summary(model2)$coefficients
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
+#summary(model2)$coefficients

### 2023-07-17 article.tex
- Please do smaller well tracked updates.

@@
 To enable direct comparisons between fuels, they were converted to tonnes of Carbon Dioxide equivalent.
+% They were converted to emissions that was expressed in TCO2E - not into the units!
 \newline
@@
 This was undertaken for data on the original scale (see Table 2) and for data as a logarithmic transform (see Table 3).
-The logarithmic transforms showed the strongest correlations, likely due to a skew caused by a greater number of smaller vineyards within the dataset (see Table 5).
+The logarithmic transformed data showed the strongest correlations, likely due to a skew caused by a greater number of smaller vineyards within the dataset (see Table 5).
 Transforming data prior to calculating the coefficients changes several things: The logarithmic transform of the data alters the interpretation of the coefficients to percentage change - a coefficient will be indicative of the change in percentage of one variable compared to the other; scaling by standard deviation also changes this interpretation to be a percentage of that variables standard deviation.
@@
 With all this in mind, when considering the logarithmically transformed variables, a coefficient of 1 would indicate that: given the change of one variable by one percentage of its standard deviation, the other variable would change by one percent of its own standard deviation.
-The importance of this is the unitless nature of these relationships and that it can be translated directly to any vineyard's case that has a well known distribution.
+The importance of this is the dimensionless nature of these relationships and that it can be translated directly to any vineyard's case that has a well known distribution.
 \newline
@@
 P-values reflected the significance of a given correlation coefficient when considering its relation to sample size via its incorporation as an element of standard error.
-Strong relationships were found to be present, as all P-values were considered significant (< 2.200E-16, see Tables 2 and 3), except for the non-transformed values for water used (see Table 4).
+Strong relationships were found to be present, as all P-values were considered significant
+% missing tables do we need these in tables?
+($< 2.200E-16$, see Tables ?? and ??)
+% This reference is messed up and next some TLC
+, except for the non-transformed values for water used (see Table 4). 
 \subsection{General Linear Models}
-To describe the response variables relative to the predictors' values, General Linear Models were used.
+General Linear Models were used to describe how response variables related to predictors' values.
 Log transformed variables were used as inputs to these models as they resulted in higher $R^2$ values and described the relationships proportionally; reflecting coefficient values as percentages of a variable's standard deviation.
-\newline
-Each model showed a strong relationship between the predictors and the response (see Table \ref{tab:modelsummary}).
-The strength of the models was measured in $R^2$, as this allowed an easy comparison between their performances and their validation.
 \begin{table}[]
@@
         \toprule
-         & $R^2$ & Adjusted $R^2$ & F-statistic & P-Value & \begin{tabular}[c]{@{}c@{}}Residual\\ Standard Error\end{tabular} & \begin{tabular}[c]{@{}c@{}}Residual Sum\\ of Squares\end{tabular} & \begin{tabular}[c]{@{}c@{}}Residual Mean\\ of Squares\end{tabular} \\ \midrule
-        Model 1 & 9.072E-01 & 9.061E-01 & 7.753E+02 & 2.200e-16 & 3.065E-01 & 4.913E+02 & 1.000E-01 \\
-        Model 2 & 7.951E-01 & 7.770E-01 & 4.403E+01 & 2.200e-16 & 4.722E-01 & 1.085E+03 & 2.200E-01 \\
-        Model 3 & 9.753E-01 & 9.748E-01 & 1.885E+03 & 2.200e-16 & 1.589E-01 & 7.111E+01 & 3.000E-02 \\
-        Model 4 & 9.669E-01 & 9.638E-01 & 3.095E+02 & 2.200e-16 & 1.904E-01 & 9.528E+01 & 4.000E-02 \\ \bottomrule
+          & \textbf{$R^2$}       & \textbf{$Adjusted R^2$} & \textbf{F-statistic} & \textbf{P-Value}     & \textbf{\begin{tabular}[c]{@{}c@{}}Residual\\ Standard Error\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Sum\\ of Squares\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Mean\\ of Squares\end{tabular}} \\ \midrule
+        \textbf{\begin{tabular}[c]{@{}c@{}}Model 1\\ Yield\end{tabular}}      & 9.072E-01            & 9.061E-01               & 7.753E+02            & 2.200e-16            & 3.065E-01                                                                  & 4.913E+02                                                                  & 1.000E-01                                                                   \\
+        \multicolumn{1}{l}{}                                                  & \multicolumn{1}{l}{} & \multicolumn{1}{l}{}    & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{}                                                       & \multicolumn{1}{l}{}                                                       & \multicolumn{1}{l}{}                                                        \\
+        \textbf{\begin{tabular}[c]{@{}c@{}}Model 2\\ Yield/Area\end{tabular}} & 7.951E-01            & 7.770E-01               & 4.403E+01            & 2.200e-16            & 4.722E-01                                                                  & 1.085E+03                                                                  & 2.200E-01                                                                   \\
+        \multicolumn{1}{l}{}                                                  & \multicolumn{1}{l}{} & \multicolumn{1}{l}{}    & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{}                                                       & \multicolumn{1}{l}{}                                                       & \multicolumn{1}{l}{}                                                        \\
+        \textbf{\begin{tabular}[c]{@{}c@{}}Model 3\\ Value\end{tabular}}      & 9.753E-01            & 9.748E-01               & 1.885E+03            & 2.200e-16            & 1.589E-01                                                                  & 7.111E+01                                                                  & 3.000E-02                                                                   \\
+        \multicolumn{1}{l}{}                                                  & \multicolumn{1}{l}{} & \multicolumn{1}{l}{}    & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{}                                                       & \multicolumn{1}{l}{}                                                       & \multicolumn{1}{l}{}                                                        \\
+        \textbf{\begin{tabular}[c]{@{}c@{}}Model 4\\ Value/Area\end{tabular}} & 9.669E-01            & 9.638E-01               & 3.095E+02            & 2.200e-16            & 1.904E-01                                                                  & 9.528E+01                                                                  & 4.000E-02                                                                   \\ \bottomrule
         \end{tabular}}
       \end{table}
+Each model showed a strong relationship between the predictors and the response (see Table \ref{tab:modelsummary}).
+The strength of the models was measured in $R^2$, as this allowed an easy comparison between their performances and their validation.
 \subsubsection{F-tests}
+% \ref{tab:ftests}
+% We need a table with all F-tests.
+%         | Yield   | ...
+% Model 1 | -2e^-16 | ...
+%
 To determine if predictors significantly related to a Model's response variable, F-tests were conducted.
-Models 1 and 2 showed significant linear relationships present between all predictors and their response variables (see Tables 6 and 7).
-Predictors in Models 3 and 4 reported the same significance for each variable except for scope 1 emissions (see Tables 8 and 9).
+Models 1 and 2 showed significant linear relationships present between all predictors and their response variables (see Table \ref{tab:ftests}).
+Predictors in Models 3 and 4 reported the same significance for each variable except for scope 1 emissions.
 Scope 1 emissions was included in all models to directly compare the response variables as ratios of vineyard size to raw values.
-Even though not significant within models 3 and 4, when using the Pearson Correlation Coefficients, scope one emissions was strongly correlated to every Model's response variable; this was especially so for Model 1 and 4 (Yield and average price per tonne as a ratio to area harvested, respectively).
+Even though not significant within models 3 and 4, when using the Pearson Correlation Coefficients scope one emissions was strongly correlated to every Model's response variable; this was especially so for Model 1 and 4 (Yield and average price per tonne as a ratio to area harvested, respectively).
 \subsubsection{t-tests}
-Predictors were also reviewed using t-tests to determine if a predictor significantly contributed to the model when accounting for other variables.
-Using t-tests allowed a more granular examination of each factor (or categorical variable) and interaction, showing which specific years and areas contributed significantly and which did not (see appendix).
+%We want two tables
+% \ref{tab:ttests}
+% \ref{fig:sigregions}
+% a table for t-tests as per the f-tests table (only continuous variables)
+% A map of significant areas contributing to each model.
+T-tests were used to determine if predictors significantly contributed to their models when accounting for other variables; this allowed a more granular examination of interactions and factors within categorical variables, showing which specific years and areas contributed significantly and which did not (see appendix).
 With regard to continuous variables, Model 1 and 2 showed water use, scope 1 emissions and area harvested were significant when accounting for other variables.
 T-tests for Model 3 showed all continuous variables except scope 1 emissions were significant.
-Model 4 showed scope 1 emissions and water use to only be significant when considered as an interaction with area harvested but not when considered on their own (see Table ?).
-%\subsubsection{Model Comparison}
+Model 4 showed scope 1 emissions and water use to only be significant when considered as an interaction with area harvested but not when considered on their own.
+\subsubsection{Model Comparison}
 % wHY ARE WE COMPARING MODELS?
-%Models were compared using $R^2$ values.
-The comparison of models performance showed that the average price per tonne of grapes describes a great deal of the relationship between predictor and response when comparing model 2 to model 4 (see Table 10).
-This relationship between yield and average price was also illustrated in the correlation values between them (see Table 2).
-
+% We compare models to see which predictors contributed relatively to each of the models.
+%
+% There were two subsets - when comparing these models what is the difference between them?
+%
+Models were compared using $R^2$ values.
+The comparison of models performance showed that the average price per tonne of grapes describes a great deal of the relationship between predictor and response when comparing Model 2 to Model 4 (see Table \ref*{tab:modelperformance}).
+% This is important but you do not illustrate why
+%
 % This section needs to be heavily revised. Why are there tails on the distributions? What does a heavy tail indicate?
@@
 
-%\begin{figure}
-%  \includegraphics[width=\linewidth]{figure1.jpg}
-%  \caption{Model 1 residuals vs fitted (left) and QQ plot (right).}
-%  \label{fig:fig1}
-%\end{figure}
+ \subsubsection{Model Coefficients}
+% A map and a table just like the t-test section.
+% ^ This does not work
 
-%\begin{figure}
-%  \includegraphics[width=\linewidth]{figure2.jpg}
-%  \caption{Model 2 residuals vs fitted (left) and QQ plot (right).}
-%  \label{fig:fig2}
-%\end{figure}
-
-%\begin{figure}
-%  \includegraphics[width=\linewidth]{figure3.jpg}
-%  \caption{Model 3 residuals vs fitted (left) and QQ plot (right).}
-%  \label{fig:fig3}
-%\end{figure}
-
-%\begin{figure}
-%  \includegraphics[width=\linewidth]{figure4.jpg}
-%  \caption{Model 4 residuals vs fitted (left) and QQ plot (right).}
-%  \label{fig:fig4}
-%\end{figure}
-
-%\begin{figure}
-%  \includegraphics[width=\linewidth]{figure5.jpg}
-%  \caption{Average Yield per hectare for each GI Region.}
-%  \label{fig:fig5}
-%\end{figure}
-
-%\begin{figure}
-%  \includegraphics[width=\linewidth]{figure6.jpg}
-%  \caption{Average Price Per Tonne for each GI Region.}
-%  \label{fig:fig6}
-%\end{figure}
-
-% It is odd to have a paragraph on its own without a label, it does look a bit lost in the article.
 The correlation between average sales price and yield was a negative trend (see table 2); the contributing factors to yield and average sales price was ???.
@@
 % ALSO THAT CATASTROPHES HAPPEN EVEN ON SMALL SCALES THis needs to be tied up better.
+
+\subsection{Productivity verse Value}
+% Graph of productivity verse value.
+different strategies are likely employed between different regions, where some regions target the mass production of cheaper grapes over quality.
 This is most notable when grouping regions by climate, especially when considering GI Regions in the 'Hot Very Dry' climate (see Figure 7).
@@
 The interaction between year and GI Region likely accounted for localised events such as bushfires, which would be impactful, but only at a local level in both time and space.
-
-%\begin{figure}
-%  \includegraphics[width=\linewidth]{figure7.jpg}
-%  \caption{A depiction of the relationship between Value of tonnes harvested per hectare and yield per hectare.Both variables have been transformed to be scaled by standard deviation with mean 0.}
-%  \label{fig:fig7}
-%\end{figure}
-
 \subsection{Model Validation}
 To validate the performance of these models k-fold cross validation was used.
-This was done using 10 folds, k=10, repeated 100 times.
-The models performed similarly to their original counter parts (see Table 11).
-
-% really this is it?
-
+This was done using 10 folds, $k=10$, repeated 100 times.
+The models performed similarly to their original counter parts (see Table \ref*{tab:kfold}).
 \begin{table}[]
-  \label{tab:tab12}
+  \label{tab:kfold}
   \caption{Model validation using k-fold cross validation, for 10 folds repeated 100 times.}
@@
   \end{table}
-
 \section{Discussion}

### 2023-07-19 data_prep.py
fixed missing 0 in m2 conversion - this did not affect results

@@ -84,7 +84,7 @@ df.loc[df["climate"]=="Unknown Climate Unknown Climate", "climate"] = np.nan
 # We change units
 #
 # changing ha to m2 for area to remove 1s and fractions for the log transform
-df["area_harvested"] = df["area_harvested"]*1000
+df["area_harvested"] = df["area_harvested"]*10000
 
 #create a flag for disease/fire/frost
 df["not_harvested"] = 0

### 2023-07-19 analysis.r
Added Violin graphs for coefs

@@ -20,7 +20,7 @@ model4 <- lm(
     ha_value ~
     area_harvested * ((scope1) + water_used) + data_year_id:giregion
     , data = df)
-
+``
 # We review our first model
 
 # MODEL 1
@@ -35,7 +35,8 @@ ggplot(
     data.frame(
         y = model1_residuals)
         , aes(sample = y)
-       ) +
+       ) +names4 <- rownames(df4)
+
        stat_qq() +
        stat_qq_line()
 ggplot(
@@ -55,9 +56,8 @@ ggplot(
 #Resampling: Cross-Validated (10 fold, repeated 100 times) 
 #Summary of sample sizes: 4768, 4767, 4768, 4769, 4768, 4769, ... 
 #Resampling results:
-#
-#  RMSE      Rsquared   MAE      
-#  0.308865  0.9044952  0.2164945
+#names4 <- rownames(df4)
+.2164945
 #
 #Tuning parameter 'intercept' was held constant at a value of TRUE
 
@@ -105,7 +105,8 @@ ggplot(
 # We add Value to the model :P
 anova(model3)
 summary(model3)
-coefs3 <- coefs(model3)
+coefs3 <- coefs(model3)names4 <- rownames(df4)
+
 coefs3_giregion <- coefs3[grep("gi*", row.names(ok)), ]
 coefs3_year <- coefs3[grep("data*", row.names(ok)), ]
 model3_residuals <- residuals(model3)
@@ -120,11 +121,32 @@ ggplot(
 ggplot(
     data.frame(
         y = model3_residuals
-        , x = model3_fitted)
-    , aes(x = x, y = y)
-    ) +
-    geom_point() +
-    geom_hline(yintercept = 0)
+        , x = model3_fitted)df3 <- data.frame(summary(model3)$coefficients)
+names3 <- rownames(df3) 
+df3[names3[grepl("^da", names3)], "class"] <- "Year"
+df3[names3[grepl("^gi", names3)], "class"] <- "GI Region"
+df3 <- df3[complete.cases(df3),]
+ggplot(df3, aes(class, Estimate)) + geom_violin()
+
+df3 <- data.frame(summary(model3)$coefficients)
+names3 <- rownames(df
+df4 <- data.frame(summary(model4)$coefficients)
+names4 <- rownames(df4)
+df4[names4[grepl("^da", names4)], "class"] <- "Year * GI Region"
+df4 <- df4[complete.cases(df4),]
+
[…]

### 2023-07-19 analysis.r
Better Violin Plots!

@@ -145,7 +145,7 @@ ggplot(df4, aes(class, Estimate)) + ggplot(df2, aes(class, Estimate)) + geom_vio
 df3[names3[grepl("^da", names3)], "class"] <- "Year"
 df3[names3[grepl("^gi", names3)], "class"] <- "GI Region"
 df3 <- df3[complete.cases(df3),]
-ggplot(df3, aes(class, Estimate)) + geom_violin()
+ggplot(df3, aes(class, Estimate)) + geom_violin() + scale_fill_brewer(palette="Dark2") + theme_minimal()
 
 # Residual standard error: 0.1688 on 1962 degrees of freedom
 #   (2877 observations deleted due to missingness)
@@ -190,17 +190,29 @@ ggplot(
     geom_point() +
     geom_hline(yintercept = 0)
 
-df4 <- data.frame(summary(model4)$coefficients)
-names4 <- rownames(df4)
-df4[names4[grepl("^da", names4)], "class"] <- "Year * GI Region"
-df4 <- df4[complete.cases(df4),]
+df1 <- data.frame(summary(model1)$coefficients)
+names1 <- rownames(df1) 
+df1[names3[grepl("^da", names1)], "class"] <- "Model 1: Year"
+df1[names1[grepl("^gi", names1)], "class"] <- "model 1: GI Region"
+df1 <- df1[complete.cases(df1),]
 
 df2 <- data.frame(summary(model2)$coefficients)
 names2 <- rownames(df2)
-df2[names2[grepl("^da", names2)], "class"] <- "Year * GI Region"
+df2[names2[grepl("^da", names2)], "class"] <- "Model 2: Year * GI Region"
 df2 <- df2[complete.cases(df2),]
 
-ggplot(df4, aes(class, Estimate)) + ggplot(df2, aes(class, Estimate)) + geom_violin() + geom_boxplot(width=0.1, fill="white") + theme_classic()
+df3 <- data.frame(summary(model3)$coefficients)
+names3 <- rownames(df3) 
+df3[names3[grepl("^da", names3)], "class"] <- "Model 3: Year"
+df3[names3[grepl("^gi", names3)], "class"] <- "model 3: GI Region"
+df3 <- df3[complete.cases(df3),]
+
+df4 <- data.frame(summary(model4)$coefficients)
+names4 <- rownames(df4)
+df4[names4[grepl("^da", names4)], "class"] <- "Model 4: Year * GI Region"
+df4 <- df4[complete.cases(df4),]
+
[…]

### 2023-07-20 analysis.r
Removal of erroneous remnants code

@@ -20,7 +20,7 @@ model4 <- lm(
     ha_value ~
     area_harvested * ((scope1) + water_used) + data_year_id:giregion
     , data = df)
-``
+
 # We review our first model
 
 # MODEL 1
@@ -35,10 +35,10 @@ ggplot(
     data.frame(
         y = model1_residuals)
         , aes(sample = y)
-       ) +names4 <- rownames(df4)
-
+       ) +
        stat_qq() +
        stat_qq_line()
+
 ggplot(
     data.frame(
         y = model1_residuals
@@ -56,8 +56,6 @@ ggplot(
 #Resampling: Cross-Validated (10 fold, repeated 100 times) 
 #Summary of sample sizes: 4768, 4767, 4768, 4769, 4768, 4769, ... 
 #Resampling results:
-#names4 <- rownames(df4)
-.2164945
 #
 #Tuning parameter 'intercept' was held constant at a value of TRUE
 
@@ -69,8 +67,8 @@ summary(model2)
 coefs2 <- coefs(model2)
 coefs2_giregion <- coefs2[grep("gi*", row.names(ok)), ]
 coefs2_year <- coefs2[grep("data*", row.names(ok)), ]
-model1_residuals <- residuals(model2)
-model1_fitted <- fitted(model2)
+model2_residuals <- residuals(model2)
+model2_fitted <- fitted(model2)
 ggplot(
     data.frame(
         y = model2_residuals)
@@ -84,7 +82,7 @@ ggplot(
         , x = model2_fitted)
     , aes(x = x, y = y)
     ) +
-    geom_point()+
+    geom_point() +
     geom_hline(yintercept = 0)
 #Linear Regression 
 #
@@ -105,7 +103,7 @@ ggplot(
 # We add Value to the model :P
 anova(model3)
 summary(model3)
-coefs3 <- coefs(model3)names4 <- rownames(df4)
+coefs3 <- coefs(model3)
 
 coefs3_giregion <- coefs3[grep("gi*", row.names(ok)), ]
 coefs3_year <- coefs3[grep("data*", row.names(ok)), ]
@@ -121,27 +119,13 @@ ggplot(
 ggplot(
     data.frame(
         y = model3_residuals
-        , x = model3_fitted)df3 <- data.frame(summary(model3)$coefficients)
-names3 <- rownames(df3) 
[…]

### 2023-07-20 article.tex
Tabular fixes

@@
-%% 
 %% Copyright 2007-2020 Elsevier Ltd
@@
       %% keywords here, in the form: keyword \sep keyword
-      keyword one \sep keyword two
+      Keyword one \sep{} keyword two
       %% PACS codes here, in the form: \PACS code \sep code
-      \PACS 0000 \sep 1111
+      \PACS{} 0000 \sep{} 1111
       %% MSC codes here, in the form: \MSC code \sep code
       %% or \MSC[2008] code \sep code (2000 is the default)
-      \MSC 0000 \sep 1111
+      \MSC{} 0000 \sep{} 1111
       \end{keyword}
@@
 \begin{table}[]
-  %\centering
-    %\begin{center}
       \caption{Summary of models; their predictors, covariates and variable interactions.}
@@
           \toprule
-           & Response & Predictors & Covariates & Interactions \\ \midrule
-          Model 1 & Yield & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope 1 Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & N/A \\
+          & \textbf{Response} & \textbf{Predictors} & \textbf{Covariates} & \textbf{Interactions} \\ \midrule
+          \textbf{Model 1} & Yield & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope 1 Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & N/A \\
           \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-          Model 2 & ${ \textrm{Yield}}\over{ \textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope 1 Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * Scope 1 Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular} \\
+          \textbf{Model 2} & ${ \textrm{Yield}}\over{ \textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope 1 Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * Scope 1 Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular} \\
           \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-          Model 3 & {$\textrm{Yield} {\times} \textrm{Average Sale Price}$} & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope 1 Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & N/A \\
+          \textbf{Model 3} & {$\textrm{Yield} {\times} \textrm{Average Sale Price}$} & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope 1 Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & N/A \\
           \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-          Model 4 & ${\textrm{Yield} \times \textrm{Average Sale Price}}\over{\textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope 1 Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * Scope 1 Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular}
+          \textbf{Model 4} & ${\textrm{Yield} \times \textrm{Average Sale Price}}\over{\textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope 1 Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * Scope 1 Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular}
           \end{tabular}}
-    %\end{adjustbox}
-    %\end{center}
 \end{table}
@@
 % missing tables do we need these in tables?
-($< 2.200E-16$, see Tables ?? and ??)
+($< 2.200E-16$, see Tables  and )
 % This reference is messed up and next some TLC
@@
 \begin{table}[]
-  \caption{Summary of models; their performance, F-statistics and Residual error.}
-      \label{tab:modelsummary}
-      \resizebox{\textwidth}{!}{\begin{tabular}{@{}cccccccc@{}}
+  \caption{Summary of models; their performance, F-statistics and Residual error.}\label{tab:modelsummary}
+      \resizebox{\textwidth}{!}{
+        \begin{tabular}{@{}cccccccc@{}}
           \toprule
-          & \textbf{$R^2$}       & \textbf{$Adjusted R^2$} & \textbf{F-statistic} & \textbf{P-Value}     & \textbf{\begin{tabular}[c]{@{}c@{}}Residual\\ Standard Error\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Sum\\ of Squares\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Mean\\ of Squares\end{tabular}} \\ \midrule
+           & $\mathbf{R^2}$ & $\mathbf{R^2}$ & \textbf{F-Statistic} & \textbf{P-Value} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual\\ Standard Error\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Sum\\ of Squares\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Mean\\ of Squares\end{tabular}} \\ \midrule
           \textbf{\begin{tabular}[c]{@{}c@{}}Model 1\\ Yield\end{tabular}} & 9.072E-01 & 9.061E-01 & 7.753E+02 & 2.200e-16 & 3.065E-01 & 4.913E+02 & 1.000E-01 \\
-        \multicolumn{1}{l}{}                                                  & \multicolumn{1}{l}{} & \multicolumn{1}{l}{}    & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{}                                                       & \multicolumn{1}{l}{}                                                       & \multicolumn{1}{l}{}                                                        \\
           \textbf{\begin{tabular}[c]{@{}c@{}}Model 2\\ Yield/Area\end{tabular}} & 7.951E-01 & 7.770E-01 & 4.403E+01 & 2.200e-16 & 4.722E-01 & 1.085E+03 & 2.200E-01 \\
-        \multicolumn{1}{l}{}                                                  & \multicolumn{1}{l}{} & \multicolumn{1}{l}{}    & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{}                                                       & \multicolumn{1}{l}{}                                                       & \multicolumn{1}{l}{}                                                        \\
           \textbf{\begin{tabular}[c]{@{}c@{}}Model 3\\ Value\end{tabular}} & 9.753E-01 & 9.748E-01 & 1.885E+03 & 2.200e-16 & 1.589E-01 & 7.111E+01 & 3.000E-02 \\
-        \multicolumn{1}{l}{}                                                  & \multicolumn{1}{l}{} & \multicolumn{1}{l}{}    & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{}                                                       & \multicolumn{1}{l}{}                                                       & \multicolumn{1}{l}{}                                                        \\
-        \textbf{\begin{tabular}[c]{@{}c@{}}Model 4\\ Value/Area\end{tabular}} & 9.669E-01            & 9.638E-01               & 3.095E+02            & 2.200e-16            & 1.904E-01                                                                  & 9.528E+01                                                                  & 4.000E-02                                                                   \\ \bottomrule
-        \end{tabular}}
+          \textbf{\begin{tabular}[c]{@{}c@{}}Model 4\\ Value / Area\end{tabular}} & 9.669E-01 & 9.638E-01 & 3.095E+02 & 2.200e-16 & 1.904E-01 & 9.528E+01 & 4.000E-02 \\ \bottomrule
+          \end{tabular}
+        }
       \end{table}
-Each model showed a strong relationship between the predictors and the response (see Table \ref{tab:modelsummary}).
-The strength of the models was measured in $R^2$, as this allowed an easy comparison between their performances and their validation.
+Each model showed a strong relationship between the predictors and the response (see Table~\ref{tab:modelsummary}).
+Model accuracy was measured in $R^2$, as this allowed an easy comparison between their performances and their validation.
 \subsubsection{F-tests}

### 2023-07-21 analysis.r
Fixed misnamed variable in violin plot.

@@ -173,7 +173,7 @@ ggplot(
 
 df1 <- data.frame(summary(model1)$coefficients)
 names1 <- rownames(df1) 
-df1[names3[grepl("^da", names1)], "class"] <- "Model 1: Year"
+df1[names1[grepl("^da", names1)], "class"] <- "Model 1: Year"
 df1[names1[grepl("^gi", names1)], "class"] <- "model 1: GI Region"
 df1 <- df1[complete.cases(df1),]
 

### 2023-07-21 analysis.r
added way tyo print violin plot

@@ -193,8 +193,10 @@ names4 <- rownames(df4)
 df4[names4[grepl("^da", names4)], "class"] <- "Model 4: Year * GI Region"
 df4 <- df4[complete.cases(df4),]
 
-ggplot(rbind(df1, df2, df3, df4), aes(class, Estimate, fill=class)) + geom_violin() + geom_boxplot(width=0.1, fill="white") + scale_color_brewer(palette="Dark2") + theme_minimal()
-
+violinplot <- ggplot(rbind(df1, df2, df3, df4), aes(class, Estimate, fill=class)) + geom_violin() + geom_boxplot(width=0.1, fill="white") + scale_color_brewer(palette="Dark2") + theme_minimal()
+pdf("violinplot.pdf")
+print(violinplot)
+dev.off()
 # Residual standard error: 0.1949 on 1791 degrees of freedom
 #   (2877 observations deleted due to missingness)
 # Multiple R-squared:  0.9662,	Adjusted R-squared:  0.962 

### 2023-07-21 article.tex
Added figure, commented out bad table.

@@
 \begin{table}[]
-      \caption{Summary of models; their predictors, covariates and variable interactions.}
-      \label{tab:tab1}
+      \caption{Summary of models; their predictors, covariates and variable interactions.}\label{tab:tab1}
         \resizebox{\textwidth}{!}{\begin{tabular}{@{}ccccc@{}}
@@
 % missing tables do we need these in tables?
-($< 2.200E-16$, see Tables  and )
+($< 2.200E-16$, see Tables  and) 
 % This reference is messed up and next some TLC
@@
           \textbf{\begin{tabular}[c]{@{}c@{}}Model 1\\ Yield\end{tabular}} & 9.072E-01 & 9.061E-01 & 7.753E+02 & 2.200e-16 & 3.065E-01 & 4.913E+02 & 1.000E-01 \\
+          \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
           \textbf{\begin{tabular}[c]{@{}c@{}}Model 2\\ Yield/Area\end{tabular}} & 7.951E-01 & 7.770E-01 & 4.403E+01 & 2.200e-16 & 4.722E-01 & 1.085E+03 & 2.200E-01 \\
+          \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
           \textbf{\begin{tabular}[c]{@{}c@{}}Model 3\\ Value\end{tabular}} & 9.753E-01 & 9.748E-01 & 1.885E+03 & 2.200e-16 & 1.589E-01 & 7.111E+01 & 3.000E-02 \\
+          \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
           \textbf{\begin{tabular}[c]{@{}c@{}}Model 4\\ Value / Area\end{tabular}} & 9.669E-01 & 9.638E-01 & 3.095E+02 & 2.200e-16 & 1.904E-01 & 9.528E+01 & 4.000E-02 \\ \bottomrule
@@
 \subsubsection{F-tests}
-% \ref{tab:ftests}
-% We need a table with all F-tests.
-%         | Yield   | ...
-% Model 1 | -2e^-16 | ...
 %
+% This table is absolutely trash.
+%
+% Kate madea good point that, although figures can be combined into a single figure. Tables cannot.
+%
+% cheat way to multi line comment
+\iffalse{}
+%
+\begin{table}[]
+  \caption{Ftests and wsutff}\label{tab:ftests}
+  \resizebox{\textwidth}{!}{
+    \begin{tabular}{@{}cllllllll@{}}
+      \toprule
+       & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Water\\ Used\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Scope 1\\ Emissions\end{tabular}}} & \textbf{Year} & \textbf{\begin{tabular}[c]{@{}l@{}}GI\\ Region\end{tabular}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\\ *\\ Scope 1\\ Emissions\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\\ *\\ Water\\ Used\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Year\\ *\\ GI\\ Region\end{tabular}}} \\ \midrule
+      \textbf{Model 1} &  &  &  &  &  &  &  &  \\
+      \begin{tabular}[c]{@{}c@{}}Total Sum\\ of Squares\end{tabular} & 3211 & 10.4 & 6.6 & 70.6 & 1507.1 &  &  &  \\
+      F-Value & 34187.908 & 110.255 & 70.56 & 83.533 & 297.163 &  &  &  \\
+      Pr(>F) & 2.20E-16 & 2.20E-16 & 2.20E-16 & 2.20E-16 & 2.20E-16 &  &  &  \\
+      \multicolumn{1}{l}{} &  &  &  &  &  &  &  &  \\
+      \textbf{Model 2} &  &  &  &  &  &  &  &  \\
+      \begin{tabular}[c]{@{}c@{}}Total Sum\\ of Squares\end{tabular} & 2856.23 & 473.32 & 33.06 &  &  & 53.47 & 0.28 & 975.61 \\
+      F-Value & 15363.1866 & 177.8109 & 2545.9357 &  &  & 287.6126 & 1.5097 & 12.3766 \\
+      Pr(>F) & <2e-16 & <2e-16 & <2e-16 &  &  & <2e-16 & 0.2192 & <2e-16 \\
+      \multicolumn{1}{l}{} &  &  &  &  &  &  &  &  \\
+      \textbf{Model 3} &  &  &  &  &  &  &  &  \\
+      \begin{tabular}[c]{@{}c@{}}Total Sum\\ of Squares\end{tabular} & 2142.46 & 0.32 & 0.04 & 13.24 & 649.83 &  &  &  \\
+      F-Value & 84905.8197 & 12.5874 & 1.4915 & 87.4771 & 515.0601 &  &  &  \\
+      Pr(>F) & <2.2E-016 & 0.0003947 & 0.2220844 & <2.2E-016 & <2.2E-016 &  &  &  \\
+      \multicolumn{1}{l}{} &  &  &  &  &  &  &  &  \\
+      \textbf{Model 4} &  &  &  &  &  &  &  &  \\
+      \begin{tabular}[c]{@{}c@{}}Total Sum\\ of Squares\end{tabular} & 1868.46 & 274.92 & 0.04 &  &  & 35.02 & 4.18 & 582.96 \\
+      F-Value & 44084.0278 & 6486.2867 & 0.8311 &  &  & 826.1925 & 98.5895 & 56.6019 \\
+      Pr(>F) & <2e-16 & <2e-16 & 0.3621 &  &  & <2e-16 & <2e-16 & <2e-16
+      \end{tabular}
+  }
+\end{table}
+%
+\fi
+%
+%
+%
+% model 1 - all sig
+% model 2 - area * water used interaction was not sig
+% model 3 - scope 1 was not sig
+% model 4 - scope 1 was not sig
 To determine if predictors significantly related to a Model's response variable, F-tests were conducted.
-Models 1 and 2 showed significant linear relationships present between all predictors and their response variables (see Table \ref{tab:ftests}).
+Aside from 3 variables, all F-tests across each model indicated a significant contribution was made at 95\% confidence (See Table
+%\ref{tab:ftests}).
+). The exceptions were scope 1 emissions in models 3 and 4, as well as the interaction between area harvested and water used in Model 3. 
+Models 1 and 2 showed significant linear relationships present between all predictors and their response variables (with P-values being less than 2.200E-16).
 Predictors in Models 3 and 4 reported the same significance for each variable except for scope 1 emissions.
 Scope 1 emissions was included in all models to directly compare the response variables as ratios of vineyard size to raw values.
-Even though not significant within models 3 and 4, when using the Pearson Correlation Coefficients scope one emissions was strongly correlated to every Model's response variable; this was especially so for Model 1 and 4 (Yield and average price per tonne as a ratio to area harvested, respectively).
+Even though not significant within models 3 and 4, when using the Pearson Correlation Coefficients scope 1 emissions was strongly correlated to every Model's response variable; this was especially so for Model 1 and 4 (Yield and average price per tonne as a ratio to area harvested, respectively).
 \subsubsection{t-tests}
 %We want two tables
+\includegraphics{violinplot.pdf}
 % \ref{tab:ttests}

### 2023-07-24 analysis.r
Adding table extraction for t-values

@@ -59,7 +59,6 @@ ggplot(
 #
 #Tuning parameter 'intercept' was held constant at a value of TRUE
 
-
 # MODEL 2
 # Lets look at ratios
 anova(model2)
@@ -171,6 +170,8 @@ ggplot(
     geom_point() +
     geom_hline(yintercept = 0)
 
+# We graph the coefficients!
+
 df1 <- data.frame(summary(model1)$coefficients)
 names1 <- rownames(df1) 
 df1[names1[grepl("^da", names1)], "class"] <- "Model 1: Year"
@@ -214,6 +215,40 @@ dev.off()
 #  RMSE     Rsquared   MAE      
 #  0.22349  0.9500051  0.1279485
 #
+
+# We look at the t-values:
+
+df1 <- data.frame(summary(model1)$coefficients)
+names1 <- rownames(df1) 
+df1[names1[grepl("^da", names1)], "class"] <- "Model 1: Year"
+df1[names1[grepl("^gi", names1)], "class"] <- "model 1: GI Region"
+df1 <- df1[complete.cases(df1),]
+
+df2 <- data.frame(summary(model2)$coefficients)
+names2 <- rownames(df2)
+df2[names2[grepl("^da", names2)], "class"] <- "Model 2: Year * GI Region"
+df2 <- df2[complete.cases(df2),]
+
+df3 <- data.frame(summary(model3)$coefficients)
+names3 <- rownames(df3) 
+df3[names3[grepl("^da", names3)], "class"] <- "Model 3: Year"
+df3[names3[grepl("^gi", names3)], "class"] <- "model 3: GI Region"
+df3 <- df3[complete.cases(df3),]
+
+df4 <- data.frame(summary(model4)$coefficients)
+names4 <- rownames(df4)
+df4[names4[grepl("^da", names4)], "class"] <- "Model 4: Year * GI Region"
+df4 <- df4[complete.cases(df4),]
+
+violinplot <- ggplot(rbind(df1, df2, df3, df4), aes(class, Estimate, fill=class)) + geom_violin() + geom_boxplot(width=0.1, fill="white") + scale_color_brewer(palette="Dark2") + theme_minimal()
+pdf("violinplot.pdf")
+print(violinplot)
+dev.off()
+
+
+
+
+
 #Tuning parameter 'intercept' was held constant at a value of TRUE
 
 # note that vtha~tha is not a great predictor of one another.

### 2023-07-24 article.tex
minor edits (end of day commit)

@@
 Before models were fit to the data, Pearson Correlation Coefficients were used to look at the existence of linear relationships between predictor variables.
-These relationships were summarised in correlation matrices to compare levels of interaction present between predictor variables.
+These relationships were summarised in correlation matrices to compare the level of interaction present between predictor variables.
 The relationships between the predictors and response variables were then modelled using General Linear Models.
@@
 %reference?
-General Linear Models also allow the exploration of interactions between predictors and easily comparable differences in the influence and magnitude of relationships.
+General Linear Models also allow the exploration of interactions between predictors and present easily comparable differences in the influence and magnitude of relationships.
 % reference?
@@
 P-values reflected the significance of a given correlation coefficient when considering its relation to sample size via its incorporation as an element of standard error.
-Strong relationships were found to be present, as all P-values were considered significant
+Strong relationships were found to be present as all P-values, except for the non-transformed values for water used (see Table ?), were considered significant
 % missing tables do we need these in tables?
-($< 2.200E-16$, see Tables  and) 
+($< 2.200E-16$, see Tables  and).
 % This reference is messed up and next some TLC
-, except for the non-transformed values for water used (see Table 4). 
 \subsection{General Linear Models}
@@
 To determine if predictors significantly related to a Model's response variable, F-tests were conducted.
-Aside from 3 variables, all F-tests across each model indicated a significant contribution was made at 95\% confidence (See Table
+Aside from 3 variables, all F-tests across each model indicated a significant contribution at 95\% confidence (See Table
 %\ref{tab:ftests}).
@@
 Even though not significant within models 3 and 4, when using the Pearson Correlation Coefficients scope 1 emissions was strongly correlated to every Model's response variable; this was especially so for Model 1 and 4 (Yield and average price per tonne as a ratio to area harvested, respectively).
-\subsubsection{t-tests}
+\subsubsection{T-tests}
 %We want two tables
-\includegraphics{violinplot.pdf}
 % \ref{tab:ttests}
@@
 T-tests were used to determine if predictors significantly contributed to their models when accounting for other variables; this allowed a more granular examination of interactions and factors within categorical variables, showing which specific years and areas contributed significantly and which did not (see appendix).
+% The range of how significant these variable were is important.
+%
+% How many were important vs how many were not?
+%
+% was it the same across each of the models? For example was one year/region always bad for every model?
+%
 With regard to continuous variables, Model 1 and 2 showed water use, scope 1 emissions and area harvested were significant when accounting for other variables.
@@
 Model 4 showed scope 1 emissions and water use to only be significant when considered as an interaction with area harvested but not when considered on their own.
-\subsubsection{Model Comparison}
+\subsubsection{Model Coefficients}
+% A map and a table just like the t-test section.
+\begin{figure}
+\caption{Violine plot of GI Region and Year coefficients for each model.}\label{fig:violin}
+\includegraphics{violinplot.pdf}
+\end{figure}
+
+The correlation between average sales price and yield was a negative trend (see table 2); the contributing factors to yield and average sales price was ???.
+Correlation values showed that water and emissions increased with yield but decreased with average sale price (see Table 4).
+In alternative attempts at models it was found that without the incorporation of GI Region or year the predictions greatly under performed.
+The possible reason behind this effect was that different strategies are likely employed between different regions, where some regions target the mass production of cheaper grapes over quality.
+% ALSO THAT CATASTROPHES HAPPEN EVEN ON SMALL SCALES THis needs to be tied up better.
+
+\subsubsection{Model Comparisons: Productivity Verse Value}
 % wHY ARE WE COMPARING MODELS?
@@
 Models were compared using $R^2$ values.
-The comparison of models performance showed that the average price per tonne of grapes describes a great deal of the relationship between predictor and response when comparing Model 2 to Model 4 (see Table \ref*{tab:modelperformance}).
+The comparison of models performance showed that the average price per tonne of grapes describes a great deal of the relationship between predictor and response when comparing Model 2 to Model 4
+% Graph of productivity verse value.
+
+(see Table \ref*{tab:modelperformance}). 
 % This is important but you do not illustrate why
@@
 
- \subsubsection{Model Coefficients}
-% A map and a table just like the t-test section.
-% ^ This does not work
-
-The correlation between average sales price and yield was a negative trend (see table 2); the contributing factors to yield and average sales price was ???.
-Correlation values showed that water and emissions increased with yield but decreased with average sale price (see Table 4).
-In alternative attempts at models it was found that without the incorporation of GI Region or year the predictions greatly under performed.
-The possible reason behind this effect was that different strategies are likely employed between different regions, where some regions target the mass production of cheaper grapes over quality.
-% ALSO THAT CATASTROPHES HAPPEN EVEN ON SMALL SCALES THis needs to be tied up better.
-
-\subsection{Productivity verse Value}
-% Graph of productivity verse value.
 different strategies are likely employed between different regions, where some regions target the mass production of cheaper grapes over quality.
@@
 The interaction between year and GI Region likely accounted for localised events such as bushfires, which would be impactful, but only at a local level in both time and space.
+
 \subsection{Model Validation}

### 2023-07-25 article.tex
Fixed counting error. Added T-test summaries.

@@
 \newline
-The first subset of data was used for Model 1 and Model 2 (see Table 1). This subset contained 5298 samples spanning the period from 2012 to 2022, covering 57 GI Regions and 1432 separate vineyards.
+The first subset of data was used for Model 1 and Model 2 (see Table 1). This subset contained 5298 samples spanning the period from 2012 to 2022, covering 55 GI Regions and 1261 separate vineyards.
 \newline
@@
 \subsubsection{F-tests}
-%
-% This table is absolutely trash.
-%
-% Kate madea good point that, although figures can be combined into a single figure. Tables cannot.
-%
-% cheat way to multi line comment
-\iffalse{}
-%
-\begin{table}[]
-  \caption{Ftests and wsutff}\label{tab:ftests}
-  \resizebox{\textwidth}{!}{
-    \begin{tabular}{@{}cllllllll@{}}
-      \toprule
-       & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Water\\ Used\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Scope 1\\ Emissions\end{tabular}}} & \textbf{Year} & \textbf{\begin{tabular}[c]{@{}l@{}}GI\\ Region\end{tabular}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\\ *\\ Scope 1\\ Emissions\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\\ *\\ Water\\ Used\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Year\\ *\\ GI\\ Region\end{tabular}}} \\ \midrule
-      \textbf{Model 1} &  &  &  &  &  &  &  &  \\
-      \begin{tabular}[c]{@{}c@{}}Total Sum\\ of Squares\end{tabular} & 3211 & 10.4 & 6.6 & 70.6 & 1507.1 &  &  &  \\
-      F-Value & 34187.908 & 110.255 & 70.56 & 83.533 & 297.163 &  &  &  \\
-      Pr(>F) & 2.20E-16 & 2.20E-16 & 2.20E-16 & 2.20E-16 & 2.20E-16 &  &  &  \\
-      \multicolumn{1}{l}{} &  &  &  &  &  &  &  &  \\
-      \textbf{Model 2} &  &  &  &  &  &  &  &  \\
-      \begin{tabular}[c]{@{}c@{}}Total Sum\\ of Squares\end{tabular} & 2856.23 & 473.32 & 33.06 &  &  & 53.47 & 0.28 & 975.61 \\
-      F-Value & 15363.1866 & 177.8109 & 2545.9357 &  &  & 287.6126 & 1.5097 & 12.3766 \\
-      Pr(>F) & <2e-16 & <2e-16 & <2e-16 &  &  & <2e-16 & 0.2192 & <2e-16 \\
-      \multicolumn{1}{l}{} &  &  &  &  &  &  &  &  \\
-      \textbf{Model 3} &  &  &  &  &  &  &  &  \\
-      \begin{tabular}[c]{@{}c@{}}Total Sum\\ of Squares\end{tabular} & 2142.46 & 0.32 & 0.04 & 13.24 & 649.83 &  &  &  \\
-      F-Value & 84905.8197 & 12.5874 & 1.4915 & 87.4771 & 515.0601 &  &  &  \\
-      Pr(>F) & <2.2E-016 & 0.0003947 & 0.2220844 & <2.2E-016 & <2.2E-016 &  &  &  \\
-      \multicolumn{1}{l}{} &  &  &  &  &  &  &  &  \\
-      \textbf{Model 4} &  &  &  &  &  &  &  &  \\
-      \begin{tabular}[c]{@{}c@{}}Total Sum\\ of Squares\end{tabular} & 1868.46 & 274.92 & 0.04 &  &  & 35.02 & 4.18 & 582.96 \\
-      F-Value & 44084.0278 & 6486.2867 & 0.8311 &  &  & 826.1925 & 98.5895 & 56.6019 \\
-      Pr(>F) & <2e-16 & <2e-16 & 0.3621 &  &  & <2e-16 & <2e-16 & <2e-16
-      \end{tabular}
-  }
-\end{table}
-%
-\fi
-%
-%
-%
-% model 1 - all sig
-% model 2 - area * water used interaction was not sig
-% model 3 - scope 1 was not sig
-% model 4 - scope 1 was not sig
 To determine if predictors significantly related to a Model's response variable, F-tests were conducted.
-Aside from 3 variables, all F-tests across each model indicated a significant contribution at 95\% confidence (See Table
-%\ref{tab:ftests}).
-). The exceptions were scope 1 emissions in models 3 and 4, as well as the interaction between area harvested and water used in Model 3. 
-Models 1 and 2 showed significant linear relationships present between all predictors and their response variables (with P-values being less than 2.200E-16).
-Predictors in Models 3 and 4 reported the same significance for each variable except for scope 1 emissions.
+Aside from 3 variables, all F-tests across each model indicated a significant contribution at 95\% confidence.
+The three exceptions were: scope 1 emissions in Model 3 (P=2.221E-01) and Model 4 (P=3.621E-01), and Model 2's interaction between area harvested and water used (P=2.192E-01).
+\newline
 Scope 1 emissions was included in all models to directly compare the response variables as ratios of vineyard size to raw values.
@@
 % A map of significant areas contributing to each model.
-T-tests were used to determine if predictors significantly contributed to their models when accounting for other variables; this allowed a more granular examination of interactions and factors within categorical variables, showing which specific years and areas contributed significantly and which did not (see appendix).
+T-tests were used to determine if predictors significantly contributed to their models when accounting for other variables; this allowed a more granular examination of interactions and factors within categorical variables, showing which specific years and areas contributed significantly and which did not (the appendix contains a comprehensive list of these values).
+
+For Models 1 and 3, depicting yield and value as totals, year played a pivitol role; with only one year in each model not being significant (2021/2022 and 2016/2017 respectively).
+Both Model 1 and 3 showed a majority of regions were significant with 32 of 54 regions being significant in Model 1 and 42 of 50 regions being significant in Model 3.
+
+For Model 1, All years were significant except for the 2021/2022 financial year (where, P=4.325E-01).
+32/54 giregions were significant (59.26\%).
+Adelaide planes was the reference region with 2012/2013 being the reference year.
+%
+For Model 3, all years were significant except for 2016/2017 (P=7.532E-01
+). 42 out of 50 regions were significant.
 % The range of how significant these variable were is important.
@@
 %
-With regard to continuous variables, Model 1 and 2 showed water use, scope 1 emissions and area harvested were significant when accounting for other variables.
+\newline
+With regard to continuous variables, Model 1 and 2 showed water use, scope 1 emissions and area harvested were significant at 95\% confidence when accounting for other variables.
 T-tests for Model 3 showed all continuous variables except scope 1 emissions were significant.

### 2023-07-25 article.tex
T-tests draft complete.

@@
 \subsubsection{T-tests}
-%We want two tables
-% \ref{tab:ttests}
-% \ref{fig:sigregions}
-% a table for t-tests as per the f-tests table (only continuous variables)
-% A map of significant areas contributing to each model.
 T-tests were used to determine if predictors significantly contributed to their models when accounting for other variables; this allowed a more granular examination of interactions and factors within categorical variables, showing which specific years and areas contributed significantly and which did not (the appendix contains a comprehensive list of these values).
-
-For Models 1 and 3, depicting yield and value as totals, year played a pivitol role; with only one year in each model not being significant (2021/2022 and 2016/2017 respectively).
-Both Model 1 and 3 showed a majority of regions were significant with 32 of 54 regions being significant in Model 1 and 42 of 50 regions being significant in Model 3.
-
-For Model 1, All years were significant except for the 2021/2022 financial year (where, P=4.325E-01).
-32/54 giregions were significant (59.26\%).
-Adelaide planes was the reference region with 2012/2013 being the reference year.
-%
-For Model 3, all years were significant except for 2016/2017 (P=7.532E-01
-). 42 out of 50 regions were significant.
-% The range of how significant these variable were is important.
-%
-% How many were important vs how many were not?
-%
-% was it the same across each of the models? For example was one year/region always bad for every model?
-%
+\newline
+For Models 1 (yield) and 3 (value) year played a pivotal role, with only one year in each model not being significant (2021/2022 and 2016/2017 respectively).
+Both Model 1 and 3 showed a majority of regions were significant with 32 of 54 regions being significant in Model 1, and 42 of 50 regions being significant in Model 3 with 95\% confidence.
+\newline
+The number of combinations of year and region meant that Models 2 and 4 had many tests (424 and 243 respectively).
+Model 2 found 62.56\% of these combinations were indicative of a significant contribution to the model at 95\% significance.
+Model 4 was found to have 88.07\% of its year/region combinations indicating a significant contribution.
+A likely reason for some combinations not being significant was a lack of samples in that particular region/year being present; with region sample sizes ranging from 1 to 1006.
 \newline

### 2023-07-25 analysis.r
Changed scales of graphs

@@ -170,34 +170,6 @@ ggplot(
     geom_point() +
     geom_hline(yintercept = 0)
 
-# We graph the coefficients!
-
-df1 <- data.frame(summary(model1)$coefficients)
-names1 <- rownames(df1) 
-df1[names1[grepl("^da", names1)], "class"] <- "Model 1: Year"
-df1[names1[grepl("^gi", names1)], "class"] <- "model 1: GI Region"
-df1 <- df1[complete.cases(df1),]
-
-df2 <- data.frame(summary(model2)$coefficients)
-names2 <- rownames(df2)
-df2[names2[grepl("^da", names2)], "class"] <- "Model 2: Year * GI Region"
-df2 <- df2[complete.cases(df2),]
-
-df3 <- data.frame(summary(model3)$coefficients)
-names3 <- rownames(df3) 
-df3[names3[grepl("^da", names3)], "class"] <- "Model 3: Year"
-df3[names3[grepl("^gi", names3)], "class"] <- "model 3: GI Region"
-df3 <- df3[complete.cases(df3),]
-
-df4 <- data.frame(summary(model4)$coefficients)
-names4 <- rownames(df4)
-df4[names4[grepl("^da", names4)], "class"] <- "Model 4: Year * GI Region"
-df4 <- df4[complete.cases(df4),]
-
-violinplot <- ggplot(rbind(df1, df2, df3, df4), aes(class, Estimate, fill=class)) + geom_violin() + geom_boxplot(width=0.1, fill="white") + scale_color_brewer(palette="Dark2") + theme_minimal()
-pdf("violinplot.pdf")
-print(violinplot)
-dev.off()
 # Residual standard error: 0.1949 on 1791 degrees of freedom
 #   (2877 observations deleted due to missingness)
 # Multiple R-squared:  0.9662,	Adjusted R-squared:  0.962 
@@ -216,7 +188,9 @@ dev.off()
 #  0.22349  0.9500051  0.1279485
 #
 
-# We look at the t-values:
+############################ #
+# We graph the coefficients! #
+##############################
 
 df1 <- data.frame(summary(model1)$coefficients)
 names1 <- rownames(df1) 
@@ -245,9 +219,40 @@ pdf("violinplot.pdf")
 print(violinplot)
 dev.off()
 
+############################
+# We look at the t-values: #
+############################
 
+df1 <- data.frame(summary(model1)$coefficients)
+names1 <- rownames(df1) 
[…]

### 2023-07-25 article.tex
looking at final results section.

@@
 \subsubsection{Model Coefficients}
-% A map and a table just like the t-test section.
 \begin{figure}
@@
 \end{figure}
-
+%
+% What are you trying to say here?
+%
+% We want to compare how much influence each variable has on the model!
+% This is prior to postulating about the why these variables matter, or have so much influence.
+% 
+% A table will help to compare the continuous variables, the others are done in the violin plot.
+%
+% What is this below paragraph?
 The correlation between average sales price and yield was a negative trend (see table 2); the contributing factors to yield and average sales price was ???.
 Correlation values showed that water and emissions increased with yield but decreased with average sale price (see Table 4).
+%
 In alternative attempts at models it was found that without the incorporation of GI Region or year the predictions greatly under performed.
@@
 % ALSO THAT CATASTROPHES HAPPEN EVEN ON SMALL SCALES THis needs to be tied up better.
-
+%
 \subsubsection{Model Comparisons: Productivity Verse Value}
@@
 Some regions appeared to produce many low quality grapes at scale compared to attempting to produce fewer higher quality grapes.
-This behaviour can be observed when reviewing Wine Australia’s annual reports, where it is apparent that warm inland regions such as the Riverland are known to only produce large amounts of lower graded grapes \cite{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
+This behaviour can be observed when reviewing Wine Australia's annual reports, where it is apparent that warm inland regions such as the Riverland are known to only produce large amounts of lower graded grapes \cite{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
 Comparatively, regions such as Tasmania only produce A grade grapes but in much smaller quantities than the Riverland.

### 2023-07-25 analysis.r
pdf outputs of figures

@@ -265,7 +265,7 @@ anova(lm(df$ha_value ~ df$ha_tonnes_grapes_harvested))
 # not the quality.
 
 # here are some plots
-ggplot(df
+yield.vs.value <- ggplot(df
     , aes(
         x = tonnes_grapes_harvested
         , y = tonnes_grapes_harvested * average_per_tonne
@@ -277,7 +277,11 @@ ggplot(df
     xlab("Yield") +
     ylab("Yield * Average Price") +
     theme_light()
-ggplot(
+pdf("yvv.pdf")
+print(yield.vs.value)
+dev.off()
+
+yield.vs.value.per.ha <- ggplot(
     df
     , aes(
         x = ha_tonnes_grapes_harvested
@@ -290,6 +294,9 @@ ggplot(
         xlab("Yield / Area Harvested") +
         ylab("Yield * Average Price / Area Harvested") +
         theme_light()
+pdf("yvvph.pdf")
+print(yield.vs.value.per.ha)
+dev.off()
 
 #######################
 # Lets make some maps #

### 2023-07-25 analysis.r
added colors and changed graphs.

@@ -26,9 +26,6 @@ model4 <- lm(
 # MODEL 1
 anova(model1)
 summary(model1)
-coefs1 <- coefs(model1)
-coefs1_giregion <- coefs1[grep("gi*", row.names(ok)), ]
-coefs1_year <- coefs1[grep("data*", row.names(ok)), ]
 model1_residuals <- residuals(model1)
 model1_fitted <- fitted(model1)
 ggplot(
@@ -53,8 +50,8 @@ ggplot(
 #   5 predictor
 #
 #No pre-processing
-#Resampling: Cross-Validated (10 fold, repeated 100 times) 
-#Summary of sample sizes: 4768, 4767, 4768, 4769, 4768, 4769, ... 
+#Resampling: Cross-Validated (10 fold, repeated 100 times)
+#Summary of sample sizes: 4768, 4767, 4768, 4769, 4768, 4769,
 #Resampling results:
 #
 #Tuning parameter 'intercept' was held constant at a value of TRUE
@@ -63,9 +60,6 @@ ggplot(
 # Lets look at ratios
 anova(model2)
 summary(model2)
-coefs2 <- coefs(model2)
-coefs2_giregion <- coefs2[grep("gi*", row.names(ok)), ]
-coefs2_year <- coefs2[grep("data*", row.names(ok)), ]
 model2_residuals <- residuals(model2)
 model2_fitted <- fitted(model2)
 ggplot(
@@ -83,17 +77,17 @@ ggplot(
     ) +
     geom_point() +
     geom_hline(yintercept = 0)
-#Linear Regression 
+#Linear Regression
 #
 #6068 samples
 #   5 predictor
 #
 #No pre-processing
-#Resampling: Cross-Validated (10 fold, repeated 100 times) 
-#Summary of sample sizes: 4768, 4768, 4768, 4769, 4769, 4768, ... 
+#Resampling: Cross-Validated (10 fold, repeated 100 times)
+#Summary of sample sizes: 4768, 4768, 4768, 4769, 4769, 4768,
 #Resampling results:
 #
-#  RMSE       Rsquared   MAE      
+#  RMSE       Rsquared   MAE
 #  0.5104218  0.7408576  0.3492901
 #
 #Tuning parameter 'intercept' was held constant at a value of TRUE
@@ -102,10 +96,6 @@ ggplot(
 # We add Value to the model :P
 anova(model3)
 summary(model3)
-coefs3 <- coefs(model3)
-
[…]

### 2023-07-26 article.tex
Added a comment about missing section in methods.
first paragraph of coefficient subsection

@@
 Values were compared in this manner to observe how economies of scale affect the use of resources.
+%
+%
+%
+\subsection{Significant Tests}
+%This briefly needs to be described. You use the results of significance tests in the results section and talk about variables significance but do not describe it in the methods.
+%
+% F-tests
+%
+% T-tests
+%
+%
 \subsection{Data}
@@
 \section{Results}
-\subsection{Exploratory Analysis}
+\subsection{Exploratory Analysis} \label{sec:exp_anal}
 Linear relationships between variables were explored using Pearson Correlation Coefficients.
@@
 \subsubsection{Model Coefficients}
+%
+% What are you trying to say here?
+% Coefficients are important as they describe the direct relationship between a variable and its response.
+They indicate whether it is a positive, or negative relationship and the magnitude of this relationship.
+The coefficients of each model describe the relationship of a predictor variable to its response when considering all other predictor variables.
+Due to the transformations of the data, coefficients are individually interpreted in the same manner as the prior regression values were (section\ref{sec:exp_anal}); although all coefficients need to be considered together.
+Unlike the regression values the coefficients range is not limited between -1 and 1, as the ratio of each variables contribution needs to be considered together.
+%The dimensionless nature of these variables each coefficient refers to its own variable's distribution's standard deviations.
+\newline
+We look at the categorical variables' and continuous variables' coefficients separately.
+This is primarily done as the categorical variables have many coefficients, one for each category, whilst continuous variables have only one.
+The coefficient for categorical variables are summarised in Figure\ref{fig:violin}.
+The continuous variables coefficients are summarised in Table\ref{tab:contvar}.
 \begin{figure}
@@
   \end{figure}
-%
-% What are you trying to say here?
 %
@@
 % What is this below paragraph?
-The correlation between average sales price and yield was a negative trend (see table 2); the contributing factors to yield and average sales price was ???.
+%The correlation between average sales price and yield was a negative trend (see table 2); the contributing factors to yield and average sales price was ???.
 Correlation values showed that water and emissions increased with yield but decreased with average sale price (see Table 4).

### 2023-07-26 article.tex
Added coef table to results.

@@
 %
+\begin{table}[]
+  \caption{Summary of models; their performance, F-statistics and Residual error.}\label{tab:contvar}
+      \resizebox{\textwidth}{!}{
+        \begin{tabular}{@{}cllllll@{}}
+          \toprule
+           & \textbf{Intercept} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Water\\ Used\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Scope 1\\ Emissions\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\\ $\ast$\\ Scope 1\\ Emissions\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\\ $\ast$\\ Water\\ Used\end{tabular}}} \\ \midrule
+          \textbf{Model 1} & -3.318E-02 & 7.418E-01 & 8.660E-02 & 6.731E-02 &  &  \\
+          \textbf{Model 2} & -6.516E-01 & 5.774E-01 & 1.079E-02 & 8.498E-02 & -4.971E-02 & -5.346E-02 \\
+          \textbf{Model 3} & 1.808E-02 & 9.713E-01 & -2.310E-02 & -6.992E-03 &  &  \\
+          \textbf{Model 4} & 6.702E-01 & -7.354E-01 & -6.732E-03 & -5.645E-03 & 2.726E-02 & 7.515E-02 \\ \bottomrule
+          \end{tabular}}
+\end{table}
 % We want to compare how much influence each variable has on the model!

### 2023-07-28 article.tex
back up for thursdays work

@@
 \section{Methods}
-We created four linear models to explore relationships between resource-use and vineyard outputs (see Table 1).
+We created four linear models to explore relationships between resource-use and vineyard outputs (see Table\ref{tab:tab1}).
 The data was sourced from Sustainable Winegrowing Australia and Wine Australia.
@@
 The second subset contained vineyards which either recorded a value for average price of sale per tonne through Sustainable Winegrowing Australia, or were within a region with an average price of sale recorded by Wine Australia; this subset was used for Models 2 and 4.
-These subsets meant that the data would be limited to samples which had recorded values for the response variables (see Table 1), where every sample had a recorded value for yield but not average price of sale per tonne.
+These subsets meant that the data would be limited to samples which had recorded values for the response variables (see Table\ref{tab:tab1}), where every sample had a recorded value for yield but not average price of sale per tonne.
 \newline
-The first subset of data was used for Model 1 and Model 2 (see Table 1). This subset contained 5298 samples spanning the period from 2012 to 2022, covering 55 GI Regions and 1261 separate vineyards.
+The first subset of data was used for Model 1 and Model 2 (see Table\ref{tab:tab1}).
+This subset contained 5298 samples spanning the period from 2012 to 2022, covering 55 GI Regions and 1261 separate vineyards.
 \newline
 The second subset of data, was limited to vineyards that recorded a value for their average sale price of grapes per tonne.
-This subset was used for Model 3 and Model 4 (see Table 1); and contained 2878 samples spanning the period from 2015 to 2022, covering 51 GI Regions and 944 separate vineyards.
+This subset was used for Model 3 and Model 4 (see Table\ref{tab:tab1}); and contained 2878 samples spanning the period from 2015 to 2022, covering 51 GI Regions and 944 separate vineyards.
 1842 of the values for average price of sale per tonne were extracted from Wine Australia surveys with the remaining 1036 being from Sustainable Winegrowing Australia's dataset.
@@
 Values for these coefficients reflect the linear relation between two variables, on a scale between -1 and 1; the magnitude and sign of a coefficient indicates the strength of the relation, and whether the relation is positive or negative respectively.
-This was undertaken for data on the original scale (see Table 2) and for data as a logarithmic transform (see Table 3).
-The logarithmic transformed data showed the strongest correlations, likely due to a skew caused by a greater number of smaller vineyards within the dataset (see Table 5).
+This was undertaken for data on the original scale (see Table ??) and for data as a logarithmic transform (see Table ??).
+The logarithmic transformed data showed the strongest correlations, likely due to a skew caused by a greater number of smaller vineyards within the dataset (see Table ??).
 Transforming data prior to calculating the coefficients changes several things: The logarithmic transform of the data alters the interpretation of the coefficients to percentage change - a coefficient will be indicative of the change in percentage of one variable compared to the other; scaling by standard deviation also changes this interpretation to be a percentage of that variables standard deviation.
@@
 P-values reflected the significance of a given correlation coefficient when considering its relation to sample size via its incorporation as an element of standard error.
-Strong relationships were found to be present as all P-values, except for the non-transformed values for water used (see Table ?), were considered significant
-% missing tables do we need these in tables?
-($< 2.200E-16$, see Tables  and).
-% This reference is messed up and next some TLC
+Strong relationships were found to be present as all P-values, except for the non-transformed values for water used, were considered significant (P \< 2.200E-16).
 \subsection{General Linear Models}
@@
 The coefficients of each model describe the relationship of a predictor variable to its response when considering all other predictor variables.
-Due to the transformations of the data, coefficients are individually interpreted in the same manner as the prior regression values were (section\ref{sec:exp_anal}); although all coefficients need to be considered together.
-Unlike the regression values the coefficients range is not limited between -1 and 1, as the ratio of each variables contribution needs to be considered together.
+Due to the transformations of the data, coefficients are individually interpreted in the same manner as the prior regression values were (see section\ref{sec:exp_anal}); although all coefficients need to be considered together.
+Unlike the regression values the coefficients range is not limited between -1 and 1, as each variable's contribution needs to be considered together.
 %The dimensionless nature of these variables each coefficient refers to its own variable's distribution's standard deviations.
 \newline
-We look at the categorical variables' and continuous variables' coefficients separately.
+We look at the coefficients of categorical and continuous variables  separately.
 This is primarily done as the categorical variables have many coefficients, one for each category, whilst continuous variables have only one.
-The coefficient for categorical variables are summarised in Figure\ref{fig:violin}.
-The continuous variables coefficients are summarised in Table\ref{tab:contvar}.
+The coefficient for categorical variables is summarised in Figure\ref{fig:violin}; illustrating the difference in the range and the affect region and year could have on each of the models.
+Comparatively, the continuous variables coefficients are summarised in Table\ref{tab:contvar}.
+In terms of magnitude, GI region has the highest possible absolute value for each model.
+An important consideration is that region and year are binary variables, such that they are only equal to zero or the coefficient (as they will present as a value of 1 multiplied by the coefficient); this means that, although region may have a strong relationship, it can be overshadowed by an extreme value of one of the continuous variables.
+The most notable difference between the continuous variables coefficients is the change from positive to negative values.
+This change occurs between the Models for Yield (Model 1 and 2) and the Models for value (Models 3 and 4); where all but the coefficient for area harvested has the opposite sign between them.
+These models also differ in an order of magnitude when looking at resource use, with the coefficients for yield being smaller than those for value.
+%
+% The resource input is different by an order of magnitude between value and yield
+%
+% Area has the strongest nfluence
+%
+% The models as a ratio of area show opposite trends - this could be indicative that smaller vineyards tend twoards more valuable crops and larger to crops of greater yield. 
+%
 \begin{figure}
-  \caption{Violine plot of GI Region and Year coefficients for each model.}\label{fig:violin}
+  \caption{Violin plots of GI Region and Year coefficients for each model.}\label{fig:violin}
   \includegraphics{violinplot.pdf}
@@
 \begin{table}[]
-  \caption{Summary of models; their performance, F-statistics and Residual error.}\label{tab:contvar}
+  \caption{Summary of each Models coefficients for continuous variables}\label{tab:contvar}
       \resizebox{\textwidth}{!}{
@@
 \end{table}
-% We want to compare how much influence each variable has on the model!
-% This is prior to postulating about the why these variables matter, or have so much influence.
-% 
-% A table will help to compare the continuous variables, the others are done in the violin plot.
-%
-% What is this below paragraph?
-%The correlation between average sales price and yield was a negative trend (see table 2); the contributing factors to yield and average sales price was ???.
-Correlation values showed that water and emissions increased with yield but decreased with average sale price (see Table 4).
-%
-In alternative attempts at models it was found that without the incorporation of GI Region or year the predictions greatly under performed.
-The possible reason behind this effect was that different strategies are likely employed between different regions, where some regions target the mass production of cheaper grapes over quality.
-% ALSO THAT CATASTROPHES HAPPEN EVEN ON SMALL SCALES THis needs to be tied up better.
-%
 \subsubsection{Model Comparisons: Productivity Verse Value}
@@
 % We compare models to see which predictors contributed relatively to each of the models.
+
 %
-% There were two subsets - when comparing these models what is the difference between them?
-%
-Models were compared using $R^2$ values.
+%Models were compared using $R^2$ values.
 The comparison of models performance showed that the average price per tonne of grapes describes a great deal of the relationship between predictor and response when comparing Model 2 to Model 4
@@
 \section{Discussion}
+% We want to compare how much influence each variable has on the model!
+% This is prior to postulating about the why these variables matter, or have so much influence.
+% 
+% A table will help to compare the continuous variables, the others are done in the violin plot.
+%
+% What is this below paragraph?
+%The correlation between average sales price and yield was a negative trend (see table 2); the contributing factors to yield and average sales price was ???.
+Correlation values showed that water and emissions increased with yield but decreased with average sale price (see Table 4).
+%
+In alternative attempts at models it was found that without the incorporation of GI Region or year the predictions greatly under performed.
+The possible reason behind this effect was that different strategies are likely employed between different regions, where some regions target the mass production of cheaper grapes over quality.
+% ALSO THAT CATASTROPHES HAPPEN EVEN ON SMALL SCALES THis needs to be tied up better.
+%
 \subsection{Limitations}

### 2023-07-28 analysis.r
draft of colours for figures

@@ -255,6 +255,10 @@ my_colors <- c(
     "#1E88E5"
     , "#1E88E5"
     , "#1E88E5"
+    # Hot
+    #, "#D81B60"
+    , "#D81B60"
+    , "#D81B60"
     # Mild
     #, "#004D40"
     , "#004d40"
@@ -265,9 +269,26 @@ my_colors <- c(
     , "#FFC107"
     , "#FFC107"
     , "#FFC107"
+)
+my_colors <- c(
+    # Cool
+    #"#1E88E5"
+    "#1E88E5"
+    , "#FFC107"
+    , "#D81B60"
     # Hot
     #, "#D81B60"
+    , "#1E88E5"
     , "#D81B60"
+    # Mild
+    #, "#004D40"
+    , "#1E88E5"
+    , "#FFC107"
+    , "#D81B60"
+    # Warm
+    #, "#FFC107"
+    , "#1E88E5"
+    , "#FFC107"
     , "#D81B60"
 )
 scale_color_manual(values = my_colors)
@@ -285,7 +306,8 @@ yield.vs.value <- ggplot(
     geom_point(size = .5) +
     xlab("Yield") +
     ylab("Yield * Average Price") +
-    theme_light()
+    theme_light() +
+    scale_color_manual(values = my_colors)
 pdf("yvv.pdf")
 print(yield.vs.value)
 dev.off()

### 2023-07-28 article.tex
Minor edits

@@
 \begin{frontmatter}
-
 %% Title, authors and addresses
-
 %% use the tnoteref command within \title for footnotes;
@@
 %% \fntext[label3]{}
-
 \title{An exploratory analysis of the influence of resource use on the yield verse quality trade-off in Australian vineyards}
-
 %% use optional labels to link authors explicitly to addresses:
@@
 %%             country={}}
-
 \affiliation[label1]{organization={QUT},
@@
 \date{20/06/2023}
-
 \begin{abstract}
 %% Text of abstract
-      Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
 \end{abstract}
-
-      %%Graphical abstract
+%%Graphical abstract`
 \begin{graphicalabstract}
   \includegraphics{graphical_abstract.jpeg}
-      \end{graphicalabstract}
+\end{graphicalabstract}'
 
@@
 \end{keyword}
-
 %%Research highlights
@@
 \end{highlights}
-
 \end{frontmatter}
@@
 \section{Results}
-\subsection{Exploratory Analysis} \label{sec:exp_anal}
+\subsection{Exploratory Analysis}\label{sec:exp_anal}
 Linear relationships between variables were explored using Pearson Correlation Coefficients.
@@
 P-values reflected the significance of a given correlation coefficient when considering its relation to sample size via its incorporation as an element of standard error.
-Strong relationships were found to be present as all P-values, except for the non-transformed values for water used, were considered significant (P \< 2.200E-16).
+Strong relationships were found to be present as all P-values, except for the non-transformed values for water used, were considered significant (P $<$ 2.200E-16).
 \subsection{General Linear Models}
@@
 They indicate whether it is a positive, or negative relationship and the magnitude of this relationship.
-The coefficients of each model describe the relationship of a predictor variable to its response when considering all other predictor variables.
-Due to the transformations of the data, coefficients are individually interpreted in the same manner as the prior regression values were (see section\ref{sec:exp_anal}); although all coefficients need to be considered together.
-Unlike the regression values the coefficients range is not limited between -1 and 1, as each variable's contribution needs to be considered together.
+The coefficients of each model describe the relationship of a predictor variable to its response when considering all other variables.
+Due to the transformations of the data, coefficients are individually interpreted in the same manner as the prior regression values were (see Section \ref{sec:exp_anal}); unlike the regression values,  coefficient ranges are not limited between -1 and 1, as each variable needs to be considered together.
 %The dimensionless nature of these variables each coefficient refers to its own variable's distribution's standard deviations.
@@
 This is primarily done as the categorical variables have many coefficients, one for each category, whilst continuous variables have only one.
-The coefficient for categorical variables is summarised in Figure\ref{fig:violin}; illustrating the difference in the range and the affect region and year could have on each of the models.
-Comparatively, the continuous variables coefficients are summarised in Table\ref{tab:contvar}.
+The coefficient for categorical variables is summarised in Figure\ref{fig:violin}; illustrating the difference in the range as well as affect region and year could have on each of the models.
+Comparatively, the continuous variables coefficients are summarised in Table \ref{tab:contvar}.
 In terms of magnitude, GI region has the highest possible absolute value for each model.
-An important consideration is that region and year are binary variables, such that they are only equal to zero or the coefficient (as they will present as a value of 1 multiplied by the coefficient); this means that, although region may have a strong relationship, it can be overshadowed by an extreme value of one of the continuous variables.
+An important consideration is that region and year are binary, such that they are only equal to zero or the coefficient (as they will present as a value of 1 which will be multiplied by the coefficient); this means that, although region may have a strong relationship, it can be overshadowed by an extreme value of one of the continuous variables.
 The most notable difference between the continuous variables coefficients is the change from positive to negative values.
-This change occurs between the Models for Yield (Model 1 and 2) and the Models for value (Models 3 and 4); where all but the coefficient for area harvested has the opposite sign between them.
+This change occurs between the Models for Yield (Model 1 and 2) and the Models for value (Models 3 and 4); where all but the coefficient for area harvested had the opposite sign.
 These models also differ in an order of magnitude when looking at resource use, with the coefficients for yield being smaller than those for value.
@@
 \end{table}
-\subsubsection{Model Comparisons: Productivity Verse Value}
+\subsubsection{Model Comparisons: Yield Verse Value}
+
 % wHY ARE WE COMPARING MODELS?
 % We compare models to see which predictors contributed relatively to each of the models.
-
 %
@@
 % Graph of productivity verse value.
-
-(see Table \ref*{tab:modelperformance}). 
+%(see Table \ref*{tab:modelperformance}). 
 % This is important but you do not illustrate why

### 2023-07-28 analysis.r
Fixed scatter plots: legend, ticks, titles.

@@ -1,4 +1,5 @@
 library(ggplot2)
+library(gridExtra)
 
 # data_prep.py has to be executed in order for df.csv to exist
 set.seed(100)
@@ -249,84 +250,118 @@ anova(lm(df$ha_value ~ df$ha_tonnes_grapes_harvested))
 # #FFC107
 # #004D40
 
-my_colors <- c(
+temp_colors <- c(
     # Cool
     #"#1E88E5"
     "#1E88E5"
-    , "#1E88E5"
-    , "#1E88E5"
     # Hot
     #, "#D81B60"
-    , "#D81B60"
     , "#D81B60"
     # Mild
     #, "#004D40"
     , "#004d40"
-    , "#004D40"
-    , "#004D40"
     # Warm
     #, "#FFC107"
     , "#FFC107"
-    , "#FFC107"
-    , "#FFC107"
 )
-my_colors <- c(
-    # Cool
+rain_colors <- c(
+    # Damp
     #"#1E88E5"
     "#1E88E5"
-    , "#FFC107"
-    , "#D81B60"
-    # Hot
+    # Dry
     #, "#D81B60"
-    , "#1E88E5"
-    , "#D81B60"
-    # Mild
-    #, "#004D40"
-    , "#1E88E5"
-    , "#FFC107"
-    , "#D81B60"
-    # Warm
+    , "#FFC107"# Very Dry
     #, "#FFC107"
-    , "#1E88E5"
-    , "#FFC107"
     , "#D81B60"
 )
-scale_color_manual(values = my_colors)
+#scale_color_manual(values = my_colors)
 
 # here are some plots
-yield.vs.value <- ggplot(
-    df[df$climate!="", ]
+temp_yield_vs_value <- ggplot(
+    df[df$temp!="", ]
     , aes(
         x = tonnes_grapes_harvested
         , y = tonnes_grapes_harvested * average_per_tonne
-        , col = climate)
+        , col = temp)
     ) +
     xlim(-2.5, 2.5) +
     ylim(-2.5, 2.5) +
     geom_point(size = .5) +
     xlab("Yield") +
-    ylab("Yield * Average Price") +
+    ylab("Yield * Averfdxage Price") +
+    ggtitle("Temperature Class") +
     theme_light() +
-    scale_color_manual(values = my_colors)
-pdf("yvv.pdf")
-print(yield.vs.value)
+    theme(panel.grid.minor = element_blank()
+        , legend.title = element_blank()
[…]

### 2023-07-28 analysis.r
Removed word class from graph titles

@@ -289,7 +289,7 @@ temp_yield_vs_value <- ggplot(
     geom_point(size = .5) +
     xlab("Yield") +
     ylab("Yield * Averfdxage Price") +
-    ggtitle("Temperature Class") +
+    ggtitle("Temperature") +
     theme_light() +
     theme(panel.grid.minor = element_blank()
         , legend.title = element_blank()
@@ -308,7 +308,7 @@ rain_yield_vs_value <- ggplot(
         geom_point(size = .5) +
         xlab("Yield") +
         ylab("Yield * Average Price") +
-        ggtitle("Rainfall Class") +
+        ggtitle("Rainfall") +
         theme_light() +
         theme(panel.grid.minor = element_blank()
             , legend.title = element_blank()
@@ -333,7 +333,7 @@ temp_yield_vs_value <- ggplot(
     geom_point(size = .5) +
     xlab("Yield / Area Harvested") +
     ylab("Yield * Average Price / Area Harvested") +
-    ggtitle("Temperature Class") +
+    ggtitle("Temperature") +
     theme_light() +
     theme(panel.grid.minor = element_blank()
         , legend.title = element_blank()
@@ -352,7 +352,7 @@ rain_yield_vs_value <- ggplot(
         geom_point(size = .5) +
         xlab("Yield / Area Harvested") +
         ylab("Yield * Average Price / Area Harvested") +
-        ggtitle("Rainfall Class") +
+        ggtitle("Rainfall") +
         theme_light() +
         theme(panel.grid.minor = element_blank()
             , legend.title = element_blank()

### 2023-07-28 article.tex
Added scatter plot figures to model comparison

@@
 \subsubsection{Model Comparisons: Yield Verse Value}
-
+%
+\begin{figure}\label{fig:yield_vs_value}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{yield_verse_value.pdf}}
+  \caption{Scatter plot of vineyard yield against the product of yield and average price per tonne. The axes are in standard deviations with points coloured by climate.}
+\end{figure}
+%
+\begin{figure}\label{fig:yield_vs_value_area}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{yield_verse_value_by_area.pdf}}
+    \caption{Scatter plot of vineyard yield against the product of yield and average price per tonne as ratios to area harvested. The axes are in standard deviations with points coloured by climate.}
+\end{figure}
 % wHY ARE WE COMPARING MODELS?

### 2023-07-28 analysis.r
map prints to pdf, added axis limits

@@ -359,7 +359,6 @@ rain_yield_vs_value <- ggplot(
             , legend.position = "top", legend.direction = "horizontal") +
         scale_color_manual(values = rain_colors)
 
-
 pdf("yield_verse_value_by_area.pdf")
 grid.arrange(temp_yield_vs_value, rain_yield_vs_value, ncol=2)
 dev.off()
@@ -386,10 +385,14 @@ v[v$GI_NAME==region , ]$ha_value <- aggs.ha_value[aggs.ha_value$Group.1==region 
 }
 
 # note these are the averages.
-plot(v, "ha_value")
+
+pdf("my_map")
+plot(v, "ha_value", ylim=c(-45, -25))
 lines(aus)
-sbar(1000, lonlat=TRUE, label="1000km")
-north(xy="bottomleft", type=2)
+#    ylim(-45, -25) +
+#    sbar(1000, lonlat=TRUE, label="1000km") +
+#    north(xy="bottomleft", type=2)
+dev.off()
 
 # below is just a way to grab the coef from a linear model.
 #summary(model2)$coefficients

### 2023-07-31 analysis.r
readable maps complete. Needs to beautified.

@@ -373,25 +373,65 @@ aus <- vect("data/STE_2021_AUST_SHP_GDA2020/STE_2021_AUST_GDA2020.shp")
 v$ha_tonnes_grapes_harvested <- 0
 v$ha_value <- 0
 
-mean.t <- function(x) {mean(x, na.rm=TRUE)}
-aggs.ha_tonnes_grapes_harvested <- aggregate(nt$ha_tonnes_grapes_harvested*1000, list(nt$giregion), FUN=mean.t)
-aggs.ha_value <- aggregate(nt$ha_value, list(nt$giregion), FUN=mean.t)
+mean.t <- function(x) {
+    mean(x, na.rm = TRUE)
+}
 
-for (region in unique(df$giregion))
-{
-v[v$GI_NAME==region , ]$ha_tonnes_grapes_harvested <-
- aggs.ha_tonnes_grapes_harvested[aggs.ha_tonnes_grapes_harvested$Group.1==region , ]$x
-v[v$GI_NAME==region , ]$ha_value <- aggs.ha_value[aggs.ha_value$Group.1==region , ]$x
+aggs.ha_tonnes_grapes_harvested <- aggregate(
+    nt$ha_tonnes_grapes_harvested * 10000
+    , list(nt$giregion), FUN = mean.t)
+aggs.ha_value <- aggregate(
+    nt$ha_value
+    , list(nt$giregion)
+    , FUN=mean.t)
+
+for (region in unique(df$giregion)){
+    v[v$GI_NAME == region
+    , ]$ha_tonnes_grapes_harvested <- aggs.ha_tonnes_grapes_harvested[
+    aggs.ha_tonnes_grapes_harvested$Group.1 == region
+    , ]$x
+    v[
+        v$GI_NAME==region
+        , ]$ha_value <- aggs.ha_value[
+            aggs.ha_value$Group.1 == region
+            , ]$x
 }
 
 # note these are the averages.
-
 pdf("my_map")
-plot(v, "ha_value", ylim=c(-45, -25))
-lines(aus)
-#    ylim(-45, -25) +
-#    sbar(1000, lonlat=TRUE, label="1000km") +
-#    north(xy="bottomleft", type=2)
+
+plot(v
+    , "ha_value"
+    , asp = 1
+   , xlim = c(113, 154)
+   , ylim = c(-45, -25)
+   , legend = "bottomleft"
+   , main = "Average Value per Hectare"
+)
+lines(aus,
+   , xlim = c(113, 154)
[…]

### 2023-07-31 analysis.r
missing .pdf at end of output file

@@ -398,7 +398,7 @@ for (region in unique(df$giregion)){
 }
 
 # note these are the averages.
-pdf("my_map")
+pdf("my_map.pdf")
 
 plot(v
     , "ha_value"

### 2023-07-31 article.tex
Added map place holder
Added pearson corr table and ref to it.

--- 2023-07-28 article.tex
@@
 \subsection{Exploratory Analysis} \label{sec:exp_anal}
+\begin{table}[]
+  \caption{Variable Pearson correlation values for logarithmically transformed values.}\label{tab:pearson_correlation}
+      \resizebox{\textwidth}{!}{
+        \begin{tabular}{@{}cccccccc@{}}
+        \toprule
+        \textbf{Variable} & \textbf{Yield} & \textbf{Area Harvested} & \textbf{Water Used} & \textbf{Scope One Emissions} & $\textbf{Yield}\over \textbf{Area}$ & \textbf{Average Sale Price} & $\textbf{Average Sale Price}\over \textbf{Area Harvested}$ \\ \midrule
+        \textbf{Yield} & 1.00E+00 & 7.44E-01 & -4.31E-03 & 7.29E-01 & 3.50E-01 & -2.26E-01 & -1.64E-01 \\
+        \textbf{Area Harvested} & 7.44E-01 & 1.00E+00 & -5.33E-03 & 8.92E-01 & 7.85E-02 & -1.18E-01 & -2.04E-01 \\
+        \textbf{Water Used} & -4.31E-03 & -5.33E-03 & 1.00E+00 & -1.93E-03 & -5.60E-03 & -3.56E-02 & -2.67E-02 \\
+        \textbf{Scope One Emissions} & 7.29E-01 & 8.92E-01 & -1.93E-03 & 1.00E+00 & 9.36E-02 & -9.42E-02 & -1.93E-01 \\
+        $\textbf{Yield}\over \textbf{Area}$ & 3.50E-01 & 7.85E-02 & -5.60E-03 & 9.36E-02 & 1.00E+00 & -4.85E-01 & -1.70E-01 \\
+        \textbf{Average Sale Price} & -2.26E-01 & -1.18E-01 & -3.56E-02 & -9.42E-02 & -4.85E-01 & 1.00E+00 & 4.73E-01 \\
+        $\textbf{Average Sale Price} \over \textbf{Area Harvested}$ & -1.64E-01 & -2.04E-01 & -2.67E-02 & -1.93E-01 & -1.70E-01 & 4.73E-01 & 1.00E+00 \\ \bottomrule
+        \end{tabular}
+}
+\end{table}
 Linear relationships between variables were explored using Pearson Correlation Coefficients.
 Values for these coefficients reflect the linear relation between two variables, on a scale between -1 and 1; the magnitude and sign of a coefficient indicates the strength of the relation, and whether the relation is positive or negative respectively.
-This was undertaken for data on the original scale (see Table ??) and for data as a logarithmic transform (see Table ??).
-The logarithmic transformed data showed the strongest correlations, likely due to a skew caused by a greater number of smaller vineyards within the dataset (see Table ??).
+This was undertaken for data on the original scale and for data as a logarithmic transform.
+The logarithmic transformed data showed the strongest correlations, likely due to a skew caused by a greater number of smaller vineyards within the dataset (see Table \ref{tab:pearson_correlation}).
 Transforming data prior to calculating the coefficients changes several things: The logarithmic transform of the data alters the interpretation of the coefficients to percentage change - a coefficient will be indicative of the change in percentage of one variable compared to the other; scaling by standard deviation also changes this interpretation to be a percentage of that variables standard deviation.

### 2023-07-31 article.tex

--- 2023-07-28 article.tex
@@
 \subsection{Exploratory Analysis}\label{sec:exp_anal}
+\begin{table}[]
+  \caption{Variable Pearson correlation values for logarithmically transformed values.}\label{tab:pearson_correlation}
+      \resizebox{\textwidth}{!}{
+        \begin{tabular}{@{}cccccccc@{}}
+        \toprule
+        \textbf{Variable} & \textbf{Yield} & \textbf{Area Harvested} & \textbf{Water Used} & \textbf{Scope One Emissions} & $\textbf{Yield}\over \textbf{Area}$ & \textbf{Average Sale Price} & $\textbf{Average Sale Price}\over \textbf{Area Harvested}$ \\ \midrule
+        \textbf{Yield} & 1.00E+00 & 7.44E-01 & -4.31E-03 & 7.29E-01 & 3.50E-01 & -2.26E-01 & -1.64E-01 \\
+        \textbf{Area Harvested} & 7.44E-01 & 1.00E+00 & -5.33E-03 & 8.92E-01 & 7.85E-02 & -1.18E-01 & -2.04E-01 \\
+        \textbf{Water Used} & -4.31E-03 & -5.33E-03 & 1.00E+00 & -1.93E-03 & -5.60E-03 & -3.56E-02 & -2.67E-02 \\
+        \textbf{Scope One Emissions} & 7.29E-01 & 8.92E-01 & -1.93E-03 & 1.00E+00 & 9.36E-02 & -9.42E-02 & -1.93E-01 \\
+        $\textbf{Yield}\over \textbf{Area}$ & 3.50E-01 & 7.85E-02 & -5.60E-03 & 9.36E-02 & 1.00E+00 & -4.85E-01 & -1.70E-01 \\
+        \textbf{Average Sale Price} & -2.26E-01 & -1.18E-01 & -3.56E-02 & -9.42E-02 & -4.85E-01 & 1.00E+00 & 4.73E-01 \\
+        $\textbf{Average Sale Price} \over \textbf{Area Harvested}$ & -1.64E-01 & -2.04E-01 & -2.67E-02 & -1.93E-01 & -1.70E-01 & 4.73E-01 & 1.00E+00 \\ \bottomrule
+        \end{tabular}
+}
+\end{table}
 Linear relationships between variables were explored using Pearson Correlation Coefficients.
 Values for these coefficients reflect the linear relation between two variables, on a scale between -1 and 1; the magnitude and sign of a coefficient indicates the strength of the relation, and whether the relation is positive or negative respectively.
-This was undertaken for data on the original scale (see Table ??) and for data as a logarithmic transform (see Table ??).
-The logarithmic transformed data showed the strongest correlations, likely due to a skew caused by a greater number of smaller vineyards within the dataset (see Table ??).
+This was undertaken for data on the original scale and for data as a logarithmic transform.
+The logarithmic transformed data showed the strongest correlations, likely due to a skew caused by a greater number of smaller vineyards within the dataset (see Table \ref{tab:pearson_correlation}).
 Transforming data prior to calculating the coefficients changes several things: The logarithmic transform of the data alters the interpretation of the coefficients to percentage change - a coefficient will be indicative of the change in percentage of one variable compared to the other; scaling by standard deviation also changes this interpretation to be a percentage of that variables standard deviation.
@@
 \end{figure}
+%
+\begin{figure}\label{fig:map}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{my_map.pdf}}
+  \caption{Map of regional average yield and value per hectare.}
+\end{figure}
+%
 % wHY ARE WE COMPARING MODELS?

### 2023-07-31 article.tex
Added table

@@
 \section{Results}
+\subsection{Data}
+Each variable was logarithmically transformed and then centred around a mean of 0.
+The values of these variables were then divided by standard deviation creating a comparable ratio intrinsic to each variable.
+Table \ref{tab:summary} shows the summary statistics of each variable, to contextualise these ratios to real values.
+
+\begin{table}[]
+  \caption{Summary statistics of each continuous variable.}\label{tab:summary}
+      \resizebox{\textwidth}{!}{
+        \begin{tabular}{@{}cllll@{}}
+          \toprule
+          \textbf{Variable} & \textbf{Mean} & \textbf{\begin{tabular}[c]{@{}l@{}}Standard\\ Deviation\end{tabular}} & \textbf{Minimum} & \textbf{Maximum} \\ \midrule
+          \textbf{Yield} & 7.757E+02 & 2.179E+03 & 1.000E+00 & 7.231E+04 \\
+          \multicolumn{1}{l}{} &  &  &  &  \\
+          \textbf{Area Harvested} & 6.670E+05 & 1.337E+06 & 7.000E+02 & 2.436E+07 \\
+          \multicolumn{1}{l}{} &  &  &  &  \\
+          \textbf{Water Used} & 7.471E+06 & 5.646E+08 & 1.000E+00 & 4.268E+10 \\
+          \multicolumn{1}{l}{} &  &  &  &  \\
+          \textbf{Scope One Emissions} & 4.173E+04 & 8.571E+04 & 6.755E+00 & 2.110E+06 \\
+          \multicolumn{1}{l}{} &  &  &  &  \\
+          $\textbf{Yield}\over \textbf{Area}$ & 1.009E+01 & 8.127E+00 & 4.000E-02 & 8.634E+01 \\
+          \multicolumn{1}{l}{} &  &  &  &  \\
+          \textbf{Average Sale Price} & 1.477E+03 & 9.216E+02 & 1.600E+02 & 2.600E+04 \\
+          \multicolumn{1}{l}{} &  &  &  &  \\
+          $\textbf{Average Sale Price} \over \textbf{Area Harvested}$ & 1.347E+02 & 5.711E+02 & 1.753E-01 & 2.979E+04 \\ \bottomrule
+          \end{tabular}
+  }
+  \end{table}
+
 \subsection{Exploratory Analysis}\label{sec:exp_anal}

### 2023-07-31 article.tex
fixing kfold table

@@
   \caption{Model validation using k-fold cross validation, for 10 folds repeated 100 times.}
-  \begin{tabular}{cccc}
-          & RMSE      & R2        & MAE       \\
-  Model 1 & 3.087E-01 & 9.045E-01 & 2.165E-01 \\
-  Model 2 & 5.104E-01 & 7.409E-01 & 3.493E-01 \\
-  Model 3 & 1.652E-01 & 9.723E-01 & 1.008E-01 \\
-  Model 4 & 2.235E-01 & 9.500E-01 & 1.279E-01
+  \resizebox{\textwidth}{!}{
+  \begin{tabular}{@{}cccc@{}}
+    \toprule
+    \textbf{} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Mean\\ Squared Error\end{tabular}} & \textbf{R2} & \textbf{\begin{tabular}[c]{@{}c@{}}Mean Average \\ Error\end{tabular}} \\ \midrule
+    \textbf{Model 1} & 3.087E-01 & 9.045E-01 & 2.165E-01 \\
+    \textbf{Model 2} & 5.104E-01 & 7.409E-01 & 3.493E-01 \\
+    \textbf{Model 3} & 1.652E-01 & 9.723E-01 & 1.008E-01 \\
+    \textbf{Model 4} & 2.235E-01 & 9.500E-01 & 1.279E-01 \\ \bottomrule
     \end{tabular}
+  }
   \end{table}

### 2023-07-31 article.tex
Minor edits - up to end of results.

@@
 In this analysis we observe relationships between yield and quality through the use of linear models.
-Although an extensive amount of research into a variety of factors' effect on grape quality and yield exists; due to the lack of long-term and in-depth data, individual effects are often studied in isolation \citep{abbalDecisionSupportSystem2016}.
+An extensive amount of research into a variety of factors' effect on grape quality and yield exists; due to the lack of long-term and in-depth data, individual effects are often studied in isolation \citep{abbalDecisionSupportSystem2016}.
 The lack of consolidated datasets also restricts the ability to gain statistical insights at large scales and across multiple regions \citep{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
@@
 \newline
-We aim to use this broad dataset to confirm the existence of a yield verse quality trade off within Australian winegrowing; one not prior confirmed explicitly across such extensive diversities.
-% you require a reference here
-In achieving this, the context of how resource-use relates to yield and quality will also be described.
-% Describing baselines is more important - it needs to be emphasised more so than something that is known but not shown. This is the heart of what your are doing sell it more.
-We link these relations to the potential for improvement through decision-making processes, whilst highlighting that the way moving forward will require the optimisation of these processes.
-The practical addition of these aims is a baseline for comparison - given a vineyard within Australia, one could extrapolate their comparative efficiency with regard to the tradeoff between invested resources, yield and quality.
+We aim to use this broad dataset to describe the relationship of input resources to the output yield and quality of vineyards.
+The practical addition of this aim is a baseline for comparison - given a vineyard within Australia, one could extrapolate their comparative efficiency with regard to the tradeoff between invested resources, yield and quality.
+In achieving this we will also confirm the existence of a yield verse quality trade off within Australian winegrowing; one not prior confirmed explicitly across such varying regions, scales and climates.
 \section{Methods}
@@
           \toprule
-           & $\mathbf{R^2}$ & $\mathbf{R^2}$ & \textbf{F-Statistic} & \textbf{P-Value} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual\\ Standard Error\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Sum\\ of Squares\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Mean\\ of Squares\end{tabular}} \\ \midrule
+           & $\mathbf{R^2}$ & \textbf{\begin{tabular}[c]{@{}c@{}} Adjusted\\ $\mathbf{R^2}$ \end{tabular}} & \textbf{F-Statistic} & \textbf{P-Value} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual\\ Standard Error\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Sum\\ of Squares\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Mean\\ of Squares\end{tabular}} \\ \midrule
           \textbf{\begin{tabular}[c]{@{}c@{}}Model 1\\ Yield\end{tabular}} & 9.072E-01 & 9.061E-01 & 7.753E+02 & 2.200e-16 & 3.065E-01 & 4.913E+02 & 1.000E-01 \\
@@
 \newline
+%b Should this be in the discussion (below para)
 Scope 1 emissions was included in all models to directly compare the response variables as ratios of vineyard size to raw values.
@@
 For Models 1 (yield) and 3 (value) year played a pivotal role, with only one year in each model not being significant (2021/2022 and 2016/2017 respectively).
-Both Model 1 and 3 showed a majority of regions were significant with 32 of 54 regions being significant in Model 1, and 42 of 50 regions being significant in Model 3 with 95\% confidence.
+Both Model 1 and 3 showed a majority of regions were significant with 32 of 54 regions being significant in Model 1, and 42 of 50 regions being significant in Model 3 at 95\% confidence.
 \newline
@@
 \newline
-With regard to continuous variables, Model 1 and 2 showed water use, scope 1 emissions and area harvested were significant at 95\% confidence when accounting for other variables.
+With regard to continuous variables: Model 1 and 2 showed all variables to be significant at 95\% confidence when accounting for other variables.
 T-tests for Model 3 showed all continuous variables except scope 1 emissions were significant.
-Model 4 showed scope 1 emissions and water use to only be significant when considered as an interaction with area harvested but not when considered on their own.
+Model 4 showed all variables aside from scope 1 emissions and water use to be significant; with scope 1 emissions and water use only being significant when considered as an interaction with area harvested but not when considered on their own.
 \subsubsection{Model Coefficients}
@@
 The coefficients of each model describe the relationship of a predictor variable to its response when considering all other variables.
-Due to the transformations of the data, coefficients are individually interpreted in the same manner as the prior regression values were (see Section \ref{sec:exp_anal}); unlike the regression values,  coefficient ranges are not limited between -1 and 1, as each variable needs to be considered together.
+Due to the transformations of the data, coefficients are individually interpreted in the same manner as the prior regression values were (see Section \ref{sec:exp_anal}); unlike the regression values,  coefficient ranges are not limited between -1 and 1.
+%, as each variable needs to be considered together.
 %The dimensionless nature of these variables each coefficient refers to its own variable's distribution's standard deviations.
@@
 We look at the coefficients of categorical and continuous variables separately.
-This is primarily done as the categorical variables have many coefficients, one for each category, whilst continuous variables have only one.
-The coefficient for categorical variables is summarised in Figure\ref{fig:violin}; illustrating the difference in the range as well as affect region and year could have on each of the models.
+This is done as the categorical variables have many coefficients, one for each category, whilst continuous variables have only one.
+The coefficient for categorical variables is summarised in Figure \ref{fig:violin}; illustrating the difference in the range as well as affect region and year could have on each model.
 Comparatively, the continuous variables coefficients are summarised in Table \ref{tab:contvar}.
@@
 The most notable difference between the continuous variables coefficients is the change from positive to negative values.
-This change occurs between the Models for Yield (Model 1 and 2) and the Models for value (Models 3 and 4); where all but the coefficient for area harvested had the opposite sign.
+This change occurs between the Models for Yield (Model 1 and 2) and the Models for value (Models 3 and 4); where all but the coefficient for area harvested had the opposite sign (see Table \ref{contvar}).
 These models also differ in an order of magnitude when looking at resource use, with the coefficients for yield being smaller than those for value.
@@
 %
-\begin{figure}
-  \caption{Violin plots of GI Region and Year coefficients for each model.}\label{fig:violin}
+\begin{figure}\label{fig:violin}
   \includegraphics{violinplot.pdf}
+  \caption{Violin plots of GI Region and Year coefficients for each model.}
   \end{figure}

### 2023-07-31 article.tex
Andrew edits

@@
 In this analysis we observe relationships between yield and quality through the use of linear models.
-An extensive amount of research into a variety of factors' effect on grape quality and yield exists; due to the lack of long-term and in-depth data, individual effects are often studied in isolation \citep{abbalDecisionSupportSystem2016}.
+An extensive amount of research into a variety of factors' effect on grape quality and yield exists; but due to the lack of long-term and in-depth data, individual effects are often studied in isolation \citep{abbalDecisionSupportSystem2016}.
 The lack of consolidated datasets also restricts the ability to gain statistical insights at large scales and across multiple regions \citep{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
@@
 Preprocessing included logarithmic transformations, centring and scaling by standard deviation.
-Variables such as scope 1, which required prior calculations were also computed using Python.
+Variables such as scope 1 emissions, which required prior calculations were also computed using Python.
 % Two values for average sale price were removed from the dataset, due to a recording of \$1. 
@@
 \newline
-was used to convert the quantity of fuel in litres, $Q$, using a prescribed Energy Content, $EC$, and emission factors of scope one, $EF1$, and scope three, $EF3$, to tonnes of Carbon Dioxide equivalent, $tCO2e$ \citep{departmentofclimatechangeenergytheenvironmentandwaterAustralianNationalGreenhouse2022}.
+was used to convert the quantity of fuel in litres, $Q$, using a prescribed Energy Content, $EC$, and emission factors of scope one, $EF1$, and scope three, $EF3$, to tonnes of Carbon Dioxide Emission equivalent, $tCO2e$ \citep{departmentofclimatechangeenergytheenvironmentandwaterAustralianNationalGreenhouse2022}.
 Emissions were calculated for total diesel, petrol, bio-diesel and LPG used.
@@
 \end{table}
+\section{stop here}
 \subsubsection{Model Comparisons: Yield Verse Value}
@@
 %
-%Models were compared using $R^2$ values.
-The comparison of models performance showed that the average price per tonne of grapes describes a great deal of the relationship between predictor and response when comparing Model 2 to Model 4
-% Graph of productivity verse value.
-%(see Table \ref*{tab:modelperformance}). 
-% This is important but you do not illustrate why
+% Although the graphs showing yield and value look similar when presented as raw values and as ratios of area. The flow of resource use is inverted.
+% Talk about how resources use relates to these graphs.
 %
-% This section needs to be heavily revised. Why are there tails on the distributions? What does a heavy tail indicate?
-% What does it look like with respect to disaster events?
-% split the discussion sections out of the results!! Even the questions I am asking are answered in the discussion, however they need to be highlighted as aspects of the models here beforehand.
- Reviewing the data to uncover reasons for this included the use of binary variables such as the utilisation of renewable energy, contractors, and the occurrence of disease, fire and frost; however none of these variables were able to explain why some vineyards produced less, or why other vineyards sold at higher prices than predicted.
-A wide variety of these influences were likely already explained within the use of year and GI Region, or the interaction of both variables.
-The change between some regions was dramatic, with particularly warmer and drier regions producing much higher volumes of grapes at lower prices (See Figures 5 and 6).
-The use of other variables and methods, specifically splines, were able to create a more normally distributed set of residuals but at a drastically reduced accuracy when comparing R2 and RSE.
-The introduction of known average prices per tonne also helped increase R2 values a small amount; it is important to not that it is common practice for wineries to purchase grapes at a regional average rate, likely resulting in much less variance within a region.
-
-different strategies are likely employed between different regions, where some regions target the mass production of cheaper grapes over quality.
-This is most notable when grouping regions by climate, especially when considering GI Regions in the 'Hot Very Dry' climate (see Figure 7).
-The effect of climate in the models was not more significant than the more granular use of GI regions.
-The interaction between year and GI Region likely accounted for localised events such as bushfires, which would be impactful, but only at a local level in both time and space.
+% Show how regions are almost inverted between avereage value vs yield per area when looking at the maps.
+%
+%
 
@@
 %
+
+%Models were compared using $R^2$ values.
+The comparison of models performance showed that the average price per tonne of grapes describes a great deal of the relationship between predictor and response when comparing Model 2 to Model 4
+% Graph of productivity verse value.
+%(see Table \ref*{tab:modelperformance}). 
+% This is important but you do not illustrate why
+%
+% This section needs to be heavily revised. Why are there tails on the distributions? What does a heavy tail indicate?
+% What does it look like with respect to disaster events?
+% split the discussion sections out of the results!! Even the questions I am asking are answered in the discussion, however they need to be highlighted as aspects of the models here beforehand.
+Reviewing the data to uncover reasons for this included the use of binary variables such as the utilisation of renewable energy, contractors, and the occurrence of disease, fire and frost; however none of these variables were able to explain why some vineyards produced less, or why other vineyards sold at higher prices than predicted.
+A wide variety of these influences were likely already explained within the use of year and GI Region, or the interaction of both variables.
+The change between some regions was dramatic, with particularly warmer and drier regions producing much higher volumes of grapes at lower prices (See Figures 5 and 6).
+The use of other variables and methods, specifically splines, were able to create a more normally distributed set of residuals but at a drastically reduced accuracy when comparing R2 and RSE.
+The introduction of known average prices per tonne also helped increase R2 values a small amount; it is important to not that it is common practice for wineries to purchase grapes at a regional average rate, likely resulting in much less variance within a region.
+
+different strategies are likely employed between different regions, where some regions target the mass production of cheaper grapes over quality.
+This is most notable when grouping regions by climate, especially when considering GI Regions in the 'Hot Very Dry' climate (see Figure 7).
+The effect of climate in the models was not more significant than the more granular use of GI regions.
+The interaction between year and GI Region likely accounted for localised events such as bushfires, which would be impactful, but only at a local level in both time and space.
+
 \subsection{Limitations}

### 2023-07-31 article.tex
minor fixes to table references

@@
 \newline
+\begin{table}[]
+  \caption{Summary of each Models coefficients for continuous variables}\label{tab:contvar}
+      \resizebox{\textwidth}{!}{
+        \begin{tabular}{@{}cllllll@{}}
+          \toprule
+           & \textbf{Intercept} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Water\\ Used\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Scope 1\\ Emissions\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\\ $\ast$\\ Scope 1\\ Emissions\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\\ $\ast$\\ Water\\ Used\end{tabular}}} \\ \midrule
+          \textbf{Model 1} & -3.318E-02 & 7.418E-01 & 8.660E-02 & 6.731E-02 &  &  \\
+          \textbf{Model 2} & -6.516E-01 & 5.774E-01 & 1.079E-02 & 8.498E-02 & -4.971E-02 & -5.346E-02 \\
+          \textbf{Model 3} & 1.808E-02 & 9.713E-01 & -2.310E-02 & -6.992E-03 &  &  \\
+          \textbf{Model 4} & 6.702E-01 & -7.354E-01 & -6.732E-03 & -5.645E-03 & 2.726E-02 & 7.515E-02 \\ \bottomrule
+          \end{tabular}}
+\end{table}
 We look at the coefficients of categorical and continuous variables separately.
@@
 The most notable difference between the continuous variables coefficients is the change from positive to negative values.
-This change occurs between the Models for Yield (Model 1 and 2) and the Models for value (Models 3 and 4); where all but the coefficient for area harvested had the opposite sign (see Table \ref{contvar}).
+This change occurs between the Models for Yield (Model 1 and 2) and the Models for value (Models 3 and 4); where all but the coefficient for area harvested had the opposite sign (see Table \ref{tab:contvar}).
 These models also differ in an order of magnitude when looking at resource use, with the coefficients for yield being smaller than those for value.
@@
 %
-\begin{table}[]
-  \caption{Summary of each Models coefficients for continuous variables}\label{tab:contvar}
-      \resizebox{\textwidth}{!}{
-        \begin{tabular}{@{}cllllll@{}}
-          \toprule
-           & \textbf{Intercept} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Water\\ Used\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Scope 1\\ Emissions\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\\ $\ast$\\ Scope 1\\ Emissions\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\\ $\ast$\\ Water\\ Used\end{tabular}}} \\ \midrule
-          \textbf{Model 1} & -3.318E-02 & 7.418E-01 & 8.660E-02 & 6.731E-02 &  &  \\
-          \textbf{Model 2} & -6.516E-01 & 5.774E-01 & 1.079E-02 & 8.498E-02 & -4.971E-02 & -5.346E-02 \\
-          \textbf{Model 3} & 1.808E-02 & 9.713E-01 & -2.310E-02 & -6.992E-03 &  &  \\
-          \textbf{Model 4} & 6.702E-01 & -7.354E-01 & -6.732E-03 & -5.645E-03 & 2.726E-02 & 7.515E-02 \\ \bottomrule
-          \end{tabular}}
-\end{table}
-\section{stop here}
 \subsubsection{Model Comparisons: Yield Verse Value}
@@
 %
-
+%Although the initial correlation between average sales price and yield was a negative trend (see table 2), these models were able to shed light on several contributing factors to yield and average sales price.
+Correlation values showed that water and emissions increased with yield but decreased with average sale price (see Table 4).
+In alternative attempts at models it was found that without the incorporation of GI Region or year the predictions greatly under performed.
+The possible reason behind this effect was that different strategies are likely employed between different regions, where some regions target the mass production of cheaper grapes over quality.
+This is most notable when grouping regions by climate, especially when considering GI Regions in the ‘Hot Very Dry’ climate (see Figure 7).
+The effect of climate when used in models was not more significant than the more granular use of GI regions.
+The interaction between year and GI Region likely accounted for localised events such as bushfires, which would be impactful, but only at a local level in both time and space.
 \subsection{Model Validation}
-To validate the performance of these models k-fold cross validation was used.
-This was done using 10 folds, $k=10$, repeated 100 times.
-The models performed similarly to their original counter parts (see Table \ref*{tab:kfold}).
 \begin{table}[]
@@
   \end{table}
+To validate the performance of these models k-fold cross validation was used.
+This was done using 10 folds, $k=10$, repeated 100 times.
+The models performed similarly to their original counter parts (see Table \ref{tab:kfold}).
 \section{Discussion}

### 2023-08-01 article.tex
start of model comparison subsection

@@
 %
-% wHY ARE WE COMPARING MODELS?
-% We compare models to see which predictors contributed relatively to each of the models.
+% Why compare models?
+
+Directly comparing response variables, how value changes with yield, also allows an indirect comparison of resource use.
+We do this through using the known relationships of the variables to their predictors.
+These relationships are described by the coefficients, and as the predictors are known to have a strong positive correlation, they will increase together.
+Resource use is described by the predictor variables (through water used and scope 1 emissions), we can observe these response variables somewhat interchangably with the predictors - although caution should be taken to look at them skeptically and through the lens of their appropriate coefficients.
+It is important to also consider the interactions of these variables, especially when comparing response variables that are a ratio of area.
+%
+Although it is adhoc, we also know that each of these predictors are positively correlated when logarithmically transformed, making it likely that the predictors will increase together - although not necessarily at the same rate.
+Observing Figure \ref{fig:yield_vs_value} shows a predictable convergence towards the mean of 0, with trends appearing in almost stratified lines.
+These linear formations are caused by regional trends, with some regions being more predictable than others, such as the the 'Hot' regions.
+\newline
+Comparing Figure \ref{fig:yield_vs_value} to Figure \ref{fig:yield_vs_value_area} shows similar trends.
+However, a not so obvious difference is that the same vineyards (or graphed points) are dispersed differently.
+In many cases these graphs are closer to mirror images of each other; this is more visible when comparing regional averages of these response variables, see Figure \ref{fig:map}.
+We can use the response variables relationship to its predictors to compare how resource use changes 
+Knowing the response variables relation to resource use
+
+%We compare models to see which predictors contributed relatively to each of the models.
 %

### 2023-08-01 analysis.r
TERRIFYING mistake with r plots.

@@ -281,14 +281,14 @@ temp_yield_vs_value <- ggplot(
     df[df$temp!="", ]
     , aes(
         x = tonnes_grapes_harvested
-        , y = tonnes_grapes_harvested * average_per_tonne
+        , y = value
         , col = temp)
     ) +
     xlim(-2.5, 2.5) +
     ylim(-2.5, 2.5) +
     geom_point(size = .5) +
     xlab("Yield") +
-    ylab("Yield * Averfdxage Price") +
+    ylab("Yield * Average Price") +
     ggtitle("Temperature") +
     theme_light() +
     theme(panel.grid.minor = element_blank()
@@ -300,7 +300,7 @@ rain_yield_vs_value <- ggplot(
     df[df$rain!="", ]
     , aes(
         x = tonnes_grapes_harvested
-        , y = tonnes_grapes_harvested * average_per_tonne
+        , y = value
         , col = rain)
         ) +
         xlim(-2.5, 2.5) +
@@ -325,7 +325,7 @@ temp_yield_vs_value <- ggplot(
     df[df$temp!="", ]
     , aes(
         x = ha_tonnes_grapes_harvested
-        , y = ha_tonnes_grapes_harvested * average_per_tonne
+        , y = ha_value
         , col = temp)
     ) +
     xlim(-2.5, 2.5) +
@@ -344,7 +344,7 @@ rain_yield_vs_value <- ggplot(
     df[df$rain!="", ]
     , aes(
         x = ha_tonnes_grapes_harvested
-        , y = ha_tonnes_grapes_harvested * average_per_tonne
+        , y = ha_value
         , col = rain)
         ) +
         xlim(-2.5, 2.5) +

### 2023-08-01 article.tex
Last result section finished

--- 2023-07-31 article.tex
@@
 %
-% wHY ARE WE COMPARING MODELS?
-% We compare models to see which predictors contributed relatively to each of the models.
+% Why compare models?
+%
+Directly comparing response variables, how crop value changes with yield, also allows an indirect comparison between the response variables and resource use.
+We do this through using known relationships of response variables to their predictors.
+These relationships are described by the coefficients.
+Resource use is described by the predictor variables (through water used and scope 1 emissions), because of this we can observe the response variables somewhat interchangeably with the predictors - although caution should be taken to view them sceptically and alongside the influence of their coefficients.
+As the predictors are known to have a strong positive correlation with each other, they will tend toward increasing and decreasing together (but not at the same rates).
+It is also important to consider the interactions of predictor variables when comparing the response variables that are ratios of area.
+Furthermore, these comparisons require the consideration of the covariates, in this case: area harvested, year and region.
+\newline
+Observing Figure \ref{fig:yield_vs_value} shows an almost discrete difference between vineyards in 'Hot' areas than other regions.
+Comparing Figure \ref{fig:yield_vs_value} to Figure \ref{fig:yield_vs_value_area} shows almost opposing trends; a not so obvious difference between the Figures, is that the difference is mostly a rotation (being a $90^\degree$ clockwise reflection).
+However, with area coming into play, many data points are scaled differently; specifically the vineyards from 'Hot' regions are then found to be on the tail end, producing large quantity of lower value grapes.
+This is more visible when comparing both graphs to the map of regional averages for response variables, see Figure \ref{fig:map}.
+There is a notable change between regional averages when looking at yield verse value.
+Through the coefficients we can deduce that: this difference is also a difference between more resources used for the raw response variables; and a difference between overall resource use and the size of the vineyard.
+Where resource use and area harvested have a combined relationship through their interaction and separate relationships as individual variables (see Table \ref{tab:contvar}).
+A notable occurrence in Figure \ref{fig:yield_vs_value_area}, is that the 'Very Dry' vineyards which produce lower yields and higher quality grapes are predominantly found in the Barossa Valley (a wine region known for its high quality Shiraz).
+This note is important as it shows climate is not exclusively the consideration, soil and other geographical phenomenon have considerable impacts on vineyard outcomes.
+
+%We compare models to see which predictors contributed relatively to each of the models.
 %

### 2023-08-01 article.tex
fixed minor latex error with \degree

@@
 Observing Figure \ref{fig:yield_vs_value} shows an almost discrete difference between vineyards in 'Hot' areas than other regions.
-Comparing Figure \ref{fig:yield_vs_value} to Figure \ref{fig:yield_vs_value_area} shows almost opposing trends; a not so obvious difference between the Figures, is that the difference is mostly a rotation (being a $90^\degree$ clockwise reflection).
+Comparing Figure \ref{fig:yield_vs_value} to Figure \ref{fig:yield_vs_value_area} shows almost opposing trends; a not so obvious difference between the Figures, is that the difference is mostly a rotation (being 90\textdegree clockwise).
 However, with area coming into play, many data points are scaled differently; specifically the vineyards from 'Hot' regions are then found to be on the tail end, producing large quantity of lower value grapes.

### 2023-08-01 article.tex
Reshuffled discussion

@@
 Observing Figure \ref{fig:yield_vs_value} shows an almost discrete difference between vineyards in 'Hot' areas than other regions.
-Comparing Figure \ref{fig:yield_vs_value} to Figure \ref{fig:yield_vs_value_area} shows almost opposing trends; a not so obvious difference between the Figures, is that the difference is mostly a rotation (being 90\textdegree clockwise).
-However, with area coming into play, many data points are scaled differently; specifically the vineyards from 'Hot' regions are then found to be on the tail end, producing large quantity of lower value grapes.
+Comparing Figure \ref{fig:yield_vs_value} to Figure \ref{fig:yield_vs_value_area} shows almost opposing trends.
+However, with area coming into play in Figure \ref{fig:yield_vs_value_area}, many data points are scaled differently; specifically the vineyards from 'Hot' regions change to be found the bottom right tail end, indicating the production of large quantity of lower value grapes.
+An un obvious difference between the Figures, is that a large amount of the difference can be explained by rotation (being $90^\circ$ clockwise from Figure \ref{fig:yield_vs_value} to \ref{fig:yield_vs_value_area}).
 This is more visible when comparing both graphs to the map of regional averages for response variables, see Figure \ref{fig:map}.
 There is a notable change between regional averages when looking at yield verse value.
-Through the coefficients we can deduce that: this difference is also a difference between more resources used for the raw response variables; and a difference between overall resource use and the size of the vineyard.
-Where resource use and area harvested have a combined relationship through their interaction and separate relationships as individual variables (see Table \ref{tab:contvar}).
+Through the coefficients we can deduce that: this difference is also a difference between more resources used for the raw response variables; and a difference between overall resource use and the size of the vineyard when considering the response variables as a ratio to area.
+Noting, resource use and area harvested have a combined relationship through their interactions, and separate relationships as individual variables (see Table \ref{tab:contvar}).
 A notable occurrence in Figure \ref{fig:yield_vs_value_area}, is that the 'Very Dry' vineyards which produce lower yields and higher quality grapes are predominantly found in the Barossa Valley (a wine region known for its high quality Shiraz).
 This note is important as it shows climate is not exclusively the consideration, soil and other geographical phenomenon have considerable impacts on vineyard outcomes.
-
 %We compare models to see which predictors contributed relatively to each of the models.
@@
 This was done using 10 folds, $k=10$, repeated 100 times.
-The models performed similarly to their original counter parts (see Table \ref{tab:kfold}).
+The models performed similarly to their original counterparts (see Table \ref{tab:kfold}).
+
 \section{Discussion}
-% We want to compare how much influence each variable has on the model!
-% This is prior to postulating about the why these variables matter, or have so much influence.
-% 
-% A table will help to compare the continuous variables, the others are done in the violin plot.
-%
+The most determining factor in the output quality and yield of a vineyard was size.
+There was a strong relationship between size and resource use, in that the overall space of a vineyard and its access to resources determined the upper limit of potential yield.
+However, size was also inversely related to the potential quality, with higher quality being related more to high resource inputs per area than overall access to resources.
+These effects were augmented by regional and yearly affects.
+Even given regional and yearly changes, there was a strong sense of smaller vineyards being easier to manage and ensure high quality when compared to larger vineyards.
+\subsection{Resource use and Yield verse Quality} 
+Literature shows that there are many on-the-ground decisions that influence both quality and yield.
+These decisions are governed by complex physical and social forces such as international market demands, disease pressures and natural disasters \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}; with many of these occurrences being highlighted throughout the past decades vintage reports \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
+ %This definitely needs to be broken up, what were these occurences, which year was significant for what?
+It is also important to consider that these reports show that the warm inland regions have seen a decline in profit during this period, as they were often compared to other regions that focused more on quality than quantity.
+This is an important consideration, as the size of some of these vineyards and resultant capactiy to produce greater volumnes would only require only a minor increase to sale price to out compete other regions.
+It is important to consider the unique challenges of winegrwoing when compared to other agricultural industries, as vineyards are vertically integrated within the wine industry; tying them to secondary and tertiary industries, such as wine production, packaging, transport and sales.
+This results in unique issues and considerations for each vineyard, where on-the-ground decisions are influenced by other wine industry's choices, such as the use of sustainable practices in vineyards as a requirement for sale in overseas markets; notably these interactions can be further complicated by some winegrowers being completely integrated into a wine company, while others are not\citep{knightFirmResourcesDevelopment2019}.
+Incorporating decisions into the model could help describe the contributing factors to regional differences beyond resource consumption and regional differences but would require incredibly granular data.
+\subsection{Regional Differences}
+Some regions appeared to produce many low quality grapes at scale whilst others focussed on producing higher quality grapes in lower volumes.
+This behaviour can also be observed when reviewing Wine Australia's annual reports, where it is apparent that some GI regions, such as the Riverland, are known for producing large amounts of lower grade grapes (grapes of lower value per tonne) \cite{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
+Comparatively other regions, such as Tasmania, only produce high quality/grade grapes but in much smaller quantities than the Riverland.
+The difference in pricing per tonne between the lowest and highest graded grapes can be greater a hundred times the difference in value per tonne.
+% ref
+Not all regions target only one grade of grape, with some producing a variety of differently graded grapes; such as the Yarra Valley, which produces a variety of different grades of grapes, from C to A
+% ref
+\newline
+Some regions are known for their quality and may have a bias in purchasers or bring greater demand regardless of similarities and differences in production of quality of grapes \citep{hallidayAustralianWineEncyclopedia2009}.
+%The Barossa is an example of this, known for its quality could also lend itself to a bias in purchasers not considering other regions that may be capable of similar quality.
+This effect could stifle the potential for market opportunities within lesser known regions.
+A further possibility is the existence of regional upper limits on potential quality, making the relationship between resource use and value have diminishing returns.
+Climate was considered to be a large determinant of the ability to grow a larger quantity of grapes, as well as a determinant in grape quality \citep{agostaRegionalClimateVariability2012}
+
 % What is this below paragraph?
@@
 %
+different strategies are likely employed between different regions, where some regions target the mass production of cheaper grapes over quality.
+This is most notable when grouping regions by climate, especially when considering GI Regions in the 'Hot Very Dry' climate (see Figure 7).
+The effect of climate in the models was not more significant than the more granular use of GI regions.
+The interaction between year and GI Region likely accounted for localised events such as bushfires, which would be impactful, but only at a local level in both time and space.
+
+\subsection{Limitations}
+Limitations included overestimating yield for models 1 and 2, (see Figures 1 and 2) and underestimating crop value in models 3 and 4 (see Figures 3 and 4).The issue of model 1 and 2 over predicting yield, may have been due to preventative measures brought on by regional pressures such as fire, frost and disease.
+Where, more resources were required to prevent these issues from spreading within a region, thus disproportionately effecting some vineyards compared to others locally.
+This type of maintenance is not well captured especially when considering that some regions, those in warmer areas are not as prone to disease as cooler climates and could potentially have lower operating costs per hectare.
+This could create a discrepancy in vineyards that utilise preventative measures in wetter regions, as opposed to those who do not, and thus expend less fuel and energy but risk disease.
+When reviewing the differences between regions it is important to consider that vineyards in Hot Very Dry areas can be hundreds of times the size of those in other regions.
+It is interesting that while area, although significantly correlated to the ratio of yield to area, was still lower than water and about the same as emissions.
+This points to economies of scale playing a role but still being only one consideration alongside the potential resources that can be used.
+The negative trend between size and average sales price could also be a side effect of mass supply verse demand, especially when looking at the level of difference in production of some vineyards (see Table 4).
+The relationships between yield, value and area are not simply about efficiently producing the most grapes; sales price and by association grape quality, are integral to the profitability, and this is strongly linked to resource-use and thus the longevity and sustainability of a vineyard.
+
 In alternative attempts at models it was found that without the incorporation of GI Region or year the predictions greatly under performed.
@@
 %
-
 %Models were compared using $R^2$ values.
@@
 The introduction of known average prices per tonne also helped increase R2 values a small amount; it is important to not that it is common practice for wineries to purchase grapes at a regional average rate, likely resulting in much less variance within a region.
-
-different strategies are likely employed between different regions, where some regions target the mass production of cheaper grapes over quality.
-This is most notable when grouping regions by climate, especially when considering GI Regions in the 'Hot Very Dry' climate (see Figure 7).
-The effect of climate in the models was not more significant than the more granular use of GI regions.
-The interaction between year and GI Region likely accounted for localised events such as bushfires, which would be impactful, but only at a local level in both time and space.
-
-\subsection{Limitations}
-Limitations included overestimating yield for models 1 and 2, (see Figures 1 and 2) and underestimating crop value in models 3 and 4 (see Figures 3 and 4).
-% this needs serious revision.
-This study investigated the general relationships between input resources of a vineyard, including fuel and water, and the outputs including yield and value.
-Some regions appeared to produce many low quality grapes at scale compared to attempting to produce fewer higher quality grapes.
-This behaviour can be observed when reviewing Wine Australia's annual reports, where it is apparent that warm inland regions such as the Riverland are known to only produce large amounts of lower graded grapes \cite{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
-Comparatively, regions such as Tasmania only produce A grade grapes but in much smaller quantities than the Riverland.
-Knowing that the difference in pricing per tonne can exceed a magnitude of 10 between grades E and A, the operations in regions that target different grades would have varied priorities.
-However, some regions such as the Yarra Valley produce a Variety of different grades of grapes, from C to A, highlighting that vineyard priorities, although may be somewhat present within regional classifications, are not necessarily aligned within a given region.
-
-The opportunity to target different grades of grapes may not always be available, with some regions being more renowned than others, and likely to be sought after regardless \citep{hallidayAustralianWineEncyclopedia2009}.
-The Barossa is an example of this, known for its quality could also lend itself to a bias in purchasers not considering other regions that may be capable of similar quality.
-This effect could stifle the potential for market opportunities within these lesser known regions.
-A further possibility is that there may be regional upper limits with the relationship between resource input and the value gained becoming no longer proportional due to diminishing returns.
-Climate was considered to be a large determinant of the ability to grow a larger quantity of grapes, as well as a determinant in grape quality \citep{agostaRegionalClimateVariability2012}; however there were vineyards in similar regions that were able to produce exceptionally better results than others (See Figure 7).
-
-The issue of model 1 and 2 over predicting yield, may have been due to preventative measures brought on by regional pressures such as fire, frost and disease.
-Where, more resources were required to prevent these issues from spreading within a region, thus disproportionately effecting some vineyards compared to others locally.
-This type of maintenance is not well captured especially when considering that some regions, those in warmer areas are not as prone to disease as cooler climates and could potentially have lower operating costs per hectare.
-This could create a discrepancy in vineyards that utilise preventative measures in wetter regions, as opposed to those who do not, and thus expend less fuel and energy but risk disease.
-When reviewing the differences between regions it is important to consider that vineyards in Hot Very Dry areas can be hundreds of times the size of those in other regions.
-It is interesting that while area, although significantly correlated to the ratio of yield to area, was still lower than water and about the same as emissions.
-This points to economies of scale playing a role but still being only one consideration alongside the potential resources that can be used.
-The negative trend between size and average sales price could also be a side effect of mass supply verse demand, especially when looking at the level of difference in production of some vineyards (see Table 4).
-The relationships between yield, value and area are not simply about efficiently producing the most grapes; sales price and by association grape quality, are integral to the profitability, and this is strongly linked to resource-use and thus the longevity and sustainability of a vineyard.
-
-Literature shows that there are many on-the-ground decisions that influence both quality and yield.
-Where these decisions are governed by complex physical and social forces such as international market demands, disease pressures and natural disasters \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}.
-Many of these occurrences being highlighted throughout the past decades vintage reports \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
-%This definitely needs to be broken up, what were these occurences, which year was significant for what?
-It is also important to consider that these reports show that the warm inland regions have seen a decline in profit during this period, as they were often compared to other regions that focused more on quality than quantity.
-This is an important consideration, as the size of some of these vineyards when considering their ratio of value to area would only require a marginal increase to out compete other regions.
-There are also differences when comparing winegrowers to other agricultural industries as they are vertically integrated within the wine industry, tying them to secondary and tertiary industries, such as wine production, packaging, transport and sales.
-This results in unique issues and considerations for each vineyard, where these on-the-ground decisions may be influenced by other wine industry's choices, such as the use of sustainable practices in vineyards as a requirement for sale in overseas markets; notably these interactions are further complicated by some winegrowers being totally integrated into wine companies, while others are not (Knight et al., 2019).
-Incorporating such decisions into the model could help describe the contributing factors to regional differences beyond resource consumption and regional differences.
 
@@
  indicated that it is possible other vineyard activities requiring fuel are not as determining factors for a vineyards grape quality.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2023-08-01 article.tex
discussion edits!

@@
 \section{Discussion}
-The most determining factor in the output quality and yield of a vineyard was size.
-There was a strong relationship between size and resource use, in that the overall space of a vineyard and its access to resources determined the upper limit of potential yield.
-However, size was also inversely related to the potential quality, with higher quality being related more to high resource inputs per area than overall access to resources.
+There was an understandably strong relationship between size and resource use, with the overall space of a vineyard and its access to resources greatly determining the upper limit of potential yield.
+However, size was also inversely related to the potential quality, with higher quality being related to high resource inputs per area; rather than to the overall expenditure of resources.
 These effects were augmented by regional and yearly affects.
-Even given regional and yearly changes, there was a strong sense of smaller vineyards being easier to manage and ensure high quality when compared to larger vineyards.
+Even given regional and yearly changes, there was a strong connection between smaller vineyards and higher quality.
+This could have been due to the easier management of smaller properties.
 \subsection{Resource use and Yield verse Quality} 
-Literature shows that there are many on-the-ground decisions that influence both quality and yield.
-These decisions are governed by complex physical and social forces such as international market demands, disease pressures and natural disasters \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}; with many of these occurrences being highlighted throughout the past decades vintage reports \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
- %This definitely needs to be broken up, what were these occurences, which year was significant for what?
-It is also important to consider that these reports show that the warm inland regions have seen a decline in profit during this period, as they were often compared to other regions that focused more on quality than quantity.
-This is an important consideration, as the size of some of these vineyards and resultant capactiy to produce greater volumnes would only require only a minor increase to sale price to out compete other regions.
-It is important to consider the unique challenges of winegrwoing when compared to other agricultural industries, as vineyards are vertically integrated within the wine industry; tying them to secondary and tertiary industries, such as wine production, packaging, transport and sales.
-This results in unique issues and considerations for each vineyard, where on-the-ground decisions are influenced by other wine industry's choices, such as the use of sustainable practices in vineyards as a requirement for sale in overseas markets; notably these interactions can be further complicated by some winegrowers being completely integrated into a wine company, while others are not\citep{knightFirmResourcesDevelopment2019}.
-Incorporating decisions into the model could help describe the contributing factors to regional differences beyond resource consumption and regional differences but would require incredibly granular data.
+ There are many on-the-ground decisions that influence both quality and yield.
+Comparing the $R^2$ values between Models 2 and 4 showed that the average price per tonne of grapes described a great deal of the relationship between resource use and yield when variables were considered as ratios of area (due to the discrepancy in $R^2$ between the two models, see Table \ref{tab:modelsummary}).
+This descrepency is likely due to different vineyard prioritisation, which can be described by the type of quality and quantity a vineyard aims to target.
+Decisions such as the prioritisation of quality over quantity, are governed by complex physical and social forces, for example: international market demands, disease pressures and natural disasters \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}; with many of these occurrences being highlighted throughout the past decades vintage reports from Wine Australia \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
+It is also important to consider that these reports show that the warm inland regions have seen a decline in profit during this period, whereas regions targetting quality did not.
+Size becomes an important consideration, as it dictates the potential capacity to produce greater volumes of grapes.
+However, given the comparison of value per area, regions with larger vineyards and larger vineyards in general, tend to underperform.
+When considering the 'Hot Very Dry' vineyards (see Figure \ref{fig:yield_vs_value_area}) These vineyards would be very competitive with only a minor increase to sale price, possibly outperforming other regions.
+\newline
+The negative trend between size and average sales price could be a side effect of supply verse demand, especially when looking at the level of difference in production of some vineyards.
+Economies of scale likely played a role in determining yield but were only one consideration alongside resource use.
+Size was also less of a determining factor when considering quality.
+It is possible that the relationship of scope 1 emissions between yield and quality was closely tied to a vineyard's area; due to requiring more fuel to cover issues (such as fixing a broken irrigation pipe), where  a larger area has the potential for issues to be further away.
+This is further cemented when noting that most irrigation systems are diesel based, with water use being a significant variable in each model and scope 1 emissions not; scope one emissions' lack of significance and contribution given its F-statistics, could be indicative that other vineyard activities requiring fuel are not as determining factors for a vineyards grape quality.
+The relationship between yield, value and area was not simply about efficiently producing the most grapes; sales price and by association grape quality, are integral to the profitability, and this is strongly linked to resource-use and thus the longevity and sustainability of a vineyard.
+\newline
+There are important considerations unique to winegrowing compared to other agricultural industries.
+The vertical integration of winegrowing within the wine industry ties winegrowers to secondary and tertiary industries, such as wine production, packaging, transport and sales.
+This results in unique issues and considerations for each vineyard, where on-the-ground decisions are influenced by other wine industry's choices, such as the use of sustainable practices in vineyards as a requirement for sale in overseas markets; notably these interactions can be further complicated by some winegrowers being completely integrated into a wine company, while others are not \citep{knightFirmResourcesDevelopment2019}.
+Incorporating decisions into the model could help describe the contributing factors to regional differences beyond resource consumption and regional differences but would require incredibly granular data and more sophisticated modelling.
 \subsection{Regional Differences}
 Some regions appeared to produce many low quality grapes at scale whilst others focussed on producing higher quality grapes in lower volumes.
-This behaviour can also be observed when reviewing Wine Australia's annual reports, where it is apparent that some GI regions, such as the Riverland, are known for producing large amounts of lower grade grapes (grapes of lower value per tonne) \cite{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
-Comparatively other regions, such as Tasmania, only produce high quality/grade grapes but in much smaller quantities than the Riverland.
-The difference in pricing per tonne between the lowest and highest graded grapes can be greater a hundred times the difference in value per tonne.
+This behaviour can also be observed when reviewing Wine Australia's annual reports, where it is apparent that some GI regions, such as the Riverland, are known for producing large amounts of lower grade (low value per tonne) grapes \cite{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
+Comparatively other regions, such as Tasmania, only produce high quality/grade grapes but in smaller quantities.
+The difference in pricing per tonne between the lowest and highest graded grapes can be greater than a hundred times the difference in value per tonne.
 % ref
-Not all regions target only one grade of grape, with some producing a variety of differently graded grapes; such as the Yarra Valley, which produces a variety of different grades of grapes, from C to A
+Not all regions target only one grade of grape, with some producing a variety of differently graded grapes; such as the Yarra Valley, which produces grades from C to A.
 % ref
@@
 This effect could stifle the potential for market opportunities within lesser known regions.
-A further possibility is the existence of regional upper limits on potential quality, making the relationship between resource use and value have diminishing returns.
-Climate was considered to be a large determinant of the ability to grow a larger quantity of grapes, as well as a determinant in grape quality \citep{agostaRegionalClimateVariability2012}
-
+A further possibility is the existence of regional upper limits on potential quality, or that there are diminishing returns in some regions when pursuing quality or quantity; however these types of relationships may be obfuscated by knowledgeable winegrowers who avoid this pitfall.
 % What is this below paragraph?
@@
 %
-different strategies are likely employed between different regions, where some regions target the mass production of cheaper grapes over quality.
-This is most notable when grouping regions by climate, especially when considering GI Regions in the 'Hot Very Dry' climate (see Figure 7).
-The effect of climate in the models was not more significant than the more granular use of GI regions.
-The interaction between year and GI Region likely accounted for localised events such as bushfires, which would be impactful, but only at a local level in both time and space.
+\newline
+Due to regional differences, different strategies are likely employed across different regions; such as some regions targeting mass production over quality.
+This is most notable when grouping regions by climate, especially when considering GI Regions in the 'Hot Very Dry' climate (see Figure \ref{fig:yield_vs_value}).
+In alternative attempts at models it was found that without the direct incorporation of GI Region or year, predictions greatly under performed.
+The effect of climate in the models was never as significant as the more granular GI regions, and always led to less accurate models.
+Although not chosen over GI region, climate was considered to be a large determinant of the ability to produce larger quantities of grapes, as well as a determinant in grape quality \citep{agostaRegionalClimateVariability2012}.
+The more granular GI Region likely explained a broader mix of geographical phenomenon, such as soil, geology and access to water resources \citep{abbalDecisionSupportSystem2016,carmonaUseParticipatoryObjectOriented2011}.
+The interaction between year and GI Region likely accounted for events such as bushfires, which would be impactful, but only at a local level, both in time and space.
 
 \subsection{Limitations}
-Limitations included overestimating yield for models 1 and 2, (see Figures 1 and 2) and underestimating crop value in models 3 and 4 (see Figures 3 and 4).The issue of model 1 and 2 over predicting yield, may have been due to preventative measures brought on by regional pressures such as fire, frost and disease.
+Limitations included overestimating yield for models 1 and 2, and underestimating crop value in models 3 and 4 (see appendix).
+The issue of model 1 and 2 over predicting yield, may have been due to preventative measures brought on by regional pressures such as fire, frost and disease.
 Where, more resources were required to prevent these issues from spreading within a region, thus disproportionately effecting some vineyards compared to others locally.
-This type of maintenance is not well captured especially when considering that some regions, those in warmer areas are not as prone to disease as cooler climates and could potentially have lower operating costs per hectare.
-This could create a discrepancy in vineyards that utilise preventative measures in wetter regions, as opposed to those who do not, and thus expend less fuel and energy but risk disease.
-When reviewing the differences between regions it is important to consider that vineyards in Hot Very Dry areas can be hundreds of times the size of those in other regions.
-It is interesting that while area, although significantly correlated to the ratio of yield to area, was still lower than water and about the same as emissions.
-This points to economies of scale playing a role but still being only one consideration alongside the potential resources that can be used.
-The negative trend between size and average sales price could also be a side effect of mass supply verse demand, especially when looking at the level of difference in production of some vineyards (see Table 4).
-The relationships between yield, value and area are not simply about efficiently producing the most grapes; sales price and by association grape quality, are integral to the profitability, and this is strongly linked to resource-use and thus the longevity and sustainability of a vineyard.
-
-In alternative attempts at models it was found that without the incorporation of GI Region or year the predictions greatly under performed.
-The possible reason behind this effect was that different strategies are likely employed between different regions, where some regions target the mass production of cheaper grapes over quality.
+This type of maintenance is not well captured especially when considering that some regions, especially those in warmer areas, are not as prone to disease as cooler climates and could potentially have lower operating costs per hectare.
+This could create a discrepancy in vineyards that utilised preventative measures in wetter regions, as opposed to those that did not, thus expending less fuel and energy but risking disease.
+When reviewing the differences between regions it is important to consider that vineyards in 'Hot Very Dry' areas can be hundreds of times the size of those in other regions.
+This limitation could be overcome by incorporating the profitability of vineyards, compare the financial success of working at different operational scales.
+\newline
+ %The possible reason behind this effect was that different strategies are likely employed between different regions, where some regions target the mass production of cheaper grapes over quality.
 % ALSO THAT CATASTROPHES HAPPEN EVEN ON SMALL SCALES THis needs to be tied up better.
-%
-%Models were compared using $R^2$ values.
-The comparison of models performance showed that the average price per tonne of grapes describes a great deal of the relationship between predictor and response when comparing Model 2 to Model 4
-% Graph of productivity verse value.
-%(see Table \ref*{tab:modelperformance}). 
-% This is important but you do not illustrate why
-%
-% This section needs to be heavily revised. Why are there tails on the distributions? What does a heavy tail indicate?
-% What does it look like with respect to disaster events?
-% split the discussion sections out of the results!! Even the questions I am asking are answered in the discussion, however they need to be highlighted as aspects of the models here beforehand.
-Reviewing the data to uncover reasons for this included the use of binary variables such as the utilisation of renewable energy, contractors, and the occurrence of disease, fire and frost; however none of these variables were able to explain why some vineyards produced less, or why other vineyards sold at higher prices than predicted.
-A wide variety of these influences were likely already explained within the use of year and GI Region, or the interaction of both variables.
-The change between some regions was dramatic, with particularly warmer and drier regions producing much higher volumes of grapes at lower prices (See Figures 5 and 6).
-The use of other variables and methods, specifically splines, were able to create a more normally distributed set of residuals but at a drastically reduced accuracy when comparing R2 and RSE.
-The introduction of known average prices per tonne also helped increase R2 values a small amount; it is important to not that it is common practice for wineries to purchase grapes at a regional average rate, likely resulting in much less variance within a region.
-
+%Reviewing data to uncover reasons for the disparity between production and quality of similar vineyards  included the use of 
+Variables such as the utilisation of renewable energy, contractors, and the occurrence of disease, fire and frost were originally explored to capture the discrepancies between similar vineyards that produced different  yields and crop values.
+However, none of these variables were significantly connected to the response variables, and did not add to model accuracy; even when considered as interactions.
+The use of other methods, specifically splines, resulted in more normally distributed residuals but at a drastically reduced overall accuracy when comparing $R^2$ and Residual Square Error.
+Attempts to fully explain small variations was always overshadowed by the dramatic differences in regional trends.
+\newline
 Having more data for each region would also be an improvement, allowing greater comparison between regions.
 More variables may also help to discern vineyards that can produce larger volumes of grapes at higher prices.
-The use of semi transparent tools such as random forests and decision trees alongside more variables and data may help to uncover the reasons for values that were under or over estimated.
+The use of semi transparent tools such as random forests and decision trees alongside more variables and data may help to uncover the reasons for values that were under or overestimated.
 These differences could be caused by the use of alternative sustainable practices in the field.
-While there is evidence to suggest that environmentally sustainable practices can reduce costs, increase efficiency, whilst improving the quality of grapes, more research is needed to link these benefits across different regions and climates \citep{baianoOverviewSustainabilityWine2021,marianiSustainableWinegrowingCurrent2015,montalvo-falconSustainabilityResearchWine2023}.
+And, while there is evidence to suggest that environmentally sustainable practices can reduce costs, increase efficiency, whilst improving the quality of grapes; more research is needed to link these benefits across different regions and climates \citep{baianoOverviewSustainabilityWine2021,marianiSustainableWinegrowingCurrent2015,montalvo-falconSustainabilityResearchWine2023}.
+\section{Conclusion}
+% This belongs in the discussion
 
-% This belongs in the discussion
-The relationship between scope one emissions and the response variables that included average sales price
-
-It is possible that the relationships between scope one emissions and the response variables were closely tied to a vineyards area. This possibility could be explained through the emissions
-
-Noting that irrigation systems use fuel
-and that the application of water was a significant variable in each model
-scope one emissions' lack of significance and contribution given its F-statistics (See Tables 7 and 8),
- indicated that it is possible other vineyard activities requiring fuel are not as determining factors for a vineyards grape quality.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 

### 2023-08-01 article.tex
draft conclusion

@@
 % This belongs in the discussion
-
+In summary, vineyard yield and crop value is well-defined by the resources used.
+However, it is important to consider a vineyard's business goal, region, external pressures and economies of scale.
+Where, larger vineyards are likely to produce greater overall yields, and have higher yield per area.
+Smaller vineyards are likely to produce more value per area, and a higher quality of grape.
+It is likely that regional constraints also contribute to the best strategy to pursue when considering quality or quantity.
 %

### 2023-08-01 article.tex
abstract

@@
 \begin{abstract}
-%% Text of abstract
+Using a ten year dataset that spans Australia we describe the relationships of input resources to the output yield and quality of grapes for wine production.
+Creating a baseline for comparing the efficiency of winegrowers with regard to the tradeoff between invested resources, yield and quality.
+Showing that economies of scale prevail for optimising crops for yield, with larger crops producing more grapes per area.
+And, showing that smaller crops utilising less resources are more closely related to quality.
 \end{abstract}

### 2023-08-02 article.tex
More Andrew edits

@@
 %% \fntext[label3]{}
-\title{An exploratory analysis of the influence of resource use on the yield verse quality trade-off in Australian vineyards}
+\title{An exploratory analysis of the influence of resource use on the yield versus quality trade-off in Australian vineyards}
 %% use optional labels to link authors explicitly to addresses:
@@
 The practical addition of this aim is a baseline for comparison - given a vineyard within Australia, one could extrapolate their comparative efficiency with regard to the tradeoff between invested resources, yield and quality.
-In achieving this we will also confirm the existence of a yield verse quality trade off within Australian winegrowing; one not prior confirmed explicitly across such varying regions, scales and climates.
+In achieving this we will also confirm the existence of a yield versus quality trade off within Australian winegrowing; one not prior confirmed explicitly across such varying regions, scales and climates.
 \section{Methods}
@@
 %
-\subsubsection{Model Comparisons: Yield Verse Value}
+\subsubsection{Model comparisons: yield versus value}
 %
@@
 However, with area coming into play in Figure \ref{fig:yield_vs_value_area}, many data points are scaled differently; specifically the vineyards from 'Hot' regions change to be found the bottom right tail end, indicating the production of large quantity of lower value grapes.
-An un obvious difference between the Figures, is that a large amount of the difference can be explained by rotation (being $90^\circ$ clockwise from Figure \ref{fig:yield_vs_value} to \ref{fig:yield_vs_value_area}).
+An inconspicuous difference between the Figures, is that a large amount of the difference can be explained by rotation (being $90^\circ$ clockwise from Figure \ref{fig:yield_vs_value} to \ref{fig:yield_vs_value_area}).
 This is more visible when comparing both graphs to the map of regional averages for response variables, see Figure \ref{fig:map}.
-There is a notable change between regional averages when looking at yield verse value.
+There is a notable change between regional averages when looking at yield versus value.
 Through the coefficients we can deduce that: this difference is also a difference between more resources used for the raw response variables; and a difference between overall resource use and the size of the vineyard when considering the response variables as a ratio to area.
@@
 \section{Discussion}
-There was an understandably strong relationship between size and resource use, with the overall space of a vineyard and its access to resources greatly determining the upper limit of potential yield.
+There was an expected strong relationship between size and resource use, with the overall space of a vineyard and its access to resources greatly determining the upper limit of potential yield.
 However, size was also inversely related to the potential quality, with higher quality being related to high resource inputs per area; rather than to the overall expenditure of resources.
-These effects were augmented by regional and yearly affects.
+Vineyard outputs were also augmented by regional and yearly affects.
 Even given regional and yearly changes, there was a strong connection between smaller vineyards and higher quality.
 This could have been due to the easier management of smaller properties.
-\subsection{Resource use and Yield verse Quality} 
+\subsection{Resource use and yield versus quality} 
  There are many on-the-ground decisions that influence both quality and yield.
@@
 Size becomes an important consideration, as it dictates the potential capacity to produce greater volumes of grapes.
-However, given the comparison of value per area, regions with larger vineyards and larger vineyards in general, tend to underperform.
+However, given the comparison of value per area, regions with larger vineyards (such as warmer in land regions) and larger vineyards in general, tend to underperform.
 When considering the 'Hot Very Dry' vineyards (see Figure \ref{fig:yield_vs_value_area}) These vineyards would be very competitive with only a minor increase to sale price, possibly outperforming other regions.
 \newline
-The negative trend between size and average sales price could be a side effect of supply verse demand, especially when looking at the level of difference in production of some vineyards.
+The negative trend between size and average sales price could be a side effect of supply versus demand, especially when looking at the level of difference in production of some vineyards.
 Economies of scale likely played a role in determining yield but were only one consideration alongside resource use.
@@
 This behaviour can also be observed when reviewing Wine Australia's annual reports, where it is apparent that some GI regions, such as the Riverland, are known for producing large amounts of lower grade (low value per tonne) grapes \cite{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
-Comparatively other regions, such as Tasmania, only produce high quality/grade grapes but in smaller quantities.
+Comparatively other regions, such as Tasmania, only produce high quality grapes but in smaller quantities.
 The difference in pricing per tonne between the lowest and highest graded grapes can be greater than a hundred times the difference in value per tonne.

### 2023-08-02 article.tex
Added highlights
fixed abstrat

@@
   country={}}
-\author[label1,label2,label3]{Bryce Polley}
+\author[label1,label2,label3]{Author}
 \date{20/06/2023}
 \begin{abstract}
-Using a ten year dataset that spans Australia we describe the relationships of input resources to the output yield and quality of grapes for wine production.
-Creating a baseline for comparing the efficiency of winegrowers with regard to the tradeoff between invested resources, yield and quality.
-Showing that economies of scale prevail for optimising crops for yield, with larger crops producing more grapes per area.
-And, showing that smaller crops utilising less resources are more closely related to quality.
+ The global focus on sustainability in agronomic industries has changed the way in which these enterprises do business.
+When strategies for a sustainable winegrowing industry are assessed, there is a trade-off between balancing the amount of resources invested and the resultant yield verses quality produced.
+In this analysis we observe relationships between yield and quality through the use of linear models.
+An extensive amount of research into a variety of factors' effect on grape quality and yield exists; but due to the lack of long-term and in-depth data, individual effects are often studied in isolation.
+The dataset used for this analysis includes data collected for the past 10 years from a multitude of vineyards located over a diverse range of Australian winegrowing regions.
+We aim to use this broad dataset to describe the relationship of input resources to the output yield and quality of vineyards.
+There was an expected strong relationship between size and resource use, with the overall space of a vineyard and its access to resources greatly determining the upper limit of potential yield.
+However, size was also negatively related to the potential quality, with higher quality being connected to high resource inputs per area; rather than to the overall expenditure of resources.
+Vineyard outputs were also augmented by regional and yearly affects.
+It is important to also consider a vineyard's business goal, region, external pressures and economies of scale.
+With regional constraints also contributing to deciding the best strategies to pursue when considering quality or quantity.
 \end{abstract}
@@
 \begin{highlights}
-  \item Research highlight 1
-  \item Research highlight 2
+  \item Comparative analysis of resource use, quality and quantity in Australian winegrowing.
+  \item Regional comparison of outcomes and resource use in Australian winegrowing regions.
+  \item Baseline models for comparing wine crops.
+  \item Analysis of national, decade long data source.
 \end{highlights}

### 2023-08-02 article.tex
Fixed error with figure references

@@
 %
-\begin{figure}\label{fig:violin}
+\begin{figure}
   \includegraphics{violinplot.pdf}
-  \caption{Violin plots of GI Region and Year coefficients for each model.}
+  \caption{Violin plots of GI Region and Year coefficients for each model.}\label{fig:violin}
   \end{figure}
@@
 %
-\begin{figure}\label{fig:yield_vs_value}
+\begin{figure}
   \resizebox{\textwidth}{!}{
   \includegraphics{yield_verse_value.pdf}}
-  \caption{Scatter plot of vineyard yield against the product of yield and average price per tonne. The axes are in standard deviations with points coloured by climate.}
+  \caption{Scatter plot of vineyard yield against the product of yield and average price per tonne. The axes are in standard deviations with points coloured by climate.}\label{fig:yield_vs_value}
 \end{figure}
 %
-\begin{figure}\label{fig:yield_vs_value_area}
+\begin{figure}
   \resizebox{\textwidth}{!}{
   \includegraphics{yield_verse_value_by_area.pdf}}
-    \caption{Scatter plot of vineyard yield against the product of yield and average price per tonne as ratios to area harvested. The axes are in standard deviations with points coloured by climate.}
+    \caption{Scatter plot of vineyard yield against the product of yield and average price per tonne as ratios to area harvested.
+The axes are in standard deviations with points coloured by climate.}\label{fig:yield_vs_value_area}
 \end{figure}
 %
-\begin{figure}\label{fig:map}
+\begin{figure}
   \resizebox{\textwidth}{!}{
   \includegraphics{my_map.pdf}}
-  \caption{Map of regional average yield and value per hectare.}
+  \caption{Map of regional average yield and value per hectare.}\label{fig:map}
 \end{figure}

### 2023-08-02 analysis.r
Holy shit the maps work

@@ -367,7 +367,7 @@ dev.off()
 # Lets make some maps #
 library(terra)
 nt <- read.csv("no_trans.csv")
-v <- vect("data/Wine_Geographical_Indications_Australia_2022/Wine_Geographical_Indications_Australia_2022.shp")
+v <- st_read("data/Wine_Geographical_Indications_Australia_2022/Wine_Geographical_Indications_Australia_2022.shp")
 aus <- vect("data/STE_2021_AUST_SHP_GDA2020/STE_2021_AUST_GDA2020.shp")
 
 v$ha_tonnes_grapes_harvested <- 0
@@ -387,12 +387,12 @@ aggs.ha_value <- aggregate(
 
 for (region in unique(df$giregion)){
     v[v$GI_NAME == region
-    , ]$ha_tonnes_grapes_harvested <- aggs.ha_tonnes_grapes_harvested[
+    , "ha_tonnes_grapes_harvested"] <- aggs.ha_tonnes_grapes_harvested[
     aggs.ha_tonnes_grapes_harvested$Group.1 == region
     , ]$x
     v[
         v$GI_NAME==region
-        , ]$ha_value <- aggs.ha_value[
+        , "ha_value"] <- aggs.ha_value[
             aggs.ha_value$Group.1 == region
             , ]$x
 }
@@ -416,6 +416,17 @@ sbar(1000
     , label = "1000km"
     , xy = "bottomright")
 
+value_map <- ggplot() +
+    geom_sf(data = v
+        , aes(fill = ha_value)) +
+    geom_sf(data=aus, fill = NA) +
+    xlim(113, 154) +
+    ylim(-45, -25) +
+    scale_fill_gradientn(colours=rev(viridis(3)),
+                         name="Average Sale Price",
+                         na.value = "grey100") +
+    theme_void()
+
 map_one <- plot(v
     , "ha_tonnes_grapes_harvested"
     , asp = 1
@@ -423,10 +434,10 @@ map_one <- plot(v
    , ylim = c(-45, -25)
    , legend = "bottomleft"
    , main = "Average Yield per Hectare"
-)
+) +
 lines(aus,
    , xlim = c(113, 154)
-   , ylim = c(-45, -25))
+   , ylim = c(-45, -25)) +
     sbar(1000
     , lonlat = TRUE
     , label = "1000km"

### 2023-08-02 analysis.r
Maps finished

@@ -398,51 +398,32 @@ for (region in unique(df$giregion)){
 }
 
 # note these are the averages.
-pdf("my_map.pdf")
-
-plot(v
-    , "ha_value"
-    , asp = 1
-   , xlim = c(113, 154)
-   , ylim = c(-45, -25)
-   , legend = "bottomleft"
-   , main = "Average Value per Hectare"
-)
-lines(aus,
-   , xlim = c(113, 154)
-   , ylim = c(-45, -25))
-sbar(1000
-    , lonlat = TRUE
-    , label = "1000km"
-    , xy = "bottomright")
-
 value_map <- ggplot() +
+        ggtitle("Average Price Per Ha ") +
     geom_sf(data = v
         , aes(fill = ha_value)) +
     geom_sf(data=aus, fill = NA) +
     xlim(113, 154) +
     ylim(-45, -25) +
-    scale_fill_gradientn(colours=rev(viridis(3)),
-                         name="Average Sale Price",
-                         na.value = "grey100") +
+    scale_fill_gradientn(colours=rev(viridis(3))
+                         , name=""
+                         , na.value = "grey100") +
     theme_void()
 
-map_one <- plot(v
-    , "ha_tonnes_grapes_harvested"
-    , asp = 1
-   , xlim = c(113, 154)
-   , ylim = c(-45, -25)
-   , legend = "bottomleft"
-   , main = "Average Yield per Hectare"
-) +
-lines(aus,
-   , xlim = c(113, 154)
-   , ylim = c(-45, -25)) +
-    sbar(1000
-    , lonlat = TRUE
-    , label = "1000km"
-    , xy = "bottomright")
+yield_map <-  ggplot() +
+        ggtitle("Average Yield (tonnes) Per Ha") +
+    geom_sf(data = v
+        , aes(fill = ha_tonnes_grapes_harvested)) +
+    geom_sf(data=aus, fill = NA) +
+    xlim(113, 154) +
+    ylim(-45, -25) +
+    scale_fill_gradientn(colours=rev(viridis(3))
+                         , name=""
+                         , na.value = "grey100") +
+    theme_void()
 
+pdf("my_map.pdf")
+grid.arrange(value_map, yield_map, ncol=2)
 dev.off()
 
 # below is just a way to grab the coef from a linear model.

### 2023-08-02 article.tex
Draft to submit to Kerrie & Kate

@@
 %%             country={}}
-\affiliation[label1]{organization={QUT},
-  addressline={},
-  city={},
-  postcode={},
-  state={QLD},
-  country={}}
-\affiliation[label2]{organization={AWRI},
-  addressline={},
-  city={},
-  postcode={},
-  state={SA},
-  country={}}
-\affiliation[label3]{organization={Food Agility CRC},
-  addressline={},
-  city={},
-  postcode={},
-  state={Vic},
-  country={}}
+%\affiliation[label1]{organization={QUT},
+%  addressline={},
+%  city={},
+%  postcode={},
+%  state={QLD},
+%  country={}}
+%\affiliation[label2]{organization={AWRI},
+%  addressline={},
+%  city={},
+%  postcode={},
+%  state={SA},
+%  country={}}
+%\affiliation[label3]{organization={Food Agility CRC},
+%  addressline={},
+%  city={},
+%  postcode={},
+%  state={Vic},
+%  country={}}
 \author[label1,label2,label3]{Author}
-\date{20/06/2023}
+\date{02/08/2023}
 \begin{abstract}
@@
 %%Graphical abstract`
-\begin{graphicalabstract}
-  \includegraphics{graphical_abstract.jpeg}
-\end{graphicalabstract}'
+%\begin{graphicalabstract}
+ % \includegraphics{graphical_abstract.jpeg}
+%\end{graphicalabstract}'
 
-\begin{keyword}
+%\begin{keyword}
 %% keywords here, in the form: keyword \sep keyword
-Keyword one \sep{} keyword two
+%Keyword one \sep{} keyword two
 %% PACS codes here, in the form: \PACS code \sep code
-\PACS{} 0000 \sep{} 1111
+%\PACS{} 0000 \sep{} 1111
 %% MSC codes here, in the form: \MSC code \sep code
 %% or \MSC[2008] code \sep code (2000 is the default)
-\MSC{} 0000 \sep{} 1111
-\end{keyword}
+%\MSC{} 0000 \sep{} 1111
+%\end{keyword}
 %%Research highlights

### 2023-08-02 analysis.r

@@ -1,5 +1,8 @@
 library(ggplot2)
 library(gridExtra)
+library(viridis)
+library(sf)
+library(cowplot)
 
 # data_prep.py has to be executed in order for df.csv to exist
 set.seed(100)
@@ -365,10 +368,9 @@ dev.off()
 
 #######################
 # Lets make some maps #
-library(terra)
 nt <- read.csv("no_trans.csv")
 v <- st_read("data/Wine_Geographical_Indications_Australia_2022/Wine_Geographical_Indications_Australia_2022.shp")
-aus <- vect("data/STE_2021_AUST_SHP_GDA2020/STE_2021_AUST_GDA2020.shp")
+aus <- st_read("data/STE_2021_AUST_SHP_GDA2020/STE_2021_AUST_GDA2020.shp")
 
 v$ha_tonnes_grapes_harvested <- 0
 v$ha_value <- 0
@@ -398,10 +400,10 @@ for (region in unique(df$giregion)){
 }
 
 # note these are the averages.
+#ggdraw() + 
 value_map <- ggplot() +
-        ggtitle("Average Price Per Ha ") +
-    geom_sf(data = v
-        , aes(fill = ha_value)) +
+    ggtitle("Average Price Per Ha ") +
+    geom_sf(data = v, aes(fill = ha_value)) +
     geom_sf(data=aus, fill = NA) +
     xlim(113, 154) +
     ylim(-45, -25) +
@@ -411,19 +413,19 @@ value_map <- ggplot() +
     theme_void()
 
 yield_map <-  ggplot() +
+    xlim(113, 154) +
+    ylim(-45, -25) +
+    theme_void() +
         ggtitle("Average Yield (tonnes) Per Ha") +
     geom_sf(data = v
         , aes(fill = ha_tonnes_grapes_harvested)) +
     geom_sf(data=aus, fill = NA) +
-    xlim(113, 154) +
-    ylim(-45, -25) +
     scale_fill_gradientn(colours=rev(viridis(3))
                          , name=""
-                         , na.value = "grey100") +
-    theme_void()
+                         , na.value = "grey100")
 
 pdf("my_map.pdf")
-grid.arrange(value_map, yield_map, ncol=2)
+grid.arrange(value_map, yield_map, ncol=2, nrow=1)
 dev.off()
 
 # below is just a way to grab the coef from a linear model.

### 2023-08-14 analysis_no_area.r
Models without area

@@ -1,0 +1,456 @@
+library(ggplot2)
+library(gridExtra)
+library(viridis)
+library(sf)
+library(cowplot)
+
+# data_prep.py has to be executed in order for df.csv to exist
+set.seed(100)
+df <- read.csv("df.csv")
+
+model1 <- lm(
+    tonnes_grapes_harvested ~
+    data_year_id + giregion + water_used + scope1
+    , data = df)
+model2 <- lm(
+    ha_tonnes_grapes_harvested ~
+     ((scope1) + water_used) + data_year_id:giregion
+    , data = df)
+model3 <- lm(
+    value ~
+    data_year_id + giregion + water_used + scope1
+    , data = df)
+model4 <- lm(
+    ha_value ~
+     ((scope1) + water_used) + data_year_id:giregion
+    , data = df)
+
+#Model 1
+#Residual standard error: 0.4381 on 5232 degrees of freedom
+#  (770 observations deleted due to missingness)
+#Multiple R-squared:  0.8104,    Adjusted R-squared:  0.8081 
+#F-statistic: 344.1 on 65 and 5232 DF,  p-value: < 2.2e-16
+
+#Model 2
+#Residual standard error: 0.5005 on 4871 degrees of freedom
+#Multiple R-squared:  0.7697,    Adjusted R-squared:  0.7495 
+#F-statistic:  38.2 on 426 and 4871 DF,  p-value: < 2.2e-16
+
+#Model 3
+#Residual standard error: 0.4343 on 2819 degrees of freedom
+#Multiple R-squared:  0.8152,    Adjusted R-squared:  0.8114 
+#F-statistic: 214.4 on 58 and 2819 DF,  p-value: < 2.2e-16
+
+#Model 4
+#Residual standard error: 0.37 on 2632 degrees of freedom
+#Multiple R-squared:  0.8748,    Adjusted R-squared:  0.8631 
+#F-statistic: 75.03 on 245 and 2632 DF,  p-value: < 2.2e-16
+
+# We review our first model
+
+# MODEL 1
+anova(model1)
+summary(model1)
+model1_residuals <- residuals(model1)
+model1_fitted <- fitted(model1)
+ggplot(
+    data.frame(
+        y = model1_residuals)
+        , aes(sample = y)
+       ) +
+       stat_qq() +
+       stat_qq_line()
+
+ggplot(
+    data.frame(
+        y = model1_residuals
+        , x = model1_fitted)
+    , aes(x = x, y = y)
+    ) +
+    geom_point() +
[…]

### 2023-08-14 article.tex
surpressed the appendix for ease
abstract edits

@@
 %% \fntext[label3]{}
-\title{An exploratory analysis of the influence of resource use on the yield versus quality trade-off in Australian vineyards}
+\title{The influence of resource use on yield versus quality trade-off in Australian vineyards}
 %% use optional labels to link authors explicitly to addresses:
@@
 \begin{abstract}
- The global focus on sustainability in agronomic industries has changed the way in which these enterprises do business.
-When strategies for a sustainable winegrowing industry are assessed, there is a trade-off between balancing the amount of resources invested and the resultant yield verses quality produced.
-In this analysis we observe relationships between yield and quality through the use of linear models.
-An extensive amount of research into a variety of factors' effect on grape quality and yield exists; but due to the lack of long-term and in-depth data, individual effects are often studied in isolation.
+ When strategies for a sustainable winegrowing industry are assessed, there is a trade-off between balancing the amount of resources invested and the resultant yield and quality of the produce.
+In this analysis we observe relationships between resource use, yield and quality through the use of statistical models.
 The dataset used for this analysis includes data collected for the past 10 years from a multitude of vineyards located over a diverse range of Australian winegrowing regions.
-We aim to use this broad dataset to describe the relationship of input resources to the output yield and quality of vineyards.
-There was an expected strong relationship between size and resource use, with the overall space of a vineyard and its access to resources greatly determining the upper limit of potential yield.
+Yield and quality (measured in sale price) was modelled to resource factors related to water usage and emissions.
+The analysis confirmed an expected strong relationship between size and resource use, with the overall space of a vineyard and its access to resources greatly determining the upper limit of potential yield.
 However, size was also negatively related to the potential quality, with higher quality being connected to high resource inputs per area; rather than to the overall expenditure of resources.
@@
 
+\if{false}
  \begin{table}[]
@@
 
+\fi
 %

### 2023-08-14 article.tex
missed abstract edits

@@
 However, size was also negatively related to the potential quality, with higher quality being connected to high resource inputs per area; rather than to the overall expenditure of resources.
-Vineyard outputs were also augmented by regional and yearly affects.
-It is important to also consider a vineyard's business goal, region, external pressures and economies of scale.
-With regional constraints also contributing to deciding the best strategies to pursue when considering quality or quantity.
+Regional and yearly effects on Vineyard outputs were also identified.
+Overall, the analysis highlighted the importance of considering a vineyard's business goal, region, external pressures and economies of scale, with regional constraints also contributing to deciding the best resource use strategies to pursue when considering quality or quantity.
 \end{abstract}

### 2023-08-14 article.tex
introduction

@@
 The global focus on sustainability in agronomic industries has changed the way in which these enterprises do business.
-When strategies for a sustainable winegrowing industry are assessed, there is a trade-off between balancing the amount of resources invested and the resultant yield verses quality produced.
-This dilemma exists across agriculture through shared fundamental considerations such as water use and nitrogen levels \citep{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016, zhuEffectsNitrogenLevel2017}.
-Quality in viticulture (the cultivation of grapes for wine production) is driven through its integration within the wine industry; with a wine's potential quality being initially defined through the chemical makeup of the grapes used in its production.
+When strategies for a sustainable winegrowing industry are assessed, there is a trade-off between balancing the amount of resources invested and the resultant yield versus quality produced.
+This dilemma exists across agriculture through shared fundamental considerations such as water use and fuel usage \citep{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016, zhuEffectsNitrogenLevel2017}.
+Quality in viticulture (the cultivation of grapes for wine production) is driven through its integration within the wine industry, with the potential quality of a wine being initially defined through the chemical makeup of the grapes used in its production.
 The consideration of sustainability within viticulture is further complicated by environmental and socio-demographic pressures.
-In the Australian context, these include: biosecurity, climate and international market demands.
+In the Australian context, these include biosecurity, climate and international market demands.
 \newline
-In this analysis we observe relationships between yield and quality through the use of linear models.
-An extensive amount of research into a variety of factors' effect on grape quality and yield exists; but due to the lack of long-term and in-depth data, individual effects are often studied in isolation \citep{abbalDecisionSupportSystem2016}.
+There is an extensive amount of research into the effects of a variety of factors on grape quality and yield
+% refeerences
+.
+However, due to the lack of long-term and in-depth data, individual factors are often studied in isolation \citep{abbalDecisionSupportSystem2016}.
 The lack of consolidated datasets also restricts the ability to gain statistical insights at large scales and across multiple regions \citep{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
 The dataset used for this analysis includes data collected for the past 10 years from a multitude of vineyards located over a diverse range of Australian winegrowing regions.
-\newline
-We aim to use this broad dataset to describe the relationship of input resources to the output yield and quality of vineyards.
-The practical addition of this aim is a baseline for comparison - given a vineyard within Australia, one could extrapolate their comparative efficiency with regard to the tradeoff between invested resources, yield and quality.
-In achieving this we will also confirm the existence of a yield versus quality trade off within Australian winegrowing; one not prior confirmed explicitly across such varying regions, scales and climates.
+We aim to use this dataset to describe the relationship of resources related to water and fuel use with the output yield and quality of the resultant product, taking into account the size and location of the vineyard.
+The practical addition of this aim is a baseline for comparison: given a vineyard within Australia, one could estimate the comparative efficiency with regard to the tradeoff between invested resources, yield and quality.
+This is the first time that such a trade off has been confirmed explicitly across such varying regions, scales and climates in the Australian winegrowing industry.
 \section{Methods}

### 2023-08-14 article.tex
discussion

@@
           & \textbf{Response} & \textbf{Predictors} & \textbf{Covariates} & \textbf{Interactions} \\ \midrule
-          \textbf{Model 1} & Yield & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope 1 Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & N/A \\
+          \textbf{Model 1} & Yield & \begin{tabular}[c]{@{}c@{}}Water Used\\ scope one Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & N/A \\
           \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-          \textbf{Model 2} & ${ \textrm{Yield}}\over{ \textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope 1 Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * Scope 1 Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular} \\
+          \textbf{Model 2} & ${ \textrm{Yield}}\over{ \textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ scope one Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * scope one Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular} \\
           \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-          \textbf{Model 3} & {$\textrm{Yield} {\times} \textrm{Average Sale Price}$} & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope 1 Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & N/A \\
+          \textbf{Model 3} & {$\textrm{Yield} {\times} \textrm{Average Sale Price}$} & \begin{tabular}[c]{@{}c@{}}Water Used\\ scope one Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & N/A \\
           \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-          \textbf{Model 4} & ${\textrm{Yield} \times \textrm{Average Sale Price}}\over{\textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope 1 Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * Scope 1 Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular}
+          \textbf{Model 4} & ${\textrm{Yield} \times \textrm{Average Sale Price}}\over{\textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ scope one Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * scope one Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular}
           \end{tabular}}
@@
 Preprocessing included logarithmic transformations, centring and scaling by standard deviation.
-Variables such as scope 1 emissions, which required prior calculations were also computed using Python.
+Variables such as scope one emissions, which required prior calculations were also computed using Python.
 % Two values for average sale price were removed from the dataset, due to a recording of \$1. 
@@
 Aside from 3 variables, all F-tests across each model indicated a significant contribution at 95\% confidence.
-The three exceptions were: scope 1 emissions in Model 3 (P=2.221E-01) and Model 4 (P=3.621E-01), and Model 2's interaction between area harvested and water used (P=2.192E-01).
+The three exceptions were: scope one emissions in Model 3 (P=2.221E-01) and Model 4 (P=3.621E-01), and Model 2's interaction between area harvested and water used (P=2.192E-01).
 \newline
 %b Should this be in the discussion (below para)
-Scope 1 emissions was included in all models to directly compare the response variables as ratios of vineyard size to raw values.
-Even though not significant within models 3 and 4, when using the Pearson Correlation Coefficients scope 1 emissions was strongly correlated to every Model's response variable; this was especially so for Model 1 and 4 (Yield and average price per tonne as a ratio to area harvested, respectively).
+scope one emissions was included in all models to directly compare the response variables as ratios of vineyard size to raw values.
+Even though not significant within models 3 and 4, when using the Pearson Correlation Coefficients scope one emissions was strongly correlated to every Model's response variable; this was especially so for Model 1 and 4 (Yield and average price per tonne as a ratio to area harvested, respectively).
 \subsubsection{T-tests}
@@
 With regard to continuous variables: Model 1 and 2 showed all variables to be significant at 95\% confidence when accounting for other variables.
-T-tests for Model 3 showed all continuous variables except scope 1 emissions were significant.
-Model 4 showed all variables aside from scope 1 emissions and water use to be significant; with scope 1 emissions and water use only being significant when considered as an interaction with area harvested but not when considered on their own.
+T-tests for Model 3 showed all continuous variables except scope one emissions were significant.
+Model 4 showed all variables aside from scope one emissions and water use to be significant; with scope one emissions and water use only being significant when considered as an interaction with area harvested but not when considered on their own.
 \subsubsection{Model Coefficients}
@@
   \toprule
-    & \textbf{Intercept} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Water\\ Used\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Scope 1\\ Emissions\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\\ $\ast$\\ Scope 1\\ Emissions\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\\ $\ast$\\ Water\\ Used\end{tabular}}} \\ \midrule
+    & \textbf{Intercept} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Water\\ Used\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}scope one\\ Emissions\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\\ $\ast$\\ scope one\\ Emissions\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\\ $\ast$\\ Water\\ Used\end{tabular}}} \\ \midrule
   \textbf{Model 1} & -3.318E-02 & 7.418E-01 & 8.660E-02 & 6.731E-02 &  &  \\
@@
 These relationships are described by the coefficients.
-Resource use is described by the predictor variables (through water used and scope 1 emissions), because of this we can observe the response variables somewhat interchangeably with the predictors - although caution should be taken to view them sceptically and alongside the influence of their coefficients.
+Resource use is described by the predictor variables (through water used and scope one emissions), because of this we can observe the response variables somewhat interchangeably with the predictors - although caution should be taken to view them sceptically and alongside the influence of their coefficients.
 As the predictors are known to have a strong positive correlation with each other, they will tend toward increasing and decreasing together (but not at the same rates).
@@
  There are many on-the-ground decisions that influence both quality and yield.
-Comparing the $R^2$ values between Models 2 and 4 showed that the average price per tonne of grapes described a great deal of the relationship between resource use and yield when variables were considered as ratios of area (due to the discrepancy in $R^2$ between the two models, see Table \ref{tab:modelsummary}).
-This descrepency is likely due to different vineyard prioritisation, which can be described by the type of quality and quantity a vineyard aims to target.
-Decisions such as the prioritisation of quality over quantity, are governed by complex physical and social forces, for example: international market demands, disease pressures and natural disasters \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}; with many of these occurrences being highlighted throughout the past decades vintage reports from Wine Australia \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
-It is also important to consider that these reports show that the warm inland regions have seen a decline in profit during this period, whereas regions targetting quality did not.
+The $R^2$ values for Models 2 and 4 showed that the average price per tonne of grapes described a great deal of the relationship between resource use and yield when variables were considered as ratios of area (due to the discrepancy in $R^2$ between the two models, see Table \ref{tab:modelsummary}).
+This discrepancy is likely due to different vineyard prioritisation, which can be described by the type of quality and quantity a vineyard aims to target.
+Decisions such as the prioritisation of quality over quantity, are governed by complex physical and social forces, for example international market demands, disease pressures and natural disasters \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}, with many of these occurrences being highlighted throughout the reports from Wine Australia \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018} over the past decade.
+It is also important to consider that these reports show that the warm inland regions have seen a decline in profit during this period, whereas regions targeting quality did not.
 Size becomes an important consideration, as it dictates the potential capacity to produce greater volumes of grapes.
 However, given the comparison of value per area, regions with larger vineyards (such as warmer in land regions) and larger vineyards in general, tend to underperform.
-When considering the 'Hot Very Dry' vineyards (see Figure \ref{fig:yield_vs_value_area}) These vineyards would be very competitive with only a minor increase to sale price, possibly outperforming other regions.
+The 'Hot Very Dry' vineyards (see Figure \ref{fig:yield_vs_value_area}) These vineyards would be very competitive with only a minor increase to sale price, possibly outperforming other regions.
 \newline
@@
 Economies of scale likely played a role in determining yield but were only one consideration alongside resource use.
+% (KM: WHere is this shown?)
+% This relationship is not shown well and should be reconsidered when changing the methods and results.
 Size was also less of a determining factor when considering quality.
-It is possible that the relationship of scope 1 emissions between yield and quality was closely tied to a vineyard's area; due to requiring more fuel to cover issues (such as fixing a broken irrigation pipe), where  a larger area has the potential for issues to be further away.
-This is further cemented when noting that most irrigation systems are diesel based, with water use being a significant variable in each model and scope 1 emissions not; scope one emissions' lack of significance and contribution given its F-statistics, could be indicative that other vineyard activities requiring fuel are not as determining factors for a vineyards grape quality.
+It is possible that the relationship of scope one emissions between yield and sale price was closely tied to a vineyard's area due to requiring more fuel to cover issues (such as fixing a broken irrigation pipe), where  a larger area has the potential for issues to be further away.
+This is further observed when noting that most irrigation systems are diesel based, with water use being a significant variable in each model and scope one emissions not; scope one emissions' lack of significance and contribution given its F-statistics, could be indicative that other vineyard activities requiring fuel are not leading factors for a vineyards grape quality.
 The relationship between yield, value and area was not simply about efficiently producing the most grapes; sales price and by association grape quality, are integral to the profitability, and this is strongly linked to resource-use and thus the longevity and sustainability of a vineyard.
@@
 This results in unique issues and considerations for each vineyard, where on-the-ground decisions are influenced by other wine industry's choices, such as the use of sustainable practices in vineyards as a requirement for sale in overseas markets; notably these interactions can be further complicated by some winegrowers being completely integrated into a wine company, while others are not \citep{knightFirmResourcesDevelopment2019}.
-Incorporating decisions into the model could help describe the contributing factors to regional differences beyond resource consumption and regional differences but would require incredibly granular data and more sophisticated modelling.
+Incorporating decisions into the model could help describe the contributing factors to regional differences beyond resource consumption and regional differences this  motivates the call for more granular data and more sophisticated modelling.
 \subsection{Regional Differences}
 Some regions appeared to produce many low quality grapes at scale whilst others focussed on producing higher quality grapes in lower volumes.
-This behaviour can also be observed when reviewing Wine Australia's annual reports, where it is apparent that some GI regions, such as the Riverland, are known for producing large amounts of lower grade (low value per tonne) grapes \cite{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
+This emperical finding is consistent with Wine Australia's annual reports, which shows that some GI regions, such as the Riverland, are known for producing large amounts of lower grade (low value per tonne) grapes \cite{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
 Comparatively other regions, such as Tasmania, only produce high quality grapes but in smaller quantities.
@@
 \newline
-Due to regional differences, different strategies are likely employed across different regions; such as some regions targeting mass production over quality.
+Due to regional differences, different strategies are also  employed across different regions, such as some regions targeting mass production over quality.
 This is most notable when grouping regions by climate, especially when considering GI Regions in the 'Hot Very Dry' climate (see Figure \ref{fig:yield_vs_value}).
-In alternative attempts at models it was found that without the direct incorporation of GI Region or year, predictions greatly under performed.
+%In the preliminary analyses alternative models found that without the direct incorporation of GI Region or year, predictions greatly under performed.
 The effect of climate in the models was never as significant as the more granular GI regions, and always led to less accurate models.
@@
 \subsection{Limitations}
-Limitations included overestimating yield for models 1 and 2, and underestimating crop value in models 3 and 4 (see appendix).
-The issue of model 1 and 2 over predicting yield, may have been due to preventative measures brought on by regional pressures such as fire, frost and disease.
-Where, more resources were required to prevent these issues from spreading within a region, thus disproportionately effecting some vineyards compared to others locally.
-This type of maintenance is not well captured especially when considering that some regions, especially those in warmer areas, are not as prone to disease as cooler climates and could potentially have lower operating costs per hectare.
+Limitations in the analyses presented in this paper included overestimating yield for models 1 and 2, and underestimating crop value in models 3 and 4 (see appendix).
+% KM: Need to see this.
+The issue of model 1 and 2 over-predicting yield may have been due to preventative measures brought on by regional pressures such as fire, frost and disease.
+More resources were required to prevent these issues from spreading within a region, thus disproportionately affecting some vineyards compared to others locally.
+This type of maintenance is not well captured in the models, especially when considering that some regions, especially those in warmer areas, are not as prone to disease as cooler climates and could potentially have lower operating costs per hectare.
 This could create a discrepancy in vineyards that utilised preventative measures in wetter regions, as opposed to those that did not, thus expending less fuel and energy but risking disease.
-When reviewing the differences between regions it is important to consider that vineyards in 'Hot Very Dry' areas can be hundreds of times the size of those in other regions.
-This limitation could be overcome by incorporating the profitability of vineyards, compare the financial success of working at different operational scales.
+When reviewing the differences between regions, it is important to consider that vineyards in 'Hot Very Dry' areas can be hundreds of times the size of those in other regions.
+This limitation could be overcome by incorporating the profitability of vineyards, comparing the financial success of working at different operational scales.
 \newline
@@
 Variables such as the utilisation of renewable energy, contractors, and the occurrence of disease, fire and frost were originally explored to capture the discrepancies between similar vineyards that produced different  yields and crop values.
-However, none of these variables were significantly connected to the response variables, and did not add to model accuracy; even when considered as interactions.
-The use of other methods, specifically splines, resulted in more normally distributed residuals but at a drastically reduced overall accuracy when comparing $R^2$ and Residual Square Error.
+However, none of these variables was significantly correlated with the response variables, and did not add to model accuracy, even when considered as interactions.
+Allowance for nonlinear relationships, specifically through splines, resulted in more normally distributed residuals but at a drastically reduced overall accuracy when comparing $R^2$ and Residual Square Error.
 Attempts to fully explain small variations was always overshadowed by the dramatic differences in regional trends.
 \newline
-Having more data for each region would also be an improvement, allowing greater comparison between regions.
+Having more data for each region would also be beneficial, allowing greater comparison between regions.
 More variables may also help to discern vineyards that can produce larger volumes of grapes at higher prices.
-The use of semi transparent tools such as random forests and decision trees alongside more variables and data may help to uncover the reasons for values that were under or overestimated.
+The use of other models such as random forests and decision trees alongside more variables and data may help to uncover the reasons for under or overestimation.
 These differences could be caused by the use of alternative sustainable practices in the field.
-And, while there is evidence to suggest that environmentally sustainable practices can reduce costs, increase efficiency, whilst improving the quality of grapes; more research is needed to link these benefits across different regions and climates \citep{baianoOverviewSustainabilityWine2021,marianiSustainableWinegrowingCurrent2015,montalvo-falconSustainabilityResearchWine2023}.
+Moreover, while there is evidence to suggest that environmentally sustainable practices can reduce costs, and increase efficiency whilst improving the quality of grapes; more research is needed to link these benefits across different regions and climates \citep{baianoOverviewSustainabilityWine2021,marianiSustainableWinegrowingCurrent2015,montalvo-falconSustainabilityResearchWine2023}.
 \section{Conclusion}
@@
 %  Water Used        & 8.660E-02 & 8.885E-03  & 9.747E+00 & 2.000E-16             \\
-%  Scope 1 Emissions & 6.731E-02 & 8.013E-03  & 8.400E+00 & 2.000E-16            
+%  scope one Emissions & 6.731E-02 & 8.013E-03  & 8.400E+00 & 2.000E-16            
 %  \end{tabular}

### 2023-08-14 article.tex
conclusion edits

@@
 In summary, vineyard yield and crop value is well-defined by the resources used.
-However, it is important to consider a vineyard's business goal, region, external pressures and economies of scale.
-Where, larger vineyards are likely to produce greater overall yields, and have higher yield per area.
-Smaller vineyards are likely to produce more value per area, and a higher quality of grape.
+However, it is important to consider a vineyard's business goal, region, external pressures and economies of scale where larger vineyards are likely to produce greater overall yields, and have higher yield per area.
+Smaller vineyards are likely to produce more value per area and a higher quality of grape.
 It is likely that regional constraints also contribute to the best strategy to pursue when considering quality or quantity.
-%
+%KM is this what the data analyses tell you? If so, say so. If not delete.
 

### 2023-08-14 starplot.r
start of radar plot

@@ -1,0 +1,52 @@
+coord_radar <- function (theta = "x", start = 0, direction = 1)
+
+ {
+        theta <- match.arg(theta, c("x", "y"))
+        r <- if (theta == "x") 
+            "y"
+        else "x"
+        ggproto("CordRadar", CoordPolar, theta = theta, r = r, start = start, 
+            direction = sign(direction),
+            is_linear = function(coord) TRUE)
+    }
+
+RadarTheme<-theme(panel.background=element_blank(),
+                      plot.title= element_text(size = 25,face=c("bold","italic")),
+                      plot.margin = unit(c(2, 2, 2, 2), "cm"),
+                      text=element_text(family="Open Sans"), aspect.ratio = 1,
+                      legend.position="bottom",legend.title=element_blank(),legend.direction="vertical",
+                      strip.text.x = element_text(size = rel(0.8)),
+                      axis.text.x = element_text(size = 15,face ="bold"),
+                      axis.ticks.y = element_blank(),
+                      axis.text.y = element_blank(),
+                      axis.line.x=element_line(size=0.5),
+                      panel.grid.major=element_line(size=0.3,linetype = 2,colour="grey"))
+
+set.seed(100)
+df <- read.csv("df.csv")
+
+dfrescale <- df[, c("water_used", "scope1", "area_harvested")]
+dfrescale <- as.data.frame(lapply(dfrescale, ggplot2:::rescale01))
+# mtcarsscaled <- as.data.frame(lapply(mtcars, ggplot2:::rescale01))
+# mtcarsscaled$model <- rownames(mtcars)
+dfrescale$rain <- df$rain
+
+dfdry <- subset(dfrescale, rain=='Very_Dry')
+# tcarsscaled1<-subset(mtcarsscaled,model=="Lotus Europa")
+dfwet <- subset(dfrescale, rain=='Damp')
+# mtcarsscaled2<-subset(mtcarsscaled,model=="Volvo 142E")
+
+dfscaled <- rbind(dfwet, dfdry)
+# mtcarsscaled<-rbind(mtcarsscaled1,mtcarsscaled2)
+
+dfmelted <- reshape2::melt(dfrescale)
+dfmelted$value<-dfmelted$value*100
+# mtcarsmelted <- reshape2::melt(mtcarsscaled)
+# mtcarsmelted$value<-mtcarsmelted$value*100
[…]

### 2023-08-15 starplot.r
first working radar plot.

@@ -1,52 +1,27 @@
-coord_radar <- function (theta = "x", start = 0, direction = 1)
-
- {
-        theta <- match.arg(theta, c("x", "y"))
-        r <- if (theta == "x") 
-            "y"
-        else "x"
-        ggproto("CordRadar", CoordPolar, theta = theta, r = r, start = start, 
-            direction = sign(direction),
-            is_linear = function(coord) TRUE)
-    }
-
-RadarTheme<-theme(panel.background=element_blank(),
-                      plot.title= element_text(size = 25,face=c("bold","italic")),
-                      plot.margin = unit(c(2, 2, 2, 2), "cm"),
-                      text=element_text(family="Open Sans"), aspect.ratio = 1,
-                      legend.position="bottom",legend.title=element_blank(),legend.direction="vertical",
-                      strip.text.x = element_text(size = rel(0.8)),
-                      axis.text.x = element_text(size = 15,face ="bold"),
-                      axis.ticks.y = element_blank(),
-                      axis.text.y = element_blank(),
-                      axis.line.x=element_line(size=0.5),
-                      panel.grid.major=element_line(size=0.3,linetype = 2,colour="grey"))
+# Library
+library(fmsb)
 
 set.seed(100)
 df <- read.csv("df.csv")
 
+mean.t <- function(x) {
+    mean(x, na.rm = TRUE)
+}
+
+selected_columns <- c("water_used", "scope1", "area_harvested")
+
 dfrescale <- df[, c("water_used", "scope1", "area_harvested")]
-dfrescale <- as.data.frame(lapply(dfrescale, ggplot2:::rescale01))
-# mtcarsscaled <- as.data.frame(lapply(mtcars, ggplot2:::rescale01))
-# mtcarsscaled$model <- rownames(mtcars)
 dfrescale$rain <- df$rain
+means <- aggregate(dfrescale, list(df$rain), FUN = mean.t)
 
-dfdry <- subset(dfrescale, rain=='Very_Dry')
-# tcarsscaled1<-subset(mtcarsscaled,model=="Lotus Europa")
-dfwet <- subset(dfrescale, rain=='Damp')
-# mtcarsscaled2<-subset(mtcarsscaled,model=="Volvo 142E")
+maxes <- c()
+mins <- c()
+for (i in selected_columns) {
[…]

### 2023-08-15 starplot.r
included all columns

@@ -8,9 +8,9 @@ mean.t <- function(x) {
     mean(x, na.rm = TRUE)
 }
 
-selected_columns <- c("water_used", "scope1", "area_harvested")
+selected_columns <- c("water_used", "scope1", "area_harvested", "average_per_tonne", "tonnes_grapes_harvested")
 
-dfrescale <- df[, c("water_used", "scope1", "area_harvested")]
+dfrescale <- df[, selected_columns]
 dfrescale$rain <- df$rain
 means <- aggregate(dfrescale, list(df$rain), FUN = mean.t)
 

### 2023-08-15 starplot.r
comments

@@ -1,27 +1,47 @@
 # Library
 library(fmsb)
 
+# read in the data
 set.seed(100)
 df <- read.csv("df.csv")
 
+# This is to pass to the aggregation function with the
+# preset argument to remove NaNs.
 mean.t <- function(x) {
     mean(x, na.rm = TRUE)
 }
 
-selected_columns <- c("water_used", "scope1", "area_harvested", "average_per_tonne", "tonnes_grapes_harvested")
+#  We grab the columns we want to display:
+selected_columns <- c("water_used"
+    , "scope1"
+    , "area_harvested"
+    , "average_per_tonne"
+    , "tonnes_grapes_harvested"
+)
 
 dfrescale <- df[, selected_columns]
-dfrescale$rain <- df$rain
-means <- aggregate(dfrescale, list(df$rain), FUN = mean.t)
+dfrescale$temp <- df$temp
+means <- aggregate(dfrescale, list(df$temp), FUN = mean.t)
 
+# The first two rows in the dataframe determine the scaling of the graph.
+# They are set to be the maximum and minimum values
+# So we go and grab them by iterating through the columns:
 maxes <- c()
 mins <- c()
 for (i in selected_columns) {
     current_min <- min(means[, i], na.rm=TRUE)
-  mins <- append(mins, current_min)
-  maxes <- append(maxes, max(means[, i], na.rm=TRUE) + current_min**2)
+    mins <- append(mins, current_min)
+    # We add the squrae of the minimum to the maximum
+    # this is to give it a bit of white space.
+    maxes <- append(maxes, max(means[, i], na.rm=TRUE) + current_min**2)
 }
+
+# We stick the two together!
 df_radar <- data.frame(rbind(maxes, mins))
 colnames(df_radar) <- selected_columns
 
-radarchart(rbind(df_radar, means[, selected_columns]))
+# And plot!
+radarchart(rbind(df_radar, means[-1, selected_columns])
+    , cglwd = 1.2
+    , plwd = 2
+    , plty = 1)

### 2023-08-15 starplot.r
star plot function created

@@ -1,6 +1,26 @@
 # Library
 library(fmsb)
 
+# We set the colours we will use:bty
+temp_colors <- c(
+    # Cool
+    "#1E88E5"
+    # Hot
+    , "#D81B60"
+    # Mild
+    , "#004d40"
+    # Warm
+    , "#FFC107"
+)
+rain_colors <- c(
+    # Damp
+    "#1E88E5"
+    # Dry
+    , "#FFC107"
+    # Very Dry
+    , "#D81B60"
+)
+
 # read in the data
 set.seed(100)
 df <- read.csv("df.csv")
@@ -19,29 +39,52 @@ selected_columns <- c("water_used"
     , "tonnes_grapes_harvested"
 )
 
-dfrescale <- df[, selected_columns]
-dfrescale$temp <- df$temp
-means <- aggregate(dfrescale, list(df$temp), FUN = mean.t)
+starplot <- function(selected_columns, df, target_col, colors_in) {
+    dfrescale <- df[, selected_columns]
+    dfrescale[, target_col] <- df[, target_col]
+    means <- aggregate(dfrescale, list(df[, target_col]), FUN = mean.t)
 
-# The first two rows in the dataframe determine the scaling of the graph.
-# They are set to be the maximum and minimum values
-# So we go and grab them by iterating through the columns:
-maxes <- c()
-mins <- c()
-for (i in selected_columns) {
-    current_min <- min(means[, i], na.rm=TRUE)
-    mins <- append(mins, current_min)
-    # We add the squrae of the minimum to the maximum
-    # this is to give it a bit of white space.
-    maxes <- append(maxes, max(means[, i], na.rm=TRUE) + current_min**2)
+    # The first two rows in the dataframe determine the scaling of the graph.
+    # They are set to be the maximum and minimum values
+    # So we go and grab them by iterating through the columns:
+    maxes <- c()
+    mins <- c()
+    for (i in selected_columns) {
+        current_min <- min(means[, i], na.rm=TRUE)
+        mins <- append(mins, current_min)
+        # We add the squrae of the minimum to the maximum
+        # this is to give it a bit of white space.
+        maxes <- append(maxes, max(means[, i], na.rm=TRUE) + current_min**2)
+    }
+
+    # We stick the two together!
+    df_radar <- data.frame(rbind(maxes, mins))
+    colnames(df_radar) <- selected_columns
+
+    # To help with the legend we name the rows
+    rownames(means) <- means$Group.1
[…]

### 2023-08-15 starplot.r
finished starplots with outputs.

@@ -1,48 +1,17 @@
 # Library
 library(fmsb)
 
-# We set the colours we will use:bty
-temp_colors <- c(
-    # Cool
-    "#1E88E5"
-    # Hot
-    , "#D81B60"
-    # Mild
-    , "#004d40"
-    # Warm
-    , "#FFC107"
-)
-rain_colors <- c(
-    # Damp
-    "#1E88E5"
-    # Dry
-    , "#FFC107"
-    # Very Dry
-    , "#D81B60"
-)
-
-# read in the data
-set.seed(100)
-df <- read.csv("df.csv")
-
 # This is to pass to the aggregation function with the
 # preset argument to remove NaNs.
 mean.t <- function(x) {
     mean(x, na.rm = TRUE)
 }
 
-#  We grab the columns we want to display:
-selected_columns <- c("water_used"
-    , "scope1"
-    , "area_harvested"
-    , "average_per_tonne"
-    , "tonnes_grapes_harvested"
-)
-
 starplot <- function(selected_columns, df, target_col, colors_in) {
     dfrescale <- df[, selected_columns]
     dfrescale[, target_col] <- df[, target_col]
-    means <- aggregate(dfrescale, list(df[, target_col]), FUN = mean.t)
+    means <- aggregate(dfrescale[, selected_columns]
+        , list(df[, target_col]), FUN = mean.t)
 
     # The first two rows in the dataframe determine the scaling of the graph.
     # They are set to be the maximum and minimum values
@@ -67,24 +36,69 @@ starplot <- function(selected_columns, df, target_col, colors_in) {
     df_radar <- rbind(df_radar, means[-1, selected_columns])
 
     # And plot!
-    return(
-        radarchart(df_radar
-            , pcol = colors_in
-            , cglwd = 1.2
-            , plwd = 2
-            , plty = 1) +
-        legend(x = 0.7
-            , y = 1
-            , legend = rownames(df_radar[-(1:2), ])
-            , bty = "n"
-            , pch = 20 
-            , col = colors
-            , text.col = "grey"
-            , cex = 1.2
-            , pt.cex = 3)
-    )
+    radarchart(df_radar
+        , pcol = colors_in
+        , cglwd = 1.2
+        , plwd = 2
+        , plty = 1)
+    legend(x = 0.7
+        , y = 1
+        , legend = rownames(df_radar[-(1:2), ])
[…]

### 2023-08-17 article.tex
backup of transplant progress

@@
 \newline
-There is an extensive amount of research into the effects of a variety of factors on grape quality and yield
-% refeerences
-.
+\newline
+There is an extensive amount of research into the effects of a variety of factors on grape quality and yield \citep{heFruitYieldPrediction2022,laurentLocalInfluenceClimate2022,liakosMachineLearningAgriculture2018}.
 However, due to the lack of long-term and in-depth data, individual factors are often studied in isolation \citep{abbalDecisionSupportSystem2016}.
@@
 \section{Methods}
-We created four linear models to explore relationships between resource-use and vineyard outputs (see Table\ref{tab:tab1}).
-The data was sourced from Sustainable Winegrowing Australia and Wine Australia.
-Variables used included: yield, average sale price, region, water use, emissions, area harvested and year.
-After fitting to the data, each model was validated using k-fold cross validation.
+\subsection{Data}
+
 \begin{table}[]
@@
     \textbf{Model 2} & ${ \textrm{Yield}}\over{ \textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ scope one Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * scope one Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular} \\
-          \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
+    {1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
     \textbf{Model 3} & {$\textrm{Yield} {\times} \textrm{Average Sale Price}$} & \begin{tabular}[c]{@{}c@{}}Water Used\\ scope one Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & N/A \\
     \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-          \textbf{Model 4} & ${\textrm{Yield} \times \textrm{Average Sale Price}}\over{\textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ scope one Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * scope one Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular}
+    \textbf{Model 4} & ${\textrm{Yield} \times \textrm{Average Sale Price}}\over{\textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ scope one Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\
+        Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * scope one Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular}
     \end{tabular}}
 \end{table}
-\subsection{Analysis}
-Before models were fit to the data, Pearson Correlation Coefficients were used to look at the existence of linear relationships between predictor variables.
-These relationships were summarised in correlation matrices to compare the level of interaction present between predictor variables.
-The relationships between the predictors and response variables were then modelled using General Linear Models.
-Both the Pearson Correlation Coefficients and General Linear Models were created using the R statistical programming language \citep{rcoreteamLanguageEnvironmentStatistical2021}.
-General Linear Models were chosen as they offer the ability to produce statistical models that are explicit in the relationships between predictors and response variables.
-%reference?
-General Linear Models also allow the exploration of interactions between predictors and present easily comparable differences in the influence and magnitude of relationships.
-% reference?
- A variety of alternate methods were also explored, including: Splines, hierarchical regression, General Additive Models, and Generalised Linear Models.
-These alternative approaches were not used as final models due to offering no further insights or improvements in accuracy.
-\newline
-The response variables of the models were yield and quality.
-Yield was defined as the total tonnes of grapes harvested.
-For the purpose of this study, quality was defined by the financial value of winegrape crops' average sale price per tonne.
-The definition of quality was an important consideration, as quality can be defined in a variety of ways, for example analysing grapes': aroma, chemical composition and color.
-% References to were this was done.
-Using sale price as a defining trait of quality was due to the market value of winegrapes being reliant on grape quality and because Wine Australia explicitly defines grape quality through the use of discrete price brackets in their annual reports
-% We need a reference to where they did this
-% Did they do this more prolifically at some point as well?
-; the generalisation made to reflect quality through using average price assumed a due diligence of those who purchased the grapes \citep{yeggeInfluenceSensoryNonsensory2001}.
-Both response variables were examined as totals and as scales of area harvested.
-Values were compared in this manner to observe how economies of scale affect the use of resources.
-%
-%
-%
-\subsection{Significant Tests}
-%This briefly needs to be described. You use the results of significance tests in the results section and talk about variables significance but do not describe it in the methods.
-%
-% F-tests
-%
-% T-tests
-%
-%
-\subsection{Data}
-Data used in this analysis was sampled by Sustainable Winegrowing Australia and Wine Australia.
+
+Data used in this analysis were obtained from Sustainable Winegrowing Australia and Wine Australia.
 Sustainable Winegrowing Australia is Australia's national wine industry sustainability program, which aims to facilitate grape-growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
@@
 \newline
-Data sampled by Wine Australia was collected via phone surveys and included: summary statistics such as yield and average price of sale per tonne; these values were summarised by region and grape varietal.
-Data recorded by Sustainable Winegrowing Australia was entered manually by winegrowers using a web based interface with some fields being optional, variables included: region, harvest year, yield, area harvested, water used and fuel used (diesel, petrol, biodiesel and LPG).
-To enable direct comparisons between fuels, they were converted to tonnes of Carbon Dioxide equivalent.
-% They were converted to emissions that was expressed in TCO2E - not into the units!
 \newline
-The inclusion of Wine Australia data was due to average sale price being an optional field in Sustainable Winegrowing Australia's dataset. 
-Regional average prices from Wine Australia were filled into values that were missing from the Sustainable Winegrowing Australia data; the common practice of purchasing grapes at regional prices was an important consideration in this decision.
-%{It would be lovely to have a reference for this but it is currently only anacdotal.}
+Predictor variables used in this analysis included yield, defined as the total tonnes of grapes harvested, and quality, defined as the average sale price.
+It is acknowledged that quality can be defined in a variety of ways, for example by the grapes': aroma, chemical composition and color \citep{kasimatiPredictingGrapeSugar2022,mejeanperrotDecisionsupportSystemPredict2022,suarezAssessingWineGrape2021}.
+Using sale price was based on the reliance of market value of winegrapes on grape quality and because Wine Australia explicitly defines grape quality through the use of discrete price brackets in their annual reports.
+The generalisation made to reflect quality through using average price assumed a due diligence of those who purchased the grapes \citep{yeggeInfluenceSensoryNonsensory2001}.
+Both response variables were examined as totals and as scales of area harvested.
+Values were compared in this manner to observe how economies of scale affect the use of resources.
+\newline
+\newline
+Data obtained from Wine Australia were collected via phone surveys and included: total tonnes purchased, average price per tonne and yearly change in price for region and grape varietal.
+Data recorded by Sustainable Winegrowing Australia was entered manually by winegrowers using a web based interface with some fields being optional.
+Required variables included: region, harvest year, yield and area harvested.
+Optional variables included average sale price, water used and fuel used (diesel, petrol, biodiesel and LPG).
+To enable direct comparisons between fuels, fuel use was converted to tonnes of Carbon Dioxide equivalent and collectively referenced to as emissions.
+\newline
+\newline
+Average sale price was an optional field in the Sustainable Winegrowing Australia's dataset. 
+Missing values were improved using regional average prices from Wine Australia.
 Two subsets of data were then created for the analysis.
-The first subset contained all vineyards and was used for Models 1 and 3.
-The second subset contained vineyards which either recorded a value for average price of sale per tonne through Sustainable Winegrowing Australia, or were within a region with an average price of sale recorded by Wine Australia; this subset was used for Models 2 and 4.
-These subsets meant that the data would be limited to samples which had recorded values for the response variables (see Table\ref{tab:tab1}), where every sample had a recorded value for yield but not average price of sale per tonne.
+The first subset contained all vineyards and was used for two models (Model 1 and Model 2, see Table \ref{tab:tab1}).
+The second subset contained vineyards which either recorded a value for average price of sale per tonne through Sustainable Winegrowing Australia, or were within a region with an average price of sale recorded by Wine Australia; this subset was used for two further models (Models 3 and 4, see Table \ref{tab:tab1}.
+These subsets meant that the data would be limited to samples which had recorded values for the response variables (see Table \ref{tab:tab1}), where every sample had a recorded value for yield but not average price of sale per tonne.
 \newline
-The first subset of data was used for Model 1 and Model 2 (see Table\ref{tab:tab1}).
-This subset contained 5298 samples spanning the period from 2012 to 2022, covering 55 GI Regions and 1261 separate vineyards.
 \newline
-The second subset of data, was limited to vineyards that recorded a value for their average sale price of grapes per tonne.
-This subset was used for Model 3 and Model 4 (see Table\ref{tab:tab1}); and contained 2878 samples spanning the period from 2015 to 2022, covering 51 GI Regions and 944 separate vineyards.
-1842 of the values for average price of sale per tonne were extracted from Wine Australia surveys with the remaining 1036 being from Sustainable Winegrowing Australia's dataset.
+The first subset of data (used for Model 1 and Model 2, see Table\ref{tab:tab1}) contained 5298 samples spanning the period from 2012 to 2022, covering 55 GI Regions and 1261 separate vineyards.
+\newline
+\newline
+The second subset of data (used for Model 3 and Model 4, see Table\ref{tab:tab1}) contained 2878 samples spanning the period from 2015 to 2022, covering 51 GI Regions and 944 separate vineyards.Average price of sale per tonne was extracted from both Wine Australia (1842 values) and Sustainable Winegrowing Australia (remaining 1036 values).
+\newline
 \newline
 Additional variables were considered for analysis but were excluded due to being either underreported or had insignificant contributions to model accuracies.
-Variables explored but not used due to low reporting values included: fertiliser, and scope 2 emissions.
-Variables considered but ultimately removed due to a lack of significant contributions to models, included: the use of renewable energy, contractor use, and pressures such as frost, fire and disease.
+Variables explored but not used due to low reporting values included fertiliser, and scope two emissions.
+Variables considered but ultimately removed due to a lack of significant contributions to models, included the use of renewable energy, contractor use, and pressures such as frost, fire and disease.
+\newline
 \newline
 Data preprocessing was conducted prior to analysis using the Python programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
-Preprocessing included logarithmic transformations, centring and scaling by standard deviation.
-Variables such as scope one emissions, which required prior calculations were also computed using Python.
-% Two values for average sale price were removed from the dataset, due to a recording of \$1. 
-% \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2012,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
-\subsection{Total Emissions}
-The equation given from the Australian National Greenhouse Accounts Factors, shown as
+Preprocessing included the conversion from fuel to scope one emissions and prior calculations for all continuous variables which included logarithmic transformations, centring and scaling by standard deviation.
+The transformation of fuel use into scope one emissions was done using the equation given from the Australian National Greenhouse Accounts Factors, shown as
+\newline
 \newline
@@
 \newline
+\newline
 was used to convert the quantity of fuel in litres, $Q$, using a prescribed Energy Content, $EC$, and emission factors of scope one, $EF1$, and scope three, $EF3$, to tonnes of Carbon Dioxide Emission equivalent, $tCO2e$ \citep{departmentofclimatechangeenergytheenvironmentandwaterAustralianNationalGreenhouse2022}.
-Emissions were calculated for total diesel, petrol, bio-diesel and LPG used.
-%\newline
-% What the hell is the below paragraph and why is it here?
-%The variables were reviewed for correlations by using a Pearson's Correlation Coefficient (see Tables 1, 2 and 3).
-This was undertaken for data on the original scale (see Table 1) and for data as a logarithmic transform (see Table 2).
-All P-values were found to be significant (< 2.200E-16), except the non-transformed values for water used (see Table 3).
-The logarithmic transforms performed the best due to a skew likely caused by a greater number of smaller vineyards within the dataset (see Table 4).
-\subsection{Region}
+\newline
+%%
+%%
+%%
+% This is where the maps go!
+%%
+%%
+%%
+\newline
 Differences in vineyard locations were captured through the use of Geographical Indicator Regions (GI Regions).
@@
 \newline
-The site of a vineyard predetermines several physical parameters such as climate, geology and soil; making location a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
+\newline
+The site of a vineyard predetermines several physical parameters such as climate, geology and soil, making location a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
 The climatic properties of each GI Region were summarised by using predefined classifications as per the \citet{sustainablewinegrowingaustraliaSustainableWinegrowingAustralia2021} user manual.
@@
 The climatic groups were used to illustrate similarities and differences occurring in areas larger than GI Regions.
+
+\subsection{Analysis}
+Pairwise Pearson Correlation Coefficients were calculated to assess the potential existence of linear relationships between the input and predicted variable.
+Four general linear models were then created (see Table \ref{tab:tab1})
+
+Both the Pearson Correlation Coefficients and General Linear Models were created using the R statistical programming language \citep{rcoreteamLanguageEnvironmentStatistical2021} and the Caret packages \citep{kuhnBuildingPredictiveModels2008}.
+General Linear Models were chosen as they offer the ability to produce statistical models that are explicit in the relationships between predictors and response variables.
+
+General Linear Models also allow the exploration of interactions between predictors and present easily comparable differences in the influence and magnitude of relationships.
+
+A variety of alternate methods were also explored, including splines, hierarchical regression, General Additive Models, and Generalised Linear Models.
+These alternative approaches were not used as final models due to offering no further insights or improvements in accuracy.
+
 \subsection{Model Validation}
@@
 % A reference for this is absolutely needed.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% This is my results to methods prison
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Values for these coefficients reflect the linear relation between two variables, on a scale between -1 and 1; the magnitude and sign of a coefficient indicates the strength of the relation, and whether the relation is positive or negative respectively.
+This was undertaken for data on the original scale and for data as a logarithmic transform.
+Transforming data prior to calculating the coefficients changes several things: The logarithmic transform of the data alters the interpretation of the coefficients to percentage change - a coefficient will be indicative of the change in percentage of one variable compared to the other; scaling by standard deviation also changes this interpretation to be a percentage of that variables standard deviation.
+When considering the logarithmically transformed variables, a coefficient of 1 would indicate that the change of one variable by one percentage of its standard deviation would correlate to the other variable changing by one percent of its own standard deviation.
+The importance of this is the dimensionless nature of these relationships and that it can be translated directly to any vineyard's case that has a well known distribution.
+
+To determine if a coefficient was indicative of a strong relationship, confidence intervals were used.
+P-values reflected the significance of a given correlation coefficient when considering its relation to sample size via its incorporation as an element of standard error.
+
+Model fit was also measured in $R^2$ and adjusted $R^2$, F statistics
+
+statistical significance p-values declared the associated was then 0.05
+
+T-tests were used to determine if predictors significantly contributed to their models when accounting for other variables; this allowed a more granular examination of interactions and factors within categorical variables, showing which specific years and areas contributed significantly and which did not (the appendix contains a comprehensive list of these values).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
 \section{Results}
-\subsection{Data}
-Each variable was logarithmically transformed and then centred around a mean of 0.
-The values of these variables were then divided by standard deviation creating a comparable ratio intrinsic to each variable.
-Table \ref{tab:summary} shows the summary statistics of each variable, to contextualise these ratios to real values.
+\subsection{Exploratory Analysis} \label{sec:exp_anal}
+ Table \ref{tab:summary} shows the summary statistics of each variable used in Models 1 to 4 (See \ref{tab:tab1}).
 
@@
           \textbf{Variable} & \textbf{Mean} & \textbf{\begin{tabular}[c]{@{}l@{}}Standard\\ Deviation\end{tabular}} & \textbf{Minimum} & \textbf{Maximum} \\ \midrule
-          \textbf{Yield} & 7.757E+02 & 2.179E+03 & 1.000E+00 & 7.231E+04 \\
+          {Yield} & 7.757E+02 & 2.179E+03 & 1.000E+00 & 7.231E+04 \\
           \multicolumn{1}{l}{} &  &  &  &  \\
-          \textbf{Area Harvested} & 6.670E+05 & 1.337E+06 & 7.000E+02 & 2.436E+07 \\
+          {Area Harvested} & 6.670E+05 & 1.337E+06 & 7.000E+02 & 2.436E+07 \\
           \multicolumn{1}{l}{} &  &  &  &  \\
-          \textbf{Water Used} & 7.471E+06 & 5.646E+08 & 1.000E+00 & 4.268E+10 \\
+          {Water Used} & 7.471E+06 & 5.646E+08 & 1.000E+00 & 4.268E+10 \\
           \multicolumn{1}{l}{} &  &  &  &  \\
-          \textbf{Scope One Emissions} & 4.173E+04 & 8.571E+04 & 6.755E+00 & 2.110E+06 \\
+          {Scope One Emissions} & 4.173E+04 & 8.571E+04 & 6.755E+00 & 2.110E+06 \\
           \multicolumn{1}{l}{} &  &  &  &  \\
-          $\textbf{Yield}\over \textbf{Area}$ & 1.009E+01 & 8.127E+00 & 4.000E-02 & 8.634E+01 \\
+          ${Yield}\over \textbf{Area}$ & 1.009E+01 & 8.127E+00 & 4.000E-02 & 8.634E+01 \\
           \multicolumn{1}{l}{} &  &  &  &  \\
-          \textbf{Average Sale Price} & 1.477E+03 & 9.216E+02 & 1.600E+02 & 2.600E+04 \\
+          {Average Sale Price} & 1.477E+03 & 9.216E+02 & 1.600E+02 & 2.600E+04 \\
           \multicolumn{1}{l}{} &  &  &  &  \\
-          $\textbf{Average Sale Price} \over \textbf{Area Harvested}$ & 1.347E+02 & 5.711E+02 & 1.753E-01 & 2.979E+04 \\ \bottomrule
+          ${Average Sale Price} \over \textbf{Area Harvested}$ & 1.347E+02 & 5.711E+02 & 1.753E-01 & 2.979E+04 \\ \bottomrule
           \end{tabular}
@@
   \end{table}
-
-\subsection{Exploratory Analysis}\label{sec:exp_anal}
 \begin{table}[]
-  \caption{Variable Pearson correlation values for logarithmically transformed values.}\label{tab:pearson_correlation}
+  \caption{Pairwise Pearson correlation coefficients for logarithmically transformed values.}\label{tab:pearson_correlation}
       \resizebox{\textwidth}{!}{
@@
 \end{table}
-Linear relationships between variables were explored using Pearson Correlation Coefficients.
-Values for these coefficients reflect the linear relation between two variables, on a scale between -1 and 1; the magnitude and sign of a coefficient indicates the strength of the relation, and whether the relation is positive or negative respectively.
-This was undertaken for data on the original scale and for data as a logarithmic transform.
-The logarithmic transformed data showed the strongest correlations, likely due to a skew caused by a greater number of smaller vineyards within the dataset (see Table \ref{tab:pearson_correlation}).
-Transforming data prior to calculating the coefficients changes several things: The logarithmic transform of the data alters the interpretation of the coefficients to percentage change - a coefficient will be indicative of the change in percentage of one variable compared to the other; scaling by standard deviation also changes this interpretation to be a percentage of that variables standard deviation.
-Scaling by standard deviation also makes the Pearson Correlation Coefficient equal to the covariance of the two variables.
-With all this in mind, when considering the logarithmically transformed variables, a coefficient of 1 would indicate that: given the change of one variable by one percentage of its standard deviation, the other variable would change by one percent of its own standard deviation.
-The importance of this is the dimensionless nature of these relationships and that it can be translated directly to any vineyard's case that has a well known distribution.
+Pearson Correlation Coefficients of the transformed, centered and scaled variables are shown in Table \ref{tab:pearson_correlation}. 
 \newline
-To determine if a coefficient was indicative of a strong relationship, confidence intervals were used.
-P-values reflected the significance of a given correlation coefficient when considering its relation to sample size via its incorporation as an element of standard error.
-Strong relationships were found to be present as all P-values, except for the non-transformed values for water used, were considered significant (P $<$ 2.200E-16).
+\newline
+All relationships were found to be statistically significant (P $<$ 2.200E-16).
 \subsection{General Linear Models}
-General Linear Models were used to describe how response variables related to predictors' values.
-Log transformed variables were used as inputs to these models as they resulted in higher $R^2$ values and described the relationships proportionally; reflecting coefficient values as percentages of a variable's standard deviation.
+
 \begin{table}[]
@@
 Each model showed a strong relationship between the predictors and the response (see Table~\ref{tab:modelsummary}).
-Model accuracy was measured in $R^2$, as this allowed an easy comparison between their performances and their validation.
-\subsubsection{F-tests}
-To determine if predictors significantly related to a Model's response variable, F-tests were conducted.
-Aside from 3 variables, all F-tests across each model indicated a significant contribution at 95\% confidence.
-The three exceptions were: scope one emissions in Model 3 (P=2.221E-01) and Model 4 (P=3.621E-01), and Model 2's interaction between area harvested and water used (P=2.192E-01).
+
+Aside from 3 variables, F-tests across each model indicated a good model fit.
+The three exceptions were: scope one emissions in Model 3 (P=0.22) and Model 4 (P=0.36), and the interaction between area harvested and water used in model 2 (P=0.22).
+Note that, scope one emissions was included in all models to directly compare the response variables as ratios of vineyard size to raw values and because it was strongly correlated to the response variable in every model; especially for Models 1 and 4 (Table \ref{tab:pearson_correlation}).
+
+A full list of regression coefficients 95\% CIs and p-values for each of the four models is provided in the appendix. % You gotta add these to the appendix in a more readable fashion
 \newline
-%b Should this be in the discussion (below para)
-scope one emissions was included in all models to directly compare the response variables as ratios of vineyard size to raw values.
-Even though not significant within models 3 and 4, when using the Pearson Correlation Coefficients scope one emissions was strongly correlated to every Model's response variable; this was especially so for Model 1 and 4 (Yield and average price per tonne as a ratio to area harvested, respectively).
-\subsubsection{T-tests}
-T-tests were used to determine if predictors significantly contributed to their models when accounting for other variables; this allowed a more granular examination of interactions and factors within categorical variables, showing which specific years and areas contributed significantly and which did not (the appendix contains a comprehensive list of these values).
 \newline
-For Models 1 (yield) and 3 (value) year played a pivotal role, with only one year in each model not being significant (2021/2022 and 2016/2017 respectively).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% See Kerries notes regarding these.
+%
+%For Models 1 (yield) and 3 (value) year played a pivotal role, with only one year in each model not being significant (2021/2022 and 2016/2017 respectively).
 Both Model 1 and 3 showed a majority of regions were significant with 32 of 54 regions being significant in Model 1, and 42 of 50 regions being significant in Model 3 at 95\% confidence.
-\newline
-The number of combinations of year and region meant that Models 2 and 4 had many tests (424 and 243 respectively).
+%\newline
+%\newline
+%The number of combinations of year and region meant that Models 2 and 4 had many tests (424 and 243 respectively).
 Model 2 found 62.56\% of these combinations were indicative of a significant contribution to the model at 95\% significance.
@@
 A likely reason for some combinations not being significant was a lack of samples in that particular region/year being present; with region sample sizes ranging from 1 to 1006.
-\newline
-With regard to continuous variables: Model 1 and 2 showed all variables to be significant at 95\% confidence when accounting for other variables.
+%\newline
+%\newline
+%With regard to continuous variables: Model 1 and 2 showed all variables to be significant at 95\% confidence when accounting for other variables.
 T-tests for Model 3 showed all continuous variables except scope one emissions were significant.
 Model 4 showed all variables aside from scope one emissions and water use to be significant; with scope one emissions and water use only being significant when considered as an interaction with area harvested but not when considered on their own.
-\subsubsection{Model Coefficients}
-%
-% What are you trying to say here?
-% Coefficients are important as they describe the direct relationship between a variable and its response.
-They indicate whether it is a positive, or negative relationship and the magnitude of this relationship.
-The coefficients of each model describe the relationship of a predictor variable to its response when considering all other variables.
-Due to the transformations of the data, coefficients are individually interpreted in the same manner as the prior regression values were (see Section \ref{sec:exp_anal}); unlike the regression values,  coefficient ranges are not limited between -1 and 1.
-%, as each variable needs to be considered together.
-%The dimensionless nature of these variables each coefficient refers to its own variable's distribution's standard deviations.
-\newline
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{table}[]
@@
 \end{table}
-We look at the coefficients of categorical and continuous variables separately.
-This is done as the categorical variables have many coefficients, one for each category, whilst continuous variables have only one.
-The coefficient for categorical variables is summarised in Figure \ref{fig:violin}; illustrating the difference in the range as well as affect region and year could have on each model.
-Comparatively, the continuous variables coefficients are summarised in Table \ref{tab:contvar}.
-In terms of magnitude, GI region has the highest possible absolute value for each model.
-An important consideration is that region and year are binary, such that they are only equal to zero or the coefficient (as they will present as a value of 1 which will be multiplied by the coefficient); this means that, although region may have a strong relationship, it can be overshadowed by an extreme value of one of the continuous variables.
-The most notable difference between the continuous variables coefficients is the change from positive to negative values.
+
+%%%%%%%%%%%%%%%%
+The regression coefficients for the categorical variables year and GI region, and their interaction under each of the four models are depcited in \ref{fig:violin}.
+The coefficients for the continuous variables are summarised in Table \ref{tab:contvar}.
+In terms of magnitude, GI region has the highest possible absolute value for each model. The most notable
+
+% Search for where you mention quality and where you mention value as this does become misleading!!!
+%
+% This is partly because of the intercept
+%
+% It depends what else is in the model.
+%
+% This direct comparisons are 
+%
+% The correlated between input vars needs to be discussed earlier.
+%
+%difference between the continuous variables coefficients is the change from positive to negative values.
 This change occurs between the Models for Yield (Model 1 and 2) and the Models for value (Models 3 and 4); where all but the coefficient for area harvested had the opposite sign (see Table \ref{tab:contvar}).
@@
 %
-\subsubsection{Model comparisons: yield versus value}
-%
+
 \begin{figure}
@@
 %
-% Why compare models?
+% KM: Are these observed or modelled values?
 %
-Directly comparing response variables, how crop value changes with yield, also allows an indirect comparison between the response variables and resource use.
-We do this through using known relationships of response variables to their predictors.
-These relationships are described by the coefficients.
-Resource use is described by the predictor variables (through water used and scope one emissions), because of this we can observe the response variables somewhat interchangeably with the predictors - although caution should be taken to view them sceptically and alongside the influence of their coefficients.
-As the predictors are known to have a strong positive correlation with each other, they will tend toward increasing and decreasing together (but not at the same rates).
-It is also important to consider the interactions of predictor variables when comparing the response variables that are ratios of area.
-Furthermore, these comparisons require the consideration of the covariates, in this case: area harvested, year and region.
-\newline
-Observing Figure \ref{fig:yield_vs_value} shows an almost discrete difference between vineyards in 'Hot' areas than other regions.
+% Replace with star charts?
+%
+%Observing Figure \ref{fig:yield_vs_value} shows an almost discrete difference between vineyards in 'Hot' areas than other regions.
 Comparing Figure \ref{fig:yield_vs_value} to Figure \ref{fig:yield_vs_value_area} shows almost opposing trends.
@@
 This is more visible when comparing both graphs to the map of regional averages for response variables, see Figure \ref{fig:map}.
-There is a notable change between regional averages when looking at yield versus value.
+
+%\newline
+%\newline
+
+%There is a notable change between regional averages when looking at yield versus value.
 Through the coefficients we can deduce that: this difference is also a difference between more resources used for the raw response variables; and a difference between overall resource use and the size of the vineyard when considering the response variables as a ratio to area.
@@
 The interaction between year and GI Region likely accounted for localised events such as bushfires, which would be impactful, but only at a local level in both time and space.
-\subsection{Model Validation}
+
 \begin{table}[]
@@
   \end{table}
-To validate the performance of these models k-fold cross validation was used.
-This was done using 10 folds, $k=10$, repeated 100 times.
-The models performed similarly to their original counterparts (see Table \ref{tab:kfold}).
 
@@
 \newline
+\newline
 The negative trend between size and average sales price could be a side effect of supply versus demand, especially when looking at the level of difference in production of some vineyards.
@@
 The relationship between yield, value and area was not simply about efficiently producing the most grapes; sales price and by association grape quality, are integral to the profitability, and this is strongly linked to resource-use and thus the longevity and sustainability of a vineyard.
+\newline
 \newline
@@
 \newline
+\newline
 Some regions are known for their quality and may have a bias in purchasers or bring greater demand regardless of similarities and differences in production of quality of grapes \citep{hallidayAustralianWineEncyclopedia2009}.
@@
 %
+\newline
 \newline
@@
 \newline
+\newline
  %The possible reason behind this effect was that different strategies are likely employed between different regions, where some regions target the mass production of cheaper grapes over quality.
@@
 Attempts to fully explain small variations was always overshadowed by the dramatic differences in regional trends.
+\newline
 \newline

### 2023-08-17 article.tex
changed to paragraph formating

@@
 In the Australian context, these include biosecurity, climate and international market demands.
-\newline
-\newline
+\par
 There is an extensive amount of research into the effects of a variety of factors on grape quality and yield \citep{heFruitYieldPrediction2022,laurentLocalInfluenceClimate2022,liakosMachineLearningAgriculture2018}.
@@
 Wine Australia is an Australian Government statutory authority governed by the Wine Australia Act 2013 \citep{WineAustraliaAct2019}.
-\newline
-\newline
+\par
 Predictor variables used in this analysis included yield, defined as the total tonnes of grapes harvested, and quality, defined as the average sale price.
@@
 Values were compared in this manner to observe how economies of scale affect the use of resources.
-\newline
-\newline
+\par
 Data obtained from Wine Australia were collected via phone surveys and included: total tonnes purchased, average price per tonne and yearly change in price for region and grape varietal.
@@
 To enable direct comparisons between fuels, fuel use was converted to tonnes of Carbon Dioxide equivalent and collectively referenced to as emissions.
-\newline
-\newline
+\par
 Average sale price was an optional field in the Sustainable Winegrowing Australia's dataset. 
@@
 These subsets meant that the data would be limited to samples which had recorded values for the response variables (see Table \ref{tab:tab1}), where every sample had a recorded value for yield but not average price of sale per tonne.
-\newline
-\newline
+\par
 The first subset of data (used for Model 1 and Model 2, see Table\ref{tab:tab1}) contained 5298 samples spanning the period from 2012 to 2022, covering 55 GI Regions and 1261 separate vineyards.
-\newline
-\newline
+\par
 The second subset of data (used for Model 3 and Model 4, see Table\ref{tab:tab1}) contained 2878 samples spanning the period from 2015 to 2022, covering 51 GI Regions and 944 separate vineyards.Average price of sale per tonne was extracted from both Wine Australia (1842 values) and Sustainable Winegrowing Australia (remaining 1036 values).
-\newline
-\newline
+\par
 Additional variables were considered for analysis but were excluded due to being either underreported or had insignificant contributions to model accuracies.
@@
 Variables considered but ultimately removed due to a lack of significant contributions to models, included the use of renewable energy, contractor use, and pressures such as frost, fire and disease.
-\newline
-\newline
+\par
 Data preprocessing was conducted prior to analysis using the Python programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
@@
 The transformation of fuel use into scope one emissions was done using the equation given from the Australian National Greenhouse Accounts Factors, shown as
-\newline
-\newline
+\par
 \begin{equation}
@@
 \end{equation}
-\newline
-\newline
+\par
 was used to convert the quantity of fuel in litres, $Q$, using a prescribed Energy Content, $EC$, and emission factors of scope one, $EF1$, and scope three, $EF3$, to tonnes of Carbon Dioxide Emission equivalent, $tCO2e$ \citep{departmentofclimatechangeenergytheenvironmentandwaterAustralianNationalGreenhouse2022}.
@@
 Both Wine Australia and Sustainable Winegrowing Australia used the same GI Region format to describe location.
-\newline
-\newline
+\par
 The site of a vineyard predetermines several physical parameters such as climate, geology and soil, making location a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
@@
 Pearson Correlation Coefficients of the transformed, centered and scaled variables are shown in Table \ref{tab:pearson_correlation}. 
-\newline
-\newline
+\par
 All relationships were found to be statistically significant (P $<$ 2.200E-16).
@@
 A full list of regression coefficients 95\% CIs and p-values for each of the four models is provided in the appendix. % You gotta add these to the appendix in a more readable fashion
-\newline
-\newline
+\par
 
@@
 The 'Hot Very Dry' vineyards (see Figure \ref{fig:yield_vs_value_area}) These vineyards would be very competitive with only a minor increase to sale price, possibly outperforming other regions.
-\newline
-\newline
+\par
 The negative trend between size and average sales price could be a side effect of supply versus demand, especially when looking at the level of difference in production of some vineyards.
@@
 The relationship between yield, value and area was not simply about efficiently producing the most grapes; sales price and by association grape quality, are integral to the profitability, and this is strongly linked to resource-use and thus the longevity and sustainability of a vineyard.
-\newline
-\newline
+\par
 There are important considerations unique to winegrowing compared to other agricultural industries.
@@
 % ref
-\newline
-\newline
+\par
 Some regions are known for their quality and may have a bias in purchasers or bring greater demand regardless of similarities and differences in production of quality of grapes \citep{hallidayAustralianWineEncyclopedia2009}.
@@
 %
-\newline
-\newline
+\par
 Due to regional differences, different strategies are also  employed across different regions, such as some regions targeting mass production over quality.
@@
 This limitation could be overcome by incorporating the profitability of vineyards, comparing the financial success of working at different operational scales.
-\newline
-\newline
+\par
  %The possible reason behind this effect was that different strategies are likely employed between different regions, where some regions target the mass production of cheaper grapes over quality.
@@
 Attempts to fully explain small variations was always overshadowed by the dramatic differences in regional trends.
-\newline
-\newline
+\par
 Having more data for each region would also be beneficial, allowing greater comparison between regions.

### 2023-08-17 article.tex

@@
 \begin{abstract}
+% The word limit needs to be checked.
 When strategies for a sustainable winegrowing industry are assessed, there is a trade-off between balancing the amount of resources invested and the resultant yield and quality of the produce.
@@
 The dataset used for this analysis includes data collected for the past 10 years from a multitude of vineyards located over a diverse range of Australian winegrowing regions.
-Yield and quality (measured in sale price) was modelled to resource factors related to water usage and emissions.
+Yield and quality (measured as a ratio of sale price to area) was modelled to resource factors related to water usage and emissions.
 The analysis confirmed an expected strong relationship between size and resource use, with the overall space of a vineyard and its access to resources greatly determining the upper limit of potential yield.

### 2023-08-17 analysis.r
working yearly trends

@@ -366,6 +366,51 @@ pdf("yield_verse_value_by_area.pdf")
 grid.arrange(temp_yield_vs_value, rain_yield_vs_value, ncol=2)
 dev.off()
 
+#####%%%%%%%%%%%%%%%%%%%%
+# Yearly variation plots
+########################
+
+temp_df <- data.frame(summary(model1)$coefficients)
+temp_df <- temp_df[grep("year", rownames(temp_df)), ]
+estimates <- data.frame(temp_df[, c("Estimate", "Std..Error")])
+#rownames(estimates) <- rownames(temp_df)
+#names(estimates) <- c("model_1_est", "model_1_er")
+
+estimates$model <- rep("model1", length(rownames(estimates)))
+
+#temp_df <- data.frame(summary(model2)$coefficients)
+#temp_df <- temp_df[grep("year", rownames(temp_df)), ]
+#estimates$model2  <- temp_df$Estimate
+temp_df <- data.frame(summary(model3)$coefficients)
+temp_df <- temp_df[grep("year", rownames(temp_df)), ]
+temp_df <- data.frame(temp_df[, c("Estimate", "Std..Error")])
+for (i in rownames(estimates)[!(rownames(estimates) %in% rownames(temp_df))]) {
+    temp_df[i,] <- NA
+}
+temp_df$model <- rep("model2", length(rownames(temp_df)))
+
+estimates <- rbind(estimates, temp_df)
+
+library(stringi)
+estimates$year <- stri_extract_first_regex(rownames(estimates), "[0-9]+")
+estimates$year <- as.integer(estimates$year)
+
+# Default line plot
+p<- ggplot(estimates, aes(x = year, y = Estimate, group = model, color=model)) + 
+  geom_line() +
+  geom_point()+
+  geom_errorbar(aes(ymin=Estimate-Std..Error, ymax=Estimate+Std..Error), width=.2,
+                 position=position_dodge(0.05))
+print(p)
+# Finished line plot
+p+labs(title="Tooth length per dose", x="Dose (mg)", y = "Length")+
+   theme_classic() +
+   scale_color_manual(values=c('#999999','#E69F00'))
+
+
+
+
+
 #######################
 # Lets make some maps #
 nt <- read.csv("no_trans.csv")

### 2023-08-17 analysis.r
Nice yearly coefficient graphs

@@ -3,6 +3,11 @@ library(gridExtra)
 library(viridis)
 library(sf)
 library(cowplot)
+library(stringi)
+
+mean.t <- function(x) {
+    mean(x, na.rm = TRUE)
+}
 
 # data_prep.py has to be executed in order for df.csv to exist
 set.seed(100)
@@ -370,46 +375,59 @@ dev.off()
 # Yearly variation plots
 ########################
 
+# We construct a data frame for the coefficients
+# We start with the first model
 temp_df <- data.frame(summary(model1)$coefficients)
 temp_df <- temp_df[grep("year", rownames(temp_df)), ]
 estimates <- data.frame(temp_df[, c("Estimate", "Std..Error")])
-#rownames(estimates) <- rownames(temp_df)
-#names(estimates) <- c("model_1_est", "model_1_er")
-
 estimates$model <- rep("model1", length(rownames(estimates)))
 
-#temp_df <- data.frame(summary(model2)$coefficients)
-#temp_df <- temp_df[grep("year", rownames(temp_df)), ]
-#estimates$model2  <- temp_df$Estimate
-temp_df <- data.frame(summary(model3)$coefficients)
+# The second model has coefficients averaged as there are so many per year.
+temp_df <- data.frame(summary(model2)$coefficients)
 temp_df <- temp_df[grep("year", rownames(temp_df)), ]
 temp_df <- data.frame(temp_df[, c("Estimate", "Std..Error")])
-for (i in rownames(estimates)[!(rownames(estimates) %in% rownames(temp_df))]) {
-    temp_df[i,] <- NA
-}
+
+temp_df$year <- stri_extract_first_regex(rownames(temp_df), "[0-9]+")
+temp_df <- aggregate(temp_df, list(temp_df$year), FUN=mean.t)
+rownames(temp_df) <- temp_df$Group.1
+temp_df <- data.frame(temp_df[, c("Estimate", "Std..Error")])
 temp_df$model <- rep("model2", length(rownames(temp_df)))
 
 estimates <- rbind(estimates, temp_df)
 
-library(stringi)
+# The third model is integrated somewhat the same as the first
+temp_df <- data.frame(summary(model3)$coefficients)
+temp_df <- temp_df[grep("year", rownames(temp_df)), ]
+temp_df <- data.frame(temp_df[, c("Estimate", "Std..Error")])
[…]

### 2023-08-17 analysis.r
added maps to separate R file.

@@ -428,64 +428,3 @@ p<- ggplot(estimates, aes(x = year, y = Estimate, group = model, color=model)) +
                  position=position_dodge(0.05)) +
     theme_classic()
 print(p)
-
-#######################
-# Lets make some maps #
-nt <- read.csv("no_trans.csv")
-v <- st_read("data/Wine_Geographical_Indications_Australia_2022/Wine_Geographical_Indications_Australia_2022.shp")
-aus <- st_read("data/STE_2021_AUST_SHP_GDA2020/STE_2021_AUST_GDA2020.shp")
-
-v$ha_tonnes_grapes_harvested <- 0
-v$ha_value <- 0
-
-aggs.ha_tonnes_grapes_harvested <- aggregate(
-    nt$ha_tonnes_grapes_harvested * 10000
-    , list(nt$giregion), FUN = mean.t)
-aggs.ha_value <- aggregate(
-    nt$ha_value
-    , list(nt$giregion)
-    , FUN=mean.t)
-
-for (region in unique(df$giregion)){
-    v[v$GI_NAME == region
-    , "ha_tonnes_grapes_harvested"] <- aggs.ha_tonnes_grapes_harvested[
-    aggs.ha_tonnes_grapes_harvested$Group.1 == region
-    , ]$x
-    v[
-        v$GI_NAME==region
-        , "ha_value"] <- aggs.ha_value[
-            aggs.ha_value$Group.1 == region
-            , ]$x
-}
-
-# note these are the averages.
-#ggdraw() + 
-value_map <- ggplot() +
-    ggtitle("Average Price Per Ha ") +
-    geom_sf(data = v, aes(fill = ha_value)) +
-    geom_sf(data=aus, fill = NA) +
-    xlim(113, 154) +
-    ylim(-45, -25) +
-    scale_fill_gradientn(colours=rev(viridis(3))
-                         , name=""
-                         , na.value = "grey100") +
-    theme_void()
-
-yield_map <-  ggplot() +
-    xlim(113, 154) +
-    ylim(-45, -25) +
-    theme_void() +
-        ggtitle("Average Yield (tonnes) Per Ha") +
-    geom_sf(data = v
-        , aes(fill = ha_tonnes_grapes_harvested)) +
-    geom_sf(data=aus, fill = NA) +
-    scale_fill_gradientn(colours=rev(viridis(3))
-                         , name=""
-                         , na.value = "grey100")
-
-pdf("my_map.pdf")
[…]

### 2023-08-17 maps.r

@@ -1,0 +1,60 @@
+#######################
+# Lets make some maps #
+nt <- read.csv("no_trans.csv")
+v <- st_read("data/Wine_Geographical_Indications_Australia_2022/Wine_Geographical_Indications_Australia_2022.shp")
+aus <- st_read("data/STE_2021_AUST_SHP_GDA2020/STE_2021_AUST_GDA2020.shp")
+
+v$ha_tonnes_grapes_harvested <- 0
+v$ha_value <- 0
+
+aggs.ha_tonnes_grapes_harvested <- aggregate(
+    nt$ha_tonnes_grapes_harvested * 10000
+    , list(nt$giregion), FUN = mean.t)
+aggs.ha_value <- aggregate(
+    nt$ha_value
+    , list(nt$giregion)
+    , FUN=mean.t)
+
+for (region in unique(df$giregion)){
+    v[v$GI_NAME == region
+    , "ha_tonnes_grapes_harvested"] <- aggs.ha_tonnes_grapes_harvested[
+    aggs.ha_tonnes_grapes_harvested$Group.1 == region
+    , ]$x
+    v[
+        v$GI_NAME==region
+        , "ha_value"] <- aggs.ha_value[
+            aggs.ha_value$Group.1 == region
+            , ]$x
+}
+
+# note these are the averages.
+#ggdraw() + 
+value_map <- ggplot() +
+    ggtitle("Average Price Per Ha ") +
+    geom_sf(data = v, aes(fill = ha_value)) +
+    geom_sf(data=aus, fill = NA) +
+    xlim(113, 154) +
+    ylim(-45, -25) +
+    scale_fill_gradientn(colours=rev(viridis(3))
+                         , name=""
+                         , na.value = "grey100") +
+    theme_void()
+
+yield_map <-  ggplot() +
+    xlim(113, 154) +
+    ylim(-45, -25) +
+    theme_void() +
+        ggtitle("Average Yield (tonnes) Per Ha") +
+    geom_sf(data = v
+        , aes(fill = ha_tonnes_grapes_harvested)) +
+    geom_sf(data=aus, fill = NA) +
+    scale_fill_gradientn(colours=rev(viridis(3))
+                         , name=""
+                         , na.value = "grey100")
+
+pdf("my_map.pdf")
+grid.arrange(value_map, yield_map, ncol=2, nrow=1)
+dev.off()
+
+# below is just a way to grab the coef from a linear model.
+#summary(model2)$coefficients

### 2023-08-17 analysis.r
added model5 (no area)

@@ -29,7 +29,10 @@ model4 <- lm(
     ha_value ~
     area_harvested * ((scope1) + water_used) + data_year_id:giregion
     , data = df)
-
+model5 <- lm(
+    ha_value ~
+    scope1 + water_used + data_year_id:giregion
+    , data = df)
 # We review our first model
 
 # MODEL 1
@@ -176,6 +179,28 @@ ggplot(
 #  RMSE     Rsquared   MAE      
 #  0.22349  0.9500051  0.1279485
 #
+
+# MODEL 1
+anova(model5)
+summary(model5)
+model5_residuals <- residuals(model5)
+model5_fitted <- fitted(model5)
+ggplot(
+    data.frame(
+        y = model5_residuals)
+        , aes(sample = y)
+       ) +
+       stat_qq() +
+       stat_qq_line()
+
+ggplot(
+    data.frame(
+        y = model5_residuals
+        , x = model5_fitted)
+    , aes(x = x, y = y)
+    ) +
+    geom_point() +
+    geom_hline(yintercept = 0)
 
 ############################ #
 # We graph the coefficients! #

### 2023-08-17 maps.r
changed map to income per vineyard by region

@@ -1,59 +1,50 @@
 #######################
 # Lets make some maps #
+
+library(ggplot2)
+library(gridExtra)
+library(viridis)
+library(sf)
+library(cowplot)
+library(stringi)
+
+mean.t <- function(x) {
+    mean(x, na.rm = TRUE)
+}
+
 nt <- read.csv("no_trans.csv")
 v <- st_read("data/Wine_Geographical_Indications_Australia_2022/Wine_Geographical_Indications_Australia_2022.shp")
 aus <- st_read("data/STE_2021_AUST_SHP_GDA2020/STE_2021_AUST_GDA2020.shp")
 
-v$ha_tonnes_grapes_harvested <- 0
-v$ha_value <- 0
+v$value <- 0
 
-aggs.ha_tonnes_grapes_harvested <- aggregate(
-    nt$ha_tonnes_grapes_harvested * 10000
+agg_value <- aggregate(
+    nt$value
     , list(nt$giregion), FUN = mean.t)
-aggs.ha_value <- aggregate(
-    nt$ha_value
-    , list(nt$giregion)
-    , FUN=mean.t)
 
-for (region in unique(df$giregion)){
+agg_value$x <- agg_value$x / 1000000
+
+for (region in unique(nt$giregion)){
     v[v$GI_NAME == region
-    , "ha_tonnes_grapes_harvested"] <- aggs.ha_tonnes_grapes_harvested[
-    aggs.ha_tonnes_grapes_harvested$Group.1 == region
+    , "value"] <- agg_value[
+    agg_value$Group.1 == region
     , ]$x
-    v[
-        v$GI_NAME==region
-        , "ha_value"] <- aggs.ha_value[
-            aggs.ha_value$Group.1 == region
-            , ]$x
 }
 
 # note these are the averages.
-#ggdraw() + 
 value_map <- ggplot() +
-    ggtitle("Average Price Per Ha ") +
-    geom_sf(data = v, aes(fill = ha_value)) +
+    ggtitle("Average Vineyard Income by Region") +
+    geom_sf(data = v, aes(fill = value)) +
     geom_sf(data=aus, fill = NA) +
     xlim(113, 154) +
     ylim(-45, -25) +
-    scale_fill_gradientn(colours=rev(viridis(3))
-                         , name=""
+    scale_fill_gradientn(colours = rev(viridis(3))
+                         , name = ""
                          , na.value = "grey100") +
     theme_void()
 
-yield_map <-  ggplot() +
[…]

### 2023-08-17 article.tex
Corrcted non-transformed PPC table to transformed

@@
           \toprule
-        \textbf{Variable} & \textbf{Yield} & \textbf{Area Harvested} & \textbf{Water Used} & \textbf{Scope One Emissions} & $\textbf{Yield}\over \textbf{Area}$ & \textbf{Average Sale Price} & $\textbf{Average Sale Price}\over \textbf{Area Harvested}$ \\ \midrule
-        \textbf{Yield} & 1.00E+00 & 7.44E-01 & -4.31E-03 & 7.29E-01 & 3.50E-01 & -2.26E-01 & -1.64E-01 \\
-        \textbf{Area Harvested} & 7.44E-01 & 1.00E+00 & -5.33E-03 & 8.92E-01 & 7.85E-02 & -1.18E-01 & -2.04E-01 \\
-        \textbf{Water Used} & -4.31E-03 & -5.33E-03 & 1.00E+00 & -1.93E-03 & -5.60E-03 & -3.56E-02 & -2.67E-02 \\
-        \textbf{Scope One Emissions} & 7.29E-01 & 8.92E-01 & -1.93E-03 & 1.00E+00 & 9.36E-02 & -9.42E-02 & -1.93E-01 \\
-        $\textbf{Yield}\over \textbf{Area}$ & 3.50E-01 & 7.85E-02 & -5.60E-03 & 9.36E-02 & 1.00E+00 & -4.85E-01 & -1.70E-01 \\
-        \textbf{Average Sale Price} & -2.26E-01 & -1.18E-01 & -3.56E-02 & -9.42E-02 & -4.85E-01 & 1.00E+00 & 4.73E-01 \\
-        $\textbf{Average Sale Price} \over \textbf{Area Harvested}$ & -1.64E-01 & -2.04E-01 & -2.67E-02 & -1.93E-01 & -1.70E-01 & 4.73E-01 & 1.00E+00 \\ \bottomrule
+          \textbf{} & \textbf{Yield} & \textbf{Area Harvested} & \textbf{Water Used} & \textbf{\begin{tabular}[c]{@{}c@{}}Scope One\\ Emissions\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Yield by\\ Area\end{tabular}} & \textbf{Average Price} & \textbf{\begin{tabular}[c]{@{}c@{}}Average Price\\ by Area\end{tabular}} \\ \midrule
+          Yield & 1.00 & 0.75 & 0.00 & 0.72 & 0.35 & 0.40 & -0.23 \\
+          Area Harvested & 0.75 & 1.00 & -0.01 & 0.88 & 0.09 & 0.86 & -0.13 \\
+          Water Used & 0.00 & -0.01 & 1.00 & 0.00 & -0.01 & 0.05 & -0.03 \\
+          \begin{tabular}[c]{@{}c@{}}Scope One\\ Emissions\end{tabular} & 0.72 & 0.88 & 0.00 & 1.00 & 0.10 & 0.75 & -0.10 \\
+          \begin{tabular}[c]{@{}c@{}}Yield by\\ Area\end{tabular} & 0.35 & 0.09 & -0.01 & 0.10 & 1.00 & -0.09 & -0.48 \\
+          Average Price & 0.40 & 0.86 & 0.05 & 0.75 & -0.09 & 1.00 & 0.07 \\
+          \begin{tabular}[c]{@{}c@{}}Average Price\\ by Area\end{tabular} & -0.23 & -0.13 & -0.03 & -0.10 & -0.48 & 0.07 & 1.00 \\ \bottomrule
           \end{tabular}

### 2023-08-17 article.tex
added model 5 to first table - changed covariates

@@
     \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-    \textbf{Model 2} & ${ \textrm{Yield}}\over{ \textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ scope one Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * scope one Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular} \\
-    {1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-    \textbf{Model 3} & {$\textrm{Yield} {\times} \textrm{Average Sale Price}$} & \begin{tabular}[c]{@{}c@{}}Water Used\\ scope one Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & N/A \\
+    \textbf{Model 2} & ${ \textrm{Yield}}\over{ \textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ scope one Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested \end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * scope one Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular} \\
     \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-    \textbf{Model 4} & ${\textrm{Yield} \times \textrm{Average Sale Price}}\over{\textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ scope one Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\
-        Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * scope one Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular}
+    \textbf{Model 3} & {$\textrm{Yield} {\times} \textrm{Average Sale Price}$} & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope One Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & N/A\\
+    \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
+    \textbf{Model 4} & ${\textrm{Yield} \times \textrm{Average Sale Price}}\over{\textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope One Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\
+        \\ \end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * Scope One Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular}
+        \\
+        \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
+     \textbf{Model 5} & ${\textrm{Yield} \times \textrm{Average Sale Price}}\over{\textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope One Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}} Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}} Year * Region\end{tabular}
         \end{tabular}}

### 2023-08-17 article.tex
stitched methods paragraphs back together

@@
 \par
-The second subset of data (used for Model 3 and Model 4, see Table\ref{tab:tab1}) contained 2878 samples spanning the period from 2015 to 2022, covering 51 GI Regions and 944 separate vineyards.Average price of sale per tonne was extracted from both Wine Australia (1842 values) and Sustainable Winegrowing Australia (remaining 1036 values).
+The second subset of data (used for Model 3, Model 4 and Model 5, see Table\ref{tab:tab1}) contained 2878 samples spanning the period from 2015 to 2022, covering 51 GI Regions and 944 separate vineyards.Average price of sale per tonne was extracted from both Wine Australia (1842 values) and Sustainable Winegrowing Australia (remaining 1036 values).
 \par
@@
 %%
-\newline
+
+\par
 Differences in vineyard locations were captured through the use of Geographical Indicator Regions (GI Regions).
@@
 The climatic groups were used to illustrate similarities and differences occurring in areas larger than GI Regions.
-
+\par
 \subsection{Analysis}
-Pairwise Pearson Correlation Coefficients were calculated to assess the potential existence of linear relationships between the input and predicted variable.
-Four general linear models were then created (see Table \ref{tab:tab1})
-
-Both the Pearson Correlation Coefficients and General Linear Models were created using the R statistical programming language \citep{rcoreteamLanguageEnvironmentStatistical2021} and the Caret packages \citep{kuhnBuildingPredictiveModels2008}.
+Pairwise Pearson Correlation Coefficients were calculated to assess the potential existence of linear relationships between the input and predicted variables.
+To determine if a coefficient was indicative of a strong relationship, confidence intervals were used.
+P-values reflected the significance of a given correlation coefficient with statistical significance being declared when the associated value was lower than 0.05.
+Pairwise Pearson Correlation Coefficients were calculated for data on the original scale and for data as a logarithmic transform.
+Transforming data prior to calculating the coefficients changes several things.
+The logarithmic transform of the data alters the interpretation of the coefficients to percentage change; a coefficient will be indicative of the change in percentage of one variable compared to the other, scaling by standard deviation also changes this interpretation to be a percentage of that variables standard deviation.
+When considering the logarithmically transformed variables, a coefficient of 1 would indicate that the change of one variable by one percentage of its standard deviation would correlate to the other variable changing by one percent of its own standard deviation.
+The importance of this is the dimensionless nature of these relationships and that it can be translated directly to any vineyard's case that has a well known distribution.
+\par
+Five general linear models were created (see Table \ref{tab:tab1}).
 General Linear Models were chosen as they offer the ability to produce statistical models that are explicit in the relationships between predictors and response variables.
-
-General Linear Models also allow the exploration of interactions between predictors and present easily comparable differences in the influence and magnitude of relationships.
-
+General Linear Models also allowed the exploration of interactions between predictors and allow for easily comparable differences in the influence and magnitude of relationships.
+Model fit was measured in $R^2$ and adjusted $R^2$ as well as F statistics.
+T-tests were used to determine if predictors significantly contributed to their models when accounting for other variables, showing which specific years and areas contributed significantly.
+Both the Pearson Correlation Coefficients and General Linear Models were created using the R statistical programming language \citep{rcoreteamLanguageEnvironmentStatistical2021} with the Caret package \citep{kuhnBuildingPredictiveModels2008}.
+\par
 A variety of alternate methods were also explored, including splines, hierarchical regression, General Additive Models, and Generalised Linear Models.
 These alternative approaches were not used as final models due to offering no further insights or improvements in accuracy.
-
+\par
 \subsection{Model Validation}
-Models were validated using K-fold cross validation calculated through the R Caret Package \citep{kuhnBuildingPredictiveModels2008}.
+Models were validated using K-fold cross validation calculated.
 K-fold cross validation works by removing a subset of data from the sample used to train models and then predicts those variables to determine how sensitive the model is to changes in the sample data.
 For this analysis each model was validated using 10 folds, repeated 100 times.
-% A reference for this is absolutely needed.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% This is my results to methods prison
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Values for these coefficients reflect the linear relation between two variables, on a scale between -1 and 1; the magnitude and sign of a coefficient indicates the strength of the relation, and whether the relation is positive or negative respectively.
-This was undertaken for data on the original scale and for data as a logarithmic transform.
-Transforming data prior to calculating the coefficients changes several things: The logarithmic transform of the data alters the interpretation of the coefficients to percentage change - a coefficient will be indicative of the change in percentage of one variable compared to the other; scaling by standard deviation also changes this interpretation to be a percentage of that variables standard deviation.
-When considering the logarithmically transformed variables, a coefficient of 1 would indicate that the change of one variable by one percentage of its standard deviation would correlate to the other variable changing by one percent of its own standard deviation.
-The importance of this is the dimensionless nature of these relationships and that it can be translated directly to any vineyard's case that has a well known distribution.
-
-To determine if a coefficient was indicative of a strong relationship, confidence intervals were used.
-P-values reflected the significance of a given correlation coefficient when considering its relation to sample size via its incorporation as an element of standard error.
-
-Model fit was also measured in $R^2$ and adjusted $R^2$, F statistics
-
-statistical significance p-values declared the associated was then 0.05
-
-T-tests were used to determine if predictors significantly contributed to their models when accounting for other variables; this allowed a more granular examination of interactions and factors within categorical variables, showing which specific years and areas contributed significantly and which did not (the appendix contains a comprehensive list of these values).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
+\par
 \section{Results}
@@
  Table \ref{tab:summary} shows the summary statistics of each variable used in Models 1 to 4 (See \ref{tab:tab1}).
-
 \begin{table}[]

### 2023-08-17 analysis.r
yearly plots and model 5

@@ -26,11 +26,11 @@ model3 <- lm(
     data_year_id + giregion + area_harvested + water_used + scope1
     , data = df)
 model4 <- lm(
-    ha_value ~
+    average_per_tonne ~
     area_harvested * ((scope1) + water_used) + data_year_id:giregion
     , data = df)
 model5 <- lm(
-    ha_value ~
+    average_per_tonne ~
     scope1 + water_used + data_year_id:giregion
     , data = df)
 # We review our first model
@@ -442,6 +442,20 @@ rownames(temp_df) <- paste0(rownames(temp_df), "a")
 
 estimates <- rbind(estimates, temp_df)
 
+# The fith model is similar to the second and requires averaging
+temp_df <- data.frame(summary(model5)$coefficients)
+temp_df <- temp_df[grep("year", rownames(temp_df)), ]
+temp_df <- data.frame(temp_df[, c("Estimate", "Std..Error")])
+
+temp_df$year <- stri_extract_first_regex(rownames(temp_df), "[0-9]+")
+temp_df <- aggregate(temp_df, list(temp_df$year), FUN=mean.t)
+rownames(temp_df) <- temp_df$Group.1
+temp_df <- data.frame(temp_df[, c("Estimate", "Std..Error")])
+temp_df$model <- rep("model4", length(rownames(temp_df)))
+rownames(temp_df) <- paste0(rownames(temp_df), "b")
+
+estimates <- rbind(estimates, temp_df)
+
 estimates$year <- stri_extract_first_regex(rownames(estimates), "[0-9]+")
 estimates$year <- as.integer(estimates$year)
 
@@ -452,4 +466,7 @@ p<- ggplot(estimates, aes(x = year, y = Estimate, group = model, color=model)) +
   geom_errorbar(aes(ymin=Estimate-Std..Error, ymax=Estimate+Std..Error), width=.2,
                  position=position_dodge(0.05)) +
     theme_classic()
+
+pdf("yearly_plots.pdf")
 print(p)
+dev.off()

### 2023-08-17 article.tex
coefficients and ppc results

@@
     \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-    \textbf{Model 4} & ${\textrm{Yield} \times \textrm{Average Sale Price}}\over{\textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope One Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\
+    \textbf{Model 4} & Average Sale Price & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope One Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\
         \\ \end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * Scope One Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular}
@@
         \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-     \textbf{Model 5} & ${\textrm{Yield} \times \textrm{Average Sale Price}}\over{\textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope One Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}} Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}} Year * Region\end{tabular}
+     \textbf{Model 5} &  Average Sale Price & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope One Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}} Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}} Year * Region\end{tabular}
         \end{tabular}}
@@
 \par
-Predictor variables used in this analysis included yield, defined as the total tonnes of grapes harvested, and quality, defined as the average sale price.
+Predictor variables used in this analysis included yield, defined as the total tonnes of grapes harvested, and quality, defined as average sale price of grapes.
 It is acknowledged that quality can be defined in a variety of ways, for example by the grapes': aroma, chemical composition and color \citep{kasimatiPredictingGrapeSugar2022,mejeanperrotDecisionsupportSystemPredict2022,suarezAssessingWineGrape2021}.
@@
 The first subset contained all vineyards and was used for two models (Model 1 and Model 2, see Table \ref{tab:tab1}).
-The second subset contained vineyards which either recorded a value for average price of sale per tonne through Sustainable Winegrowing Australia, or were within a region with an average price of sale recorded by Wine Australia; this subset was used for two further models (Models 3 and 4, see Table \ref{tab:tab1}.
+The second subset contained vineyards which either recorded a value for average price of sale per tonne through Sustainable Winegrowing Australia, or were within a region with an average price of sale recorded by Wine Australia; this subset was used for three further models (Models 3, 4 and 5, see Table \ref{tab:tab1}.
 These subsets meant that the data would be limited to samples which had recorded values for the response variables (see Table \ref{tab:tab1}), where every sample had a recorded value for yield but not average price of sale per tonne.
@@
 was used to convert the quantity of fuel in litres, $Q$, using a prescribed Energy Content, $EC$, and emission factors of scope one, $EF1$, and scope three, $EF3$, to tonnes of Carbon Dioxide Emission equivalent, $tCO2e$ \citep{departmentofclimatechangeenergytheenvironmentandwaterAustralianNationalGreenhouse2022}.
-\newline
-%%
-%%
-%%
-% This is where the maps go!
-%%
-%%
-%%
+\par
+
+\begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{my_map.pdf}}
+  \caption{Map of vineyard average income for each of the used GI Regions.}\label{fig:map}
+\end{figure}%%
 
@@
 \section{Results}
-\subsection{Exploratory Analysis} \label{sec:exp_anal}
- Table \ref{tab:summary} shows the summary statistics of each variable used in Models 1 to 4 (See \ref{tab:tab1}).
+\subsection{Exploratory Analysis}\label{sec:exp_anal}
+ Table \ref{tab:summary} shows the summary statistics of each variable in its original units.
+The range of these values shows the level of difference between some vineyards, with operations differing by orders of magnitude in size, yield and average price of sale (See Table \ref{tab:tab1}).
+\par
 \begin{table}[]
   \caption{Summary statistics of each continuous variable.}\label{tab:summary}
-      \resizebox{\textwidth}{!}{
+      % \resizebox{\textwidth}{!}{
         \begin{tabular}{@{}cllll@{}}
@@
           \textbf{Variable} & \textbf{Mean} & \textbf{\begin{tabular}[c]{@{}l@{}}Standard\\ Deviation\end{tabular}} & \textbf{Minimum} & \textbf{Maximum} \\ \midrule
-          {Yield} & 7.757E+02 & 2.179E+03 & 1.000E+00 & 7.231E+04 \\
+          {Yield (tonnes)} & 7.757E+02 & 2.179E+03 & 1.000E+00 & 7.231E+04 \\
           \multicolumn{1}{l}{} &  &  &  &  \\
-          {Area Harvested} & 6.670E+05 & 1.337E+06 & 7.000E+02 & 2.436E+07 \\
+          {Area Harvested ($m^2$)} & 6.670E+05 & 1.337E+06 & 7.000E+02 & 2.436E+07 \\
           \multicolumn{1}{l}{} &  &  &  &  \\
-          {Water Used} & 7.471E+06 & 5.646E+08 & 1.000E+00 & 4.268E+10 \\
+          {Water Used (ML)} & 7.471E+06 & 5.646E+08 & 1.000E+00 & 4.268E+10 \\
           \multicolumn{1}{l}{} &  &  &  &  \\
-          {Scope One Emissions} & 4.173E+04 & 8.571E+04 & 6.755E+00 & 2.110E+06 \\
+          {Scope One Emissions ($tCO_{2}e$)} & 4.173E+04 & 8.571E+04 & 6.755E+00 & 2.110E+06 \\
           \multicolumn{1}{l}{} &  &  &  &  \\
-          ${Yield}\over \textbf{Area}$ & 1.009E+01 & 8.127E+00 & 4.000E-02 & 8.634E+01 \\
+          $\mbox{Yield (tonnes)}\over \mbox{Area (m}^2)$ & 1.009E+01 & 8.127E+00 & 4.000E-02 & 8.634E+01 \\
           \multicolumn{1}{l}{} &  &  &  &  \\
-          {Average Sale Price} & 1.477E+03 & 9.216E+02 & 1.600E+02 & 2.600E+04 \\
+          {Average Sale Price (AUD)} & 1.477E+03 & 9.216E+02 & 1.600E+02 & 2.600E+04 \\
           \multicolumn{1}{l}{} &  &  &  &  \\
-          ${Average Sale Price} \over \textbf{Area Harvested}$ & 1.347E+02 & 5.711E+02 & 1.753E-01 & 2.979E+04 \\ \bottomrule
+          $\mbox{Average Sale Price (AUD)} \over \mbox{Area Harvested (m}^2)$ & 1.347E+02 & 5.711E+02 & 1.753E-01 & 2.979E+04 \\ \bottomrule
           \end{tabular}
-  }
+  % }
   \end{table}
+
+Pearson Correlation Coefficients of the transformed, centered and scaled variables are shown in Table \ref{tab:pearson_correlation}.
+All correlations were found to be statistically significant (P $<$ 2.200E-16).
+Except for 'average price' all variables are positively correlated.
+It can be considered that more resources and area are likely to lead to greater yields.
+However, the negative correlation between resource use and average sale price, although not as strong, indicates the possibility that resource consumption alone is not determining factor for grape quality.
+\par
 \begin{table}[]
@@
           \toprule
-          \textbf{} & \textbf{Yield} & \textbf{Area Harvested} & \textbf{Water Used} & \textbf{\begin{tabular}[c]{@{}c@{}}Scope One\\ Emissions\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Yield by\\ Area\end{tabular}} & \textbf{Average Price} & \textbf{\begin{tabular}[c]{@{}c@{}}Average Price\\ by Area\end{tabular}} \\ \midrule
-          Yield & 1.00 & 0.75 & 0.00 & 0.72 & 0.35 & 0.40 & -0.23 \\
-          Area Harvested & 0.75 & 1.00 & -0.01 & 0.88 & 0.09 & 0.86 & -0.13 \\
-          Water Used & 0.00 & -0.01 & 1.00 & 0.00 & -0.01 & 0.05 & -0.03 \\
-          \begin{tabular}[c]{@{}c@{}}Scope One\\ Emissions\end{tabular} & 0.72 & 0.88 & 0.00 & 1.00 & 0.10 & 0.75 & -0.10 \\
-          \begin{tabular}[c]{@{}c@{}}Yield by\\ Area\end{tabular} & 0.35 & 0.09 & -0.01 & 0.10 & 1.00 & -0.09 & -0.48 \\
-          Average Price & 0.40 & 0.86 & 0.05 & 0.75 & -0.09 & 1.00 & 0.07 \\
-          \begin{tabular}[c]{@{}c@{}}Average Price\\ by Area\end{tabular} & -0.23 & -0.13 & -0.03 & -0.10 & -0.48 & 0.07 & 1.00 \\ \bottomrule
-          \end{tabular}
-}
+          \textbf{} & \multicolumn{1}{c}{\textbf{Yield}} & \multicolumn{1}{c}{\textbf{Area Harvested}} & \multicolumn{1}{c}{\textbf{Water Used}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Scope One\\ Emissions\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Yield by\\ Area\end{tabular}}} & \multicolumn{1}{c}{\textbf{Average Price}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Average Price\\ by Area\end{tabular}}} \\ \midrule
+          Yield & 1.00 & 0.88 & 0.82 & 0.76 & 0.96 & -0.46 & -0.88 \\
+          Area Harvested & 0.88 & 1.00 & 0.78 & 0.83 & 0.73 & -0.19 & -0.81 \\
+          Water Used & 0.82 & 0.78 & 1.00 & 0.67 & 0.76 & -0.49 & -0.82 \\
+          \begin{tabular}[c]{@{}c@{}}Scope One\\ Emissions\end{tabular} & 0.76 & 0.83 & 0.67 & 1.00 & 0.65 & -0.16 & -0.67 \\
+          \begin{tabular}[c]{@{}c@{}}Yield by\\ Area\end{tabular} & 0.96 & 0.73 & 0.76 & 0.65 & 1.00 & -0.54 & -0.84 \\
+          Average Price & -0.46 & -0.19 & -0.49 & -0.16 & -0.54 & 1.00 & 0.72 \\
+          \begin{tabular}[c]{@{}c@{}}Average Price\\ by Area\end{tabular} & -0.88 & -0.81 & -0.82 & -0.67 & -0.84 & 0.72 & 1.00 \\ \bottomrule
+          \end{tabular}}
 \end{table}
-Pearson Correlation Coefficients of the transformed, centered and scaled variables are shown in Table \ref{tab:pearson_correlation}. 
-\par
-All relationships were found to be statistically significant (P $<$ 2.200E-16).
+
 \subsection{General Linear Models}
 
+Each model had a high $R^2$ value, indicating that a most of the variance within the data was described by the models (see Table~\ref{tab:modelsummary}).
+The models were were found to be a good fit, with overall F-tests being statistically significant (P $<$ 2.200E-16).
+And, aside from 3 variables, F-tests across each model's variables were also significant (with all being at least, P $<$ 0.05).
+The three exceptions were: scope one emissions in Model 3 (P=0.22) and Model 4 (P=0.0.39), and the interaction between area harvested and water used in model 2 (P=0.22).
+Note that, scope one emissions was included in all models to directly compare the response variables as ratios of vineyard size to raw values and because it was strongly correlated to the response variable in every model (except model 5); especially for Models 1 and 4 (Table \ref{tab:pearson_correlation}).
+A full list of regression coefficients 95\% CIs and p-values for each of the four models is provided in the appendix.
+% You gotta add these to the appendix in a more readable fashion
+\par
 \begin{table}[]
@@
            & $\mathbf{R^2}$ & \textbf{\begin{tabular}[c]{@{}c@{}} Adjusted\\ $\mathbf{R^2}$ \end{tabular}} & \textbf{F-Statistic} & \textbf{P-Value} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual\\ Standard Error\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Sum\\ of Squares\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Mean\\ of Squares\end{tabular}} \\ \midrule
-          \textbf{\begin{tabular}[c]{@{}c@{}}Model 1\\ Yield\end{tabular}} & 9.072E-01 & 9.061E-01 & 7.753E+02 & 2.200e-16 & 3.065E-01 & 4.913E+02 & 1.000E-01 \\
+          \textbf{\begin{tabular}[c]{@{}c@{}}Model 1\\ \end{tabular}} & 9.072E-01 & 9.061E-01 & 7.753E+02 & 2.200e-16 & 3.065E-01 & 4.913E+02 & 1.000E-01 \\
           \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-          \textbf{\begin{tabular}[c]{@{}c@{}}Model 2\\ Yield/Area\end{tabular}} & 7.951E-01 & 7.770E-01 & 4.403E+01 & 2.200e-16 & 4.722E-01 & 1.085E+03 & 2.200E-01 \\
+          \textbf{\begin{tabular}[c]{@{}c@{}}Model 2\\ \end{tabular}} & 7.951E-01 & 7.770E-01 & 4.403E+01 & 2.200e-16 & 4.722E-01 & 1.085E+03 & 2.200E-01 \\
           \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-          \textbf{\begin{tabular}[c]{@{}c@{}}Model 3\\ Value\end{tabular}} & 9.753E-01 & 9.748E-01 & 1.885E+03 & 2.200e-16 & 1.589E-01 & 7.111E+01 & 3.000E-02 \\
+          \textbf{\begin{tabular}[c]{@{}c@{}}Model 3\\ \end{tabular}} & 9.753E-01 & 9.748E-01 & 1.885E+03 & 2.200e-16 & 1.589E-01 & 7.111E+01 & 3.000E-02 \\
           \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-          \textbf{\begin{tabular}[c]{@{}c@{}}Model 4\\ Value / Area\end{tabular}} & 9.669E-01 & 9.638E-01 & 3.095E+02 & 2.200e-16 & 1.904E-01 & 9.528E+01 & 4.000E-02 \\ \bottomrule
+          \textbf{\begin{tabular}[c]{@{}c@{}}Model 4\\ \end{tabular}} & 9.669E-01 & 9.638E-01 & 3.095E+02 & 2.200e-16 & 1.904E-01 & 9.528E+01 & 4.000E-02 \\
+          \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
+          \textbf{\begin{tabular}[c]{@{}c@{}}Model 5\\ \end{tabular}} & 0.9089 & 0.9004 & 107.2 & 2.200e-16 & 0.3155 & 262.04 & 0.10 \\ \bottomrule
           \end{tabular}
@@
 \end{table}
-Each model showed a strong relationship between the predictors and the response (see Table~\ref{tab:modelsummary}).
 
-Aside from 3 variables, F-tests across each model indicated a good model fit.
-The three exceptions were: scope one emissions in Model 3 (P=0.22) and Model 4 (P=0.36), and the interaction between area harvested and water used in model 2 (P=0.22).
-Note that, scope one emissions was included in all models to directly compare the response variables as ratios of vineyard size to raw values and because it was strongly correlated to the response variable in every model; especially for Models 1 and 4 (Table \ref{tab:pearson_correlation}).
-
-A full list of regression coefficients 95\% CIs and p-values for each of the four models is provided in the appendix. % You gotta add these to the appendix in a more readable fashion
-\par
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% See Kerries notes regarding these.
-%
-%For Models 1 (yield) and 3 (value) year played a pivotal role, with only one year in each model not being significant (2021/2022 and 2016/2017 respectively).
-Both Model 1 and 3 showed a majority of regions were significant with 32 of 54 regions being significant in Model 1, and 42 of 50 regions being significant in Model 3 at 95\% confidence.
-%\newline
-%\newline
-%The number of combinations of year and region meant that Models 2 and 4 had many tests (424 and 243 respectively).
-Model 2 found 62.56\% of these combinations were indicative of a significant contribution to the model at 95\% significance.
-Model 4 was found to have 88.07\% of its year/region combinations indicating a significant contribution.
-A likely reason for some combinations not being significant was a lack of samples in that particular region/year being present; with region sample sizes ranging from 1 to 1006.
-%\newline
-%\newline
-%With regard to continuous variables: Model 1 and 2 showed all variables to be significant at 95\% confidence when accounting for other variables.
-T-tests for Model 3 showed all continuous variables except scope one emissions were significant.
-Model 4 showed all variables aside from scope one emissions and water use to be significant; with scope one emissions and water use only being significant when considered as an interaction with area harvested but not when considered on their own.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Models' continuous variable's coefficient values are summarised in Table \ref{tab:contvar}.
+Model 1 showed all coefficients except for the intercept were significantly contributing to the model (P < 0.05).
+Model 2's coefficients were all statistically significant.
+However, for Models 3, 4 and 5 Scope one emissions did not significantly contribute.
+And, Model 4 only saw statistically significant contributions from the intercept and water use.
+Although the coefficient for water use was statistically significant for each model, it did not have the highest value, instead area harvested, being an order of magnitude greater dominated the models.
+Model 5 was able to achieve a similar $R^2$ to Model 4 without area harvested, having stronger influences from water use and scope one emissions.
 
@@
   \textbf{Model 3} & 1.808E-02 & 9.713E-01 & -2.310E-02 & -6.992E-03 &  &  \\
-  \textbf{Model 4} & 6.702E-01 & -7.354E-01 & -6.732E-03 & -5.645E-03 & 2.726E-02 & 7.515E-02 \\ \bottomrule
+  \textbf{Model 4} & 6.702E-01 & -7.354E-01 & -6.732E-03 & -5.645E-03 & 2.726E-02 & 7.515E-02 \\
+  \textbf{Model 5} & 1.2008636941 &  & -0.040359875 & -0.017057894 &  &  \\
+   \bottomrule
 \end{tabular}}
@@
 In terms of magnitude, GI region has the highest possible absolute value for each model. The most notable
+
+\begin{figure}
+  \includegraphics{violinplot.pdf}
+  \caption{Violin plots of GI Region and Year coefficients for each model.}\label{fig:violin}
+  \end{figure}
 
@@
 %
-\begin{figure}
-  \includegraphics{violinplot.pdf}
-  \caption{Violin plots of GI Region and Year coefficients for each model.}\label{fig:violin}
-  \end{figure}
+
 %
@@
 %
-\begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{my_map.pdf}}
-  \caption{Map of regional average yield and value per hectare.}\label{fig:map}
-\end{figure}
+
 %

### 2023-08-17 article.tex
text for yearly interactions

@@
 %%%%%%%%%%%%%%%%
-The regression coefficients for the categorical variables year and GI region, and their interaction under each of the four models are depcited in \ref{fig:violin}.
-The coefficients for the continuous variables are summarised in Table \ref{tab:contvar}.
-In terms of magnitude, GI region has the highest possible absolute value for each model. The most notable
+The regression coefficients for the categorical variables year and GI region, and their interaction under each of the four models are depicted in \ref{fig:yearly}.
+The first year for a model's data is used as the baseline.
+The Adelaide Hills is used as the regional baseline with the interaction between year and region using the first year and the Adelaide Hills as the baseline.
+Region and year contributed, in some but not all cases, more than the other variables.
+However, some years are not significant, as they are not statistically different from 0, given their error.
+Models 4 and 5 appear very similar, however this is likely due to the averaging generalising the finer differences between specific year and region combinations.
+No trend is apparent for any model, with only a slight drop for 2019 in Models 1 and 2, and a slight increase after 2018 for Models 3, 4 and 5.
 
 \begin{figure}
-  \includegraphics{violinplot.pdf}
-  \caption{Violin plots of GI Region and Year coefficients for each model.}\label{fig:violin}
+  \includegraphics{yearly_plots.pdf}
+  \caption{Model Coefficient values for GI Region and Year, with standard error bars. The average of Models 2, 4 and 5 were used in place of every combination of region and year.}\label{fig:yearly}
   \end{figure}
@@
 \section{Discussion}
+%Interestingly all but Model 2 do not appear to capture the introduction of Tariffs to Australian exports, which had a pronounced effect within the sale and production of wine in Australia .
+This needs a wine aus ref and to be put in the discussion
 There was an expected strong relationship between size and resource use, with the overall space of a vineyard and its access to resources greatly determining the upper limit of potential yield.

### 2023-08-17 analysis.r
fixes to yearly interaction plot

@@ -451,7 +451,7 @@ temp_df$year <- stri_extract_first_regex(rownames(temp_df), "[0-9]+")
 temp_df <- aggregate(temp_df, list(temp_df$year), FUN=mean.t)
 rownames(temp_df) <- temp_df$Group.1
 temp_df <- data.frame(temp_df[, c("Estimate", "Std..Error")])
-temp_df$model <- rep("model4", length(rownames(temp_df)))
+temp_df$model <- rep("model5", length(rownames(temp_df)))
 rownames(temp_df) <- paste0(rownames(temp_df), "b")
 
 estimates <- rbind(estimates, temp_df)
@@ -463,6 +463,7 @@ estimates$year <- as.integer(estimates$year)
 p<- ggplot(estimates, aes(x = year, y = Estimate, group = model, color=model)) + 
 #  geom_line() +
   geom_point()+
+  geom_line()+
   geom_errorbar(aes(ymin=Estimate-Std..Error, ymax=Estimate+Std..Error), width=.2,
                  position=position_dodge(0.05)) +
     theme_classic()

### 2023-08-17 article.tex
spiderplot desc

@@
 
-%%%%%%%%%%%%%%%%
 The regression coefficients for the categorical variables year and GI region, and their interaction under each of the four models are depicted in \ref{fig:yearly}.
@@
 No trend is apparent for any model, with only a slight drop for 2019 in Models 1 and 2, and a slight increase after 2018 for Models 3, 4 and 5.
-
+\par
 \begin{figure}
@@
 
-%
+Broad regional differences are summarised in Figure \ref{fig:spider}.
+The most notable difference is between vineyards within 'Very Hot' and 'Very Dry' (inland) areas where little emphasis is put on achieving high average sale prices, with the focus being on the larger scale yield.
+Water Use changes dramatically between these regions as well, with water being a driving force in the mass production of grapes but not necessarily the quality.
+The warmer and drier regions tend to also cater to larger vineyards, with greater areas.
+
+\begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{starplots.pdf}}
+  \caption{Radar plot of climatic profile's resource use, yield and average sale price.}\label{fig:spider}
+\end{figure}
+
+\begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{yield_verse_value.pdf}}
+  \caption{Scatter plot of vineyard yield against the product of yield and average price per tonne. The axes are in standard deviations with points coloured by climate.}\label{fig:yield_vs_value}
+\end{figure}
 

### 2023-08-17 analysis.r
results draft.

@@ -358,14 +358,14 @@ temp_yield_vs_value <- ggplot(
     df[df$temp!="", ]
     , aes(
         x = ha_tonnes_grapes_harvested
-        , y = ha_value
+        , y = ha_average_per_tonne
         , col = temp)
     ) +
     xlim(-2.5, 2.5) +
     ylim(-2.5, 2.5) +
     geom_point(size = .5) +
     xlab("Yield / Area Harvested") +
-    ylab("Yield * Average Price / Area Harvested") +
+    ylab("Average Price / Area Harvested") +
     ggtitle("Temperature") +
     theme_light() +
     theme(panel.grid.minor = element_blank()
@@ -377,14 +377,14 @@ rain_yield_vs_value <- ggplot(
     df[df$rain!="", ]
     , aes(
         x = ha_tonnes_grapes_harvested
-        , y = ha_value
+        , y = ha_average_per_tonne
         , col = rain)
         ) +
         xlim(-2.5, 2.5) +
         ylim(-2.5, 2.5) +
         geom_point(size = .5) +
         xlab("Yield / Area Harvested") +
-        ylab("Yield * Average Price / Area Harvested") +
+        ylab("Average Price / Area Harvested") +
         ggtitle("Rainfall") +
         theme_light() +
         theme(panel.grid.minor = element_blank()

### 2023-08-17 article.tex
backup of pdfs

@@
 Broad regional differences are summarised in Figure \ref{fig:spider}.
-The most notable difference is between vineyards within 'Very Hot' and 'Very Dry' (inland) areas where little emphasis is put on achieving high average sale prices, with the focus being on the larger scale yield.
+The most notable difference is between vineyards within 'Hot' and 'Very Dry' (inland) areas where little emphasis is put on achieving high average sale prices, with the focus being on the larger scale yield.
 Water Use changes dramatically between these regions as well, with water being a driving force in the mass production of grapes but not necessarily the quality.
 The warmer and drier regions tend to also cater to larger vineyards, with greater areas.
-
+\par
 \begin{figure}
@@
 
-\begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{yield_verse_value.pdf}}
-  \caption{Scatter plot of vineyard yield against the product of yield and average price per tonne. The axes are in standard deviations with points coloured by climate.}\label{fig:yield_vs_value}
-\end{figure}
-
-\begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{yield_verse_value.pdf}}
-  \caption{Scatter plot of vineyard yield against the product of yield and average price per tonne. The axes are in standard deviations with points coloured by climate.}\label{fig:yield_vs_value}
-\end{figure}
-%
+Figure \ref{fig:yield_vs_value_area} further shows the emphasis that 'Hot' areas have on high yields with low average sale price compared with other regions.
+Scaling average price and yield by area shows a strong negative trend, trading quantity for higher sales prices.
+\par
 \begin{figure}
@@
 \end{figure}
+
+Table \ref{tab:kfold} shows the validation results of each of the models.
+The $R^2$ measures of fit show similar results to the original models, with a slight decrease.
+Indicating that the models are robust and consistent.
+\par
+\begin{table}[]
+  \label{tab:kfold}
+  \caption{Model validation using k-fold cross validation, for 10 folds repeated 100 times.}
+  \resizebox{\textwidth}{!}{
+  \begin{tabular}{@{}cccc@{}}
+    \toprule
+    \textbf{} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Mean\\ Squared Error\end{tabular}} & \textbf{R2} & \textbf{\begin{tabular}[c]{@{}c@{}}Mean Average \\ Error\end{tabular}} \\ \midrule
+    \textbf{Model 1} & 3.087E-01 & 9.045E-01 & 2.165E-01 \\
+    \textbf{Model 2} & 5.104E-01 & 7.409E-01 & 3.493E-01 \\
+    \textbf{Model 3} & 1.652E-01 & 9.723E-01 & 1.008E-01 \\
+    \textbf{Model 4} & 2.235E-01 & 9.500E-01 & 1.279E-01 \\ \bottomrule
+    \end{tabular}
+  }
+  \end{table}
+
+\section{Discussion}
+%Interestingly all but Model 2 do not appear to capture the introduction of Tariffs to Australian exports, which had a pronounced effect within the sale and production of wine in Australia .
+This needs a wine aus ref and to be put in the discussion
+
 %
@@
 %
-%Observing Figure \ref{fig:yield_vs_value} shows an almost discrete difference between vineyards in 'Hot' areas than other regions.
-Comparing Figure \ref{fig:yield_vs_value} to Figure \ref{fig:yield_vs_value_area} shows almost opposing trends.
-However, with area coming into play in Figure \ref{fig:yield_vs_value_area}, many data points are scaled differently; specifically the vineyards from 'Hot' regions change to be found the bottom right tail end, indicating the production of large quantity of lower value grapes.
-An inconspicuous difference between the Figures, is that a large amount of the difference can be explained by rotation (being $90^\circ$ clockwise from Figure \ref{fig:yield_vs_value} to \ref{fig:yield_vs_value_area}).
-This is more visible when comparing both graphs to the map of regional averages for response variables, see Figure \ref{fig:map}.
+%
 
@@
 
-\begin{table}[]
-  \label{tab:kfold}
-  \caption{Model validation using k-fold cross validation, for 10 folds repeated 100 times.}
-  \resizebox{\textwidth}{!}{
-  \begin{tabular}{@{}cccc@{}}
-    \toprule
-    \textbf{} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Mean\\ Squared Error\end{tabular}} & \textbf{R2} & \textbf{\begin{tabular}[c]{@{}c@{}}Mean Average \\ Error\end{tabular}} \\ \midrule
-    \textbf{Model 1} & 3.087E-01 & 9.045E-01 & 2.165E-01 \\
-    \textbf{Model 2} & 5.104E-01 & 7.409E-01 & 3.493E-01 \\
-    \textbf{Model 3} & 1.652E-01 & 9.723E-01 & 1.008E-01 \\
-    \textbf{Model 4} & 2.235E-01 & 9.500E-01 & 1.279E-01 \\ \bottomrule
-    \end{tabular}
-  }
-  \end{table}
-
-\section{Discussion}
-%Interestingly all but Model 2 do not appear to capture the introduction of Tariffs to Australian exports, which had a pronounced effect within the sale and production of wine in Australia .
-This needs a wine aus ref and to be put in the discussion
 There was an expected strong relationship between size and resource use, with the overall space of a vineyard and its access to resources greatly determining the upper limit of potential yield.

### 2023-08-18 analysis.r
added giregion and year variable for tracking effects

@@ -19,7 +19,7 @@ model1 <- lm(
     , data = df)
 model2 <- lm(
     ha_tonnes_grapes_harvested ~
-    area_harvested * ((scope1) + water_used) + data_year_id:giregion
+    area_harvested * ((scope1) + water_used) + data_year_id:giregion + data_year_id + giregion
     , data = df)
 model3 <- lm(
     value ~
@@ -27,11 +27,11 @@ model3 <- lm(
     , data = df)
 model4 <- lm(
     average_per_tonne ~
-    area_harvested * ((scope1) + water_used) + data_year_id:giregion
+    area_harvested * ((scope1) + water_used) + data_year_id:giregion + data_year_id + giregion
     , data = df)
 model5 <- lm(
     average_per_tonne ~
-    scope1 + water_used + data_year_id:giregion
+    scope1 + water_used + data_year_id:giregion + data_year_id + giregion
     , data = df)
 # We review our first model
 

### 2023-08-18 article.tex
changed coef table

@@
 \usepackage{booktabs}
+\usepackage{multirow}
 
@@
 \par
-The first subset of data (used for Model 1 and Model 2, see Table\ref{tab:tab1}) contained 5298 samples spanning the period from 2012 to 2022, covering 55 GI Regions and 1261 separate vineyards.
+The first subset of data (used for Model 1 and Model 2, see Table \ref{tab:tab1}) contained 5298 samples spanning the period from 2012 to 2022, covering 55 GI Regions and 1261 separate vineyards.
 \par
-The second subset of data (used for Model 3, Model 4 and Model 5, see Table\ref{tab:tab1}) contained 2878 samples spanning the period from 2015 to 2022, covering 51 GI Regions and 944 separate vineyards.Average price of sale per tonne was extracted from both Wine Australia (1842 values) and Sustainable Winegrowing Australia (remaining 1036 values).
+The second subset of data (used for Model 3, Model 4 and Model 5, see Table \ref{tab:tab1}) contained 2878 samples spanning the period from 2015 to 2022, covering 51 GI Regions and 944 separate vineyards.Average price of sale per tonne was extracted from both Wine Australia (1842 values) and Sustainable Winegrowing Australia (remaining 1036 values).
 \par
@@
   \resizebox{\textwidth}{!}{
-\begin{tabular}{@{}cllllll@{}}
+    \begin{tabular}{@{}cccccccc@{}}
+      \begin{tabular}{@{}cccccccc@{}}
         \toprule
-    & \textbf{Intercept} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Water\\ Used\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}scope one\\ Emissions\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\\ $\ast$\\ scope one\\ Emissions\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Area\\ Harvested\\ $\ast$\\ Water\\ Used\end{tabular}}} \\ \midrule
-  \textbf{Model 1} & -3.318E-02 & 7.418E-01 & 8.660E-02 & 6.731E-02 &  &  \\
-  \textbf{Model 2} & -6.516E-01 & 5.774E-01 & 1.079E-02 & 8.498E-02 & -4.971E-02 & -5.346E-02 \\
-  \textbf{Model 3} & 1.808E-02 & 9.713E-01 & -2.310E-02 & -6.992E-03 &  &  \\
-  \textbf{Model 4} & 6.702E-01 & -7.354E-01 & -6.732E-03 & -5.645E-03 & 2.726E-02 & 7.515E-02 \\
-  \textbf{Model 5} & 1.2008636941 &  & -0.040359875 & -0.017057894 &  &  \\
-   \bottomrule
-\end{tabular}}
+        \textbf{} & \textbf{} & \textbf{Intercept} & \textbf{Area Harvested} & \textbf{Water Used} & \textbf{\begin{tabular}[c]{@{}c@{}}Scope One\\ Emissions\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Area Harvested \\ *\\ Scope One Emissions\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Area Harvested \\ * \\ Water Used\end{tabular}} \\ \midrule
+        \multirow{2}{*}{Model 1} & Coefficient & -0.0332 & 0.7418 & 0.0866 & 0.0673 &  &  \\
+         & Std Error & 0.0196 & 0.0100 & 0.0089 & 0.0080 &  &  \\
+         &  &  &  &  &  &  &  \\
+         \multirow{2}{*}{Model 2} & Coefficient & 0.1696 & 0.5774 & 0.1079 & 0.0850 & -0.0497 & -0.0535 \\
+         & Std Error & 0.0591 & 0.0148 & 0.0131 & 0.0117 & 0.0081 & 0.0084 \\
+         &  &  &  &  &  &  &  \\
+         \multirow{2}{*}{Model 3} & Coefficient & 0.0181 & 0.9713 & -0.0231 & -0.0070 &  &  \\
+         & Std Error & 0.0130 & 0.0072 & 0.0069 & 0.0057 &  &  \\
+         &  &  &  &  &  &  &  \\
+         \multirow{2}{*}{Model 4} & Coefficient & 0.1450 & 0.0024 & -0.0466 & -0.0170 & 0.0115 & 0.0014 \\
+         & Std Error & 0.0528 & 0.0150 & 0.0143 & 0.0118 & 0.0079 & 0.0083 \\
+         &  &  &  &  &  &  &  \\
+         \multirow{2}{*}{Model 5} & Coefficient & 0.1517 &  & -0.0404 & -0.0171 &  &  \\
+         & Std Error & 0.0527 &  & 0.0113 & 0.0097 &  &  \\ \bottomrule
+        \end{tabular}
+
+      \end{tabular}      
+  }
 \end{table}
@@
 Table \ref{tab:kfold} shows the validation results of each of the models.
-The $R^2$ measures of fit show similar results to the original models, with a slight decrease.
+The $R^2$ measures of fit show similar results to the initial models, with a slight decrease.
 Indicating that the models are robust and consistent.
+%
+%
+% At least summary statistics of the baseline model.
+%
+%
 \par
@@
   \caption{Model validation using k-fold cross validation, for 10 folds repeated 100 times.}
-  \resizebox{\textwidth}{!}{
+  % \resizebox{\textwidth}{!}{
   \begin{tabular}{@{}cccc@{}}
@@
     \end{tabular}
-  }
+  % }
   \end{table}

### 2023-08-19 maps.r
Aud added as units

@@ -39,7 +39,7 @@ value_map <- ggplot() +
     xlim(113, 154) +
     ylim(-45, -25) +
     scale_fill_gradientn(colours = rev(viridis(3))
-                         , name = ""
+                         , name = "$1m AUD"
                          , na.value = "grey100") +
     theme_void()
 

### 2023-08-19 analysis.r
Added coefficients for models 2, 4, 5.

@@ -407,15 +407,19 @@ temp_df <- temp_df[grep("year", rownames(temp_df)), ]
 estimates <- data.frame(temp_df[, c("Estimate", "Std..Error")])
 estimates$model <- rep("model1", length(rownames(estimates)))
 
+# R > temp_df[grep("^data_year_id[[:digit:]][[:digit:]][[:digit:]][[:digit:]]/[[:digit:]][[:digit:]][[:digit:]][[:digit:]]$", rownames(temp_df)), ]
+ 
+
+
 # The second model has coefficients averaged as there are so many per year.
 temp_df <- data.frame(summary(model2)$coefficients)
-temp_df <- temp_df[grep("year", rownames(temp_df)), ]
+temp_df <- temp_df[grep("^data_year_id[[:digit:]][[:digit:]][[:digit:]][[:digit:]]/[[:digit:]][[:digit:]][[:digit:]][[:digit:]]$", rownames(temp_df)), ]
 temp_df <- data.frame(temp_df[, c("Estimate", "Std..Error")])
 
-temp_df$year <- stri_extract_first_regex(rownames(temp_df), "[0-9]+")
-temp_df <- aggregate(temp_df, list(temp_df$year), FUN=mean.t)
-rownames(temp_df) <- temp_df$Group.1
-temp_df <- data.frame(temp_df[, c("Estimate", "Std..Error")])
+# temp_df$year <- stri_extract_first_regex(rownames(temp_df), "[0-9]+")
+# temp_df <- aggregate(temp_df, list(temp_df$year), FUN=mean.t)
+# rownames(temp_df) <- temp_df$Group.1
+# temp_df <- data.frame(temp_df[, c("Estimate", "Std..Error")])
 temp_df$model <- rep("model2", length(rownames(temp_df)))
 
 estimates <- rbind(estimates, temp_df)
@@ -430,29 +434,29 @@ estimates <- rbind(estimates, temp_df)
 
 # The fourth model is similar to the second and requires averaging
 temp_df <- data.frame(summary(model4)$coefficients)
-temp_df <- temp_df[grep("year", rownames(temp_df)), ]
+temp_df <- temp_df[grep("^data_year_id[[:digit:]][[:digit:]][[:digit:]][[:digit:]]/[[:digit:]][[:digit:]][[:digit:]][[:digit:]]$", rownames(temp_df)), ]
[…]

### 2023-08-19 article.tex
fixed numbers with new models

@@
     \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-    \textbf{Model 2} & ${ \textrm{Yield}}\over{ \textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ scope one Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested \end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * scope one Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular} \\
+    \textbf{Model 2} & ${ \textrm{Yield}}\over{ \textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ scope one Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * scope one Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular} \\
     \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
@@
     \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-    \textbf{Model 4} & Average Sale Price & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope One Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\
-        \\ \end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * Scope One Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular}
+    \textbf{Model 4} & Average Sale Price & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope One Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * Scope One Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular}
         \\
@@
            & $\mathbf{R^2}$ & \textbf{\begin{tabular}[c]{@{}c@{}} Adjusted\\ $\mathbf{R^2}$ \end{tabular}} & \textbf{F-Statistic} & \textbf{P-Value} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual\\ Standard Error\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Sum\\ of Squares\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Mean\\ of Squares\end{tabular}} \\ \midrule
-          \textbf{\begin{tabular}[c]{@{}c@{}}Model 1\\ \end{tabular}} & 9.072E-01 & 9.061E-01 & 7.753E+02 & 2.200e-16 & 3.065E-01 & 4.913E+02 & 1.000E-01 \\
+          \textbf{\begin{tabular}[c]{@{}c@{}}Model 1\\ \end{tabular}} & 0.9072 & 0.9061 & 775.3 & 2.200e-16 & 0.3065 & 491.3 & 0.1 \\
           \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-          \textbf{\begin{tabular}[c]{@{}c@{}}Model 2\\ \end{tabular}} & 7.951E-01 & 7.770E-01 & 4.403E+01 & 2.200e-16 & 4.722E-01 & 1.085E+03 & 2.200E-01 \\
+          \textbf{\begin{tabular}[c]{@{}c@{}}Model 2\\ \end{tabular}} & 0.8291 & 0.8141 & 55.07 & 2.200e-16 & 0.4312
+           & 905.03 & 0.19 \\
           \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-          \textbf{\begin{tabular}[c]{@{}c@{}}Model 3\\ \end{tabular}} & 9.753E-01 & 9.748E-01 & 1.885E+03 & 2.200e-16 & 1.589E-01 & 7.111E+01 & 3.000E-02 \\
+          \textbf{\begin{tabular}[c]{@{}c@{}}Model 3\\ \end{tabular}} &0.9753 &0.9748 & 1885 & 2.200e-16 & 0.1589 & 71.11 & 0.03 \\
           \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-          \textbf{\begin{tabular}[c]{@{}c@{}}Model 4\\ \end{tabular}} & 9.669E-01 & 9.638E-01 & 3.095E+02 & 2.200e-16 & 1.904E-01 & 9.528E+01 & 4.000E-02 \\
+          \textbf{\begin{tabular}[c]{@{}c@{}}Model 4\\ \end{tabular}} & 0.9091 & 0.9006 & 106.1 & 2.200e-16 & 0.3153  & 261.41 & 0.10 \\
           \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\

### 2023-08-20 article.tex
merge it ship it

@@
 \par
-Differences in vineyard locations were captured through the use of Geographical Indicator Regions (GI Regions).
+Differences in vineyard locations were captured through the use of Geographical Indicator Regions (GI Regions, see Figure \ref{fig:map}).
 Each GI Region has its own unique mixture of climatic and geophysical properties that describes a unique winegrowing region within Australia; these regions were predefined by Wine Australia \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
@@
 
-Pearson Correlation Coefficients of the transformed, centered and scaled variables are shown in Table \ref{tab:pearson_correlation}.
-All correlations were found to be statistically significant (P $<$ 2.200E-16).
-Except for 'average price' all variables are positively correlated.
-It can be considered that more resources and area are likely to lead to greater yields.
-However, the negative correlation between resource use and average sale price, although not as strong, indicates the possibility that resource consumption alone is not determining factor for grape quality.
+Pearson Correlation Coefficients of the transformed, centred and scaled variables are shown in Table \ref{tab:pearson_correlation}.
+All correlations were found to be statistically significant (P $<$ 2.200E-16), and except for 'average price' all variables were positively correlated.
+With water use, area harvested and emissions being positively correlated to yield, it can be considered that more resources and area are likely to lead to greater yields.
+Average sale price's negative correlation to yield, water use, area and scope one emissions, indicated that size and fuel separately were not the determining factor for grape quality.
+The negative correlations are not causal relationships, that using more water does not cause lower quality, but relative are measures indicating that using greater amounts of water than others may lead lower quality.
 \par

### 2023-08-20 article.tex
fixed discussion and conclusion

@@
 Each model had a high $R^2$ value, indicating that a most of the variance within the data was described by the models (see Table~\ref{tab:modelsummary}).
-The models were were found to be a good fit, with overall F-tests being statistically significant (P $<$ 2.200E-16).
+The models were found to be a good fit, with overall F-tests being statistically significant (P $<$ 2.200E-16).
 And, aside from 3 variables, F-tests across each model's variables were also significant (with all being at least, P $<$ 0.05).
@@
 However, some years are not significant, as they are not statistically different from 0, given their error.
-Models 4 and 5 appear very similar, however this is likely due to the averaging generalising the finer differences between specific year and region combinations.
-No trend is apparent for any model, with only a slight drop for 2019 in Models 1 and 2, and a slight increase after 2018 for Models 3, 4 and 5.
+Models 4 and 5 are very similar, indicating that the exclusion of area does not greatly affect the contribution from yearly influence.
+Models 4 and 5 have the most prominent trends, showing an increase in yearly effects over time, with Model 3 also increasing from 2016 to 2018 but plateau afterwards.
+Models 1 and 2 do not show a clear trend but do drop during 2017 and 2018 after increasing in the first 3 years.
 \par
@@
   \includegraphics{yearly_plots.pdf}
-  \caption{Model Coefficient values for GI Region and Year, with standard error bars. The average of Models 2, 4 and 5 were used in place of every combination of region and year.}\label{fig:yearly}
+  \caption{Model Coefficient values for Year, with standard error bars.}\label{fig:yearly}
   \end{figure}
@@
 
-Broad regional differences are summarised in Figure \ref{fig:spider}.
-The most notable difference is between vineyards within 'Hot' and 'Very Dry' (inland) areas where little emphasis is put on achieving high average sale prices, with the focus being on the larger scale yield.
+Regional differences are summarised in Figure \ref{fig:spider}.
+The most notable difference is between vineyards within 'Hot' and 'Very Dry' regions (warm inland regions), where little emphasis is put on achieving high average sale prices, instead focussing on larger scale yield.
 Water Use changes dramatically between these regions as well, with water being a driving force in the mass production of grapes but not necessarily the quality.
@@
   \includegraphics{yield_verse_value_by_area.pdf}}
-    \caption{Scatter plot of vineyard yield against the product of yield and average price per tonne as ratios to area harvested.
+    \caption{Scatter plot of vineyard yield against the average sale price as ratios to area harvested.
 The axes are in standard deviations with points coloured by climate.}\label{fig:yield_vs_value_area}
@@
 \section{Discussion}
-%Interestingly all but Model 2 do not appear to capture the introduction of Tariffs to Australian exports, which had a pronounced effect within the sale and production of wine in Australia .
-This needs a wine aus ref and to be put in the discussion
-
-%
-
-%
-% KM: Are these observed or modelled values?
-%
-% Replace with star charts?
-%
-%
-
-%\newline
-%\newline
-
-%There is a notable change between regional averages when looking at yield versus value.
-Through the coefficients we can deduce that: this difference is also a difference between more resources used for the raw response variables; and a difference between overall resource use and the size of the vineyard when considering the response variables as a ratio to area.
-Noting, resource use and area harvested have a combined relationship through their interactions, and separate relationships as individual variables (see Table \ref{tab:contvar}).
-A notable occurrence in Figure \ref{fig:yield_vs_value_area}, is that the 'Very Dry' vineyards which produce lower yields and higher quality grapes are predominantly found in the Barossa Valley (a wine region known for its high quality Shiraz).
-This note is important as it shows climate is not exclusively the consideration, soil and other geographical phenomenon have considerable impacts on vineyard outcomes.
-%We compare models to see which predictors contributed relatively to each of the models.
-%
-% Although the graphs showing yield and value look similar when presented as raw values and as ratios of area. The flow of resource use is inverted.
-% Talk about how resources use relates to these graphs.
-%
-% Show how regions are almost inverted between avereage value vs yield per area when looking at the maps.
-%
-%
-%Although the initial correlation between average sales price and yield was a negative trend (see table 2), these models were able to shed light on several contributing factors to yield and average sales price.
-Correlation values showed that water and emissions increased with yield but decreased with average sale price (see Table 4).
-In alternative attempts at models it was found that without the incorporation of GI Region or year the predictions greatly under performed.
-The possible reason behind this effect was that different strategies are likely employed between different regions, where some regions target the mass production of cheaper grapes over quality.
-This is most notable when grouping regions by climate, especially when considering GI Regions in the ‘Hot Very Dry’ climate (see Figure 7).
-The effect of climate when used in models was not more significant than the more granular use of GI regions.
-The interaction between year and GI Region likely accounted for localised events such as bushfires, which would be impactful, but only at a local level in both time and space.
 
@@
 This could have been due to the easier management of smaller properties.
-\subsection{Resource use and yield versus quality} 
+\par
+Scope one emissions' lack of significance and contribution given its F-statistics, could be indicative that other vineyard activities requiring fuel are not leading factors for a vineyards grape quality.
+The relationship between yield, value and area was not simply about efficiently producing the most grapes.
+It is possible that the relationship of scope one emissions between yield and sale price was closely tied to a vineyard's area due to requiring more fuel to address more issues over greater distances.
+It is difficult to discern the connection of scope one emissions directly, as fuel can be used for a broad category of activities.
+\par
 There are many on-the-ground decisions that influence both quality and yield.
-The $R^2$ values for Models 2 and 4 showed that the average price per tonne of grapes described a great deal of the relationship between resource use and yield when variables were considered as ratios of area (due to the discrepancy in $R^2$ between the two models, see Table \ref{tab:modelsummary}).
-This discrepancy is likely due to different vineyard prioritisation, which can be described by the type of quality and quantity a vineyard aims to target.
-Decisions such as the prioritisation of quality over quantity, are governed by complex physical and social forces, for example international market demands, disease pressures and natural disasters \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}, with many of these occurrences being highlighted throughout the reports from Wine Australia \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018} over the past decade.
-It is also important to consider that these reports show that the warm inland regions have seen a decline in profit during this period, whereas regions targeting quality did not.
-Size becomes an important consideration, as it dictates the potential capacity to produce greater volumes of grapes.
-However, given the comparison of value per area, regions with larger vineyards (such as warmer in land regions) and larger vineyards in general, tend to underperform.
-The 'Hot Very Dry' vineyards (see Figure \ref{fig:yield_vs_value_area}) These vineyards would be very competitive with only a minor increase to sale price, possibly outperforming other regions.
-\par
-The negative trend between size and average sales price could be a side effect of supply versus demand, especially when looking at the level of difference in production of some vineyards.
-Economies of scale likely played a role in determining yield but were only one consideration alongside resource use.
-% (KM: WHere is this shown?)
-% This relationship is not shown well and should be reconsidered when changing the methods and results.
-Size was also less of a determining factor when considering quality.
-It is possible that the relationship of scope one emissions between yield and sale price was closely tied to a vineyard's area due to requiring more fuel to cover issues (such as fixing a broken irrigation pipe), where  a larger area has the potential for issues to be further away.
-This is further observed when noting that most irrigation systems are diesel based, with water use being a significant variable in each model and scope one emissions not; scope one emissions' lack of significance and contribution given its F-statistics, could be indicative that other vineyard activities requiring fuel are not leading factors for a vineyards grape quality.
-The relationship between yield, value and area was not simply about efficiently producing the most grapes; sales price and by association grape quality, are integral to the profitability, and this is strongly linked to resource-use and thus the longevity and sustainability of a vineyard.
-\par
+The decision to prioritise quality over quantity, is governed by complex physical and social forces, for example international market demands, disease pressures and natural disasters \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}, with many of these occurrences being highlighted throughout the reports from Wine Australia \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018} over the past decade.
+However, the changes in the coefficients are not reflective of many known occurrences, such as the 2020 bush fires, which had higher values for coefficients than prior years; During the 2020 bush fires 40,000 tonnes of grapes were lost across 18 different wine regions due to bush fires and smoke taint; the predicted incidence of wildfires is expected to increase \citep{canadellMultidecadalIncreaseForest2021}.
+\par In comparison to countrywide pressures such as drought, this damage made up only 3\% of the total amount of grapes for that year; although acknowledged as a considerable loss on an individual basis, it was deemed to be only a minor national concern by Wine Australia when compared to other environmental pressures such as drought \citep{wineaustraliaNationalVintageReport2020}
+
 There are important considerations unique to winegrowing compared to other agricultural industries.
@@
 Incorporating decisions into the model could help describe the contributing factors to regional differences beyond resource consumption and regional differences this  motivates the call for more granular data and more sophisticated modelling.
-\subsection{Regional Differences}
+\par
 Some regions appeared to produce many low quality grapes at scale whilst others focussed on producing higher quality grapes in lower volumes.
@@
 The difference in pricing per tonne between the lowest and highest graded grapes can be greater than a hundred times the difference in value per tonne.
-% ref
 Not all regions target only one grade of grape, with some producing a variety of differently graded grapes; such as the Yarra Valley, which produces grades from C to A.
-% ref
-\par
-Some regions are known for their quality and may have a bias in purchasers or bring greater demand regardless of similarities and differences in production of quality of grapes \citep{hallidayAustralianWineEncyclopedia2009}.
-%The Barossa is an example of this, known for its quality could also lend itself to a bias in purchasers not considering other regions that may be capable of similar quality.
 This effect could stifle the potential for market opportunities within lesser known regions.
 A further possibility is the existence of regional upper limits on potential quality, or that there are diminishing returns in some regions when pursuing quality or quantity; however these types of relationships may be obfuscated by knowledgeable winegrowers who avoid this pitfall.
-% What is this below paragraph?
-%The correlation between average sales price and yield was a negative trend (see table 2); the contributing factors to yield and average sales price was ???.
-Correlation values showed that water and emissions increased with yield but decreased with average sale price (see Table 4).
-%
 \par
@@
 The interaction between year and GI Region likely accounted for events such as bushfires, which would be impactful, but only at a local level, both in time and space.
-
-\subsection{Limitations}
+\par
+Climatic pressures are an important consideration for growers, especially those in warmer and drier regions.
+The Wine Australia reports also show that warm inland regions have seen a decline in profit over the past decade, whereas regions targeting quality did not.
+The warm inland regions also tend to contain larger vineyards, making up for lower sale prices with larger yields.
+Considering the negative correlation of average price to area, for this strategy to work economies of scale become and important factor.
+Given the large quantities of grapes that can be produced by some vineyards, even at low margins there is the potential to be profitable.
+However, the increasing climatic pressures mixed with the requirement for larger volumes of water, make the sustainability of some vineyards come into question.
+Furthermore, intensive farming in general is known to jeopardise the sustainability of an operation through the degradation of soil and waterways \citep{capelloEffectsTractorPasses2019,linHydropedologySynergisticIntegration2012,pisciottaGroundwaterNitrateRisk2015}.
+There are established methods that can help to mitigate these affects, such as the use of cover crops and crop rotation.
+However, it has become more apparent that the active reduction of grape yield, through methods such as thinning, can help increase the quality of grapes and improve soil health \citep{condursoEffectsClusterThinning2016,wangChangesGlobalAroma2019}.
+\par
 Limitations in the analyses presented in this paper included overestimating yield for models 1 and 2, and underestimating crop value in models 3 and 4 (see appendix).
@@
 \par
- %The possible reason behind this effect was that different strategies are likely employed between different regions, where some regions target the mass production of cheaper grapes over quality.
-% ALSO THAT CATASTROPHES HAPPEN EVEN ON SMALL SCALES THis needs to be tied up better.
-%Reviewing data to uncover reasons for the disparity between production and quality of similar vineyards  included the use of 
 Variables such as the utilisation of renewable energy, contractors, and the occurrence of disease, fire and frost were originally explored to capture the discrepancies between similar vineyards that produced different  yields and crop values.
@@
 \section{Conclusion}
-% This belongs in the discussion
-In summary, vineyard yield and crop value is well-defined by the resources used.
-However, it is important to consider a vineyard's business goal, region, external pressures and economies of scale where larger vineyards are likely to produce greater overall yields, and have higher yield per area.
-Smaller vineyards are likely to produce more value per area and a higher quality of grape.
-It is likely that regional constraints also contribute to the best strategy to pursue when considering quality or quantity.
-%KM is this what the data analyses tell you? If so, say so. If not delete.
+
+This study delved into the relationships between resource use, grape quality and yield.
+The findings underscore the multifaceted nature of vineyard management, where the interplay of size, resource allocation, climate, and regional influences collectively shape both the quantity and quality of grape yields.
+The investigation demonstrated a strong correlation between vineyard size and resource use, highlighting the pivotal role of vineyard size and  accessibility to resources, has in dictating the upper limits of potential grape yields.
+However, a noteworthy inverse relationship between vineyard size and potential grape quality was revealed.
+Quality was not solely tied to the overall expenditure of resources, but rather to the efficient allocation of resources per unit area.
+This emphasises that factors beyond sheer scale contribute significantly to the final quality of the grapes produced.
+Moreover, regional and yearly variations exhibited substantial effects on vineyard outputs, impacting both quantity and quality.
+The consistent connection observed between smaller vineyards and higher grape quality suggests that the management of smaller properties might be more streamlined and effective, enabling greater a greater quality of grape to be produced.
 

### 2023-08-20 analysis.r

@@ -463,14 +463,16 @@ estimates <- rbind(estimates, temp_df)
 estimates$year <- stri_extract_first_regex(rownames(estimates), "[0-9]+")
 estimates$year <- as.integer(estimates$year)
 
+scaleFUN <- function(x) sprintf("%.0f", x)
 # Default line plot
 p<- ggplot(estimates, aes(x = year, y = Estimate, group = model, color=model)) + 
 #  geom_line() +
-  geom_point()+
-  geom_line()+
-  geom_errorbar(aes(ymin=Estimate-Std..Error, ymax=Estimate+Std..Error), width=.2,
-                 position=position_dodge(0.05)) +
-    theme_classic()
+    geom_point()+
+    geom_line()+
+    geom_errorbar(aes(ymin=Estimate-Std..Error, ymax=Estimate+Std..Error), width=.2,
+        position=position_dodge(0.05)) +
+    theme_classic() +
+    scale_x_continuous(labels=scaleFUN)
 
 pdf("yearly_plots.pdf")
 print(p)

### 2023-08-20 article.tex
Final re4ad through with andrew

@@
           \multicolumn{1}{l}{} &  &  &  &  \\
-          {Area Harvested ($m^2$)} & 6.670E+05 & 1.337E+06 & 7.000E+02 & 2.436E+07 \\
+          {Area Harvested (ha)} & 6.670E+01 & 1.337E+02 & 7.000E-02 & 2.436E+03 \\
           \multicolumn{1}{l}{} &  &  &  &  \\
@@
           \multicolumn{1}{l}{} &  &  &  &  \\
-          $\mbox{Yield (tonnes)}\over \mbox{Area (m}^2)$ & 1.009E+01 & 8.127E+00 & 4.000E-02 & 8.634E+01 \\
+          $\mbox{Yield (tonnes)}\over \mbox{Area harvested (ha)}$ & 1.009E+01 & 8.127E+00 & 4.000E-02 & 8.634E+01 \\
           \multicolumn{1}{l}{} &  &  &  &  \\
@@
           \multicolumn{1}{l}{} &  &  &  &  \\
-          $\mbox{Average Sale Price (AUD)} \over \mbox{Area Harvested (m}^2)$ & 1.347E+02 & 5.711E+02 & 1.753E-01 & 2.979E+04 \\ \bottomrule
+          $\mbox{Average Sale Price (AUD)} \over \mbox{Area Harvested (ha)}$ & 1.347E+02 & 5.711E+02 & 1.753E-01 & 2.979E+04 \\ \bottomrule
           \end{tabular}
@@
 Average sale price's negative correlation to yield, water use, area and scope one emissions, indicated that size and fuel separately were not the determining factor for grape quality.
-The negative correlations are not causal relationships, that using more water does not cause lower quality, but relative are measures indicating that using greater amounts of water than others may lead lower quality.
+The negative correlations are not causal relationships (using more water does not cause lower quality) but relative are measures indicating that using greater amounts of water than others may lead to lower quality.
 \par
@@
 
-The regression coefficients for the categorical variables year and GI region, and their interaction under each of the four models are depicted in \ref{fig:yearly}.
+The regression coefficients for the year for each model is depicted in Figure \ref{fig:yearly}.
 The first year for a model's data is used as the baseline.
@@
   \resizebox{\textwidth}{!}{
-  \includegraphics{starplots.pdf}}
+  \includegraphics{starplots.png}}
   \caption{Radar plot of climatic profile's resource use, yield and average sale price.}\label{fig:spider}
@@
     \textbf{} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Mean\\ Squared Error\end{tabular}} & \textbf{R2} & \textbf{\begin{tabular}[c]{@{}c@{}}Mean Average \\ Error\end{tabular}} \\ \midrule
-    \textbf{Model 1} & 3.087E-01 & 9.045E-01 & 2.165E-01 \\
-    \textbf{Model 2} & 5.104E-01 & 7.409E-01 & 3.493E-01 \\
-    \textbf{Model 3} & 1.652E-01 & 9.723E-01 & 1.008E-01 \\
-    \textbf{Model 4} & 2.235E-01 & 9.500E-01 & 1.279E-01 \\ \bottomrule
+    \textbf{Model 1} & .309 & .905 & .2165 \\
+    \textbf{Model 2} & .457 & .7921 & .313 \\
+    \textbf{Model 3} & .165 & .972 & .101 \\
+    \textbf{Model 4} & .348 & .878 & .182 \\ 
+    \textbf{Model 5} & .348 & .878 & .183 \\ \bottomrule
     \end{tabular}
@@
 \par
-There are many on-the-ground decisions that influence both quality and yield.
-The decision to prioritise quality over quantity, is governed by complex physical and social forces, for example international market demands, disease pressures and natural disasters \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}, with many of these occurrences being highlighted throughout the reports from Wine Australia \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018} over the past decade.
-However, the changes in the coefficients are not reflective of many known occurrences, such as the 2020 bush fires, which had higher values for coefficients than prior years; During the 2020 bush fires 40,000 tonnes of grapes were lost across 18 different wine regions due to bush fires and smoke taint; the predicted incidence of wildfires is expected to increase \citep{canadellMultidecadalIncreaseForest2021}.
-\par In comparison to countrywide pressures such as drought, this damage made up only 3\% of the total amount of grapes for that year; although acknowledged as a considerable loss on an individual basis, it was deemed to be only a minor national concern by Wine Australia when compared to other environmental pressures such as drought \citep{wineaustraliaNationalVintageReport2020}
-
 There are important considerations unique to winegrowing compared to other agricultural industries.
@@
 This results in unique issues and considerations for each vineyard, where on-the-ground decisions are influenced by other wine industry's choices, such as the use of sustainable practices in vineyards as a requirement for sale in overseas markets; notably these interactions can be further complicated by some winegrowers being completely integrated into a wine company, while others are not \citep{knightFirmResourcesDevelopment2019}.
-Incorporating decisions into the model could help describe the contributing factors to regional differences beyond resource consumption and regional differences this  motivates the call for more granular data and more sophisticated modelling.
+Incorporating decisions into the model could help describe the contributing factors to regional differences beyond resource consumption, motivating the call for more granular data and more sophisticated modelling.
 \par
-Some regions appeared to produce many low quality grapes at scale whilst others focussed on producing higher quality grapes in lower volumes.
-This emperical finding is consistent with Wine Australia's annual reports, which shows that some GI regions, such as the Riverland, are known for producing large amounts of lower grade (low value per tonne) grapes \cite{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
-Comparatively other regions, such as Tasmania, only produce high quality grapes but in smaller quantities.
-The difference in pricing per tonne between the lowest and highest graded grapes can be greater than a hundred times the difference in value per tonne.
-Not all regions target only one grade of grape, with some producing a variety of differently graded grapes; such as the Yarra Valley, which produces grades from C to A.
-This effect could stifle the potential for market opportunities within lesser known regions.
-A further possibility is the existence of regional upper limits on potential quality, or that there are diminishing returns in some regions when pursuing quality or quantity; however these types of relationships may be obfuscated by knowledgeable winegrowers who avoid this pitfall.
-\par
-Due to regional differences, different strategies are also  employed across different regions, such as some regions targeting mass production over quality.
-This is most notable when grouping regions by climate, especially when considering GI Regions in the 'Hot Very Dry' climate (see Figure \ref{fig:yield_vs_value}).
-%In the preliminary analyses alternative models found that without the direct incorporation of GI Region or year, predictions greatly under performed.
-The effect of climate in the models was never as significant as the more granular GI regions, and always led to less accurate models.
-Although not chosen over GI region, climate was considered to be a large determinant of the ability to produce larger quantities of grapes, as well as a determinant in grape quality \citep{agostaRegionalClimateVariability2012}.
-The more granular GI Region likely explained a broader mix of geographical phenomenon, such as soil, geology and access to water resources \citep{abbalDecisionSupportSystem2016,carmonaUseParticipatoryObjectOriented2011}.
-The interaction between year and GI Region likely accounted for events such as bushfires, which would be impactful, but only at a local level, both in time and space.
+There are many on-the-ground decisions that influence both quality and yield.
+The decision to prioritise quality over quantity, is governed by complex physical and social forces, for example international market demands, disease pressures and natural disasters \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}, with many of these occurrences being highlighted throughout the reports from Wine Australia \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018} over the past decade.
+However, the changes in the coefficients (see Figure \ref{fig:yearly}) are not reflective of many known occurrences, such as the 2020 bush fires, which had higher values for coefficients than prior years; During the 2020 bush fires 40,000 tonnes of grapes were lost across 18 different wine regions due to bush fires and smoke taint.
+In comparison to countrywide pressures such as drought, this damage made up only 3\% of the total amount of grapes for that year; although acknowledged as a considerable loss on an individual basis, it was deemed to be only a minor national concern by Wine Australia when compared to other environmental pressures such as drought \citep{wineaustraliaNationalVintageReport2020}
 \par
@@
 However, it has become more apparent that the active reduction of grape yield, through methods such as thinning, can help increase the quality of grapes and improve soil health \citep{condursoEffectsClusterThinning2016,wangChangesGlobalAroma2019}.
+\par
+Some regions appeared to produce many low quality grapes at scale whilst others focussed on producing higher quality grapes in lower volumes.
+This emperical finding is consistent with Wine Australia's annual reports, which shows that some GI regions, such as the Riverland, are known for producing large amounts of lower grade (low value per tonne) grapes \citep{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
+Comparatively, other regions, such as Tasmania, only produce high quality grapes but in smaller quantities.
+The difference in pricing per tonne between the lowest and highest graded grapes can be greater than a hundred times the difference in value per tonne.
+Not all regions target only one grade of grape, with some producing a variety of differently graded grapes; such as the Yarra Valley, which produces grades from C to A.
+This effect could stifle the potential for market opportunities within lesser known regions.
+A further possibility is the existence of regional upper limits on potential quality, or that there are diminishing returns in some regions when pursuing quality or quantity; however these types of relationships may be obfuscated by knowledgeable winegrowers who avoid this pitfall.
+\par
+Due to regional differences, different strategies are also  employed across different regions, such as some regions targeting mass production over quality.
+This is most notable when grouping regions by climate, especially when considering GI Regions in the 'Hot Very Dry' climate (see Figure \ref{fig:yield_vs_value_area}).
+%In the preliminary analyses alternative models found that without the direct incorporation of GI Region or year, predictions greatly under performed.
+The effect of climate in the models was never as significant as the more granular GI regions, and always led to less accurate models.
+Although not chosen over GI region, climate was considered to be a large determinant of the ability to produce larger quantities of grapes, as well as a determinant in grape quality \citep{agostaRegionalClimateVariability2012}.
+The more granular GI Region likely explained a broader mix of geographical phenomenon, such as soil, geology and access to water resources \citep{abbalDecisionSupportSystem2016,carmonaUseParticipatoryObjectOriented2011}.
+The interaction between year and GI Region likely accounted for events such as bushfires, which would be impactful, but only at a local level, both in time and space.
 \par
@@
 This study delved into the relationships between resource use, grape quality and yield.
-The findings underscore the multifaceted nature of vineyard management, where the interplay of size, resource allocation, climate, and regional influences collectively shape both the quantity and quality of grape yields.
-The investigation demonstrated a strong correlation between vineyard size and resource use, highlighting the pivotal role of vineyard size and  accessibility to resources, has in dictating the upper limits of potential grape yields.
-However, a noteworthy inverse relationship between vineyard size and potential grape quality was revealed.
+The findings underscore the multifaceted nature of vineyard management, where the interplay of size, resource allocation, climate, and regional influences collectively shape both the quality and quantity of grape yields.
 Quality was not solely tied to the overall expenditure of resources, but rather to the efficient allocation of resources per unit area.
 This emphasises that factors beyond sheer scale contribute significantly to the final quality of the grapes produced.
-Moreover, regional and yearly variations exhibited substantial effects on vineyard outputs, impacting both quantity and quality.
-The consistent connection observed between smaller vineyards and higher grape quality suggests that the management of smaller properties might be more streamlined and effective, enabling greater a greater quality of grape to be produced.
+Moreover, regional and yearly variations exhibited substantial effects on vineyard outputs, impacting both quality and  quantity.
+The connection observed between smaller vineyards and higher grape quality suggests that the management of smaller properties might be more streamlined and effective, enabling a greater quality of grape to be produced.
 

### 2023-08-20 kfold.r

@@ -1,0 +1,78 @@
+library(caret)
+
+# data_prep.py has to be executed in order for df.csv to exist
+set.seed(100)
+df <- read.csv("df.csv")
+
+#specify the cross-validation method
+ctrl <- trainControl(method = "repeatedcv", number = 10, repeats=100)
+#fit a regression model and use k-fold CV to evaluate performance
+model <- train(tonnes_grapes_harvested ~
+    data_year_id + giregion + area_harvested + water_used + scope1
+    , data = df
+    , method = "lm"
+    , trControl = ctrl
+    , na.action = na.omit)
+
+
+sink("model1.txt")
+print(model)
+sink()
+
+
+#specify the cross-validation method
+ctrl <- trainControl(method = "repeatedcv", number = 10, repeats=100)
+#fit a regression model and use k-fold CV to evaluate performance
+model <- train(ha_tonnes_grapes_harvested ~
+    area_harvested * ((scope1) + water_used) + data_year_id:giregion + data_year_id + giregion
+    , data = df
+    , method = "lm"
+    , trControl = ctrl
+    , na.action = na.omit)
+
+sink("model2.txt")
+print(model)
+sink()
+
+#specify the cross-validation method
+ctrl <- trainControl(method = "repeatedcv", number = 10, repeats=100)
+#fit a regression model and use k-fold CV to evaluate performance
+model <- train(value ~
+    data_year_id + giregion + area_harvested + water_used + scope1
+    , data = df
+    , method = "lm"
+    , trControl = ctrl
+    , na.action = na.omit)
+
+sink("model3.txt")
+print(model)
+sink()
+
+#specify the cross-validation method
+ctrl <- trainControl(method = "repeatedcv", number = 10, repeats=100)
+#fit a regression model and use k-fold CV to evaluate performance
+model <- train(average_per_tonne ~
+    area_harvested * ((scope1) + water_used) + data_year_id:giregion + data_year_id + giregion
+    , data = df
+    , method = "lm"
+    , trControl = ctrl
+    , na.action = na.omit)
+
+
+sink("model4.txt")
+print(model)
+sink()
+
+#specify the cross-validation method
+ctrl <- trainControl(method = "repeatedcv", number = 10, repeats=100)
+#fit a regression model and use k-fold CV to evaluate performance
+model <- train(average_per_tonne ~
+    scope1 + water_used + data_year_id:giregion + data_year_id + giregion
+    , data = df
[…]

### 2023-08-20 starplot.r

@@ -98,7 +98,10 @@ rain_colors <- c(
 )
 
 # Two example starplots
-pdf("starplots.pdf")
+png("starplotst.png")
 starplot(new_columns, df, "rain", rain_colors)
+dev.off()
+# Two example starplots
+png("starplotsr.png")
 starplot(new_columns, df, "temp", temp_colors)
 dev.off()

### 2023-08-21 article.tex
added appendix

@@
 
+\usepackage{csvsimple}
+\usepackage{longtable}
 \journal{Geography and Sustainability}
@@
 Note that, scope one emissions was included in all models to directly compare the response variables as ratios of vineyard size to raw values and because it was strongly correlated to the response variable in every model (except model 5); especially for Models 1 and 4 (Table \ref{tab:pearson_correlation}).
-A full list of regression coefficients 95\% CIs and p-values for each of the four models is provided in the appendix.
-% You gotta add these to the appendix in a more readable fashion
+%A full list of regression coefficients 95\% CIs and p-values for each of the four models is provided in the appendix. % You gotta add these to the appendix in a more readable fashion
 \par
@@
 
-\if{false}
- \begin{table}[]
-  \centering
-  \caption{Summary of models, their predictors, covariates and variable interactions.}
-  \label{tab:tab2}
-  \begin{tabulary}{\textwidth}{CCCCCCCC}
-  Variable                             & Yield      & Area       & Water Used & Scope One Emissions & $\textrm{Yield}\over\textrm{Area}$ & Average Price Per Tonne & ${\textrm{Average Price per tonne}\over\textrm{Area}}$ \\
-  Yield                                & 1.000E+00  & 7.440E-01  & -4.309E-03 & 7.290E-01           & 3.500E-01            & -2.262E-01              & -1.644E-01                           \\
-  Area                                 & 7.440E-01  & 1.000E+00  & -5.331E-03 & 8.921E-01           & 7.854E-02            & -1.178E-01              & -2.042E-01                           \\
-  Water Used                           & -4.309E-03 & -5.331E-03 & 1.000E+00  & -1.929E-03          & -5.600E-03           & -3.562E-02              & -2.669E-02                           \\
-  Scope One Emissions                  & 7.290E-01  & 8.921E-01  & -1.929E-03 & 1.000E+00           & 9.357E-02            & -9.422E-02              & -1.933E-01                           \\
-  $\textrm{Yield}\over\textrm{Area}$                 & 3.500E-01  & 7.854E-02  & -5.600E-03 & 9.357E-02           & 1.000E+00            & -4.849E-01              & -1.698E-01                           \\
-  Average Price Per Tonne              & -2.262E-01 & -1.178E-01 & -3.562E-02 & -9.422E-02          & -4.849E-01           & 1.000E+00               & 4.732E-01                            \\
-  ${\textrm{Average Price per tonne}\over\textrm{Area}}$ & -1.644E-01 & -2.042E-01 & -2.669E-02 & -1.933E-01          & -1.698E-01           & 4.732E-01               & 1.000E+00
-  \end{tabulary}
-  \end{table}
-
-  \begin{table}[]
-    \caption{Pearson correlation coefficients for each logarithmically transformed variable.}
-    \label{tab:tab3}
-    \begin{tabular}{clllllll}
-    Variable                                           & \multicolumn{1}{c}{Yield} & \multicolumn{1}{c}{Area} & \multicolumn{1}{c}{Water Used} & \multicolumn{1}{c}{Scope One Emissions} & \multicolumn{1}{c}{$\textrm{Yield}\over\textrm{Area}$} & \multicolumn{1}{c}{Average Price Per Tonne} & \multicolumn{1}{c}{${\textrm{Average Price per tonne}\over\textrm{Area}}$} \\
-    Yield                                              & 1.000E+00                 & 8.822E-01                & 8.245E-01                      & 7.617E-01                               & 9.353E-01                                          & -4.591E-01                                  & -8.918E-01                                                             \\
-    Area                                               & 8.822E-01                 & 1.000E+00                & 7.750E-01                      & 8.311E-01                               & 6.742E-01                                          & -1.911E-01                                  & -8.474E-01                                                             \\
-    Water Used                                         & 8.245E-01                 & 7.750E-01                & 1.000E+00                      & 6.668E-01                               & 7.292E-01                                          & -4.881E-01                                  & -8.300E-01                                                             \\
-    Scope One Emissions                                & 7.617E-01                 & 8.311E-01                & 6.668E-01                      & 1.000E+00                               & 6.086E-01                                          & -1.559E-01                                  & -7.063E-01                                                             \\
-    $\textrm{Yield}\over\textrm{Area}$                     & 9.353E-01                 & 6.742E-01                & 7.292E-01                      & 6.086E-01                               & 1.000E+00                                          & -5.625E-01                                  & -8.076E-01                                                             \\
-    Average Price Per Tonne                            & -4.591E-01                & -1.911E-01               & -4.881E-01                     & -1.559E-01                              & -5.625E-01                                         & 1.000E+00                                   & 6.592E-01                                                              \\
-    ${\textrm{Average Price per tonne}\over\textrm{Area}}$ & -8.918E-01                & -8.474E-01               & -8.300E-01                     & -7.063E-01                              & -8.076E-01                                         & 6.592E-01                                   & 1.000E+00
-    \end{tabular}
-    \end{table}
+\section{Appendix}
+\par
 
@@
       \caption{P-values for the non-transformed water used variable's Pearson correlation coefficients.}
-    \label{tab:tab4}
       \begin{tabular}{cl}
@@
 
-    \begin{table}[]
-      \caption{Summary statistics for each variable on the original scale..}
-      \label{tab:tab5}
-      \begin{tabular}{clllllll}
-      Variable &
-        \multicolumn{1}{c}{Yield} &
-        \multicolumn{1}{c}{Area} &
-        \multicolumn{1}{c}{Water Used} &
-        \multicolumn{1}{c}{Scope One Emissions} &
-        \multicolumn{1}{c}{$\textrm{Yield}\over\textrm{Area}$} &
-        \multicolumn{1}{c}{Average Price Per Tonne} &
-        \multicolumn{1}{c}{${\textrm{Average Price per tonne}\over\textrm{Area}}$} \\
-      Yield                                              & 1.000E+00  & 8.822E-01  & 8.245E-01  & 7.617E-01  & 9.353E-01  & -4.591E-01 & -8.918E-01 \\
-      Area                                               & 8.822E-01  & 1.000E+00  & 7.750E-01  & 8.311E-01  & 6.742E-01  & -1.911E-01 & -8.474E-01 \\
-      Water Used                                         & 8.245E-01  & 7.750E-01  & 1.000E+00  & 6.668E-01  & 7.292E-01  & -4.881E-01 & -8.300E-01 \\
-      Scope One Emissions                                & 7.617E-01  & 8.311E-01  & 6.668E-01  & 1.000E+00  & 6.086E-01  & -1.559E-01 & -7.063E-01 \\
-      $\textrm{Yield}\over\textrm{Area}$                     & 9.353E-01  & 6.742E-01  & 7.292E-01  & 6.086E-01  & 1.000E+00  & -5.625E-01 & -8.076E-01 \\
-      Average Price Per Tonne                            & -4.591E-01 & -1.911E-01 & -4.881E-01 & -1.559E-01 & -5.625E-01 & 1.000E+00  & 6.592E-01  \\
-      ${\textrm{Average Price per tonne}\over\textrm{Area}}$ & -8.918E-01 & -8.474E-01 & -8.300E-01 & -7.063E-01 & -8.076E-01 & 6.592E-01  & 1.000E+00 
-      \end{tabular}
-      \end{table}
-% should these tables all go into the appendix solely or should they be summarised as well?
-%\begin{table}[]
-%  \begin{tabular}{lllll}
-%                    & Coef      & Std. Error & t-value   & Pr(\textgreater{}|t|) \\
-%  Area Harvested    & 7.418E-01 & 1.003E-02  & 7.390E+01 & 2.000E-16             \\
-%  Water Used        & 8.660E-02 & 8.885E-03  & 9.747E+00 & 2.000E-16             \\
-%  scope one Emissions & 6.731E-02 & 8.013E-03  & 8.400E+00 & 2.000E-16            
-%  \end{tabular}
-%  \end{table}
-%
-\begin{table}[]
-  \label{tab:tab6}
-  \caption{Model 1 ANOVA summarising variable significance at the .5 level.}
-  \begin{tabular}{llllll}
-  Variable       & Df & Sum Sq    & Mean Sq   & F Value   & Pr(\textgreater{}F) \\
-  Year                & 9  & 7.060E+01 & 7.800E+00 & 8.353E+01 & \textless 2.20E-16 \\
-  GI Region           & 54 & 1.507E+03 & 2.790E+01 & 2.972E+02 & \textless 2.20E-16 \\
-  Area Harvested      & 1  & 3.211E+03 & 3.211E+03 & 3.419E+04 & \textless 2.20E-16 \\
-  Water Used          & 1  & 1.040E+01 & 1.040E+01 & 1.103E+02 & \textless 2.20E-16 \\
-  Scope One Emissions & 1  & 6.600E+00 & 6.600E+00 & 7.056E+01 & \textless 2.20E-16
-  \end{tabular}
-\end{table}
-\begin{table}[]
-  \label{tab:tab7}
-  \caption{Model 2 ANOVA summarising variable significance at the .5 level.}
-  \begin{tabular}{llllll}
-  Variable                    & Df  & Sum Sq    & Mean Sq   & F Value   & Pr(\textgreater{}F)    \\
-  Area Harvested              & 1   & 2.407E+03 & 2.407E+03 & 1.080E+04 & \textless 2.20E-16 *** \\
-  Scope One Emissions         & 1   & 3.989E+01 & 3.989E+01 & 1.789E+02 & \textless 2.20E-16 *** \\
-  Water Used                  & 1   & 5.500E+02 & 5.500E+02 & 2.467E+03 & \textless 2.20E-16 *** \\
-  Area Harvested*Scope One Emissions & 1 & 6.921E+01 & 6.921E+01 & 3.104E+02 & \textless 2.20E-16 *** \\
-  Area Harvested * Water Used & 1   & 1.040E+00 & 1.040E+00 & 4.686E+00 & 3.045E-02 **           \\
-  Year * GI Region            & 424 & 1.144E+03 & 2.700E+00 & 1.210E+01 & \textless 2.20E-16 ***
-  \end{tabular}
-\end{table}
-%
-\begin{table}[]
-  \caption{Model 3 ANOVA summarising variable significance at the .5 level.}
-  \label{tab:tab8}
-  \begin{tabular}{llllll}
-  Variable            & Df & Sum Sq    & Mean Sq   & F Value   & Pr(\textgreater{}F)    \\
-  Year                & 6  & 1.324E+01 & 2.210E+00 & 8.748E+01 & \textless 2.20E-16 *** \\
-  GI Region           & 50 & 6.498E+02 & 1.300E+01 & 5.151E+02 & \textless 2.20E-16 *** \\
-  Area Harvested      & 1  & 2.142E+03 & 2.142E+03 & 8.491E+04 & \textless 2.20E-16 *** \\
-  Water Used          & 1  & 3.200E-01 & 3.200E-01 & 1.259E+01 & 3.947E-04 **           \\
-  Scope One Emissions & 1  & 4.000E-02 & 4.000E-02 & 1.492E+00 & 2.221E-01             
-  \end{tabular}
-\end{table}
+\begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{model1_qq.pdf}}
+  \caption{QQ-plot of Model 1.}
+\end{figure}
 
-\begin{table}[]
-\label{tab:tab9}
-\caption{Model 4 ANOVA summarising variable significance at the .5 level.}
-\begin{tabular}{llllll}
-Variable            & Df  & Sum Sq    & Mean Sq   & F Value   & Pr(\textgreater{}F)    \\
-Area Harvested      & 1   & 2.066E+03 & 2.066E+03 & 5.700E+04 & \textless 2.20E-16 *** \\
-Scope One Emissions & 1   & 6.000E-02 & 6.000E-02 & 1.569E+00 & 2.105E-01              \\
-Water Used          & 1   & 2.014E+02 & 2.014E+02 & 5.557E+03 & \textless 2.20E-16 *** \\
-Area Harvested*Scope One Emissions & 1 & 5.246E+01 & 5.246E+01 & 1.448E+03 & \textless 2.20E-16 *** \\
-Area Harvested * Water Used        & 1 & 7.270E+00 & 7.270E+00 & 2.005E+02 & \textless 2.20E-16 *** \\
-Year * GI Region    & 243 & 4.546E+02 & 1.870E+00 & 5.162E+01 & \textless 2.20E-16 ***
-\end{tabular}
-\end{table}
+\begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{model1_rvf.pdf}}
+  \caption{Residuals vs fitted values for Model 1.}
+\end{figure}
 
-\begin{table}[]
-\label{tab:tab 10}
-\caption{Comparison of Model Residuals}
-\begin{tabular}{llll}
-      & Df   & Sum Sq    & Mean Sq   \\
-Model 1 & 5231 & 4.913E+02 & 1.000E-01 \\
-Model 2 & 4868 & 1.085E+03 & 2.200E-01 \\
-Model 3 & 2818 & 7.111E+01 & 3.000E-02 \\
-Model 4 & 2629 & 9.528E+01 & 4.000E-02
-\end{tabular}
-\end{table}
+\begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{model2_qq.pdf}}
+  \caption{QQ-plot of Model 2.}
+\end{figure}
 
-\begin{table}[]
-\caption{Comparison of Model performance.}
-\label{tab:tab 11}
-\begin{tabular}{llllll}
-        & RSE       & R2        & Adjusted 			R2 & F-statistic & P-Value           \\
-Model 1 & 3.065E-01 & 9.072E-01 & 9.061E-01      & 7.753E+02   & \textless 2.2e-16 \\
-Model 2 & 4.722E-01 & 7.951E-01 & 7.770E-01      & 4.403E+01   & \textless 2.2e-16 \\
-Model 3 & 1.589E-01 & 9.753E-01 & 9.748E-01      & 1.885E+03   & \textless 2.2e-16 \\
-Model 4 & 1.904E-01 & 9.669E-01 & 9.638E-01      & 3.095E+02   & \textless 2.2e-16
-\end{tabular}
-\end{table}
+\begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{model2_rvf.pdf}}
+  \caption{Residuals vs fitted values for Model 2.}
+\end{figure}
 
-\fi
+\begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{model3_qq.pdf}}
+  \caption{QQ-plot of Model 3.}
+\end{figure}
+
+\begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{model3_rvf.pdf}}
+  \caption{Residuals vs fitted values for Model 3.}
+\end{figure}
+
+\begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{model4_qq.pdf}}
+  \caption{QQ-plot of Model 4.}
+\end{figure}
+
+\begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{model4_rvf.pdf}}
+  \caption{Residuals vs fitted values for Model 4.}
+\end{figure}
+\begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{model5_qq.pdf}}
+  \caption{QQ-plot of Model 5.}
+\end{figure}
+
+\begin{figure}
+  \resizebox{\textwidth}{!}{
+  \includegraphics{model5_rvf.pdf}}
+  \caption{Residuals vs fitted values for Model 5.}
+\end{figure}
+
+% \subsection{Complete list of coefficients for Model 1.}
+% \csvautolongtable[]{model1_coef.csv}
+
+%   \subsection{Complete list of coefficients for Model 2.}
+% \csvautolongtable[]{model2_coef.csv}
+
+%   \subsection{Complete list of coefficients for Model 3.}
+% \csvautolongtable[]{model3_coef.csv}
+
+%   \subsection{Complete list of coefficients for Model 4.}
+% \csvautolongtable[]{model4_coef.csv}
+
+%   \subsection{Complete list of coefficients for Model 5.}
+% \csvautolongtable[]{model5_coefs.csv}
+
 %

### 2023-08-22 article.tex
Kate Edits

@@
 In this analysis we observe relationships between resource use, yield and quality through the use of statistical models.
-The dataset used for this analysis includes data collected for the past 10 years from a multitude of vineyards located over a diverse range of Australian winegrowing regions.
+The dataset used for this analysis includes data collected for the past 10 years from 1261 vineyards located over a diverse range of Australian winegrowing regions.
 Yield and quality (measured as a ratio of sale price to area) was modelled to resource factors related to water usage and emissions.
-The analysis confirmed an expected strong relationship between size and resource use, with the overall space of a vineyard and its access to resources greatly determining the upper limit of potential yield.
-However, size was also negatively related to the potential quality, with higher quality being connected to high resource inputs per area; rather than to the overall expenditure of resources.
+The analysis confirmed an expected strong relationship between size and resource use, with the overall space of a vineyard and its access to resources greatly determining the upper limit of yield.
+However, size was also negatively related to the average sale price of grapes, we find that higher quality being connected to high resource inputs per area; rather than to the overall expenditure of resources.
 Regional and yearly effects on Vineyard outputs were also identified.
@@
 Quality in viticulture (the cultivation of grapes for wine production) is driven through its integration within the wine industry, with the potential quality of a wine being initially defined through the chemical makeup of the grapes used in its production.
-The consideration of sustainability within viticulture is further complicated by environmental and socio-demographic pressures.
-In the Australian context, these include biosecurity, climate and international market demands.
+The consideration of sustainability within viticulture is further complicated by environmental and socio-demographic pressures \citep{santiago-brownSustainabilityAssessmentWineGrape2015}.
+In the Australian context, these include biosecurity, climate and international market demands \citep{canadellMultidecadalIncreaseForest2021,longbottomRoleVineyardPractices2015,oliverReviewSoilPhysical2013}.
 \par
@@
 However, due to the lack of long-term and in-depth data, individual factors are often studied in isolation \citep{abbalDecisionSupportSystem2016}.
-The lack of consolidated datasets also restricts the ability to gain statistical insights at large scales and across multiple regions \citep{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
-The dataset used for this analysis includes data collected for the past 10 years from a multitude of vineyards located over a diverse range of Australian winegrowing regions.
-We aim to use this dataset to describe the relationship of resources related to water and fuel use with the output yield and quality of the resultant product, taking into account the size and location of the vineyard.
+The lack of consolidated datasets restricts the ability to gain statistical insights at large scales and across multiple regions, as a result broader studies are lacking \citep{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
+The dataset used for this analysis includes data spanning 10 years from a multitude of vineyards located over a diverse range of Australian winegrowing regions.
+We use this dataset to describe the relationship of resources related to water and fuel use with the output yield and quality of the resultant product, taking into account the size and location of the vineyard.
 The practical addition of this aim is a baseline for comparison: given a vineyard within Australia, one could estimate the comparative efficiency with regard to the tradeoff between invested resources, yield and quality.
@@
 \par
-Predictor variables used in this analysis included yield, defined as the total tonnes of grapes harvested, and quality, defined as average sale price of grapes.
+Predicted variables in this analysis were yield, defined as the total tonnes of grapes harvested, and quality, defined as average sale price of grapes.
 It is acknowledged that quality can be defined in a variety of ways, for example by the grapes': aroma, chemical composition and color \citep{kasimatiPredictingGrapeSugar2022,mejeanperrotDecisionsupportSystemPredict2022,suarezAssessingWineGrape2021}.
@@
 Required variables included: region, harvest year, yield and area harvested.
+Harvest year and region were recorded as categorical variables.
 Optional variables included average sale price, water used and fuel used (diesel, petrol, biodiesel and LPG).
+The dataset was limited to respondents that recorded values for area harvested, water used, yield and fuel use.
 To enable direct comparisons between fuels, fuel use was converted to tonnes of Carbon Dioxide equivalent and collectively referenced to as emissions.
@@
 The first subset contained all vineyards and was used for two models (Model 1 and Model 2, see Table \ref{tab:tab1}).
-The second subset contained vineyards which either recorded a value for average price of sale per tonne through Sustainable Winegrowing Australia, or were within a region with an average price of sale recorded by Wine Australia; this subset was used for three further models (Models 3, 4 and 5, see Table \ref{tab:tab1}.
+The second subset contained vineyards which either recorded a value for average price of sale per tonne through Sustainable Winegrowing Australia, or were within a region with an average price of sale recorded by Wine Australia; this subset was used for three further models (Models 3, 4 and 5, see Table \ref{tab:tab1}).
 These subsets meant that the data would be limited to samples which had recorded values for the response variables (see Table \ref{tab:tab1}), where every sample had a recorded value for yield but not average price of sale per tonne.
@@
 Preprocessing included the conversion from fuel to scope one emissions and prior calculations for all continuous variables which included logarithmic transformations, centring and scaling by standard deviation.
-The transformation of fuel use into scope one emissions was done using the equation given from the Australian National Greenhouse Accounts Factors, shown as
+We converted multiple emission sources into scope one emissions using the equation given from the Australian National Greenhouse Accounts Factors \citep{agdeeNationalGreenhouseAccounts2021}, shown as
 \par
@@
 \par
-Differences in vineyard locations were captured through the use of Geographical Indicator Regions (GI Regions, see Figure \ref{fig:map}).
-Each GI Region has its own unique mixture of climatic and geophysical properties that describes a unique winegrowing region within Australia; these regions were predefined by Wine Australia \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
-Both Wine Australia and Sustainable Winegrowing Australia used the same GI Region format to describe location.
+The site of a vineyard predetermines several physical parameters such as climate, geology and soil, making location a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
+Differences in vineyard locations were captured through the use of Geographical Indicator Regions (GI Regions, see Figure \ref{fig:map}) defined by Wine Australia \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
+Each GI Region has its own unique mixture of climatic and geophysical properties that describes a unique winegrowing region within Australia and is a protected trademark under the Wine Australia act \citet{WineAustraliaAct2019}.
+Both Wine Australia and Sustainable Winegrowing Australia used the same GI Region categorical variable format to describe location.
 \par
-The site of a vineyard predetermines several physical parameters such as climate, geology and soil, making location a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
  The climatic properties of each GI Region were summarised by using predefined classifications as per the \citet{sustainablewinegrowingaustraliaSustainableWinegrowingAustralia2021} user manual.
@@
 Five general linear models were created (see Table \ref{tab:tab1}).
+Both the Pearson Correlation Coefficients and General Linear Models were created using the R statistical programming language \citep{rcoreteamLanguageEnvironmentStatistical2021} with the Caret package \citep{kuhnBuildingPredictiveModels2008}.
 General Linear Models were chosen as they offer the ability to produce statistical models that are explicit in the relationships between predictors and response variables.
@@
 T-tests were used to determine if predictors significantly contributed to their models when accounting for other variables, showing which specific years and areas contributed significantly.
-Both the Pearson Correlation Coefficients and General Linear Models were created using the R statistical programming language \citep{rcoreteamLanguageEnvironmentStatistical2021} with the Caret package \citep{kuhnBuildingPredictiveModels2008}.
 \par
@@
 Models were validated using K-fold cross validation calculated.
-K-fold cross validation works by removing a subset of data from the sample used to train models and then predicts those variables to determine how sensitive the model is to changes in the sample data.
+K-fold cross validation works by removing a subset of data from the sample used to train models and then predicts those variables to determine how sensitive the model is to changes in the sample data .
 For this analysis each model was validated using 10 folds, repeated 100 times.
@@
 \par
-Limitations in the analyses presented in this paper included overestimating yield for models 1 and 2, and underestimating crop value in models 3 and 4 (see appendix).
-% KM: Need to see this.
-The issue of model 1 and 2 over-predicting yield may have been due to preventative measures brought on by regional pressures such as fire, frost and disease.
+We identified two main limiations to our linear modeling.
+First model 1 and 2 over-predicting yield may have been due to preventative measures brought on by regional pressures such as fire, frost and disease.
 More resources were required to prevent these issues from spreading within a region, thus disproportionately affecting some vineyards compared to others locally.
@@
 \par
+The second limitations was the lack of further explanatory variables to help link models to causal affects.
 Variables such as the utilisation of renewable energy, contractors, and the occurrence of disease, fire and frost were originally explored to capture the discrepancies between similar vineyards that produced different yields and crop values.
@@
 Attempts to fully explain small variations was always overshadowed by the dramatic differences in regional trends.
-\par
 Having more data for each region would also be beneficial, allowing greater comparison between regions.
 More variables may also help to discern vineyards that can produce larger volumes of grapes at higher prices.
+\par
 The use of other models such as random forests and decision trees alongside more variables and data may help to uncover the reasons for under or overestimation.

### 2023-08-22 article.tex
fixed wine australia act ref problem

@@
 Sustainable Winegrowing Australia is Australia's national wine industry sustainability program, which aims to facilitate grape-growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
-Wine Australia is an Australian Government statutory authority governed by the Wine Australia Act 2013 \citep{WineAustraliaAct2019}.
+Wine Australia is an Australian Government statutory authority governed by the Wine Australia Act \citep{attorney-generalsdepartmentWineAustraliaCorporation2010}.
 \par
@@
 Differences in vineyard locations were captured through the use of Geographical Indicator Regions (GI Regions, see Figure \ref{fig:map}) defined by Wine Australia \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
-Each GI Region has its own unique mixture of climatic and geophysical properties that describes a unique winegrowing region within Australia and is a protected trademark under the Wine Australia act \citet{WineAustraliaAct2019}.
+Each GI Region has its own unique mixture of climatic and geophysical properties that describes a unique winegrowing region within Australia and is a protected trademark under the Wine Australia act \citep{attorney-generalsdepartmentWineAustraliaCorporation2010}.
 Both Wine Australia and Sustainable Winegrowing Australia used the same GI Region categorical variable format to describe location.

### 2023-11-16 article.tex
all edits done but the introduction

@@
 %% \fntext[label3]{}
-\title{The influence of resource use on yield versus quality trade-off in Australian vineyards}
+\title{The influence of resource use on yield versus sale price trade-off in Australian vineyards}
 %% use optional labels to link authors explicitly to addresses:
@@
 % The word limit needs to be checked.
-When strategies for a sustainable winegrowing industry are assessed, there is a trade-off between balancing the amount of resources invested and the resultant yield and quality of the produce.
-In this analysis we observe relationships between resource use, yield and quality through the use of statistical models.
+When strategies for a sustainable winegrowing industry are assessed, there is a trade-off between balancing the amount of resources invested and the resultant yield and sale price of the produce.
+In this analysis we observe relationships between resource use, yield and sale price through the use of statistical models.
 The dataset used for this analysis includes data collected for the past 10 years from 1261 vineyards located over a diverse range of Australian winegrowing regions.
-Yield and quality (measured as a ratio of sale price to area) was modelled to resource factors related to water usage and emissions.
-The analysis confirmed an expected strong relationship between size and resource use, with the overall space of a vineyard and its access to resources greatly determining the upper limit of yield.
-However, size was also negatively related to the average sale price of grapes, we find that higher quality being connected to high resource inputs per area; rather than to the overall expenditure of resources.
-Regional and yearly effects on Vineyard outputs were also identified.
-Overall, the analysis highlighted the importance of considering a vineyard's business goal, region, external pressures and economies of scale, with regional constraints also contributing to deciding the best resource use strategies to pursue when considering quality or quantity.
+Yield and sale price was modelled to resource factors related to water use and Green House Gas (GHG) emissions.
+The analysis confirmed a strong relationship between area and resource use, with the overall area of a vineyard and its access to resources greatly determining the upper limit of yield.
+However, area was also negatively related to the average sale price of grapes, we find that higher average sale prices were connected to high resource inputs per area; rather than to the overall expenditure of resources.
+Regional and temporal effects on vineyard yield and average sales price were also identified.
+Overall, the analysis highlighted the importance of considering a vineyard's business goal, region, external pressures and economies of scale, when considering whether to pursue higher yields or average sales prices.
 \end{abstract}
@@
 \begin{highlights}
-  \item Comparative analysis of resource use, quality and quantity in Australian winegrowing.
+  \item Comparative analysis of resource use, average sale price and quantity in Australian winegrowing.
   \item Regional comparison of outcomes and resource use in Australian winegrowing regions.
-  \item Baseline models for comparing wine crops.
+  \item Baseline models for comparing vineyards.
   \item Analysis of national, decade long data source.
@@
 The global focus on sustainability in agronomic industries has changed the way in which these enterprises do business. 
-When strategies for a sustainable winegrowing industry are assessed, there is a trade-off between balancing the amount of resources invested and the resultant yield versus quality produced.
-This dilemma exists across agriculture through shared fundamental considerations such as water use and fuel usage \citep{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016, zhuEffectsNitrogenLevel2017}.
-Quality in viticulture (the cultivation of grapes for wine production) is driven through its integration within the wine industry, with the potential quality of a wine being initially defined through the chemical makeup of the grapes used in its production.
+% When strategies for a sustainable winegrowing industry are assessed, there is a trade-off between balancing the amount of resources invested and the resultant yield versus quality produced. 
+% TODO: fix paper justifiction.
+% NOTE: This introduction does just need a bit of work.
+
+% Using sale price was based on the reliance of market value of winegrapes on grape quality and because Wine Australia explicitly defines grape quality through the use of discrete price brackets in their annual reports.
+The generalisation made to reflect quality through using average price assumed a due diligence of those who purchased the grapes \citep{yeggeInfluenceSensoryNonsensory2001}.
+
+A dilemma exists across agriculture through shared fundamental considerations such as water use and fuel usage \citep{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016, zhuEffectsNitrogenLevel2017}.
+The average price of grapes for wine production is driven through its integration within the wine industry.
+An important connection between grapes and their price is the grapes perceived quality, which initially defines a wines potential through the grapes chemical makeup \citep{blackTerpenoidsTheirRole2015,schreierFlavorCompositionWines1979}.
+Grapes of higher of percieved higher quality or from particularly famour regions will garner different prices \citep{wineaustraliaNationalVintageReport2021}.
 The consideration of sustainability within viticulture is further complicated by environmental and socio-demographic pressures \citep{santiago-brownSustainabilityAssessmentWineGrape2015}.
@@
 \par
-There is an extensive amount of research into the effects of a variety of factors on grape quality and yield \citep{heFruitYieldPrediction2022,laurentLocalInfluenceClimate2022,liakosMachineLearningAgriculture2018}.
+There is an extensive amount of research into the varied effects of factors on grape quality and yield \citep{heFruitYieldPrediction2022,laurentLocalInfluenceClimate2022,liakosMachineLearningAgriculture2018}.
 However, due to the lack of long-term and in-depth data, individual factors are often studied in isolation \citep{abbalDecisionSupportSystem2016}.
@@
 The dataset used for this analysis includes data spanning 10 years from a multitude of vineyards located over a diverse range of Australian winegrowing regions.
-We use this dataset to describe the relationship of resources related to water and fuel use with the output yield and quality of the resultant product, taking into account the size and location of the vineyard.
-The practical addition of this aim is a baseline for comparison: given a vineyard within Australia, one could estimate the comparative efficiency with regard to the tradeoff between invested resources, yield and quality.
+We use this dataset to describe the relationship of resources related to water and fuel use with the output yield and average sale price of the resultant product, taking into account the size and location of the vineyard.
+The practical addition of this aim is a baseline for comparison: given a vineyard within Australia, one could estimate the comparative efficiency with regard to the tradeoff between invested resources, yield and average sale price.
 This is the first time that such a trade off has been confirmed explicitly across such varying regions, scales and climates in the Australian winegrowing industry.
@@
 
-Data used in this analysis were obtained from Sustainable Winegrowing Australia and Wine Australia.
+Data used in these analyses were obtained from Sustainable Winegrowing Australia and Wine Australia.
 Sustainable Winegrowing Australia is Australia's national wine industry sustainability program, which aims to facilitate grape-growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
+Data collected by Sustainable Winegrowing Australia is entered voluntarily by winegrowers at the end of harvest.
 Wine Australia is an Australian Government statutory authority governed by the Wine Australia Act \citep{attorney-generalsdepartmentWineAustraliaCorporation2010}.
+Data collected by Wine Australia is publicly available.
 \par
-Predicted variables in this analysis were yield, defined as the total tonnes of grapes harvested, and quality, defined as average sale price of grapes.
-It is acknowledged that quality can be defined in a variety of ways, for example by the grapes': aroma, chemical composition and color \citep{kasimatiPredictingGrapeSugar2022,mejeanperrotDecisionsupportSystemPredict2022,suarezAssessingWineGrape2021}.
-Using sale price was based on the reliance of market value of winegrapes on grape quality and because Wine Australia explicitly defines grape quality through the use of discrete price brackets in their annual reports.
-The generalisation made to reflect quality through using average price assumed a due diligence of those who purchased the grapes \citep{yeggeInfluenceSensoryNonsensory2001}.
+Predicted variables in this analysis were yield, defined as the total tonnes of grapes harvested, and average sale price of grapes.
 Both response variables were examined as totals and as scales of area harvested.
@@
 \par
-The first subset of data (used for Model 1 and Model 2, see Table \ref{tab:tab1}) contained 5298 samples spanning the period from 2012 to 2022, covering 55 GI Regions and 1261 separate vineyards.
+The first subset of data (used for Model 1 and Model 2, see Table \ref{tab:tab1}) contained 5298 samples spanning the period from 2012 to 2022, covering 55 GI Regions and 1261 discrete vineyards.
 \par
@@
 Additional variables were considered for analysis but were excluded due to being either underreported or had insignificant contributions to model accuracies.
-Variables explored but not used due to low reporting values included fertiliser, and scope two emissions.
+Variables explored but not used due to low reporting values included fertiliser, electricity, and scope two emissions.
 Variables considered but ultimately removed due to a lack of significant contributions to models, included the use of renewable energy, contractor use, and pressures such as frost, fire and disease.
@@
   \includegraphics{my_map.pdf}}
-  \caption{Map of vineyard average income for each of the used GI Regions.}\label{fig:map}
+  \caption{Map of each GI Regions average income for a vineyard of that region (average grape sale price per tonne $\times$ total tonnes yielded).}\label{fig:map}
 \end{figure}%%
@@
 \par
-The site of a vineyard predetermines several physical parameters such as climate, geology and soil, making location a widely considered key determinant of grape yield and quality \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
+The site of a vineyard predetermines several physical parameters such as climate, geology and soil, making location a widely considered key determinant of grape yield and average sale price \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
 Differences in vineyard locations were captured through the use of Geographical Indicator Regions (GI Regions, see Figure \ref{fig:map}) defined by Wine Australia \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
@@
           \multicolumn{1}{l}{} &  &  &  &  \\
-          {Average Sale Price (AUD)} & 1.477E+03 & 9.216E+02 & 1.600E+02 & 2.600E+04 \\
+          {Average Sale Price (AUD/tonne)} & 1.477E+03 & 9.216E+02 & 1.600E+02 & 2.600E+04 \\
           \multicolumn{1}{l}{} &  &  &  &  \\
-          $\mbox{Average Sale Price (AUD)} \over \mbox{Area Harvested (ha)}$ & 1.347E+02 & 5.711E+02 & 1.753E-01 & 2.979E+04 \\ \bottomrule
+          $\mbox{Average Sale Price (AUD/tonne)} \over \mbox{Area Harvested (ha)}$ & 1.347E+02 & 5.711E+02 & 1.753E-01 & 2.979E+04 \\ \bottomrule
           \end{tabular}
@@
 With water use, area harvested and emissions being positively correlated to yield, it can be considered that more resources and area are likely to lead to greater yields.
-Average sale price's negative correlation to yield, water use, area and scope one emissions, indicated that size and fuel separately were not the determining factor for grape quality.
-The negative correlations are not causal relationships (using more water does not cause lower quality) but relative are measures indicating that using greater amounts of water than others may lead to lower quality.
+Average sale price's negative correlation to yield, water use, area and scope one emissions, indicated that size and fuel separately were not the determining factor for average sale price.
+The negative correlations are not causal relationships but relative (using more water does not cause lower sale prices).
 \par
@@
 
-Each model had a high $R^2$ value, indicating that a most of the variance within the data was described by the models (see Table~\ref{tab:modelsummary}).
+Each model had a high $R^2$ value, indicating that most of the variance within the data was described by the models (see Table~\ref{tab:modelsummary}).
 The models were found to be a good fit, with overall F-tests being statistically significant (P $<$ 2.200E-16).
@@
 Model 2's coefficients were all statistically significant.
-However, for Models 3, 4 and 5 Scope one emissions did not significantly contribute.
+However, for Models 3, 4 and 5 scope one emissions did not significantly contribute.
 And, Model 4 only saw statistically significant contributions from the intercept and water use.
 Although the coefficient for water use was statistically significant for each model, it did not have the highest value, instead area harvested, being an order of magnitude greater dominated the models.
-Model 5 was able to achieve a similar $R^2$ to Model 4 without area harvested, having stronger influences from water use and scope one emissions.
+Model 5 achieved a similar $R^2$ to Model 4 without area harvested, having stronger influences from water use and scope one emissions.
 
@@
 Regional differences are summarised in Figure \ref{fig:spider}.
-The most notable difference is between vineyards within 'Hot' and 'Very Dry' regions (warm inland regions), where little emphasis is put on achieving high average sale prices, instead focussing on larger scale yield.
-Water Use changes dramatically between these regions as well, with water being a driving force in the mass production of grapes but not necessarily the quality.
-The warmer and drier regions tend to also cater to larger vineyards, with greater areas.
+The most notable difference is between vineyards within 'Hot' and 'Very Dry' regions (warm inland regions), where high average sale prices are historically low, and yield is high.
+Water Use changes dramatically between these regions as well, with water being a driving force for yield but not necessarily average sale price.
+The warmer and drier regions tend to also have larger vineyards.
 \par
@@
   \includegraphics{starplots.png}}
-  \caption{Radar plot of climatic profile's resource use, yield and average sale price.}\label{fig:spider}
+  \caption{Radar plot of climatic profile's resource use, yield and average sale price.
+The left reflects vineyards in different climatic temperatures.
+The right reflects vineyards in different rainfall climates.}\label{fig:spider}
 \end{figure}
@@
 
-There was an expected strong relationship between size and resource use, with the overall space of a vineyard and its access to resources greatly determining the upper limit of potential yield.
-However, size was also inversely related to the potential quality, with higher quality being related to high resource inputs per area; rather than to the overall expenditure of resources.
-Vineyard outputs were also augmented by regional and yearly affects.
-Even given regional and yearly changes, there was a strong connection between smaller vineyards and higher quality.
-This could have been due to the easier management of smaller properties.
+There was an expected strong relationship between size and resource use, with the overall area of a vineyard and its access to resources greatly determining the upper limit of potential yield.
+However, size was also inversely related to the potential average sale price, with higher sales prices being related to high resource inputs per area; rather than to the overall expenditure of resources.
+Vineyard yields and sales price changed greatly by region and year.
+Even given regional and yearly changes, there was a strong connection between smaller vineyards and higher sales prices.
+This could have been due to more attention available when managing smaller properties.
 \par
-Scope one emissions' lack of significance and contribution given its F-statistics, could be indicative that other vineyard activities requiring fuel are not leading factors for a vineyards grape quality.
+The lack of significance of scope one emissions and its contribution to models, given its F-statistics, could be indicative that other vineyard activities requiring fuel are not leading factors for a vineyards avereage sale price.
 The relationship between yield, value and area was not simply about efficiently producing the most grapes.
@@
 \par
-There are many on-the-ground decisions that influence both quality and yield.
-The decision to prioritise quality over quantity, is governed by complex physical and social forces, for example international market demands, disease pressures and natural disasters \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}, with many of these occurrences being highlighted throughout the reports from Wine Australia \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018} over the past decade.
+There are many on-the-ground decisions that influence both sales price and yield.
+The decision to prioritise average sale price over quantity, is governed by complex physical and social forces, for example international market demands, disease pressures and natural disasters \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}, with many of these occurrences being highlighted throughout the reports from Wine Australia \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018} over the past decade.
 However, the changes in the coefficients (see Figure \ref{fig:yearly}) are not reflective of many known occurrences, such as the 2020 bush fires, which had higher values for coefficients than prior years; During the 2020 bush fires 40,000 tonnes of grapes were lost across 18 different wine regions due to bush fires and smoke taint.
@@
 Climatic pressures are an important consideration for growers, especially those in warmer and drier regions.
-The Wine Australia reports also show that warm inland regions have seen a decline in profit over the past decade, whereas regions targeting quality did not.
-The warm inland regions also tend to contain larger vineyards, making up for lower sale prices with larger yields.
-Considering the negative correlation of average price to area, for this strategy to work economies of scale become and important factor.
+The Wine Australia reports also show that warm inland regions have seen a decline in profit over the past decade, whereas regions with lower avereage sales prices did not \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
+Vineyards in warm inland regions also tend to contain larger vineyards, making up for lower sale prices with larger yields.
+Considering the negative correlation of average price to area, for this strategy to work, economies of scale become and important factor.
 Given the large quantities of grapes that can be produced by some vineyards, even at low margins there is the potential to be profitable.
@@
 Furthermore, intensive farming in general is known to jeopardise the sustainability of an operation through the degradation of soil and waterways \citep{capelloEffectsTractorPasses2019,linHydropedologySynergisticIntegration2012,pisciottaGroundwaterNitrateRisk2015}.
-There are established methods that can help to mitigate these affects, such as the use of cover crops and crop rotation.
-However, it has become more apparent that the active reduction of grape yield, through methods such as thinning, can help increase the quality of grapes and improve soil health \citep{condursoEffectsClusterThinning2016,wangChangesGlobalAroma2019}.
+There are established methods that can help to mitigate these effect, such as the use of cover crops and midrow crop rotation.
+However, it has become more apparent that the active reduction of grape yield, through methods such as thinning, can help improve soil health \citep{condursoEffectsClusterThinning2016,wangChangesGlobalAroma2019}.
 \par
-Some regions appeared to produce many low quality grapes at scale whilst others focussed on producing higher quality grapes in lower volumes.
-This emperical finding is consistent with Wine Australia's annual reports, which shows that some GI regions, such as the Riverland, are known for producing large amounts of lower grade (low value per tonne) grapes \citep{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
-Comparatively, other regions, such as Tasmania, only produce high quality grapes but in smaller quantities.
+Some regions appeared to produce grapes of lower average sale price at scale whilst others focussed on producing higher priced grapes in lower volumes.
+This empirical finding is consistent with Wine Australia's annual reports, which shows that some GI regions, such as the Riverland, are known for producing large amounts of lower grade (low value per tonne) grapes \citep{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
+Comparatively, other regions, such as Tasmania, only produce grapes of higher sales price but in smaller quantities.
 The difference in pricing per tonne between the lowest and highest graded grapes can be greater than a hundred times the difference in value per tonne.
@@
 This effect could stifle the potential for market opportunities within lesser known regions.
-A further possibility is the existence of regional upper limits on potential quality, or that there are diminishing returns in some regions when pursuing quality or quantity; however these types of relationships may be obfuscated by knowledgeable winegrowers who avoid this pitfall.
+A further possibility is the existence of regional upper limits on potential sales price, or that there are diminishing returns in some regions when pursuing higher sales prices or quantity; however these types of relationships may be obfuscated by knowledgeable winegrowers who avoid this pitfall.
 \par
-Due to regional differences, different strategies are also  employed across different regions, such as some regions targeting mass production over quality.
+Due to regional differences, different strategies are also  employed across different regions, such as some regions targeting mass production over higher sales price.
 This is most notable when grouping regions by climate, especially when considering GI Regions in the 'Hot Very Dry' climate (see Figure \ref{fig:yield_vs_value_area}).
@@
 The effect of climate in the models was never as significant as the more granular GI regions, and always led to less accurate models.
-Although not chosen over GI region, climate was considered to be a large determinant of the ability to produce larger quantities of grapes, as well as a determinant in grape quality \citep{agostaRegionalClimateVariability2012}.
+Although not chosen over GI region, climate was considered to be a large determinant of the ability to produce larger quantities of grapes, as well as a determinant in grape sale price \citep{agostaRegionalClimateVariability2012}.
 The more granular GI Region likely explained a broader mix of geographical phenomenon, such as soil, geology and access to water resources \citep{abbalDecisionSupportSystem2016,carmonaUseParticipatoryObjectOriented2011}.
@@
 First model 1 and 2 over-predicting yield may have been due to preventative measures brought on by regional pressures such as fire, frost and disease.
-More resources were required to prevent these issues from spreading within a region, thus disproportionately affecting some vineyards compared to others locally.
-This type of maintenance is not well captured in the models, especially when considering that some regions, especially those in warmer areas, are not as prone to disease as cooler climates and could potentially have lower operating costs per hectare.
+More fuel and water was likely used to prevent these issues from spreading within a region, thus disproportionately affecting some vineyards compared to others locally.
+This type of maintenance is not well captured in the models, especially when considering that some regions, especially those in warmer areas, are not as prone to disease as cooler climates and could potentially have lower fuel and water use per hectare.
 This could create a discrepancy in vineyards that utilised preventative measures in wetter regions, as opposed to those that did not, thus expending less fuel and energy but risking disease.
@@
 Having more data for each region would also be beneficial, allowing greater comparison between regions.
-More variables may also help to discern vineyards that can produce larger volumes of grapes at higher prices.
+More variables, such as soil quality and health, may also help to discern vineyards that can produce larger volumes of grapes at higher prices.
 \par
@@
 
-This study delved into the relationships between resource use, grape quality and yield.
-The findings underscore the multifaceted nature of vineyard management, where the interplay of size, resource allocation, climate, and regional influences collectively shape both the quality and quantity of grape yields.
-Quality was not solely tied to the overall expenditure of resources, but rather to the efficient allocation of resources per unit area.
-This emphasises that factors beyond sheer scale contribute significantly to the final quality of the grapes produced.
-Moreover, regional and yearly variations exhibited substantial effects on vineyard outputs, impacting both quality and  quantity.
-The connection observed between smaller vineyards and higher grape quality suggests that the management of smaller properties might be more streamlined and effective, enabling a greater quality of grape to be produced.
+This study delved into the relationships between resource use, grape sales price and yield.
+The findings underscore the multifaceted nature of vineyard management, where the interplay of size, resource allocation, climate, and regional influences collectively shape both the expected sale price and the quantity of grape yields.
+Average sales price of grapes was not solely tied to the overall expenditure of resources, but rather to the efficient allocation of resources per unit area.
+This emphasises that factors beyond sheer scale contribute significantly to the final sale price of grapes produced.
+Moreover, regional and yearly variations exhibited substantial effects on vineyard outputs, impacting both sales price and quantity.
+The connection observed between smaller vineyards and higher grape sale price suggests that the management of smaller properties might be more streamlined and effective.
 

### 2023-11-16 article.tex
Donezo!!

@@
 The global focus on sustainability in agronomic industries has changed the way in which these enterprises do business.
-% When strategies for a sustainable winegrowing industry are assessed, there is a trade-off between balancing the amount of resources invested and the resultant yield versus quality produced. 
-% TODO: fix paper justifiction.
-% NOTE: This introduction does just need a bit of work.
-
-% Using sale price was based on the reliance of market value of winegrapes on grape quality and because Wine Australia explicitly defines grape quality through the use of discrete price brackets in their annual reports.
-The generalisation made to reflect quality through using average price assumed a due diligence of those who purchased the grapes \citep{yeggeInfluenceSensoryNonsensory2001}.
-
-A dilemma exists across agriculture through shared fundamental considerations such as water use and fuel usage \citep{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016, zhuEffectsNitrogenLevel2017}.
+A dilemma exists across agriculture through shared fundamental considerations of resource use such as water and fuel, and the resultant yield and crop value that is produced\citep{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016, zhuEffectsNitrogenLevel2017}.
 The average price of grapes for wine production is driven through its integration within the wine industry.
 An important connection between grapes and their price is the grapes perceived quality, which initially defines a wines potential through the grapes chemical makeup \citep{blackTerpenoidsTheirRole2015,schreierFlavorCompositionWines1979}.
-Grapes of higher of percieved higher quality or from particularly famour regions will garner different prices \citep{wineaustraliaNationalVintageReport2021}.
-The consideration of sustainability within viticulture is further complicated by environmental and socio-demographic pressures \citep{santiago-brownSustainabilityAssessmentWineGrape2015}.
-In the Australian context, these include biosecurity, climate and international market demands \citep{canadellMultidecadalIncreaseForest2021,longbottomRoleVineyardPractices2015,oliverReviewSoilPhysical2013}.
+Grapes of higher of perceived quality or grapes from particularly famous regions are likely to have higher prices \citep{wineaustraliaNationalVintageReport2021}.
+Grape quality is connected to the market value of winegrapes, with Wine Australia explicitly defining grape quality through the use of discrete price brackets in their annual reports \citep{winemakersfederationofaustraliaNationalVintageReport2018}.
+Although it is also important to note that the generalisation made to reflect quality through using average price assumes a due diligence of those who purchasing the grapes \citep{yeggeInfluenceSensoryNonsensory2001}.
+The economic sustainability of a vineyard is tied to this market culture driven by the wine industry.
+The consideration of sustainability within viticulture is also subject to environmental and socio-demographic pressures \citep{santiago-brownSustainabilityAssessmentWineGrape2015}.
+In the Australian context, these pressures include biosecurity and climate and international market changes \citep{canadellMultidecadalIncreaseForest2021,longbottomRoleVineyardPractices2015,oliverReviewSoilPhysical2013}.
 \par
 There is an extensive amount of research into the varied effects of factors on grape quality and yield \citep{heFruitYieldPrediction2022,laurentLocalInfluenceClimate2022,liakosMachineLearningAgriculture2018}.
-However, due to the lack of long-term and in-depth data, individual factors are often studied in isolation \citep{abbalDecisionSupportSystem2016}.
+With there being a lack of research on grape sale price and its driving factors due to the lack of long-term and in-depth data.
+Furthermore, individual factors are often studied in isolation with yield and sales price not appearing together \citep{abbalDecisionSupportSystem2016}.
 The lack of consolidated datasets restricts the ability to gain statistical insights at large scales and across multiple regions, as a result broader studies are lacking \citep{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.

### 2023-11-20 article.tex
ship it

@@
 The dataset used for this analysis includes data collected for the past 10 years from 1261 vineyards located over a diverse range of Australian winegrowing regions.
-Yield and sale price was modelled to resource factors related to water use and Green House Gas (GHG) emissions.
+Yield and sale price was modelled to the resource use factors water use and Green House Gas (GHG) emissions.
 The analysis confirmed a strong relationship between area and resource use, with the overall area of a vineyard and its access to resources greatly determining the upper limit of yield.
@@
 Regional and temporal effects on vineyard yield and average sales price were also identified.
-Overall, the analysis highlighted the importance of considering a vineyard's business goal, region, external pressures and economies of scale, when considering whether to pursue higher yields or average sales prices.
+Overall, the analysis highlighted the importance of considering a vineyard's business goal, region, external pressures and economies of scale, when considering whether to pursue higher yields or higher average sales prices.
 \end{abstract}
@@
 The economic sustainability of a vineyard is tied to this market culture driven by the wine industry.
-The consideration of sustainability within viticulture is also subject to environmental and socio-demographic pressures \citep{santiago-brownSustainabilityAssessmentWineGrape2015}.
+The consideration of sustainability within viticulture is also subject to environmental and sociodemographic pressures \citep{santiago-brownSustainabilityAssessmentWineGrape2015}.
 In the Australian context, these pressures include biosecurity and climate and international market changes \citep{canadellMultidecadalIncreaseForest2021,longbottomRoleVineyardPractices2015,oliverReviewSoilPhysical2013}.
@@
 We use this dataset to describe the relationship of resources related to water and fuel use with the output yield and average sale price of the resultant product, taking into account the size and location of the vineyard.
-The practical addition of this aim is a baseline for comparison: given a vineyard within Australia, one could estimate the comparative efficiency with regard to the tradeoff between invested resources, yield and average sale price.
-This is the first time that such a trade off has been confirmed explicitly across such varying regions, scales and climates in the Australian winegrowing industry.
+The practical addition of this aim is a baseline for comparison: given a vineyard within Australia, one could estimate the comparative efficiency with regard to the trade-off between invested resources, yield and average sale price.
+This is the first time that such a trade-off has been confirmed explicitly across such varying regions, scales and climates in the Australian winegrowing industry.
 \section{Methods}
@@
 Sustainable Winegrowing Australia is Australia's national wine industry sustainability program, which aims to facilitate grape-growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
-Data collected by Sustainable Winegrowing Australia is entered voluntarily by winegrowers at the end of harvest.
 Wine Australia is an Australian Government statutory authority governed by the Wine Australia Act \citep{attorney-generalsdepartmentWineAustraliaCorporation2010}.
 Data collected by Wine Australia is publicly available.
-\par
-Predicted variables in this analysis were yield, defined as the total tonnes of grapes harvested, and average sale price of grapes.
+The predicted variables in this analysis where yield, defined as the total tonnes of grapes harvested, and average sale price of grapes in Australian dollars per tonne.
 Both response variables were examined as totals and as scales of area harvested.
@@
 \par
-Data obtained from Wine Australia were collected via phone surveys and included: total tonnes purchased, average price per tonne and yearly change in price for region and grape varietal.
-Data recorded by Sustainable Winegrowing Australia was entered manually by winegrowers using a web based interface with some fields being optional.
-Required variables included: region, harvest year, yield and area harvested.
-Harvest year and region were recorded as categorical variables.
-Optional variables included average sale price, water used and fuel used (diesel, petrol, biodiesel and LPG).
-The dataset was limited to respondents that recorded values for area harvested, water used, yield and fuel use.
+
+Data recorded by Sustainable Winegrowing Australia was entered manually by winegrowers voluntarily using a web based interface.
+For each model, vineyards were only included if they recorded all the variables used for the corresponding model (see Table \ref{tab:modelsummary}).
+Each vineyard had at least recorded region, harvest year, yield and area harvested.
+Other variables that were used but not present for every single vineyard were average sale price, water used and fuel used (diesel, petrol, biodiesel and LPG).
 To enable direct comparisons between fuels, fuel use was converted to tonnes of Carbon Dioxide equivalent and collectively referenced to as emissions.
+All variables were continuous except for harvest year and region, which were categorical variables.
 \par
-Average sale price was an optional field in the Sustainable Winegrowing Australia's dataset.
-Missing values were improved using regional average prices from Wine Australia.
-Two subsets of data were then created for the analysis.
+As data from Sustainable Wine Australia was voluntarily given, missing values were improved using regional average prices from the Wine Australia data.
+Data obtained from Wine Australia were collected via phone surveys and included: total tonnes purchased, average price per tonne and yearly change in price for region and grape varietal; with the data being publicly available.
+\par
+As the models used similar variables the original data was only split into two distinct subsets that could be used across the different models.
 The first subset contained all vineyards and was used for two models (Model 1 and Model 2, see Table \ref{tab:tab1}).
 The second subset contained vineyards which either recorded a value for average price of sale per tonne through Sustainable Winegrowing Australia, or were within a region with an average price of sale recorded by Wine Australia; this subset was used for three further models (Models 3, 4 and 5, see Table \ref{tab:tab1}).
-These subsets meant that the data would be limited to samples which had recorded values for the response variables (see Table \ref{tab:tab1}), where every sample had a recorded value for yield but not average price of sale per tonne.
+These subsets meant that the data would be limited to samples which had recorded values for the response variables (see Table \ref{tab:tab1}), as every sample/vineyard had a recorded value for yield but not average price of sale per tonne.
 \par
@@
 \par
-The second subset of data (used for Model 3, Model 4 and Model 5, see Table \ref{tab:tab1}) contained 2878 samples spanning the period from 2015 to 2022, covering 51 GI Regions and 944 separate vineyards.Average price of sale per tonne was extracted from both Wine Australia (1842 values) and Sustainable Winegrowing Australia (remaining 1036 values).
+The second subset of data (used for Model 3, Model 4 and Model 5, see Table \ref{tab:tab1}) contained 2878 samples spanning the period from 2015 to 2022, covering 51 GI Regions and 944 separate vineyards.
+Average price of sale per tonne was extracted from both Wine Australia (1842 values) and Sustainable Winegrowing Australia (remaining 1036 values).
 \par
@@
 Preprocessing included the conversion from fuel to scope one emissions and prior calculations for all continuous variables which included logarithmic transformations, centring and scaling by standard deviation.
-We converted multiple emission sources into scope one emissions using the equation given from the Australian National Greenhouse Accounts Factors \citep{agdeeNationalGreenhouseAccounts2021}, shown as
+We converted multiple emission sources into scope one emissions using the equation given from the Australian National Greenhouse Accounts Factors \citep{agdeeNationalGreenhouseAccounts2021}.
+The calculation conducted using
 \par
@@
 \par
-was used to convert the quantity of fuel in litres, $Q$, using a prescribed Energy Content, $EC$, and emission factors of scope one, $EF1$, and scope three, $EF3$, to tonnes of Carbon Dioxide Emission equivalent, $tCO2e$ \citep{departmentofclimatechangeenergytheenvironmentandwaterAustralianNationalGreenhouse2022}.
+where emissions was the the product of the quantity of fuel in litres, $Q$, a prescribed Energy Content, $EC$, and emission factors (as given by the Australian National Greenhouse Accounts factors) of scope one, $EF1$, and scope three, $EF3$, to tonnes of Carbon Dioxide Emission equivalent, $tCO2e$ \citep{departmentofclimatechangeenergytheenvironmentandwaterAustralianNationalGreenhouse2022}.
 \par
-
 \begin{figure}
@@
 \par
-The site of a vineyard predetermines several physical parameters such as climate, geology and soil, making location a widely considered key determinant of grape yield and average sale price \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
-Differences in vineyard locations were captured through the use of Geographical Indicator Regions (GI Regions, see Figure \ref{fig:map}) defined by Wine Australia \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
+Differences in vineyard locations were captured through the use of Geographical Indicator Regions defined by Wine Australia \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
+Although vineyards generally differed in income, there were pronounced differences between regions as shown in Figure \ref{fig:map}.
+These differences were likely due to the site of a vineyard being important as it predetermines several physical parameters such as climate, geology and soil, making location a widely considered key determinant of grape yield and average sale price \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
 Each GI Region has its own unique mixture of climatic and geophysical properties that describes a unique winegrowing region within Australia and is a protected trademark under the Wine Australia act \citep{attorney-generalsdepartmentWineAustraliaCorporation2010}.
@@
 Models were validated using K-fold cross validation calculated.
-K-fold cross validation works by removing a subset of data from the sample used to train models and then predicts those variables to determine how sensitive the model is to changes in the sample data .
+K-fold cross validation works by removing a subset of data from the sample used to train models and then predicts those variables to determine how sensitive the model is to changes in the sample data.
 For this analysis each model was validated using 10 folds, repeated 100 times.
@@
 \begin{table}[]
-  \caption{Summary of each Models coefficients for continuous variables}\label{tab:contvar}
+  \caption{Summary of each Model's coefficients for continuous variables}\label{tab:contvar}
   \resizebox{\textwidth}{!}{
@@
         \toprule
-        \textbf{} & \textbf{} & \textbf{Intercept} & \textbf{Area Harvested} & \textbf{Water Used} & \textbf{\begin{tabular}[c]{@{}c@{}}Scope One\\ Emissions\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Area Harvested \\ *\\ Scope One Emissions\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Area Harvested \\ * \\ Water Used\end{tabular}} \\ \midrule
+        \textbf{} & \textbf{} & \textbf{Intercept} & \textbf{Area Harvested} & \textbf{Water Used} & \textbf{\begin{tabular}[c]{@{}c@{}}Scope One\\ Emissions\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Area Harvested \\ ** \\ Scope One Emissions\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Area Harvested \\ ** \\ Water Used\end{tabular}} \\ \midrule
         \multirow{2}{*}{Model 1} & Coefficient & -0.0332 & 0.7418 & 0.0866 & 0.0673 &  &  \\
@@
 The regression coefficients for the year for each model is depicted in Figure \ref{fig:yearly}.
-The first year for a model's data is used as the baseline.
+The first year for a model's data is used as the baseline; 2012 for Model 1 and 2, and 2015 for Model 3, 4 and 5.
 The Adelaide Hills is used as the regional baseline with the interaction between year and region using the first year and the Adelaide Hills as the baseline.
@@
 Models 4 and 5 have the most prominent trends, showing an increase in yearly effects over time, with Model 3 also increasing from 2016 to 2018 but plateau afterwards.
-Models 1 and 2 do not show a clear trend but do drop during 2017 and 2018 after increasing in the first 3 years.
+Model 1 and 2 do not show a clear trend but do drop during 2017 and 2018 after increasing in the first 3 years.
 \par
@@
 
-Regional differences are summarised in Figure \ref{fig:spider}.
+Regional differences by temperature and rainfall are summarised in Figure \ref{fig:spider}.
 The most notable difference is between vineyards within 'Hot' and 'Very Dry' regions (warm inland regions), where high average sale prices are historically low, and yield is high.
@@
 The warmer and drier regions tend to also have larger vineyards.
+These regional differences are further shown in Figure \ref{fig:yield_vs_value_area}, as a ratio of vineyard area.
+Where, again there is a pronounced difference in 'Hot' and 'Very Dry' regions producing more per area with lower average sale prices per area than other regions.
 \par
@@
 There was an expected strong relationship between size and resource use, with the overall area of a vineyard and its access to resources greatly determining the upper limit of potential yield.
-However, size was also inversely related to the potential average sale price, with higher sales prices being related to high resource inputs per area; rather than to the overall expenditure of resources.
+However, size was also inversely related to the potential average sale price.
+Higher average sales prices were also related to high resource inputs per area instead of the overall expenditure of resources.
 Vineyard yields and sales price changed greatly by region and year.
@@
 \par
-The lack of significance of scope one emissions and its contribution to models, given its F-statistics, could be indicative that other vineyard activities requiring fuel are not leading factors for a vineyards avereage sale price.
+The lack of significance of scope one emissions and its contribution to models, given its F-statistics, could be indicative that other vineyard activities requiring fuel are not leading factors for a vineyards average sale price.
 The relationship between yield, value and area was not simply about efficiently producing the most grapes.
@@
 There are many on-the-ground decisions that influence both sales price and yield.
-The decision to prioritise average sale price over quantity, is governed by complex physical and social forces, for example international market demands, disease pressures and natural disasters \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}, with many of these occurrences being highlighted throughout the reports from Wine Australia \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018} over the past decade.
+The decision to prioritise average sale price over quantity, is governed by complex physical and social forces, for example international market demands, disease pressures and natural disasters \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}, with many of these occurrences being highlighted throughout the reports from Wine Australia over the past decade \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
 However, the changes in the coefficients (see Figure \ref{fig:yearly}) are not reflective of many known occurrences, such as the 2020 bush fires, which had higher values for coefficients than prior years; During the 2020 bush fires 40,000 tonnes of grapes were lost across 18 different wine regions due to bush fires and smoke taint.
@@
 Climatic pressures are an important consideration for growers, especially those in warmer and drier regions.
-The Wine Australia reports also show that warm inland regions have seen a decline in profit over the past decade, whereas regions with lower avereage sales prices did not \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
+The Wine Australia reports also show that warm inland regions have seen a decline in profit over the past decade, whereas regions with lower average sales prices did not \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
 Vineyards in warm inland regions also tend to contain larger vineyards, making up for lower sale prices with larger yields.
@@
 Comparatively, other regions, such as Tasmania, only produce grapes of higher sales price but in smaller quantities.
-The difference in pricing per tonne between the lowest and highest graded grapes can be greater than a hundred times the difference in value per tonne.
-Not all regions target only one grade of grape, with some producing a variety of differently graded grapes; such as the Yarra Valley, which produces grades from C to A.
-This effect could stifle the potential for market opportunities within lesser known regions.
-A further possibility is the existence of regional upper limits on potential sales price, or that there are diminishing returns in some regions when pursuing higher sales prices or quantity; however these types of relationships may be obfuscated by knowledgeable winegrowers who avoid this pitfall.
+The difference in pricing per tonne between the lowest and highest regional average sales prices was almost a hundred times, showing that region had a profound influence.
+Some regions also had a mixture of high and low average sales price and yield such as the Yarra Valley, showing that some regions were not solely locked into chosing a focus on high yields or high average sales price.
+A further possibility is the existence of regional upper limits on potential sales price, or that there are diminishing returns in some regions when pursuing higher sales prices or quantity; however these types of relationships may be obfuscated by knowledgeable winegrowers who avoid such pitfalls.
 \par
-Due to regional differences, different strategies are also  employed across different regions, such as some regions targeting mass production over higher sales price.
+Due to regional differences, different strategies are  employed, such as some regions targeting mass production over higher sales price.
 This is most notable when grouping regions by climate, especially when considering GI Regions in the 'Hot Very Dry' climate (see Figure \ref{fig:yield_vs_value_area}).
+Figure \ref{fig:spider} also shows that comparatively 'Warm' and 'Dry' Regions manage their resources incredibly efficiently having generally larger areas using similar resources too those in 'Cool' regions but having areas comparable to regions in 'Hot' climates.
+The coefficients for Model 4 also show a greater benefit of resource use per area when producing grapes at higher average sale prices, showing higher resource use per area reflective of higher average sales price (see Table \ref{tab:summary}).
 %In the preliminary analyses alternative models found that without the direct incorporation of GI Region or year, predictions greatly under performed.
@@
 \par
-We identified two main limiations to our linear modeling.
+We identified two main limitations to our linear modelling.
 First model 1 and 2 over-predicting yield may have been due to preventative measures brought on by regional pressures such as fire, frost and disease.
@@
 The findings underscore the multifaceted nature of vineyard management, where the interplay of size, resource allocation, climate, and regional influences collectively shape both the expected sale price and the quantity of grape yields.
-Average sales price of grapes was not solely tied to the overall expenditure of resources, but rather to the efficient allocation of resources per unit area.
+Average sales price of grapes was not solely tied to the overall expenditure of resources, but rather to the efficient allocation of resources per area.
 This emphasises that factors beyond sheer scale contribute significantly to the final sale price of grapes produced.

### 2023-12-13 article.tex
saving km edits in progress

@@
 The dataset used for this analysis includes data collected for the past 10 years from 1261 vineyards located over a diverse range of Australian winegrowing regions.
-Yield and sale price was modelled to the resource use factors water use and Green House Gas (GHG) emissions.
+Yield and sale price were evaluated with respect to the resource use factors water use and Green House Gas (GHG) emissions.
 The analysis confirmed a strong relationship between area and resource use, with the overall area of a vineyard and its access to resources greatly determining the upper limit of yield.
-However, area was also negatively related to the average sale price of grapes, we find that higher average sale prices were connected to high resource inputs per area; rather than to the overall expenditure of resources.
+However, area was also negatively related to the average sale price of grapes; we find that higher average sale prices were connected to high resource inputs per area, rather than to the overall expenditure of resources.
 Regional and temporal effects on vineyard yield and average sales price were also identified.
@@
 The global focus on sustainability in agronomic industries has changed the way in which these enterprises do business.
-A dilemma exists across agriculture through shared fundamental considerations of resource use such as water and fuel, and the resultant yield and crop value that is produced\citep{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016, zhuEffectsNitrogenLevel2017}.
+A dilemma exists across agriculture through shared fundamental considerations of resource use such as water and fuel, and the resultant yield and crop value that are produced \citep{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016, zhuEffectsNitrogenLevel2017}.
 The average price of grapes for wine production is driven through its integration within the wine industry.
-An important connection between grapes and their price is the grapes perceived quality, which initially defines a wines potential through the grapes chemical makeup \citep{blackTerpenoidsTheirRole2015,schreierFlavorCompositionWines1979}.
-Grapes of higher of perceived quality or grapes from particularly famous regions are likely to have higher prices \citep{wineaustraliaNationalVintageReport2021}.
+An important connection between grapes and their price is the grape's perceived quality, which initially defines a wine's potential through the grape's chemical makeup \citep{blackTerpenoidsTheirRole2015,schreierFlavorCompositionWines1979}.
+Grapes of higher perceived quality or grapes from well known regions are likely to have higher prices \citep{wineaustraliaNationalVintageReport2021}.
 Grape quality is connected to the market value of winegrapes, with Wine Australia explicitly defining grape quality through the use of discrete price brackets in their annual reports \citep{winemakersfederationofaustraliaNationalVintageReport2018}.
-Although it is also important to note that the generalisation made to reflect quality through using average price assumes a due diligence of those who purchasing the grapes \citep{yeggeInfluenceSensoryNonsensory2001}.
-The economic sustainability of a vineyard is tied to this market culture driven by the wine industry.
+It is important to note that the generalisation made to reflect quality through using average price assumes a due diligence of those who purchase the grapes \citep{yeggeInfluenceSensoryNonsensory2001}.
+The economic sustainability of a vineyard is tied to this market culture, driven by the wine industry.
 The consideration of sustainability within viticulture is also subject to environmental and sociodemographic pressures \citep{santiago-brownSustainabilityAssessmentWineGrape2015}.
@@
 There is an extensive amount of research into the varied effects of factors on grape quality and yield \citep{heFruitYieldPrediction2022,laurentLocalInfluenceClimate2022,liakosMachineLearningAgriculture2018}.
-With there being a lack of research on grape sale price and its driving factors due to the lack of long-term and in-depth data.
+Due to the lack of long-term and in-depth data there is a lack of research on grape sale price and its driving factors.
 Furthermore, individual factors are often studied in isolation with yield and sales price not appearing together \citep{abbalDecisionSupportSystem2016}.
-The lack of consolidated datasets restricts the ability to gain statistical insights at large scales and across multiple regions, as a result broader studies are lacking \citep{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
+The lack of consolidated datasets restricts the ability to gain statistical insights at large scales and across multiple regions.
+As a result, broader studies are lacking \citep{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
 The dataset used for this analysis includes data spanning 10 years from a multitude of vineyards located over a diverse range of Australian winegrowing regions.
 We use this dataset to describe the relationship of resources related to water and fuel use with the output yield and average sale price of the resultant product, taking into account the size and location of the vineyard.
-The practical addition of this aim is a baseline for comparison: given a vineyard within Australia, one could estimate the comparative efficiency with regard to the trade-off between invested resources, yield and average sale price.
+In a practical sense, this aim is a baseline for comparison: given a vineyard within Australia, one could estimate the comparative efficiency with regard to the trade-off between invested resources, yield and average sale price.
 This is the first time that such a trade-off has been confirmed explicitly across such varying regions, scales and climates in the Australian winegrowing industry.
@@
 Data used in these analyses were obtained from Sustainable Winegrowing Australia and Wine Australia.
-Sustainable Winegrowing Australia is Australia's national wine industry sustainability program, which aims to facilitate grape-growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
+Sustainable Winegrowing Australia is Australia's national wine industry sustainability program, which aims to facilitate grape  growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
 Wine Australia is an Australian Government statutory authority governed by the Wine Australia Act \citep{attorney-generalsdepartmentWineAustraliaCorporation2010}.
-Data collected by Wine Australia is publicly available.
-The predicted variables in this analysis where yield, defined as the total tonnes of grapes harvested, and average sale price of grapes in Australian dollars per tonne.
-Both response variables were examined as totals and as scales of area harvested.
+Data collected by Wine Australia are publicly available.
+The response (predicted) variables in this analysis were yield, defined as the total tonnes of grapes harvested, and average sale price of grapes in Australian dollars per tonne.
+Both response variables were examined as totals and also relative to area harvested (see Table \ref{tab:tab1}).
 Values were compared in this manner to observe how economies of scale affect the use of resources.
 \par
-
-Data recorded by Sustainable Winegrowing Australia was entered manually by winegrowers voluntarily using a web based interface.
-For each model, vineyards were only included if they recorded all the variables used for the corresponding model (see Table \ref{tab:modelsummary}).
+Data recorded by Sustainable Winegrowing Australia were entered manually by winegrowers voluntarily using a web based interface.
+For each model, vineyards were only included if they recorded all the variables used for the corresponding model (see Table \ref{tab:tab1}).
 Each vineyard had at least recorded region, harvest year, yield and area harvested.
 Other variables that were used but not present for every single vineyard were average sale price, water used and fuel used (diesel, petrol, biodiesel and LPG).
-To enable direct comparisons between fuels, fuel use was converted to tonnes of Carbon Dioxide equivalent and collectively referenced to as emissions.
-All variables were continuous except for harvest year and region, which were categorical variables.
+To enable direct comparisons between fuels, fuel use was converted to tonnes of carbon dioxide equivalent and collectively referenced to as emissions.
+All variables were continuous except for harvest year and region, which were categorical variables (Table \ref{tab:tab1}).
 \par
-As data from Sustainable Wine Australia was voluntarily given, missing values were improved using regional average prices from the Wine Australia data.
+As data from Sustainable Wine Australia were voluntarily given, missing values were improved using regional average prices from the Wine Australia data.
 Data obtained from Wine Australia were collected via phone surveys and included: total tonnes purchased, average price per tonne and yearly change in price for region and grape varietal; with the data being publicly available.
 \par
-As the models used similar variables the original data was only split into two distinct subsets that could be used across the different models.
+The dataset was split into two distinct subsets that could be used across the different models.
 The first subset contained all vineyards and was used for two models (Model 1 and Model 2, see Table \ref{tab:tab1}).
@@
 The calculation conducted using
-\par
 \begin{equation}
@@
 \end{equation}
-\par
 where emissions was the the product of the quantity of fuel in litres, $Q$, a prescribed Energy Content, $EC$, and emission factors (as given by the Australian National Greenhouse Accounts factors) of scope one, $EF1$, and scope three, $EF3$, to tonnes of Carbon Dioxide Emission equivalent, $tCO2e$ \citep{departmentofclimatechangeenergytheenvironmentandwaterAustralianNationalGreenhouse2022}.
@@
   \includegraphics{my_map.pdf}}
-  \caption{Map of each GI Regions average income for a vineyard of that region (average grape sale price per tonne $\times$ total tonnes yielded).}\label{fig:map}
+  \caption{Map of each GI Regions' average income for a vineyard of that region (average grape sale price per tonne $\times$ total tonnes yielded).}\label{fig:map}
 \end{figure}%%
@@
 \subsection{Analysis}
-Pairwise Pearson Correlation Coefficients were calculated to assess the potential existence of linear relationships between the input and predicted variables.
-To determine if a coefficient was indicative of a strong relationship, confidence intervals were used.
-P-values reflected the significance of a given correlation coefficient with statistical significance being declared when the associated value was lower than 0.05.
-Pairwise Pearson Correlation Coefficients were calculated for data on the original scale and for data as a logarithmic transform.
+Pairwise Pearson Correlation Coefficients (PPCC) were calculated to assess the potential existence of linear relationships between the input and predicted variables.
+To determine if a coefficient was indicative of a strong relationship, both the magnitude of the value and the associated confidence intervals were evaluated.
+P-values reflected the significance of a given correlation coefficient with statistical significance being declared when the corresponding P-value was lower than 0.05.
+PPCC analyses were undertaken for data on the original scale and for data as a logarithmic transform.
 Transforming data prior to calculating the coefficients changes several things.
@@
 \par
-Five general linear models were created (see Table \ref{tab:tab1}).
-Both the Pearson Correlation Coefficients and General Linear Models were created using the R statistical programming language \citep{rcoreteamLanguageEnvironmentStatistical2021} with the Caret package \citep{kuhnBuildingPredictiveModels2008}.
-General Linear Models were chosen as they offer the ability to produce statistical models that are explicit in the relationships between predictors and response variables.
+Five general linear models (GLMs) were created (see Table \ref{tab:tab1}).
+GLMs were chosen as they offer the ability to produce statistical models that are explicit in the relationships between predictors and response variables.
 General Linear Models also allowed the exploration of interactions between predictors and allow for easily comparable differences in the influence and magnitude of relationships.
-Model fit was measured in $R^2$ and adjusted $R^2$ as well as F statistics.
-T-tests were used to determine if predictors significantly contributed to their models when accounting for other variables, showing which specific years and areas contributed significantly.
+Model fit was measured in terms of $R^2$ and adjusted $R^2$ as well as F statistics, and t-tests were used to determine if predictors significantly contributed to the models when accounting for other variables, showing which specific years and areas contributed significantly.
+Both the PPCC and GLM analyses were created using the R statistical programming language \citep{rcoreteamLanguageEnvironmentStatistical2021} with the Caret package \citep{kuhnBuildingPredictiveModels2008}.
 \par
-A variety of alternate methods were also explored, including splines, hierarchical regression, General Additive Models, and Generalised Linear Models.
-These alternative approaches were not used as final models due to offering no further insights or improvements in accuracy.
+A variety of alternate methods were also explored, including splines, hierarchical regression, general additive models, and generalised linear models.
+These alternative approaches were not used as final models due to offering little further insights or improvements in accuracy.
 \par
 \subsection{Model Validation}
-Models were validated using K-fold cross validation calculated.
-K-fold cross validation works by removing a subset of data from the sample used to train models and then predicts those variables to determine how sensitive the model is to changes in the sample data.
-For this analysis each model was validated using 10 folds, repeated 100 times.
+Models were validated using K-fold cross validation.
+This was performed by removing a subset of data from the sample used to train models and then predicting those variables to determine how sensitive the model is to changes in the sample data.
+For this analysis, each model was validated using 10 folds, repeated 100 times.
 \par
@@
 
-Pearson Correlation Coefficients of the transformed, centred and scaled variables are shown in Table \ref{tab:pearson_correlation}.
-All correlations were found to be statistically significant (P $<$ 2.200E-16), and except for 'average price' all variables were positively correlated.
+PPCCs values for the transformed, centred and scaled variables 
+% # TODO: CENTRED AND SCALED VARIABLES NEED TO BE DEFINED!!! 
+are shown in Table \ref{tab:pearson_correlation}.
+All correlations were found to be statistically significant (p $<$ 2.200E-16), and except for Average Sale Price all variables were positively correlated.
 With water use, area harvested and emissions being positively correlated to yield, it can be considered that more resources and area are likely to lead to greater yields.
-Average sale price's negative correlation to yield, water use, area and scope one emissions, indicated that size and fuel separately were not the determining factor for average sale price.
-The negative correlations are not causal relationships but relative (using more water does not cause lower sale prices).
+The negative correlations between Average Sale Price and Yield, Water Use, Area and Scope One Emissions each indicated that Area Harvested and Fuel Use separately were not the determining factor for average sale price.
+The negative correlations are associative, not causal relationships (using more water does not cause lower sale prices).
 \par
@@
 
-Each model had a high $R^2$ value, indicating that most of the variance within the data was described by the models (see Table~\ref{tab:modelsummary}).
-The models were found to be a good fit, with overall F-tests being statistically significant (P $<$ 2.200E-16).
-And, aside from 3 variables, F-tests across each model's variables were also significant (with all being at least, P $<$ 0.05).
-The three exceptions were: scope one emissions in Model 3 (P=0.22) and Model 4 (P=0.0.39), and the interaction between area harvested and water used in model 2 (P=0.22).
-Note that, scope one emissions was included in all models to directly compare the response variables as ratios of vineyard size to raw values and because it was strongly correlated to the response variable in every model (except model 5); especially for Models 1 and 4 (Table \ref{tab:pearson_correlation}).
+Each model had a high $R^2$ value, indicating that most of the variance within the data was described by the models (see Table~\ref{tab:modelsummary}) and statistical significance of F-tests (p $<$ 2.200E-16).
+Aside from 3 variables, all regression parameters were also statistically significant from zero, F-tests across each model's variables were (p $<$ 0.05).
+The three exceptions were: scope one emissions in Model 3 (p=0.22) and Model 4 (p=0.0.39), and the interaction between area harvested and water used in model 2 (p=0.22).
+Note that scope one emissions was included in all models to directly compare the response variables as ratios of vineyard size to raw values and because it was strongly correlated to the response variable in every model (except model 5) especially for Models 1 and 4 (Table \ref{tab:pearson_correlation}).
 %A full list of regression coefficients 95\% CIs and p-values for each of the four models is provided in the appendix. % You gotta add these to the appendix in a more readable fashion
@@
 
-Models' continuous variable's coefficient values are summarised in Table \ref{tab:contvar}.
-Model 1 showed all coefficients except for the intercept were significantly contributing to the model (P < 0.05).
+Coefficients related to continuous variables are summarised in Table \ref{tab:contvar}.
+In Model 1 all coefficients except for the intercept significantly contributed to the model (p $<$ 0.05).
 Model 2's coefficients were all statistically significant.

### 2023-12-17 article.tex
edits done (pending table fixes)

@@
 Coefficients related to continuous variables are summarised in Table \ref{tab:contvar}.
-In Model 1 all coefficients except for the intercept significantly contributed to the model (p $<$ 0.05).
-Model 2's coefficients were all statistically significant.
-However, for Models 3, 4 and 5 scope one emissions did not significantly contribute.
-And, Model 4 only saw statistically significant contributions from the intercept and water use.
-Although the coefficient for water use was statistically significant for each model, it did not have the highest value, instead area harvested, being an order of magnitude greater dominated the models.
+In Model 1 all coefficients except for the intercept significantly contributed to the model (p $<$ 0.05), and in Model 2 all coefficients were statistically significant (p $<$ 0.05).
+However, scope one emissions did not significantly contribute to models 3, 4 and 5.
+Model 4 only had statistically significant contributions from the intercept and water use.
+Although the coefficient for water use was statistically significant for each model, it did not have the highest value; instead, area harvested, being an order of magnitude greater, dominated the models.
 Model 5 achieved a similar $R^2$ to Model 4 without area harvested, having stronger influences from water use and scope one emissions.
@@
 
-The regression coefficients for the year for each model is depicted in Figure \ref{fig:yearly}.
-The first year for a model's data is used as the baseline; 2012 for Model 1 and 2, and 2015 for Model 3, 4 and 5.
-The Adelaide Hills is used as the regional baseline with the interaction between year and region using the first year and the Adelaide Hills as the baseline.
+The regression coefficients for the Year variable of each model is depicted in Figure \ref{fig:yearly}.
+For each model, the first year for a model's was used as the baseline: 2012 for Models 1 and 2, and 2015 for Models 3, 4 and 5.
+Adelaide Hills is used as the regional baseline with the interaction between year and region using the first year and the Adelaide Hills as the baseline.
 Region and year contributed, in some but not all cases, more than the other variables.
-However, some years are not significant, as they are not statistically different from 0, given their error.
-Models 4 and 5 are very similar, indicating that the exclusion of area does not greatly affect the contribution from yearly influence.
-Models 4 and 5 have the most prominent trends, showing an increase in yearly effects over time, with Model 3 also increasing from 2016 to 2018 but plateau afterwards.
-Model 1 and 2 do not show a clear trend but do drop during 2017 and 2018 after increasing in the first 3 years.
+However, the coefficients varied substantially over the years, and were not significantly different from zero in some years.
+Models 4 and 5 are very similar in Figure \ref{fig:yearly}, indicating that the exclusion of area does not greatly affect the contribution from yearly influence.
+Models 4 and 5 have the most prominent trends, showing an increase in yearly effects over time, with Model 3 also increasing from 2016 to 2018, but plateauing afterwards.
+Model 1 and 2 do not show a clear trend, but do drop during 2017 and 2018 after increasing in the first 3 years.
 \par
@@
 The warmer and drier regions tend to also have larger vineyards.
-These regional differences are further shown in Figure \ref{fig:yield_vs_value_area}, as a ratio of vineyard area.
-Where, again there is a pronounced difference in 'Hot' and 'Very Dry' regions producing more per area with lower average sale prices per area than other regions.
+These regional differences are further shown in Figure \ref{fig:yield_vs_value_area}, as a ratio of vineyard; where there is a pronounced difference in 'Hot' and 'Very Dry' regions producing more per area with lower average sale prices per area than other regions.
 \par
@@
 
-Figure \ref{fig:yield_vs_value_area} further shows the emphasis that 'Hot' areas have on high yields with low average sale price compared with other regions.
+Figure \ref{fig:yield_vs_value_area} further shows the emphasis that 'Hot' areas have on high yields, with low average sale price compared with other regions.
 Scaling average price and yield by area shows a strong negative trend, trading quantity for higher sales prices.
@@
 Table \ref{tab:kfold} shows the validation results of each of the models.
-The $R^2$ measures of fit show similar results to the initial models, with a slight decrease.
-Indicating that the models are robust and consistent.
+The $R^2$ measures of fit show similar results to the initial models, with a slight decrease as expected, indicating that the models are robust and consistent.
 %
@@
 We identified two main limitations to our linear modelling.
-First model 1 and 2 over-predicting yield may have been due to preventative measures brought on by regional pressures such as fire, frost and disease.
+First, model 1 and 2 over-predicting yield may have been due to preventative measures brought on by regional pressures such as fire, frost and disease.
 More fuel and water was likely used to prevent these issues from spreading within a region, thus disproportionately affecting some vineyards compared to others locally.

### 2023-12-18 article.tex
mardi edits

@@
 The consideration of sustainability within viticulture is also subject to environmental and sociodemographic pressures \citep{santiago-brownSustainabilityAssessmentWineGrape2015}.
-In the Australian context, these pressures include biosecurity and climate and international market changes \citep{canadellMultidecadalIncreaseForest2021,longbottomRoleVineyardPractices2015,oliverReviewSoilPhysical2013}.
+In the Australian context, these pressures include biosecurity, climate and international market changes \citep{canadellMultidecadalIncreaseForest2021,longbottomRoleVineyardPractices2015,oliverReviewSoilPhysical2013}.
 \par
@@
 Additional variables were considered for analysis but were excluded due to being either underreported or had insignificant contributions to model accuracies.
-Variables explored but not used due to low reporting values included fertiliser, electricity, and scope two emissions.
-Variables considered but ultimately removed due to a lack of significant contributions to models, included the use of renewable energy, contractor use, and pressures such as frost, fire and disease.
+Variables explored but not used due to low reporting values included fertiliser (kg), electricity (Kw/h), and scope two emissions (TCO$^2$ equivalent).
+Variables considered but ultimately removed due to a lack of significant contributions to models, included the use of renewable energy, use of contractors, and pressures such as frost, fire and disease.
 \par
 Data preprocessing was conducted prior to analysis using the Python programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
-Preprocessing included the conversion from fuel to scope one emissions and prior calculations for all continuous variables which included logarithmic transformations, centring and scaling by standard deviation.
+Preprocessing included the conversion of fuel to scope one emissions and prior calculations for all continuous variables which included logarithmic transformations, centring and scaling by standard deviation.
 We converted multiple emission sources into scope one emissions using the equation given from the Australian National Greenhouse Accounts Factors \citep{agdeeNationalGreenhouseAccounts2021}.
@@
   \includegraphics{my_map.pdf}}
-  \caption{Map of each GI Regions' average income for a vineyard of that region (average grape sale price per tonne $\times$ total tonnes yielded).}\label{fig:map}
+  \caption{Map of each GI Regions' (outlined by Wine Australia) average income for a vineyard of that region (average grape sale price per tonne $\times$ total tonnes yielded).}\label{fig:map}
 \end{figure}%%
@@
 The negative correlations between Average Sale Price and Yield, Water Use, Area and Scope One Emissions each indicated that Area Harvested and Fuel Use separately were not the determining factor for average sale price.
-The negative correlations are associative, not causal relationships (using more water does not cause lower sale prices).
+The negative correlations are associative, not causal relationships (i.e using more water does not cause lower sale prices).
 \par
@@
 The Wine Australia reports also show that warm inland regions have seen a decline in profit over the past decade, whereas regions with lower average sales prices did not \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
-Vineyards in warm inland regions also tend to contain larger vineyards, making up for lower sale prices with larger yields.
-Considering the negative correlation of average price to area, for this strategy to work, economies of scale become and important factor.
+Vineyards in warm inland regions also tend to be larger, making up for lower sale prices with larger yields.
+Considering the negative correlation of average price to area, for this strategy to work, economies of scale become an important factor.
 Given the large quantities of grapes that can be produced by some vineyards, even at low margins there is the potential to be profitable.
@@
 Furthermore, intensive farming in general is known to jeopardise the sustainability of an operation through the degradation of soil and waterways \citep{capelloEffectsTractorPasses2019,linHydropedologySynergisticIntegration2012,pisciottaGroundwaterNitrateRisk2015}.
-There are established methods that can help to mitigate these effect, such as the use of cover crops and midrow crop rotation.
-However, it has become more apparent that the active reduction of grape yield, through methods such as thinning, can help improve soil health \citep{condursoEffectsClusterThinning2016,wangChangesGlobalAroma2019}.
+There are established methods that can help to mitigate these effects, such as the use of cover crops, midrow crop rotation and efficient irrigation.
 \par
-Some regions appeared to produce grapes of lower average sale price at scale whilst others focussed on producing higher priced grapes in lower volumes.
-This empirical finding is consistent with Wine Australia's annual reports, which shows that some GI regions, such as the Riverland, are known for producing large amounts of lower grade (low value per tonne) grapes \citep{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
-Comparatively, other regions, such as Tasmania, only produce grapes of higher sales price but in smaller quantities.
+Some regions appeared to produce grapes of lower average sale price at scale whilst others achieved higher priced grapes in lower volumes.
+This empirical finding is consistent with Wine Australia's annual reports, which shows that some GI regions are known for producing large amounts of lower grade (low value per tonne) grapes \citep{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
+Comparatively, other regions only produce grapes of higher sales price but in smaller quantities.
 The difference in pricing per tonne between the lowest and highest regional average sales prices was almost a hundred times, showing that region had a profound influence.
-Some regions also had a mixture of high and low average sales price and yield such as the Yarra Valley, showing that some regions were not solely locked into chosing a focus on high yields or high average sales price.
+Some regions also had a mixture of high and low average sales price and yield showing regional variability in pricing which may be explained by varieties produced.
 A further possibility is the existence of regional upper limits on potential sales price, or that there are diminishing returns in some regions when pursuing higher sales prices or quantity; however these types of relationships may be obfuscated by knowledgeable winegrowers who avoid such pitfalls.
@@
 This is most notable when grouping regions by climate, especially when considering GI Regions in the 'Hot Very Dry' climate (see Figure \ref{fig:yield_vs_value_area}).
-Figure \ref{fig:spider} also shows that comparatively 'Warm' and 'Dry' Regions manage their resources incredibly efficiently having generally larger areas using similar resources too those in 'Cool' regions but having areas comparable to regions in 'Hot' climates.
+Figure \ref{fig:spider} also shows that comparatively 'Warm' and 'Dry' Regions manage their resources incredibly efficiently having generally larger areas using similar resources to those in 'Cool' regions but having areas comparable to regions in 'Hot' climates.
 The coefficients for Model 4 also show a greater benefit of resource use per area when producing grapes at higher average sale prices, showing higher resource use per area reflective of higher average sales price (see Table \ref{tab:summary}).
@@
 Having more data for each region would also be beneficial, allowing greater comparison between regions.
-More variables, such as soil quality and health, may also help to discern vineyards that can produce larger volumes of grapes at higher prices.
 \par
@@
 Moreover, regional and yearly variations exhibited substantial effects on vineyard outputs, impacting both sales price and quantity.
-The connection observed between smaller vineyards and higher grape sale price suggests that the management of smaller properties might be more streamlined and effective.
+The connection observed between smaller vineyards and higher grape sale price suggests that the management of smaller properties might be more efficient and profitable.
 

### 2023-12-18 article.tex
fixed sf and notation in tables

@@
           \textbf{Variable} & \textbf{Mean} & \textbf{\begin{tabular}[c]{@{}l@{}}Standard\\ Deviation\end{tabular}} & \textbf{Minimum} & \textbf{Maximum} \\ \midrule
-          {Yield (tonnes)} & 7.757E+02 & 2.179E+03 & 1.000E+00 & 7.231E+04 \\
+          {Yield (tonnes)} & 7.76E+02 & 2.18E+03 & 1.00E+00 & 7.23E+04 \\
           \multicolumn{1}{l}{} &  &  &  &  \\
-          {Area Harvested (ha)} & 6.670E+01 & 1.337E+02 & 7.000E-02 & 2.436E+03 \\
+          {Area Harvested (ha)} & 6.67E+01 & 1.34E+02 & 7.00E-02 & 2.44E+03 \\
           \multicolumn{1}{l}{} &  &  &  &  \\
-          {Water Used (ML)} & 7.471E+06 & 5.646E+08 & 1.000E+00 & 4.268E+10 \\
+          {Water Used (ML)} & 7.47E+06 & 5.65E+08 & 1.00E+00 & 4.27E+10 \\
           \multicolumn{1}{l}{} &  &  &  &  \\
-          {Scope One Emissions ($tCO_{2}e$)} & 4.173E+04 & 8.571E+04 & 6.755E+00 & 2.110E+06 \\
+          {Scope One Emissions ($tCO_{2}e$)} & 4.17E+04 & 8.57E+04 & 6.76E+00 & 2.11E+06 \\
           \multicolumn{1}{l}{} &  &  &  &  \\
-          $\mbox{Yield (tonnes)}\over \mbox{Area harvested (ha)}$ & 1.009E+01 & 8.127E+00 & 4.000E-02 & 8.634E+01 \\
+          $\mbox{Yield (tonnes)}\over \mbox{Area harvested (ha)}$ & 1.01E+01 & 8.13E+00 & 4.00E-02 & 8.63E+01 \\
           \multicolumn{1}{l}{} &  &  &  &  \\
-          {Average Sale Price (AUD/tonne)} & 1.477E+03 & 9.216E+02 & 1.600E+02 & 2.600E+04 \\
+          {Average Sale Price (AUD/tonne)} & 1.48E+03 & 9.22E+02 & 1.60E+02 & 2.60E+04 \\
           \multicolumn{1}{l}{} &  &  &  &  \\
-          $\mbox{Average Sale Price (AUD/tonne)} \over \mbox{Area Harvested (ha)}$ & 1.347E+02 & 5.711E+02 & 1.753E-01 & 2.979E+04 \\ \bottomrule
+          $\mbox{Average Sale Price (AUD/tonne)} \over \mbox{Area Harvested (ha)}$ & 1.35E+02 & 5.71E+02 & 1.75E-01 & 2.98E+04 \\ \bottomrule
           \end{tabular}
@@
 
-PPCCs values for the transformed, centred and scaled variables 
-% # TODO: CENTRED AND SCALED VARIABLES NEED TO BE DEFINED!!! 
-are shown in Table \ref{tab:pearson_correlation}.
+PPCCs values for the transformed, centred and scaled variables are shown in Table \ref{tab:pearson_correlation}.
+The process of centring and scaling variables changes them so that their mean is centred on 0 and all values are scaled to the variables standard deviation, such that 1 is equal to the variables standard deviation.
 All correlations were found to be statistically significant (p $<$ 2.200E-16), and except for Average Sale Price all variables were positively correlated.
@@
 
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+% \usepackage{multirow}
 \begin{table}[]
@@
   \begin{tabular}{@{}cccccccc@{}}
-      \begin{tabular}{@{}cccccccc@{}}
   \toprule
   \textbf{} & \textbf{} & \textbf{Intercept} & \textbf{Area Harvested} & \textbf{Water Used} & \textbf{\begin{tabular}[c]{@{}c@{}}Scope One\\ Emissions\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Area Harvested \\ ** \\ Scope One Emissions\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Area Harvested \\ ** \\ Water Used\end{tabular}} \\ \midrule
-        \multirow{2}{*}{Model 1} & Coefficient & -0.0332 & 0.7418 & 0.0866 & 0.0673 &  &  \\
-         & Std Error & 0.0196 & 0.0100 & 0.0089 & 0.0080 &  &  \\
+  \multirow{2}{*}{Model 1} & Coefficient & -3.32E-02 & 7.42E-01 & 8.66E-02 & 6.73E-02 &  &  \\
+   & Std Error & 1.96E-02 & 1.00E-02 & 8.90E-03 & 8.00E-03 &  &  \\
    &  &  &  &  &  &  &  \\
-         \multirow{2}{*}{Model 2} & Coefficient & 0.1696 & 0.5774 & 0.1079 & 0.0850 & -0.0497 & -0.0535 \\
-         & Std Error & 0.0591 & 0.0148 & 0.0131 & 0.0117 & 0.0081 & 0.0084 \\
+  \multirow{2}{*}{Model 2} & Coefficient & 1.70E-01 & 5.77E-01 & 1.08E-01 & 8.50E-02 & -4.97E-02 & -5.35E-02 \\
+   & Std Error & 5.91E-02 & 1.48E-02 & 1.31E-02 & 1.17E-02 & 8.10E-03 & 8.40E-03 \\
    &  &  &  &  &  &  &  \\
-         \multirow{2}{*}{Model 3} & Coefficient & 0.0181 & 0.9713 & -0.0231 & -0.0070 &  &  \\
-         & Std Error & 0.0130 & 0.0072 & 0.0069 & 0.0057 &  &  \\
+  \multirow{2}{*}{Model 3} & Coefficient & 1.81E-02 & 9.71E-01 & -2.31E-02 & -7.00E-03 &  &  \\
+   & Std Error & 1.30E-02 & 7.20E-03 & 6.90E-03 & 5.70E-03 &  &  \\
    &  &  &  &  &  &  &  \\
-         \multirow{2}{*}{Model 4} & Coefficient & 0.1450 & 0.0024 & -0.0466 & -0.0170 & 0.0115 & 0.0014 \\
-         & Std Error & 0.0528 & 0.0150 & 0.0143 & 0.0118 & 0.0079 & 0.0083 \\
+  \multirow{2}{*}{Model 4} & Coefficient & 1.45E-01 & 2.40E-03 & -4.66E-02 & -1.70E-02 & 1.15E-02 & 1.40E-03 \\
+   & Std Error & 5.28E-02 & 1.50E-02 & 1.43E-02 & 1.18E-02 & 7.90E-03 & 8.30E-03 \\
    &  &  &  &  &  &  &  \\
-         \multirow{2}{*}{Model 5} & Coefficient & 0.1517 &  & -0.0404 & -0.0171 &  &  \\
-         & Std Error & 0.0527 &  & 0.0113 & 0.0097 &  &  \\ \bottomrule
-        \end{tabular}
-
+  \multirow{2}{*}{Model 5} & Coefficient & 1.52E-01 &  & -4.04E-02 & -1.71E-02 &  &  \\
+   & Std Error & 5.27E-02 &  & 1.13E-02 & 9.70E-03 &  &  \\ \cmidrule(l){2-8} 
   \end{tabular}

### 2023-12-18 Cover Letter/main.tex
cover letter

\documentclass{article}

\usepackage{charter} 
\usepackage{ragged2e}
\usepackage{graphicx}
\usepackage[
	a4paper,    % Paper size
	top=1in,    % Top margin
	bottom=1in, % Bottom margin
	left=1in,   % Left margin
	right=1in,  % Right margin
	%showframe  % Uncomment to show frames around the margins for debugging purposes
]{geometry}

\setlength{\parindent}{0pt}     % Paragraph indentation 
\setlength{\parskip}{1em}       % Vertical space between paragraphs

\usepackage{graphicx}       % Required for including images

\usepackage{fancyhdr}       % Required for customizing headers and footers

\fancypagestyle{firstpage}{%
	\fancyhf{} % Clear default headers/footers
	\renewcommand{\headrulewidth}{0pt} % No header rule
	\renewcommand{\footrulewidth}{1pt} % Footer rule thickness
}

\fancypagestyle{subsequentpages}{%
	\fancyhf{} % Clear default headers/footers
	\renewcommand{\headrulewidth}{1pt} % Header rule thickness
	\renewcommand{\footrulewidth}{1pt} % Footer rule thickness
}

\AtBeginDocument{\thispagestyle{firstpage}} % Use the first page headers/footers style on the first page
\pagestyle{subsequentpages} % Use the subsequent pages headers/footers style on subsequent pages

%----------------------------------------------------------------------------------------

\begin{document}

%----------------------------------------------------------------------------------------
%	FIRST PAGE HEADER
%----------------------------------------------------------------------------------------

\includegraphics[scale=0.5]{logo.jpg} % Logo

\vspace{-1em} % Pull the rule closer to the logo

\rule{\linewidth}{1pt} % Horizontal rule

\bigskip\bigskip % Vertical whitespace

%----------------------------------------------------------------------------------------
%	YOUR NAME AND CONTACT INFORMATION
%----------------------------------------------------------------------------------------

\hfill
\begin{tabular}{l @{}}
	$29^{th}$ October, 2023\\ % Date
	\textbf{Bryce Ronald William Polley $|$ PhD Candidate}\\
	Queensland University of Technology \\
	School of Mathematical Sciences\\~\\
	Address: Level 8 Y Block, 2 George Street, \\
	Brisbane, Queensland, Australia, 4000\\
	Phone: +61 400 310 105 \\
	Email: bryce.polley@awri.com.au
\end{tabular}

\bigskip % Vertical whitespace

%----------------------------------------------------------------------------------------
%	ADDRESSEE AND GREETING
%----------------------------------------------------------------------------------------

\begin{tabular}{@{} l}
	\textbf{Professor Bojie Fu,}\\
	Editor-in-Chief \\
	Geography and Sustainability\\
\end{tabular}

\bigskip % Vertical whitespace

Dear Professor Bojie Fu,
\\~\\
\parskip % Vertical whitespace

%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------
\centering \textbf{Submission of Manuscript for Consideration by \\Geography and Sustainability}\\
\bigskip % Vertical whitespace
\justifying
\newline \noindent
We wish to submit a new manuscript entitled "\textit{The influence of resource use on yield versus sale price trade-off in Australian vineyards}" for consideration by Geography and Sustainability.
\newline \noindent
In this manuscript we explore natural resource use in agricultural wine production, emphasising sustainability in the trade-off between the sale price and quantity of product. We explore the economic and environmental risks induced by producing grapes at greater yields. We find that lower grape sale price is not solely related to the overall expenditure of resources but rather to efficient management, as well as regional and seasonal variations. We use novel, extensive nationwide data from Australia, collected over a decade by Sustainable Winegrowing Australia. We construct and analyse several models of the multifaceted, multi-resource nature of vineyard management. 
\newline \noindent
We believe that this manuscript is appropriate for publication by Geography and Sustainability as it demonstrates the sustainability of different winegrowing production strategies across winegrowing regions in Australia. These results illustrate the relationship between human interaction and the environment as driven by agricultural management strategies and the resources required to achieve the necessary outcomes to maintain economic sustainability using environmental resources. Thank you for your consideration of this manuscript.
\bigskip \noindent% Vertical whitespace

\includegraphics[scale=0.1]{signature_block.png}
\newline \noindent
Bryce Polley (corresponding author) on behalf of all coauthors: Kate Helmsted, Mardi Longbottom, Madeline Mitchell, Kerrie Mengersen\\
\newline \noindent
\textit{Corresponding Author}, \textit{Queensland University of Technology}

\end{document}

### 2024-03-10 Cover Letter/main.tex
backing up article edits

@@
 \begin{tabular}{l @{}}
-	$29^{th}$ October, 2023\\ % Date
-	\textbf{Bryce Ronald William Polley $|$ PhD Candidate}\\
+	$26^{th}$ January, 2023\\ % Date
+	\textbf{Bryce Ronald William Boyd $|$ PhD Candidate}\\
 	Queensland University of Technology \\
@@
 \newline \noindent
-Bryce Polley (corresponding author) on behalf of all coauthors: Kate Helmsted, Mardi Longbottom, Madeline Mitchell, Kerrie Mengersen\\
+Bryce Boyd (corresponding author) on behalf of all coauthors: Kate Helmstedt, Mardi Longbottom, Madeline Mitchell, Kerrie Mengersen\\
 \newline \noindent

### 2024-03-10 titlepage.tex
backing up article edits

--- 2023-12-18 article.tex
@@
 \usepackage{multirow}
-
 \usepackage{csvsimple}
@@
 \journal{Geography and Sustainability}
-%1.Full Length Article A full-length article should be a substantial and in-depth research study regarding a particular state of issue through several techniques or approaches. 
-% The main text should be approximately 6,000 words in length, but it should not exceed 8,000 words (excluding abstract, references, tables, figures, and appendices).
-%A maximum of 250 words abstract and up to 10 displayed items (figures and tables) is allowed. A full-length article should include an Introduction,
-%Materials and methods, Results, Discussion, Conclusions, and References, which can be accompanied by Supplementary material.
 
 \begin{document}
-\begin{linenumbers}
 \begin{frontmatter}
-%% Title, authors and addresses
-%% use the tnoteref command within \title for footnotes;
-%% use the tnotetext command for theassociated footnote;
-%% use the fnref command within \author or \affiliation for footnotes;
-%% use the fntext command for theassociated footnote;
-%% use the corref command within \author for corresponding author footnotes;
-%% use the cortext command for theassociated footnote;
-%% use the ead command for the email address,
-%% and the form \ead[url] for the home page:
-%% \title{Title\tnoteref{label1}}
-%% \tnotetext[label1]{}
-%% \author{Name\corref{cor1}\fnref{label2}}
-%% \ead{email address}
-%% \ead[url]{home page}
-%% \fntext[label2]{}
-%% \cortext[cor1]{}
-%% \affiliation{organization={},
-%%            addressline={}, 
-%%            city={},
-%%            postcode={}, 
-%%            state={},
-%%            country={}}
-%% \fntext[label3]{}
+
 \title{The influence of resource use on yield versus sale price trade-off in Australian vineyards}
-%% use optional labels to link authors explicitly to addresses:
-%% \author[label1,label2]{}
-%% \affiliation[label1]{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
-%%
-%% \affiliation[label2]{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
-%\affiliation[label1]{organization={QUT},
-%  addressline={},
-%  city={},
-%  postcode={},
-%  state={QLD},
-%  country={}}
-%\affiliation[label2]{organization={AWRI},
-%  addressline={},
-%  city={},
-%  postcode={},
-%  state={SA},
-%  country={}}
-%\affiliation[label3]{organization={Food Agility CRC},
-%  addressline={},
-%  city={},
-%  postcode={},
-%  state={Vic},
-%  country={}}
-\author[label1,label2,label3]{Author}
-\date{02/08/2023}
+
+\affiliation[label1]{organization={Queensland University of Technology}}
+\affiliation[label2]{organization={Australian Wine Research Institute}}
+\affiliation[label3]{organization={Food Agility CRC}}
+ \author[label1,label2,label3]{Bryce Boyd}
+ \author[label1]{Kate Helmstedt}
+ \author[label2]{Mardi Longbottom}
+ \author[label1]{Kerrie Mengersen}
+
+ \begin{highlights}
+    \item Comparative analysis of resource use, average sale price and quantity in Australian winegrowing.
+    \item Regional comparison of outcomes and resource use in Australian winegrowing regions.
+    \item Baseline models for comparing vineyards.
+    \item Analysis of national, decade long data source.
+  \end{highlights}
+
  \begin{abstract}
@@
     %%Research highlights
-\begin{highlights}
-  \item Comparative analysis of resource use, average sale price and quantity in Australian winegrowing.
-  \item Regional comparison of outcomes and resource use in Australian winegrowing regions.
-  \item Baseline models for comparing vineyards.
-  \item Analysis of national, decade long data source.
-\end{highlights}
+
 \end{frontmatter}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                main text             %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Introduction}
-The global focus on sustainability in agronomic industries has changed the way in which these enterprises do business.
-A dilemma exists across agriculture through shared fundamental considerations of resource use such as water and fuel, and the resultant yield and crop value that are produced \citep{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016, zhuEffectsNitrogenLevel2017}.
-The average price of grapes for wine production is driven through its integration within the wine industry.
-An important connection between grapes and their price is the grape's perceived quality, which initially defines a wine's potential through the grape's chemical makeup \citep{blackTerpenoidsTheirRole2015,schreierFlavorCompositionWines1979}.
-Grapes of higher perceived quality or grapes from well known regions are likely to have higher prices \citep{wineaustraliaNationalVintageReport2021}.
-Grape quality is connected to the market value of winegrapes, with Wine Australia explicitly defining grape quality through the use of discrete price brackets in their annual reports \citep{winemakersfederationofaustraliaNationalVintageReport2018}.
-It is important to note that the generalisation made to reflect quality through using average price assumes a due diligence of those who purchase the grapes \citep{yeggeInfluenceSensoryNonsensory2001}.
-The economic sustainability of a vineyard is tied to this market culture, driven by the wine industry.
-The consideration of sustainability within viticulture is also subject to environmental and sociodemographic pressures \citep{santiago-brownSustainabilityAssessmentWineGrape2015}.
-In the Australian context, these pressures include biosecurity, climate and international market changes \citep{canadellMultidecadalIncreaseForest2021,longbottomRoleVineyardPractices2015,oliverReviewSoilPhysical2013}.
-\par
-There is an extensive amount of research into the varied effects of factors on grape quality and yield \citep{heFruitYieldPrediction2022,laurentLocalInfluenceClimate2022,liakosMachineLearningAgriculture2018}.
-Due to the lack of long-term and in-depth data there is a lack of research on grape sale price and its driving factors.
-Furthermore, individual factors are often studied in isolation with yield and sales price not appearing together \citep{abbalDecisionSupportSystem2016}.
-The lack of consolidated datasets restricts the ability to gain statistical insights at large scales and across multiple regions.
-As a result, broader studies are lacking \citep{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
-The dataset used for this analysis includes data spanning 10 years from a multitude of vineyards located over a diverse range of Australian winegrowing regions.
-We use this dataset to describe the relationship of resources related to water and fuel use with the output yield and average sale price of the resultant product, taking into account the size and location of the vineyard.
-In a practical sense, this aim is a baseline for comparison: given a vineyard within Australia, one could estimate the comparative efficiency with regard to the trade-off between invested resources, yield and average sale price.
-This is the first time that such a trade-off has been confirmed explicitly across such varying regions, scales and climates in the Australian winegrowing industry.
-\section{Methods}
-\subsection{Data}
-
-\begin{table}[]
-\caption{Summary of models; their predictors, covariates and variable interactions.}\label{tab:tab1}
-  \resizebox{\textwidth}{!}{\begin{tabular}{@{}ccccc@{}}
-    \toprule
-    & \textbf{Response} & \textbf{Predictors} & \textbf{Covariates} & \textbf{Interactions} \\ \midrule
-    \textbf{Model 1} & Yield & \begin{tabular}[c]{@{}c@{}}Water Used\\ scope one Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & N/A \\
-    \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-    \textbf{Model 2} & ${ \textrm{Yield}}\over{ \textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ scope one Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * scope one Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular} \\
-    \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-    \textbf{Model 3} & {$\textrm{Yield} {\times} \textrm{Average Sale Price}$} & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope One Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & N/A\\
-    \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-    \textbf{Model 4} & Average Sale Price & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope One Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * Scope One Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular}
-        \\
-        \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-     \textbf{Model 5} &  Average Sale Price & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope One Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}} Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}} Year * Region\end{tabular}
-        \end{tabular}}
-\end{table}
-
-Data used in these analyses were obtained from Sustainable Winegrowing Australia and Wine Australia.
-Sustainable Winegrowing Australia is Australia's national wine industry sustainability program, which aims to facilitate grape  growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
-Wine Australia is an Australian Government statutory authority governed by the Wine Australia Act \citep{attorney-generalsdepartmentWineAustraliaCorporation2010}.
-Data collected by Wine Australia are publicly available.
-The response (predicted) variables in this analysis were yield, defined as the total tonnes of grapes harvested, and average sale price of grapes in Australian dollars per tonne.
-Both response variables were examined as totals and also relative to area harvested (see Table \ref{tab:tab1}).
-Values were compared in this manner to observe how economies of scale affect the use of resources.
-\par
-Data recorded by Sustainable Winegrowing Australia were entered manually by winegrowers voluntarily using a web based interface.
-For each model, vineyards were only included if they recorded all the variables used for the corresponding model (see Table \ref{tab:tab1}).
-Each vineyard had at least recorded region, harvest year, yield and area harvested.
-Other variables that were used but not present for every single vineyard were average sale price, water used and fuel used (diesel, petrol, biodiesel and LPG).
-To enable direct comparisons between fuels, fuel use was converted to tonnes of carbon dioxide equivalent and collectively referenced to as emissions.
-All variables were continuous except for harvest year and region, which were categorical variables (Table \ref{tab:tab1}).
-\par
-As data from Sustainable Wine Australia were voluntarily given, missing values were improved using regional average prices from the Wine Australia data.
-Data obtained from Wine Australia were collected via phone surveys and included: total tonnes purchased, average price per tonne and yearly change in price for region and grape varietal; with the data being publicly available.
-\par
-The dataset was split into two distinct subsets that could be used across the different models.
-The first subset contained all vineyards and was used for two models (Model 1 and Model 2, see Table \ref{tab:tab1}).
-The second subset contained vineyards which either recorded a value for average price of sale per tonne through Sustainable Winegrowing Australia, or were within a region with an average price of sale recorded by Wine Australia; this subset was used for three further models (Models 3, 4 and 5, see Table \ref{tab:tab1}).
-These subsets meant that the data would be limited to samples which had recorded values for the response variables (see Table \ref{tab:tab1}), as every sample/vineyard had a recorded value for yield but not average price of sale per tonne.
-\par
-The first subset of data (used for Model 1 and Model 2, see Table \ref{tab:tab1}) contained 5298 samples spanning the period from 2012 to 2022, covering 55 GI Regions and 1261 discrete vineyards.
-\par
-The second subset of data (used for Model 3, Model 4 and Model 5, see Table \ref{tab:tab1}) contained 2878 samples spanning the period from 2015 to 2022, covering 51 GI Regions and 944 separate vineyards.
-Average price of sale per tonne was extracted from both Wine Australia (1842 values) and Sustainable Winegrowing Australia (remaining 1036 values).
-\par
-Additional variables were considered for analysis but were excluded due to being either underreported or had insignificant contributions to model accuracies.
-Variables explored but not used due to low reporting values included fertiliser (kg), electricity (Kw/h), and scope two emissions (TCO$^2$ equivalent).
-Variables considered but ultimately removed due to a lack of significant contributions to models, included the use of renewable energy, use of contractors, and pressures such as frost, fire and disease.
-\par
-Data preprocessing was conducted prior to analysis using the Python programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
-Preprocessing included the conversion of fuel to scope one emissions and prior calculations for all continuous variables which included logarithmic transformations, centring and scaling by standard deviation.
-We converted multiple emission sources into scope one emissions using the equation given from the Australian National Greenhouse Accounts Factors \citep{agdeeNationalGreenhouseAccounts2021}.
-The calculation conducted using
-\begin{equation}
-\label{(1)}
-    tCO_{2}e={{Q \times EC \times EF1 + EF3}\over{1000}},
-\end{equation}
-where emissions was the the product of the quantity of fuel in litres, $Q$, a prescribed Energy Content, $EC$, and emission factors (as given by the Australian National Greenhouse Accounts factors) of scope one, $EF1$, and scope three, $EF3$, to tonnes of Carbon Dioxide Emission equivalent, $tCO2e$ \citep{departmentofclimatechangeenergytheenvironmentandwaterAustralianNationalGreenhouse2022}.
-\par
-\begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{my_map.pdf}}
-  \caption{Map of each GI Regions' (outlined by Wine Australia) average income for a vineyard of that region (average grape sale price per tonne $\times$ total tonnes yielded).}\label{fig:map}
-\end{figure}%%
-
-\par
-Differences in vineyard locations were captured through the use of Geographical Indicator Regions defined by Wine Australia \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
-Although vineyards generally differed in income, there were pronounced differences between regions as shown in Figure \ref{fig:map}.
-These differences were likely due to the site of a vineyard being important as it predetermines several physical parameters such as climate, geology and soil, making location a widely considered key determinant of grape yield and average sale price \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
-Each GI Region has its own unique mixture of climatic and geophysical properties that describes a unique winegrowing region within Australia and is a protected trademark under the Wine Australia act \citep{attorney-generalsdepartmentWineAustraliaCorporation2010}.
-Both Wine Australia and Sustainable Winegrowing Australia used the same GI Region categorical variable format to describe location.
-\par
- The climatic properties of each GI Region were summarised by using predefined classifications as per the \citet{sustainablewinegrowingaustraliaSustainableWinegrowingAustralia2021} user manual.
-The user manual describes climates by rainfall and temperature, creating supersets of Regions of similar climatic properties.
-The climatic groups were used to illustrate similarities and differences occurring in areas larger than GI Regions.
-\par
-\subsection{Analysis}
-Pairwise Pearson Correlation Coefficients (PPCC) were calculated to assess the potential existence of linear relationships between the input and predicted variables.
-To determine if a coefficient was indicative of a strong relationship, both the magnitude of the value and the associated confidence intervals were evaluated.
-P-values reflected the significance of a given correlation coefficient with statistical significance being declared when the corresponding P-value was lower than 0.05.
-PPCC analyses were undertaken for data on the original scale and for data as a logarithmic transform.
-Transforming data prior to calculating the coefficients changes several things.
-The logarithmic transform of the data alters the interpretation of the coefficients to percentage change; a coefficient will be indicative of the change in percentage of one variable compared to the other, scaling by standard deviation also changes this interpretation to be a percentage of that variables standard deviation.
-When considering the logarithmically transformed variables, a coefficient of 1 would indicate that the change of one variable by one percentage of its standard deviation would correlate to the other variable changing by one percent of its own standard deviation.
-The importance of this is the dimensionless nature of these relationships and that it can be translated directly to any vineyard's case that has a well known distribution.
-\par
-Five general linear models (GLMs) were created (see Table \ref{tab:tab1}).
-GLMs were chosen as they offer the ability to produce statistical models that are explicit in the relationships between predictors and response variables.
-General Linear Models also allowed the exploration of interactions between predictors and allow for easily comparable differences in the influence and magnitude of relationships.
-Model fit was measured in terms of $R^2$ and adjusted $R^2$ as well as F statistics, and t-tests were used to determine if predictors significantly contributed to the models when accounting for other variables, showing which specific years and areas contributed significantly.
-Both the PPCC and GLM analyses were created using the R statistical programming language \citep{rcoreteamLanguageEnvironmentStatistical2021} with the Caret package \citep{kuhnBuildingPredictiveModels2008}.
-\par
-A variety of alternate methods were also explored, including splines, hierarchical regression, general additive models, and generalised linear models.
-These alternative approaches were not used as final models due to offering little further insights or improvements in accuracy.
-\par
-\subsection{Model Validation}
-Models were validated using K-fold cross validation.
-This was performed by removing a subset of data from the sample used to train models and then predicting those variables to determine how sensitive the model is to changes in the sample data.
-For this analysis, each model was validated using 10 folds, repeated 100 times.
-\par
-\section{Results}
-\subsection{Exploratory Analysis}\label{sec:exp_anal}
- Table \ref{tab:summary} shows the summary statistics of each variable in its original units.
-The range of these values shows the level of difference between some vineyards, with operations differing by orders of magnitude in size, yield and average price of sale (See Table \ref{tab:tab1}).
-\par
-\begin{table}[]
-  \caption{Summary statistics of each continuous variable.}\label{tab:summary}
-      % \resizebox{\textwidth}{!}{
-        \begin{tabular}{@{}cllll@{}}
-          \toprule
-          \textbf{Variable} & \textbf{Mean} & \textbf{\begin{tabular}[c]{@{}l@{}}Standard\\ Deviation\end{tabular}} & \textbf{Minimum} & \textbf{Maximum} \\ \midrule
-          {Yield (tonnes)} & 7.757E+02 & 2.179E+03 & 1.000E+00 & 7.231E+04 \\
-          \multicolumn{1}{l}{} &  &  &  &  \\
-          {Area Harvested (ha)} & 6.670E+01 & 1.337E+02 & 7.000E-02 & 2.436E+03 \\
-          \multicolumn{1}{l}{} &  &  &  &  \\
-          {Water Used (ML)} & 7.471E+06 & 5.646E+08 & 1.000E+00 & 4.268E+10 \\
-          \multicolumn{1}{l}{} &  &  &  &  \\
-          {Scope One Emissions ($tCO_{2}e$)} & 4.173E+04 & 8.571E+04 & 6.755E+00 & 2.110E+06 \\
-          \multicolumn{1}{l}{} &  &  &  &  \\
-          $\mbox{Yield (tonnes)}\over \mbox{Area harvested (ha)}$ & 1.009E+01 & 8.127E+00 & 4.000E-02 & 8.634E+01 \\
-          \multicolumn{1}{l}{} &  &  &  &  \\
-          {Average Sale Price (AUD/tonne)} & 1.477E+03 & 9.216E+02 & 1.600E+02 & 2.600E+04 \\
-          \multicolumn{1}{l}{} &  &  &  &  \\
-          $\mbox{Average Sale Price (AUD/tonne)} \over \mbox{Area Harvested (ha)}$ & 1.347E+02 & 5.711E+02 & 1.753E-01 & 2.979E+04 \\ \bottomrule
-          \end{tabular}
-  % }
-  \end{table}
-
-PPCCs values for the transformed, centred and scaled variables 
-% # TODO: CENTRED AND SCALED VARIABLES NEED TO BE DEFINED!!! 
-are shown in Table \ref{tab:pearson_correlation}.
-All correlations were found to be statistically significant (p $<$ 2.200E-16), and except for Average Sale Price all variables were positively correlated.
-With water use, area harvested and emissions being positively correlated to yield, it can be considered that more resources and area are likely to lead to greater yields.
-The negative correlations between Average Sale Price and Yield, Water Use, Area and Scope One Emissions each indicated that Area Harvested and Fuel Use separately were not the determining factor for average sale price.
-The negative correlations are associative, not causal relationships (i.e using more water does not cause lower sale prices).
-\par
-\begin{table}[]
-  \caption{Pairwise Pearson correlation coefficients for logarithmically transformed values.}\label{tab:pearson_correlation}
-      \resizebox{\textwidth}{!}{
-      \begin{tabular}{@{}cccccccc@{}}
-          \toprule
-          \textbf{} & \multicolumn{1}{c}{\textbf{Yield}} & \multicolumn{1}{c}{\textbf{Area Harvested}} & \multicolumn{1}{c}{\textbf{Water Used}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Scope One\\ Emissions\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Yield by\\ Area\end{tabular}}} & \multicolumn{1}{c}{\textbf{Average Price}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Average Price\\ by Area\end{tabular}}} \\ \midrule
-          Yield & 1.00 & 0.88 & 0.82 & 0.76 & 0.96 & -0.46 & -0.88 \\
-          Area Harvested & 0.88 & 1.00 & 0.78 & 0.83 & 0.73 & -0.19 & -0.81 \\
-          Water Used & 0.82 & 0.78 & 1.00 & 0.67 & 0.76 & -0.49 & -0.82 \\
-          \begin{tabular}[c]{@{}c@{}}Scope One\\ Emissions\end{tabular} & 0.76 & 0.83 & 0.67 & 1.00 & 0.65 & -0.16 & -0.67 \\
-          \begin{tabular}[c]{@{}c@{}}Yield by\\ Area\end{tabular} & 0.96 & 0.73 & 0.76 & 0.65 & 1.00 & -0.54 & -0.84 \\
-          Average Price & -0.46 & -0.19 & -0.49 & -0.16 & -0.54 & 1.00 & 0.72 \\
-          \begin{tabular}[c]{@{}c@{}}Average Price\\ by Area\end{tabular} & -0.88 & -0.81 & -0.82 & -0.67 & -0.84 & 0.72 & 1.00 \\ \bottomrule
-          \end{tabular}}
-\end{table}
-
-\subsection{General Linear Models}
-
-Each model had a high $R^2$ value, indicating that most of the variance within the data was described by the models (see Table~\ref{tab:modelsummary}) and statistical significance of F-tests (p $<$ 2.200E-16).
-Aside from 3 variables, all regression parameters were also statistically significant from zero, F-tests across each model's variables were (p $<$ 0.05).
-The three exceptions were: scope one emissions in Model 3 (p=0.22) and Model 4 (p=0.0.39), and the interaction between area harvested and water used in model 2 (p=0.22).
-Note that scope one emissions was included in all models to directly compare the response variables as ratios of vineyard size to raw values and because it was strongly correlated to the response variable in every model (except model 5) especially for Models 1 and 4 (Table \ref{tab:pearson_correlation}).
-%A full list of regression coefficients 95\% CIs and p-values for each of the four models is provided in the appendix. % You gotta add these to the appendix in a more readable fashion
-\par
-\begin{table}[]
-  \caption{Summary of models; their performance, F-statistics and Residual error.}\label{tab:modelsummary}
-      \resizebox{\textwidth}{!}{
-        \begin{tabular}{@{}cccccccc@{}}
-          \toprule
-           & $\mathbf{R^2}$ & \textbf{\begin{tabular}[c]{@{}c@{}} Adjusted\\ $\mathbf{R^2}$ \end{tabular}} & \textbf{F-Statistic} & \textbf{P-Value} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual\\ Standard Error\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Sum\\ of Squares\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Mean\\ of Squares\end{tabular}} \\ \midrule
-          \textbf{\begin{tabular}[c]{@{}c@{}}Model 1\\ \end{tabular}} & 0.9072 & 0.9061 & 775.3 & 2.200e-16 & 0.3065 & 491.3 & 0.1 \\
-          \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-          \textbf{\begin{tabular}[c]{@{}c@{}}Model 2\\ \end{tabular}} & 0.8291 & 0.8141 & 55.07 & 2.200e-16 & 0.4312
-           & 905.03 & 0.19 \\
-          \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-          \textbf{\begin{tabular}[c]{@{}c@{}}Model 3\\ \end{tabular}} &0.9753 &0.9748 & 1885 & 2.200e-16 & 0.1589 & 71.11 & 0.03 \\
-          \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-          \textbf{\begin{tabular}[c]{@{}c@{}}Model 4\\ \end{tabular}} & 0.9091 & 0.9006 & 106.1 & 2.200e-16 & 0.3153  & 261.41 & 0.10 \\
-          \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
-          \textbf{\begin{tabular}[c]{@{}c@{}}Model 5\\ \end{tabular}} & 0.9089 & 0.9004 & 107.2 & 2.200e-16 & 0.3155 & 262.04 & 0.10 \\ \bottomrule
-          \end{tabular}
-        }
-\end{table}
-
-Coefficients related to continuous variables are summarised in Table \ref{tab:contvar}.
-In Model 1 all coefficients except for the intercept significantly contributed to the model (p $<$ 0.05), and in Model 2 all coefficients were statistically significant (p $<$ 0.05).
-However, scope one emissions did not significantly contribute to models 3, 4 and 5.
-Model 4 only had statistically significant contributions from the intercept and water use.
-Although the coefficient for water use was statistically significant for each model, it did not have the highest value; instead, area harvested, being an order of magnitude greater, dominated the models.
-Model 5 achieved a similar $R^2$ to Model 4 without area harvested, having stronger influences from water use and scope one emissions.
-
-\begin{table}[]
-  \caption{Summary of each Model's coefficients for continuous variables}\label{tab:contvar}
-  \resizebox{\textwidth}{!}{
-    \begin{tabular}{@{}cccccccc@{}}
-      \begin{tabular}{@{}cccccccc@{}}
-        \toprule
-        \textbf{} & \textbf{} & \textbf{Intercept} & \textbf{Area Harvested} & \textbf{Water Used} & \textbf{\begin{tabular}[c]{@{}c@{}}Scope One\\ Emissions\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Area Harvested \\ ** \\ Scope One Emissions\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Area Harvested \\ ** \\ Water Used\end{tabular}} \\ \midrule
-        \multirow{2}{*}{Model 1} & Coefficient & -0.0332 & 0.7418 & 0.0866 & 0.0673 &  &  \\
-         & Std Error & 0.0196 & 0.0100 & 0.0089 & 0.0080 &  &  \\
-         &  &  &  &  &  &  &  \\
-         \multirow{2}{*}{Model 2} & Coefficient & 0.1696 & 0.5774 & 0.1079 & 0.0850 & -0.0497 & -0.0535 \\
-         & Std Error & 0.0591 & 0.0148 & 0.0131 & 0.0117 & 0.0081 & 0.0084 \\
-         &  &  &  &  &  &  &  \\
-         \multirow{2}{*}{Model 3} & Coefficient & 0.0181 & 0.9713 & -0.0231 & -0.0070 &  &  \\
-         & Std Error & 0.0130 & 0.0072 & 0.0069 & 0.0057 &  &  \\
-         &  &  &  &  &  &  &  \\
-         \multirow{2}{*}{Model 4} & Coefficient & 0.1450 & 0.0024 & -0.0466 & -0.0170 & 0.0115 & 0.0014 \\
-         & Std Error & 0.0528 & 0.0150 & 0.0143 & 0.0118 & 0.0079 & 0.0083 \\
-         &  &  &  &  &  &  &  \\
-         \multirow{2}{*}{Model 5} & Coefficient & 0.1517 &  & -0.0404 & -0.0171 &  &  \\
-         & Std Error & 0.0527 &  & 0.0113 & 0.0097 &  &  \\ \bottomrule
-        \end{tabular}
-
-      \end{tabular}      
-  }
-\end{table}
-
-The regression coefficients for the Year variable of each model is depicted in Figure \ref{fig:yearly}.
-For each model, the first year for a model's was used as the baseline: 2012 for Models 1 and 2, and 2015 for Models 3, 4 and 5.
-Adelaide Hills is used as the regional baseline with the interaction between year and region using the first year and the Adelaide Hills as the baseline.
-Region and year contributed, in some but not all cases, more than the other variables.
-However, the coefficients varied substantially over the years, and were not significantly different from zero in some years.
-Models 4 and 5 are very similar in Figure \ref{fig:yearly}, indicating that the exclusion of area does not greatly affect the contribution from yearly influence.
-Models 4 and 5 have the most prominent trends, showing an increase in yearly effects over time, with Model 3 also increasing from 2016 to 2018, but plateauing afterwards.
-Model 1 and 2 do not show a clear trend, but do drop during 2017 and 2018 after increasing in the first 3 years.
-\par
-\begin{figure}
-  \includegraphics{yearly_plots.pdf}
-  \caption{Model Coefficient values for Year, with standard error bars.}\label{fig:yearly}
-  \end{figure}
-
-% Search for where you mention quality and where you mention value as this does become misleading!!!
-%
-% This is partly because of the intercept
-%
-% It depends what else is in the model.
-%
-% This direct comparisons are 
-%
-% The correlated between input vars needs to be discussed earlier.
-%
-%difference between the continuous variables coefficients is the change from positive to negative values.
-This change occurs between the Models for Yield (Model 1 and 2) and the Models for value (Models 3 and 4); where all but the coefficient for area harvested had the opposite sign (see Table \ref{tab:contvar}).
-These models also differ in an order of magnitude when looking at resource use, with the coefficients for yield being smaller than those for value.
-%
-% The resource input is different by an order of magnitude between value and yield
-%
-% Area has the strongest nfluence
-%
-% The models as a ratio of area show opposite trends - this could be indicative that smaller vineyards tend twoards more valuable crops and larger to crops of greater yield. 
-%
-
-Regional differences by temperature and rainfall are summarised in Figure \ref{fig:spider}.
-The most notable difference is between vineyards within 'Hot' and 'Very Dry' regions (warm inland regions), where high average sale prices are historically low, and yield is high.
-Water Use changes dramatically between these regions as well, with water being a driving force for yield but not necessarily average sale price.
-The warmer and drier regions tend to also have larger vineyards.
-These regional differences are further shown in Figure \ref{fig:yield_vs_value_area}, as a ratio of vineyard; where there is a pronounced difference in 'Hot' and 'Very Dry' regions producing more per area with lower average sale prices per area than other regions.
-\par
-\begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{starplots.png}}
-  \caption{Radar plot of climatic profile's resource use, yield and average sale price.
-The left reflects vineyards in different climatic temperatures.
-The right reflects vineyards in different rainfall climates.}\label{fig:spider}
-\end{figure}
-
-Figure \ref{fig:yield_vs_value_area} further shows the emphasis that 'Hot' areas have on high yields, with low average sale price compared with other regions.
-Scaling average price and yield by area shows a strong negative trend, trading quantity for higher sales prices.
-\par
-\begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{yield_verse_value_by_area.pdf}}
-    \caption{Scatter plot of vineyard yield against the average sale price as ratios to area harvested.
-The axes are in standard deviations with points coloured by climate.}\label{fig:yield_vs_value_area}
-\end{figure}
-
-Table \ref{tab:kfold} shows the validation results of each of the models.
-The $R^2$ measures of fit show similar results to the initial models, with a slight decrease as expected, indicating that the models are robust and consistent.
-%
-%
-% At least summary statistics of the baseline model.
-%
-%
-\par
-\begin{table}[]
-  \label{tab:kfold}
-  \caption{Model validation using k-fold cross validation, for 10 folds repeated 100 times.}
-  % \resizebox{\textwidth}{!}{
-  \begin{tabular}{@{}cccc@{}}
-    \toprule
-    \textbf{} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Mean\\ Squared Error\end{tabular}} & \textbf{R2} & \textbf{\begin{tabular}[c]{@{}c@{}}Mean Average \\ Error\end{tabular}} \\ \midrule
-    \textbf{Model 1} & .309 & .905 & .2165 \\
-    \textbf{Model 2} & .457 & .7921 & .313 \\
-    \textbf{Model 3} & .165 & .972 & .101 \\
-    \textbf{Model 4} & .348 & .878 & .182 \\ 
-    \textbf{Model 5} & .348 & .878 & .183 \\ \bottomrule
-    \end{tabular}
-  % }
-  \end{table}
-
-\section{Discussion}
-
-There was an expected strong relationship between size and resource use, with the overall area of a vineyard and its access to resources greatly determining the upper limit of potential yield.
-However, size was also inversely related to the potential average sale price.
-Higher average sales prices were also related to high resource inputs per area instead of the overall expenditure of resources.
-Vineyard yields and sales price changed greatly by region and year.
-Even given regional and yearly changes, there was a strong connection between smaller vineyards and higher sales prices.
-This could have been due to more attention available when managing smaller properties.
-\par
-The lack of significance of scope one emissions and its contribution to models, given its F-statistics, could be indicative that other vineyard activities requiring fuel are not leading factors for a vineyards average sale price.
-The relationship between yield, value and area was not simply about efficiently producing the most grapes.
-It is possible that the relationship of scope one emissions between yield and sale price was closely tied to a vineyard's area due to requiring more fuel to address more issues over greater distances.
-It is difficult to discern the connection of scope one emissions directly, as fuel can be used for a broad category of activities.
-\par
-There are important considerations unique to winegrowing compared to other agricultural industries.
-The vertical integration of winegrowing within the wine industry ties winegrowers to secondary and tertiary industries, such as wine production, packaging, transport and sales.
-This results in unique issues and considerations for each vineyard, where on-the-ground decisions are influenced by other wine industry's choices, such as the use of sustainable practices in vineyards as a requirement for sale in overseas markets; notably these interactions can be further complicated by some winegrowers being completely integrated into a wine company, while others are not \citep{knightFirmResourcesDevelopment2019}.
-Incorporating decisions into the model could help describe the contributing factors to regional differences beyond resource consumption, motivating the call for more granular data and more sophisticated modelling.
-\par
-There are many on-the-ground decisions that influence both sales price and yield.
-The decision to prioritise average sale price over quantity, is governed by complex physical and social forces, for example international market demands, disease pressures and natural disasters \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}, with many of these occurrences being highlighted throughout the reports from Wine Australia over the past decade \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
-However, the changes in the coefficients (see Figure \ref{fig:yearly}) are not reflective of many known occurrences, such as the 2020 bush fires, which had higher values for coefficients than prior years; During the 2020 bush fires 40,000 tonnes of grapes were lost across 18 different wine regions due to bush fires and smoke taint.
-In comparison to countrywide pressures such as drought, this damage made up only 3\% of the total amount of grapes for that year; although acknowledged as a considerable loss on an individual basis, it was deemed to be only a minor national concern by Wine Australia when compared to other environmental pressures such as drought \citep{wineaustraliaNationalVintageReport2020}
-\par
-Climatic pressures are an important consideration for growers, especially those in warmer and drier regions.
-The Wine Australia reports also show that warm inland regions have seen a decline in profit over the past decade, whereas regions with lower average sales prices did not \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
-Vineyards in warm inland regions also tend to be larger, making up for lower sale prices with larger yields.
-Considering the negative correlation of average price to area, for this strategy to work, economies of scale become an important factor.
-Given the large quantities of grapes that can be produced by some vineyards, even at low margins there is the potential to be profitable.
-However, the increasing climatic pressures mixed with the requirement for larger volumes of water, make the sustainability of some vineyards come into question.
-Furthermore, intensive farming in general is known to jeopardise the sustainability of an operation through the degradation of soil and waterways \citep{capelloEffectsTractorPasses2019,linHydropedologySynergisticIntegration2012,pisciottaGroundwaterNitrateRisk2015}.
-There are established methods that can help to mitigate these effects, such as the use of cover crops, midrow crop rotation and efficient irrigation.
-\par
-Some regions appeared to produce grapes of lower average sale price at scale whilst others achieved higher priced grapes in lower volumes.
-This empirical finding is consistent with Wine Australia's annual reports, which shows that some GI regions are known for producing large amounts of lower grade (low value per tonne) grapes \citep{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
-Comparatively, other regions only produce grapes of higher sales price but in smaller quantities.
-The difference in pricing per tonne between the lowest and highest regional average sales prices was almost a hundred times, showing that region had a profound influence.
-Some regions also had a mixture of high and low average sales price and yield showing regional variability in pricing which may be explained by varieties produced.
-A further possibility is the existence of regional upper limits on potential sales price, or that there are diminishing returns in some regions when pursuing higher sales prices or quantity; however these types of relationships may be obfuscated by knowledgeable winegrowers who avoid such pitfalls.
-\par
-Due to regional differences, different strategies are  employed, such as some regions targeting mass production over higher sales price.
-This is most notable when grouping regions by climate, especially when considering GI Regions in the 'Hot Very Dry' climate (see Figure \ref{fig:yield_vs_value_area}).
-Figure \ref{fig:spider} also shows that comparatively 'Warm' and 'Dry' Regions manage their resources incredibly efficiently having generally larger areas using similar resources to those in 'Cool' regions but having areas comparable to regions in 'Hot' climates.
-The coefficients for Model 4 also show a greater benefit of resource use per area when producing grapes at higher average sale prices, showing higher resource use per area reflective of higher average sales price (see Table \ref{tab:summary}).
-%In the preliminary analyses alternative models found that without the direct incorporation of GI Region or year, predictions greatly under performed.
-The effect of climate in the models was never as significant as the more granular GI regions, and always led to less accurate models.
-Although not chosen over GI region, climate was considered to be a large determinant of the ability to produce larger quantities of grapes, as well as a determinant in grape sale price \citep{agostaRegionalClimateVariability2012}.
-The more granular GI Region likely explained a broader mix of geographical phenomenon, such as soil, geology and access to water resources \citep{abbalDecisionSupportSystem2016,carmonaUseParticipatoryObjectOriented2011}.
-The interaction between year and GI Region likely accounted for events such as bushfires, which would be impactful, but only at a local level, both in time and space.
-\par
-We identified two main limitations to our linear modelling.
-First, model 1 and 2 over-predicting yield may have been due to preventative measures brought on by regional pressures such as fire, frost and disease.
-More fuel and water was likely used to prevent these issues from spreading within a region, thus disproportionately affecting some vineyards compared to others locally.
-This type of maintenance is not well captured in the models, especially when considering that some regions, especially those in warmer areas, are not as prone to disease as cooler climates and could potentially have lower fuel and water use per hectare.
-This could create a discrepancy in vineyards that utilised preventative measures in wetter regions, as opposed to those that did not, thus expending less fuel and energy but risking disease.
-When reviewing the differences between regions, it is important to consider that vineyards in 'Hot Very Dry' areas can be hundreds of times the size of those in other regions.
-This limitation could be overcome by incorporating the profitability of vineyards, comparing the financial success of working at different operational scales.
-\par
-The second limitations was the lack of further explanatory variables to help link models to causal affects.
-Variables such as the utilisation of renewable energy, contractors, and the occurrence of disease, fire and frost were originally explored to capture the discrepancies between similar vineyards that produced different yields and crop values.
-However, none of these variables was significantly correlated with the response variables, and did not add to model accuracy, even when considered as interactions.
-Allowance for nonlinear relationships, specifically through splines, resulted in more normally distributed residuals but at a drastically reduced overall accuracy when comparing $R^2$ and Residual Square Error.
-Attempts to fully explain small variations was always overshadowed by the dramatic differences in regional trends.
-Having more data for each region would also be beneficial, allowing greater comparison between regions.
-\par
-The use of other models such as random forests and decision trees alongside more variables and data may help to uncover the reasons for under or overestimation.
-These differences could be caused by the use of alternative sustainable practices in the field.
-Moreover, while there is evidence to suggest that environmentally sustainable practices can reduce costs, and increase efficiency whilst improving the quality of grapes; more research is needed to link these benefits across different regions and climates \citep{baianoOverviewSustainabilityWine2021,marianiSustainableWinegrowingCurrent2015,montalvo-falconSustainabilityResearchWine2023}.
-\section{Conclusion}
-
-This study delved into the relationships between resource use, grape sales price and yield.
-The findings underscore the multifaceted nature of vineyard management, where the interplay of size, resource allocation, climate, and regional influences collectively shape both the expected sale price and the quantity of grape yields.
-Average sales price of grapes was not solely tied to the overall expenditure of resources, but rather to the efficient allocation of resources per area.
-This emphasises that factors beyond sheer scale contribute significantly to the final sale price of grapes produced.
-Moreover, regional and yearly variations exhibited substantial effects on vineyard outputs, impacting both sales price and quantity.
-The connection observed between smaller vineyards and higher grape sale price suggests that the management of smaller properties might be more efficient and profitable.
-
-\bibliography{references}
-\bibliographystyle{elsarticle-harv}
-
-%% The Appendices part is started with the command \appendix;
-%% appendix sections are then done as normal sections
- \appendix
-
-\section{Appendix}
-\par
-
-\begin{table}[]
-      \caption{P-values for the non-transformed water used variable's Pearson correlation coefficients.}
-      \begin{tabular}{cl}
-      Variable                                           & Water Used \\
-      Yield                                              & 7.538E-01  \\
-      Area                                               & 6.981E-01  \\
-      Scope One Emissions                                & 8.883E-01  \\
-      $\textrm{Yield}\over\textrm{Area}$                     & 6.836E-01  \\
-      Average Price Per Tonne                            & 5.600E-02  \\
-      ${\textrm{Average Price per tonne}\over\textrm{Area}}$ & 1.522E-01 
-      \end{tabular}
-\end{table}
-
-\begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{model1_qq.pdf}}
-  \caption{QQ-plot of Model 1.}
-\end{figure}
-
-\begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{model1_rvf.pdf}}
-  \caption{Residuals vs fitted values for Model 1.}
-\end{figure}
-
-\begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{model2_qq.pdf}}
-  \caption{QQ-plot of Model 2.}
-\end{figure}
-
-\begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{model2_rvf.pdf}}
-  \caption{Residuals vs fitted values for Model 2.}
-\end{figure}
-
-\begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{model3_qq.pdf}}
-  \caption{QQ-plot of Model 3.}
-\end{figure}
-
-\begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{model3_rvf.pdf}}
-  \caption{Residuals vs fitted values for Model 3.}
-\end{figure}
-
-\begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{model4_qq.pdf}}
-  \caption{QQ-plot of Model 4.}
-\end{figure}
-
-\begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{model4_rvf.pdf}}
-  \caption{Residuals vs fitted values for Model 4.}
-\end{figure}
-\begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{model5_qq.pdf}}
-  \caption{QQ-plot of Model 5.}
-\end{figure}
-
-\begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{model5_rvf.pdf}}
-  \caption{Residuals vs fitted values for Model 5.}
-\end{figure}
-
-% \subsection{Complete list of coefficients for Model 1.}
-% \csvautolongtable[]{model1_coef.csv}
-
-%   \subsection{Complete list of coefficients for Model 2.}
-% \csvautolongtable[]{model2_coef.csv}
-
-%   \subsection{Complete list of coefficients for Model 3.}
-% \csvautolongtable[]{model3_coef.csv}
-
-%   \subsection{Complete list of coefficients for Model 4.}
-% \csvautolongtable[]{model4_coef.csv}
-
-%   \subsection{Complete list of coefficients for Model 5.}
-% \csvautolongtable[]{model5_coefs.csv}
-
-%
-%% \section{}
-%% \label{}
-
-%% If you have bibdatabase file and want bibtex to generate the
-%% bibitems, please use
-%%
-
-%% else use the following coding to input the bibitems directly in the
-%% TeX file.
-
-\end{linenumbers}
 \end{document}
-
-  \endinput

### 2024-05-15 article.tex
backing up article edits
some edits!!

@@
 \usepackage{longtable}
-\journal{Geography and Sustainability}
+\journal{ Journal of Agriculture and Food Research}
 %1.Full Length Article A full-length article should be a substantial and in-depth research study regarding a particular state of issue through several techniques or approaches. 
@@
 \author[label1,label2,label3]{Author}
-\date{02/08/2023}
+% \date{02/08/2023}
 \begin{abstract}
 % The word limit needs to be checked.
-When strategies for a sustainable winegrowing industry are assessed, there is a trade-off between balancing the amount of resources invested and the resultant yield and sale price of the produce.
+
+Strategies for achieving sustainability in the winegrowing industry requires balancing resource investment against the economic outcomes of resultant yields and sale price of the produce.
+Although there has been much research into the forecasting of outcomes and resource use, little has been done to illustrate the effects they have on one another and the consequential economic outcomes.
 In this analysis we observe relationships between resource use, yield and sale price through the use of statistical models.
@@
 \begin{highlights}
-  \item Comparative analysis of resource use, average sale price and quantity in Australian winegrowing.
-  \item Regional comparison of outcomes and resource use in Australian winegrowing regions.
-  \item Baseline models for comparing vineyards.
-  \item Analysis of national, decade long data source.
+\item Found that higher resource inputs per area are associated with higher sale prices, rather than overall resource expenditure.
+\item Identified significant differences in yield and sale prices across regions, highlighting the influence of localized conditions on vineyard profitability.
+\item Developed baseline models that demonstrate the critical relationship between vineyard area and both resource usage and economic outcomes.
+\item Analysis of long-term data offering a comprehensive view of industry trends highlighting resource use trends and economic outcomes.
 \end{highlights}
@@
 The climatic groups were used to illustrate similarities and differences occurring in areas larger than GI Regions.
+These classifications were literal descriptors used by industry members to discern between different weather types within winegrowing regions.
+Other climatic descriptors were explored, such as the  Köppen climate classification, hoiwever this did not offer a granular enough index to highlight any trends or clustering within the data.
 \par
@@
 \section{Results}
-\subsection{Exploratory Analysis}\label{sec:exp_anal}
+\subsection{Summary Statistics}\label{sec:exp_anal}
  Table \ref{tab:summary} shows the summary statistics of each variable in its original units.
@@
   \includegraphics{yearly_plots.pdf}
-  \caption{Model Coefficient values for Year, with standard error bars.}\label{fig:yearly}
+  \caption{Model Coefficient values for Year, with standard error bars.
+The coefficiens represent the influence that each year had on the model outcomes (see Table 1 for model outcomes and their predictors).
+Each years influence is comparative to the first year in the models, 2012.}t\label{fig:yearly}
   \end{figure}
@@
   \caption{Radar plot of climatic profile's resource use, yield and average sale price.
+Each point shows the overal contribution of a variable to the vineyards outcomes for that climate.
 The left reflects vineyards in different climatic temperatures.

### 2024-05-20 Cover Letter/main.tex
fixed cover letter for plos one instead of climate

@@
 \begin{tabular}{l @{}}
-	$26^{th}$ January, 2023\\ % Date
+	$17^{th}$ May, 2024\\ % Date
 	\textbf{Bryce Ronald William Boyd $|$ PhD Candidate}\\
@@
 \begin{tabular}{@{} l}
-	\textbf{Professor Bojie Fu,}\\
-	Editor-in-Chief \\
-	Geography and Sustainability\\
+	\textbf{Editorial board,}\\
+	PLOS ONE\\
 \end{tabular}
@@
 
-Dear Professor Bojie Fu,
+Dear members of the editorial board,
 \\~\\
@@
 %----------------------------------------------------------------------------------------
-\centering \textbf{Submission of Manuscript for Consideration by \\Geography and Sustainability}\\
+\centering \textbf{Submission of Manuscript for Consideration by \\PLOS ONE}\\
 \bigskip % Vertical whitespace
@@
 \newline \noindent
-We wish to submit a new manuscript entitled "\textit{The influence of resource use on yield versus sale price trade-off in Australian vineyards}" for consideration by Geography and Sustainability.
+We wish to submit a new manuscript entitled "\textit{The influence of resource use on yield versus sale price trade-off in Australian vineyards}" for consideration by PLOS ONE.
 \newline \noindent
@@
 We explore the economic and environmental risks induced by producing grapes at greater yields.
-We find that lower grape sale price is not solely related to the overall expenditure of resources but rather to efficient management, as well as regional and seasonal variations.
+We find that lower grape sale price is not solely related to the overall expenditure of resources but rather to efficient management, as well as regional and climatic variations.
 We use novel, extensive nationwide data from Australia, collected over a decade by Sustainable Winegrowing Australia.
@@
 \newline \noindent
-We believe that this manuscript is appropriate for publication by Geography and Sustainability as it demonstrates the sustainability of different winegrowing production strategies across winegrowing regions in Australia.
+We believe that this manuscript is appropriate for publication by PLOS ONE as it demonstrates the sustainability of different winegrowing production strategies across Australian winegrowing regions.
 These results illustrate the relationship between human interaction and the environment as driven by agricultural management strategies and the resources required to achieve the necessary outcomes to maintain economic sustainability using environmental resources.
@@
 \newline \noindent
-Bryce Boyd (corresponding author) on behalf of all coauthors: Kate Helmstedt, Mardi Longbottom, Madeline Mitchell, Kerrie Mengersen\\
+Bryce Boyd (corresponding author) on behalf of all coauthors: Kate Helmstedt, Mardi Longbottom, Kerrie Mengersen\\
 \newline \noindent

### 2024-05-20 article.tex
fixed title (what were you even doing??)

@@
-%% Copyright 2007-2020 Elsevier Ltd
-%% 
-%% This file is part of the 'Elsarticle Bundle'.
-%% ---------------------------------------------
-%% 
-%% It may be distributed under the conditions of the LaTeX Project Public
-%% License, either version 1.2 of this license or (at your option) any
-%% later version.  The latest version of this license is in
-%%    http://www.latex-project.org/lppl.txt
-%% and version 1.2 or later is part of all distributions of LaTeX
-%% version 1999/12/01 or later.
-%% 
-%% The list of all files belonging to the 'Elsarticle Bundle' is
-%% given in the file `manifest.txt'.
-%% 
-%% Template article for Elsevier's document class `elsarticle'
-%% with harvard style bibliographic references
+% Template for PLoS
+% Version 3.6 Aug 2022
+%
+% % % % % % % % % % % % % % % % % % % % % %
+%
+% -- IMPORTANT NOTE
+%
+% This template contains comments intended 
+% to minimize problems and delays during our production 
+% process. Please follow the template instructions
+% whenever possible.
+%
+% % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Once your paper is accepted for publication, 
+% PLEASE REMOVE ALL TRACKED CHANGES in this file 
+% and leave only the final text of your manuscript. 
+% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
+% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
+%
+%
+% There are no restrictions on package use within the LaTeX files except that no packages listed in the template may be deleted.
+%
+% Please do not include colors or graphics in the text.
+%
+% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
+%
+% % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- FIGURES AND TABLES
+%
+% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
+%
+% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
+% - Figures should be uploaded separately from your manuscript file. 
+% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
+% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
+% For figure citations, please use "Fig" instead of "Figure".
+% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
+%
+% Tables should be cell-based and may not contain:
+% - spacing/line breaks within cells to alter layout or alignment
+% - do not nest tabular environments (no tabular environments within tabular environments)
+% - no graphics or colored text (cell background color/shading OK)
+% See http://journals.plos.org/plosone/s/tables for table guidelines.
+%
+% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
+%
+% IMPORTANT
+% Below are a few tips to help format your equations and other special characters according to our specifications.
+For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
+%
+% For inline equations, please be sure to include all portions of an equation in the math environment.
+For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
+%
+% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
+%
+% Please add line breaks to long display equations when possible in order to fit size of the column. 
+%
+% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
+%
+% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
+%
+% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
+%
+% % % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Please contact latex@plos.org with any questions.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
 
-\documentclass[review,12pt,authoryear]{elsarticle}
+\documentclass[10pt,letterpaper]{article}
+\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
 
-%% Use the option review to obtain double line spacing
-%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
+% amsmath and amssymb packages, useful for mathematical formulas and symbols
+\usepackage{amsmath,amssymb}
 
-%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
-%% for a journal layout:
-%% \documentclass[final,1p,times,authoryear]{elsarticle}
-%% \documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
-%% \documentclass[final,3p,times,authoryear]{elsarticle}
-%% \documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
-%% \documentclass[final,5p,times,authoryear]{elsarticle}
-%% \documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}
+% Use adjustwidth environment to exceed column width (see example table in text)
+\usepackage{changepage}
 
-%% For including figures, graphicx.sty has been loaded in
-%% elsarticle.cls. If you prefer to use the old commands
-%% please give \usepackage{epsfig}
+% textcomp package and marvosym package for additional characters
+\usepackage{textcomp,marvosym}
 
-%% The amssymb package provides various useful mathematical symbols
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
+
+% Use nameref to cite supporting information files (see Supporting Information section for more info)
+\usepackage{nameref,hyperref}
+
+% line numbers
+\usepackage[right]{lineno}
+
+% ligatures disabled
+\usepackage[nopatch=eqnum]{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
+
+% color can be used to apply background shading to table cells only
+\usepackage[table]{xcolor}
+
+% array package and thick rules for tables
+
+% create "+" rule type for thick vertical lines
+\newcolumntype{+}{!{\vrule width 2pt}}
+
+% create \thickcline for thick horizontal lines of variable length
+\newlength\savedwidth
+\newcommand\thickcline[1]{%
+  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+  \cline{#1}%
+  \noalign{\vskip\arrayrulewidth}%
+  \noalign{\global\arrayrulewidth\savedwidth}%
+}
+
+% \thickhline command for thick horizontal lines that span the table
+\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+\hline
+\noalign{\global\arrayrulewidth\savedwidth}}
+
+% Remove comment for double spacing
+%\usepackage{setspace} 
+%\doublespacing
+
+% Text layout
+\raggedright
+\setlength{\parindent}{0.5cm}
+\textwidth 5.25in 
+\textheight 8.75in
+
+% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
+% Captions will be left justified
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
+\renewcommand{\figurename}{Fig}
+
+% Use the PLoS provided BiBTeX style
+\bibliographystyle{plos2015}
+
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
+
+% Header and Footer with logo
+\usepackage{lastpage,fancyhdr,graphicx}
+\usepackage{epstopdf}
+%\pagestyle{myheadings}
+\pagestyle{fancy}
+\fancyhf{}
+%\setlength{\headheight}{27.023pt}
+%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+\rfoot{\thepage/\pageref{LastPage}}
+\renewcommand{\headrulewidth}{0pt}
+\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+\fancyheadoffset[L]{2.25in}
+\fancyfootoffset[L]{2.25in}
+\lfoot{\today}
+
+%% Include all macros below
+
+% \newcommand{\lorem}{{\bf LOREM}}
+% \newcommand{\ipsum}{{\bf IPSUM}}
+
+%% END MACROS SECTION
+
+% Original draft packages
+\usepackage{array}
 \usepackage{amssymb}
-%% The amsthm package provides extended theorem environments
-%% \usepackage{amsthm}
-
-%% The lineno packages adds line numbers. Start line numbering with
-%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
-%% for the whole article with \linenumbers.
-\usepackage{lineno}
-
-% for adjusting table width automatically
 \usepackage{adjustbox}
@@
 \usepackage{multirow}
-
 \usepackage{csvsimple}
 \usepackage{longtable}
-\journal{ Journal of Agriculture and Food Research}
-%1.Full Length Article A full-length article should be a substantial and in-depth research study regarding a particular state of issue through several techniques or approaches. 
-% The main text should be approximately 6,000 words in length, but it should not exceed 8,000 words (excluding abstract, references, tables, figures, and appendices).
-%A maximum of 250 words abstract and up to 10 displayed items (figures and tables) is allowed. A full-length article should include an Introduction,
-%Materials and methods, Results, Discussion, Conclusions, and References, which can be accompanied by Supplementary material.
 
 \begin{document}
-\begin{linenumbers}
-\begin{frontmatter}
-%% Title, authors and addresses
-%% use the tnoteref command within \title for footnotes;
-%% use the tnotetext command for theassociated footnote;
-%% use the fnref command within \author or \affiliation for footnotes;
-%% use the fntext command for theassociated footnote;
-%% use the corref command within \author for corresponding author footnotes;
-%% use the cortext command for theassociated footnote;
-%% use the ead command for the email address,
-%% and the form \ead[url] for the home page:
-%% \title{Title\tnoteref{label1}}
-%% \tnotetext[label1]{}
-%% \author{Name\corref{cor1}\fnref{label2}}
-%% \ead{email address}
-%% \ead[url]{home page}
-%% \fntext[label2]{}
-%% \cortext[cor1]{}
-%% \affiliation{organization={},
-%%            addressline={}, 
-%%            city={},
-%%            postcode={}, 
-%%            state={},
-%%            country={}}
-%% \fntext[label3]{}
-\title{The influence of resource use on yield versus sale price trade-off in Australian vineyards}
-%% use optional labels to link authors explicitly to addresses:
-%% \author[label1,label2]{}
-%% \affiliation[label1]{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
-%%
-%% \affiliation[label2]{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
-%\affiliation[label1]{organization={QUT},
-%  addressline={},
-%  city={},
-%  postcode={},
-%  state={QLD},
-%  country={}}
-%\affiliation[label2]{organization={AWRI},
-%  addressline={},
-%  city={},
-%  postcode={},
-%  state={SA},
-%  country={}}
-%\affiliation[label3]{organization={Food Agility CRC},
-%  addressline={},
-%  city={},
-%  postcode={},
-%  state={Vic},
-%  country={}}
-\author[label1,label2,label3]{Author}
-% \date{02/08/2023}
-\begin{abstract}
+
+\vspace*{0.2in}
+
+% Title must be 250 characters or less.
+\begin{flushleft}
+{\Large
+\textbf\newline{The influence of resource use on yield versus sale price trade-off in Australian vineyards}
+% \title{The influence of resource use on yield versus sale price trade-off in Australian vineyards}
+% Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+}
+\newline
+% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
+\\
+Bryce Boyd\textsuperscript{1,2,3,*},
+Kate Helmstedt\textsuperscript{1},
+Mardi Longbottom\textsuperscript{2},
+Kerrie Mengersen\textsuperscript{1}
+% {\textpilcrow}
+\\
+\bigskip
+\textbf{1} Affiliation Faculty of Mathematical Sciences, Queensland University of Technology, Brisbane, QLD, Australia
+\\
+\textbf{2} Affiliation Australian Wine Research Institute, Adelaide, SA, Australia
+\\
+\textbf{3} Affiliation Food Agility CRC, Sydney, NSW, Australia
+\\
+\bigskip
+
+% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
+% 
+% Remove or comment out the author notes below if they aren't used.
+%
+
+% % Primary Equal Contribution Note
+% \Yinyang These authors contributed equally to this work.
+
+% % Additional Equal Contribution Note
+% % Also use this double-dagger symbol for special authorship notes, such as senior authorship.
+% \ddag These authors also contributed equally to this work.
+
+% % Current address notes
+% \textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
+% % \textcurrency b Insert second current address 
+% % \textcurrency c Insert third current address
+
+% % Deceased author note
+% \dag Deceased
+
+% % Group/Consortium Author Note
+% \textpilcrow Membership list can be found in the Acknowledgments section.
+
+% % Use the asterisk to denote corresponding authorship and provide email address in note below.
+* Bryce.polley@awri.com.au
+
+\end{flushleft}
+
+\section*{Abstract}
 % The word limit needs to be checked.
@@
 Overall, the analysis highlighted the importance of considering a vineyard's business goal, region, external pressures and economies of scale, when considering whether to pursue higher yields or higher average sales prices.
-\end{abstract}
-%%Graphical abstract`
-%\begin{graphicalabstract}
- % \includegraphics{graphical_abstract.jpeg}
-%\end{graphicalabstract}'
 
-%\begin{keyword}
-%% keywords here, in the form: keyword \sep keyword
-%Keyword one \sep{} keyword two
-%% PACS codes here, in the form: \PACS code \sep code
-%\PACS{} 0000 \sep{} 1111
-%% MSC codes here, in the form: \MSC code \sep code
-%% or \MSC[2008] code \sep code (2000 is the default)
-%\MSC{} 0000 \sep{} 1111
-%\end{keyword}
-%%Research highlights
-\begin{highlights}
-\item Found that higher resource inputs per area are associated with higher sale prices, rather than overall resource expenditure.
-\item Identified significant differences in yield and sale prices across regions, highlighting the influence of localized conditions on vineyard profitability.
-\item Developed baseline models that demonstrate the critical relationship between vineyard area and both resource usage and economic outcomes.
-\item Analysis of long-term data offering a comprehensive view of industry trends highlighting resource use trends and economic outcomes.
-\end{highlights}
-\end{frontmatter}
+% \begin{highlights}
+% \item Found that higher resource inputs per area are associated with higher sale prices, rather than overall resource expenditure.
+% \item Identified significant differences in yield and sale prices across regions, highlighting the influence of localized conditions on vineyard profitability.
+% \item Developed baseline models that demonstrate the critical relationship between vineyard area and both resource usage and economic outcomes.
+% \item Analysis of long-term data offering a comprehensive view of industry trends highlighting resource use trends and economic outcomes.
+% \end{highlights}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\linenumbers
 
-\section{Introduction}
-The global focus on sustainability in agronomic industries has changed the way in which these enterprises do business.
-A dilemma exists across agriculture through shared fundamental considerations of resource use such as water and fuel, and the resultant yield and crop value that are produced \citep{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016, zhuEffectsNitrogenLevel2017}.
+\section*{Introduction}
+
+The focus on sustainability in agronomic industries has changed the way in which these enterprises do business.
+The diverse nature of vineyard operations results in complicated interactions with both environmental and economic outcomes.
+The interactions between different factors can put an environmentally advantageous decision at odds with an economic one, and conversely economic decisions can be detrimental to the environment \cite{beaumelleBiodiversityConservationEcosystem2023}.
+As a result many studies are including more diverse ranges of data and scales.
+However, studies tend to focus on specific aspects, such as reducing chemical use or improving soil health, often focusing on singular relationships or variable outcomes \cite{barriguinhaVineyardYieldEstimation2021,heFruitYieldPrediction2022,laurentReviewIssuesMethods2021}.
+Other analyses, look at life cycles or cost benefit analyses, reflecting a specific management decision's outcome \cite{abbottAWRIAssessingEnvironmental2016,beauchetInterannualVariabilityEnvironmental2019,ferraraLifeCycleAssessment2018} or their cost benefits by utilising different scenarios to assess specific practices \cite{fincoCombiningPrecisionViticulture2022a,singhBibliometricReviewUse2022}.
+It has been noted that some studies have relied too heavily on single points of engagement when relaying the idea of sustainability and vineyard outcomes, especially when the inclusion of more variables or data could have better described a situation as a whole \cite{baianoOverviewSustainabilityWine2021,santiago-brownSustainabilityAssessmentWineGrape2015}.
+The changing nature of sustainability also calls for measures to be more than simple snapshots of data \cite{baianoOverviewSustainabilityWine2021,montalvo-falconSustainabilityResearchWine2023,santiago-brownSustainabilityAssessmentWineGrape2015}.
+Given recent studies there is still growing interest from the winegrowing industry in further studies, with several recent reviews calling for more research to demonstrate the potential trade-offs between different operational approaches across different climates and regions \cite{baianoOverviewSustainabilityWine2021,marianiSustainableWinegrowingCurrent2015,montalvo-falconSustainabilityResearchWine2023}.
+
+A dilemma exists across agriculture through shared fundamental considerations of resource use such as water and fuel, and the resultant yield and crop value that are produced \cite{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016, zhuEffectsNitrogenLevel2017}.
 The average price of grapes for wine production is driven through its integration within the wine industry.
-An important connection between grapes and their price is the grape's perceived quality, which initially defines a wine's potential through the grape's chemical makeup \citep{blackTerpenoidsTheirRole2015,schreierFlavorCompositionWines1979}.
-Grapes of higher perceived quality or grapes from well known regions are likely to have higher prices \citep{wineaustraliaNationalVintageReport2021}.
-Grape quality is connected to the market value of winegrapes, with Wine Australia explicitly defining grape quality through the use of discrete price brackets in their annual reports \citep{winemakersfederationofaustraliaNationalVintageReport2018}.
-It is important to note that the generalisation made to reflect quality through using average price assumes a due diligence of those who purchase the grapes \citep{yeggeInfluenceSensoryNonsensory2001}.
+An important connection between grapes and their price is the grape's perceived quality, which initially defines a wine's potential through the grape's chemical makeup \cite{blackTerpenoidsTheirRole2015,schreierFlavorCompositionWines1979}.
+Grapes of higher perceived quality or grapes from well known regions are likely to have higher prices \cite{wineaustraliaNationalVintageReport2021}.
+Grape quality is connected to the market value of winegrapes, with Wine Australia explicitly defining grape quality through the use of discrete price brackets in their annual reports \cite{winemakersfederationofaustraliaNationalVintageReport2018}.
+It is important to note that the generalisation made to reflect quality through using average price assumes a due diligence of those who purchase the grapes \cite{yeggeInfluenceSensoryNonsensory2001}.
 The economic sustainability of a vineyard is tied to this market culture, driven by the wine industry.
-The consideration of sustainability within viticulture is also subject to environmental and sociodemographic pressures \citep{santiago-brownSustainabilityAssessmentWineGrape2015}.
-In the Australian context, these pressures include biosecurity, climate and international market changes \citep{canadellMultidecadalIncreaseForest2021,longbottomRoleVineyardPractices2015,oliverReviewSoilPhysical2013}.
+The consideration of sustainability within viticulture is also subject to environmental and sociodemographic pressures \cite{santiago-brownSustainabilityAssessmentWineGrape2015}.
+In the Australian context, these pressures include biosecurity, climate and international market changes \cite{canadellMultidecadalIncreaseForest2021,longbottomRoleVineyardPractices2015,oliverReviewSoilPhysical2013}.
 \par
-There is an extensive amount of research into the varied effects of factors on grape quality and yield \citep{heFruitYieldPrediction2022,laurentLocalInfluenceClimate2022,liakosMachineLearningAgriculture2018}.
+There is an extensive amount of research into the varied effects of factors on grape quality and yield \cite{heFruitYieldPrediction2022,laurentLocalInfluenceClimate2022,liakosMachineLearningAgriculture2018}.
 Due to the lack of long-term and in-depth data there is a lack of research on grape sale price and its driving factors.
-Furthermore, individual factors are often studied in isolation with yield and sales price not appearing together \citep{abbalDecisionSupportSystem2016}.
+Furthermore, individual factors are often studied in isolation with yield and sales price not appearing together \cite{abbalDecisionSupportSystem2016}.
 The lack of consolidated datasets restricts the ability to gain statistical insights at large scales and across multiple regions.
-As a result, broader studies are lacking \citep{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
+As a result, broader studies are lacking \cite{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
 The dataset used for this analysis includes data spanning 10 years from a multitude of vineyards located over a diverse range of Australian winegrowing regions.
@@
 This is the first time that such a trade-off has been confirmed explicitly across such varying regions, scales and climates in the Australian winegrowing industry.
-\section{Methods}
-\subsection{Data}
+\section*{Methods}
+\subsection*{Data}
 
@@
 Data used in these analyses were obtained from Sustainable Winegrowing Australia and Wine Australia.
-Sustainable Winegrowing Australia is Australia's national wine industry sustainability program, which aims to facilitate grape  growers and winemakers in demonstrating and improving their sustainability \citep{swaSustainableWingrowingAustralia2022}.
-Wine Australia is an Australian Government statutory authority governed by the Wine Australia Act \citep{attorney-generalsdepartmentWineAustraliaCorporation2010}.
+Sustainable Winegrowing Australia is Australia's national wine industry sustainability program, which aims to facilitate grape  growers and winemakers in demonstrating and improving their sustainability \cite{swaSustainableWingrowingAustralia2022}.
+Wine Australia is an Australian Government statutory authority governed by the Wine Australia Act \cite{attorney-generalsdepartmentWineAustraliaCorporation2010}.
 Data collected by Wine Australia are publicly available.
@@
 \par
-Data preprocessing was conducted prior to analysis using the Python programming language \citep{g.vanrossumPythonTutorialTechnical1995}.
+Data preprocessing was conducted prior to analysis using the Python programming language \cite{g.vanrossumPythonTutorialTechnical1995}.
 Preprocessing included the conversion of fuel to scope one emissions and prior calculations for all continuous variables which included logarithmic transformations, centring and scaling by standard deviation.
-We converted multiple emission sources into scope one emissions using the equation given from the Australian National Greenhouse Accounts Factors \citep{agdeeNationalGreenhouseAccounts2021}.
+We converted multiple emission sources into scope one emissions using the equation given from the Australian National Greenhouse Accounts Factors \cite{agdeeNationalGreenhouseAccounts2021}.
 The calculation conducted using
@@
 \end{equation}
-where emissions was the the product of the quantity of fuel in litres, $Q$, a prescribed Energy Content, $EC$, and emission factors (as given by the Australian National Greenhouse Accounts factors) of scope one, $EF1$, and scope three, $EF3$, to tonnes of Carbon Dioxide Emission equivalent, $tCO2e$ \citep{departmentofclimatechangeenergytheenvironmentandwaterAustralianNationalGreenhouse2022}.
+where emissions was the the product of the quantity of fuel in litres, $Q$, a prescribed Energy Content, $EC$, and emission factors (as given by the Australian National Greenhouse Accounts factors) of scope one, $EF1$, and scope three, $EF3$, to tonnes of Carbon Dioxide Emission equivalent, $tCO2e$ \cite{departmentofclimatechangeenergytheenvironmentandwaterAustralianNationalGreenhouse2022}.
 \par
@@
 \par
-Differences in vineyard locations were captured through the use of Geographical Indicator Regions defined by Wine Australia \citep{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
+Differences in vineyard locations were captured through the use of Geographical Indicator Regions defined by Wine Australia \cite{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
 Although vineyards generally differed in income, there were pronounced differences between regions as shown in Figure \ref{fig:map}.
-These differences were likely due to the site of a vineyard being important as it predetermines several physical parameters such as climate, geology and soil, making location a widely considered key determinant of grape yield and average sale price \citep{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
-Each GI Region has its own unique mixture of climatic and geophysical properties that describes a unique winegrowing region within Australia and is a protected trademark under the Wine Australia act \citep{attorney-generalsdepartmentWineAustraliaCorporation2010}.
+These differences were likely due to the site of a vineyard being important as it predetermines several physical parameters such as climate, geology and soil, making location a widely considered key determinant of grape yield and average sale price \cite{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
+Each GI Region has its own unique mixture of climatic and geophysical properties that describes a unique winegrowing region within Australia and is a protected trademark under the Wine Australia act \cite{attorney-generalsdepartmentWineAustraliaCorporation2010}.
 Both Wine Australia and Sustainable Winegrowing Australia used the same GI Region categorical variable format to describe location.
 \par
- The climatic properties of each GI Region were summarised by using predefined classifications as per the \citet{sustainablewinegrowingaustraliaSustainableWinegrowingAustralia2021} user manual.
+ The climatic properties of each GI Region were summarised by using predefined classifications as per the \cite{sustainablewinegrowingaustraliaSustainableWinegrowingAustralia2021} user manual.
 The user manual describes climates by rainfall and temperature, creating supersets of Regions of similar climatic properties.
@@
 \par
-\subsection{Analysis}
+\subsection*{Analysis}
 Pairwise Pearson Correlation Coefficients (PPCC) were calculated to assess the potential existence of linear relationships between the input and predicted variables.
@@
 Model fit was measured in terms of $R^2$ and adjusted $R^2$ as well as F statistics, and t-tests were used to determine if predictors significantly contributed to the models when accounting for other variables, showing which specific years and areas contributed significantly.
-Both the PPCC and GLM analyses were created using the R statistical programming language \citep{rcoreteamLanguageEnvironmentStatistical2021} with the Caret package \citep{kuhnBuildingPredictiveModels2008}.
+Both the PPCC and GLM analyses were created using the R statistical programming language \cite{rcoreteamLanguageEnvironmentStatistical2021} with the Caret package \cite{kuhnBuildingPredictiveModels2008}.
 \par
@@
 \par
-\subsection{Model Validation}
+\subsection*{Model Validation}
 Models were validated using K-fold cross validation.
@@
 \par
-\section{Results}
-\subsection{Summary Statistics}\label{sec:exp_anal}
+\section*{Results}
+\subsection*{Summary Statistics}\label{sec:exp_anal}
  Table \ref{tab:summary} shows the summary statistics of each variable in its original units.
@@
 
-\subsection{General Linear Models}
+\subsection*{General Linear Models}
 
@@
 
-\section{Discussion}
+\section*{Discussion}
 
@@
 The vertical integration of winegrowing within the wine industry ties winegrowers to secondary and tertiary industries, such as wine production, packaging, transport and sales.
-This results in unique issues and considerations for each vineyard, where on-the-ground decisions are influenced by other wine industry's choices, such as the use of sustainable practices in vineyards as a requirement for sale in overseas markets; notably these interactions can be further complicated by some winegrowers being completely integrated into a wine company, while others are not \citep{knightFirmResourcesDevelopment2019}.
+This results in unique issues and considerations for each vineyard, where on-the-ground decisions are influenced by other wine industry's choices, such as the use of sustainable practices in vineyards as a requirement for sale in overseas markets; notably these interactions can be further complicated by some winegrowers being completely integrated into a wine company, while others are not \cite{knightFirmResourcesDevelopment2019}.
 Incorporating decisions into the model could help describe the contributing factors to regional differences beyond resource consumption, motivating the call for more granular data and more sophisticated modelling.
@@
 There are many on-the-ground decisions that influence both sales price and yield.
-The decision to prioritise average sale price over quantity, is governed by complex physical and social forces, for example international market demands, disease pressures and natural disasters \citep{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}, with many of these occurrences being highlighted throughout the reports from Wine Australia over the past decade \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
+The decision to prioritise average sale price over quantity, is governed by complex physical and social forces, for example international market demands, disease pressures and natural disasters \cite{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}, with many of these occurrences being highlighted throughout the reports from Wine Australia over the past decade \cite{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
 However, the changes in the coefficients (see Figure \ref{fig:yearly}) are not reflective of many known occurrences, such as the 2020 bush fires, which had higher values for coefficients than prior years; During the 2020 bush fires 40,000 tonnes of grapes were lost across 18 different wine regions due to bush fires and smoke taint.
-In comparison to countrywide pressures such as drought, this damage made up only 3\% of the total amount of grapes for that year; although acknowledged as a considerable loss on an individual basis, it was deemed to be only a minor national concern by Wine Australia when compared to other environmental pressures such as drought \citep{wineaustraliaNationalVintageReport2020}
+In comparison to countrywide pressures such as drought, this damage made up only 3\% of the total amount of grapes for that year; although acknowledged as a considerable loss on an individual basis, it was deemed to be only a minor national concern by Wine Australia when compared to other environmental pressures such as drought \cite{wineaustraliaNationalVintageReport2020}
 \par
 Climatic pressures are an important consideration for growers, especially those in warmer and drier regions.
-The Wine Australia reports also show that warm inland regions have seen a decline in profit over the past decade, whereas regions with lower average sales prices did not \citep{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
+The Wine Australia reports also show that warm inland regions have seen a decline in profit over the past decade, whereas regions with lower average sales prices did not \cite{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
 Vineyards in warm inland regions also tend to be larger, making up for lower sale prices with larger yields.
@@
 However, the increasing climatic pressures mixed with the requirement for larger volumes of water, make the sustainability of some vineyards come into question.
-Furthermore, intensive farming in general is known to jeopardise the sustainability of an operation through the degradation of soil and waterways \citep{capelloEffectsTractorPasses2019,linHydropedologySynergisticIntegration2012,pisciottaGroundwaterNitrateRisk2015}.
+Furthermore, intensive farming in general is known to jeopardise the sustainability of an operation through the degradation of soil and waterways \cite{capelloEffectsTractorPasses2019,linHydropedologySynergisticIntegration2012,pisciottaGroundwaterNitrateRisk2015}.
 There are established methods that can help to mitigate these effects, such as the use of cover crops, midrow crop rotation and efficient irrigation.
@@
 Some regions appeared to produce grapes of lower average sale price at scale whilst others achieved higher priced grapes in lower volumes.
-This empirical finding is consistent with Wine Australia's annual reports, which shows that some GI regions are known for producing large amounts of lower grade (low value per tonne) grapes \citep{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
+This empirical finding is consistent with Wine Australia's annual reports, which shows that some GI regions are known for producing large amounts of lower grade (low value per tonne) grapes \cite{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
 Comparatively, other regions only produce grapes of higher sales price but in smaller quantities.
@@
 The effect of climate in the models was never as significant as the more granular GI regions, and always led to less accurate models.
-Although not chosen over GI region, climate was considered to be a large determinant of the ability to produce larger quantities of grapes, as well as a determinant in grape sale price \citep{agostaRegionalClimateVariability2012}.
-The more granular GI Region likely explained a broader mix of geographical phenomenon, such as soil, geology and access to water resources \citep{abbalDecisionSupportSystem2016,carmonaUseParticipatoryObjectOriented2011}.
+Although not chosen over GI region, climate was considered to be a large determinant of the ability to produce larger quantities of grapes, as well as a determinant in grape sale price \cite{agostaRegionalClimateVariability2012}.
+The more granular GI Region likely explained a broader mix of geographical phenomenon, such as soil, geology and access to water resources \cite{abbalDecisionSupportSystem2016,carmonaUseParticipatoryObjectOriented2011}.
 The interaction between year and GI Region likely accounted for events such as bushfires, which would be impactful, but only at a local level, both in time and space.
@@
 These differences could be caused by the use of alternative sustainable practices in the field.
-Moreover, while there is evidence to suggest that environmentally sustainable practices can reduce costs, and increase efficiency whilst improving the quality of grapes; more research is needed to link these benefits across different regions and climates \citep{baianoOverviewSustainabilityWine2021,marianiSustainableWinegrowingCurrent2015,montalvo-falconSustainabilityResearchWine2023}.
-\section{Conclusion}
+Moreover, while there is evidence to suggest that environmentally sustainable practices can reduce costs, and increase efficiency whilst improving the quality of grapes; more research is needed to link these benefits across different regions and climates \cite{baianoOverviewSustainabilityWine2021,marianiSustainableWinegrowingCurrent2015,montalvo-falconSustainabilityResearchWine2023}.
+\section*{Conclusion}
 
@@
 
+\nolinenumbers
+
+% \bibliographystyle{plos2015}
 \bibliography{references}
-\bibliographystyle{elsarticle-harv}
 
@@
 
-\section{Appendix}
+\section*{Appendix}
 \par
@@
 
-% \subsection{Complete list of coefficients for Model 1.}
+% \subsection*{Complete list of coefficients for Model 1.}
 % \csvautolongtable[]{model1_coef.csv}
 
-%   \subsection{Complete list of coefficients for Model 2.}
+%   \subsection*{Complete list of coefficients for Model 2.}
 % \csvautolongtable[]{model2_coef.csv}
 
-%   \subsection{Complete list of coefficients for Model 3.}
+%   \subsection*{Complete list of coefficients for Model 3.}
 % \csvautolongtable[]{model3_coef.csv}
 
-%   \subsection{Complete list of coefficients for Model 4.}
+%   \subsection*{Complete list of coefficients for Model 4.}
 % \csvautolongtable[]{model4_coef.csv}
 
-%   \subsection{Complete list of coefficients for Model 5.}
+%   \subsection*{Complete list of coefficients for Model 5.}
 % \csvautolongtable[]{model5_coefs.csv}
@@
 %
-%% \section{}
+%% \section*{}
 %% \label{}
@@
 
-\end{linenumbers}
 \end{document}
-
-  \endinput

### 2024-05-20 Cover Letter/main.tex
updated cover letter to include information about second paper submission

@@
 
-\includegraphics[scale=0.5]{logo.jpg} % Logo
+\includegraphics[scale=0.3]{logo.jpg} % Logo
 
@@
 \begin{tabular}{l @{}}
-	$17^{th}$ May, 2024\\ % Date
+	17 May, 2024\\ % Date
 	\textbf{Bryce Ronald William Boyd $|$ PhD Candidate}\\
@@
 
-\bigskip % Vertical whitespace
+\newline % Vertical whitespace
 
@@
 \begin{tabular}{@{} l}
-	\textbf{Editorial board,}\\
+	\textbf{Editorial Board,}\\
 	PLOS ONE\\
@@
 
-\bigskip % Vertical whitespace
-
-Dear members of the editorial board,
+\newline % Vertical whitespace
+Dear Members of the Editorial Board,
 \\~\\
-\parskip % Vertical whitespace
 
@@
 %----------------------------------------------------------------------------------------
-\centering \textbf{Submission of Manuscript for Consideration by \\PLOS ONE}\\
+\centering \textbf{Submission of Manuscript for Consideration by PLOS ONE}\\
 \bigskip % Vertical whitespace
 \justifying
+We wish to submit a new manuscript entitled "\textit{The Influence of Resource Use on Yield versus Sale Price Trade-Off in Australian Vineyards}" for consideration by PLOS ONE.
 \newline \noindent
-We wish to submit a new manuscript entitled "\textit{The influence of resource use on yield versus sale price trade-off in Australian vineyards}" for consideration by PLOS ONE.
-\newline \noindent
-In this manuscript we explore natural resource use in agricultural wine production, emphasising sustainability in the trade-off between the sale price and quantity of product.
-We explore the economic and environmental risks induced by producing grapes at greater yields.
+In this manuscript, we explore natural resource use in agricultural wine production, emphasizing sustainability in the trade-off between the sale price and quantity of product.
+We examine the economic and environmental risks induced by producing grapes at greater yields.
 We find that lower grape sale price is not solely related to the overall expenditure of resources but rather to efficient management, as well as regional and climatic variations.
 We use novel, extensive nationwide data from Australia, collected over a decade by Sustainable Winegrowing Australia.
-We construct and analyse several models of the multifaceted, multi-resource nature of vineyard management.
+We construct and analyze several models of the multifaceted, multi-resource nature of vineyard management.
+\newline \noindent
+The research presented in this paper is part of my PhD program at QUT, where I am investigating sustainability in viticulture.
+This submission represents the first paper in a series of studies we are conducting.
+Our aim is to produce several papers using different transects of the data to explore various aspects of this topic comprehensively.
+We are excited about the potential findings and their implications for sustainability and informing decisions in vineyard management.
+\newline \noindent
+Additionally, I would like to inform you that another paper, titled "An Analysis of Underlying Relationships Between Factors Related to Operating Costs and Revenue in Australian Vineyards," is currently under review by the Journal of Agriculture and Food Research.
+In that paper, we delve deeper into these trends from an economic perspective, highlighting the interconnectedness of the variables and how they impact vineyard revenue and operating costs.
 \newline \noindent
@@
 Thank you for your consideration of this manuscript.
-\bigskip \noindent% Vertical whitespace
-
-\includegraphics[scale=0.1]{signature_block.png}
+\newline \noindent% Vertical whitespace
+\includegraphics[scale=0.05]{signature_block.png}
 \newline \noindent

### 2024-05-28 article.tex
Fixed Figures for PLOS One Submission

@@
 The diverse nature of vineyard operations results in complicated interactions with both environmental and economic outcomes.
-The interactions between different factors can put an environmentally advantageous decision at odds with an economic one, and conversely economic decisions can be detrimental to the environment \cite{beaumelleBiodiversityConservationEcosystem2023}.
+The interactions between different factors can put an environmentally advantageous decision at odds with an economic one, and conversely economic decisions can be detrimental to the environment~\cite{beaumelleBiodiversityConservationEcosystem2023}.
 As a result many studies are including more diverse ranges of data and scales.
-However, studies tend to focus on specific aspects, such as reducing chemical use or improving soil health, often focusing on singular relationships or variable outcomes \cite{barriguinhaVineyardYieldEstimation2021,heFruitYieldPrediction2022,laurentReviewIssuesMethods2021}.
-Other analyses, look at life cycles or cost benefit analyses, reflecting a specific management decision's outcome \cite{abbottAWRIAssessingEnvironmental2016,beauchetInterannualVariabilityEnvironmental2019,ferraraLifeCycleAssessment2018} or their cost benefits by utilising different scenarios to assess specific practices \cite{fincoCombiningPrecisionViticulture2022a,singhBibliometricReviewUse2022}.
-It has been noted that some studies have relied too heavily on single points of engagement when relaying the idea of sustainability and vineyard outcomes, especially when the inclusion of more variables or data could have better described a situation as a whole \cite{baianoOverviewSustainabilityWine2021,santiago-brownSustainabilityAssessmentWineGrape2015}.
-The changing nature of sustainability also calls for measures to be more than simple snapshots of data \cite{baianoOverviewSustainabilityWine2021,montalvo-falconSustainabilityResearchWine2023,santiago-brownSustainabilityAssessmentWineGrape2015}.
-Given recent studies there is still growing interest from the winegrowing industry in further studies, with several recent reviews calling for more research to demonstrate the potential trade-offs between different operational approaches across different climates and regions \cite{baianoOverviewSustainabilityWine2021,marianiSustainableWinegrowingCurrent2015,montalvo-falconSustainabilityResearchWine2023}.
+However, studies tend to focus on specific aspects, such as reducing chemical use or improving soil health, often focusing on singular relationships or variable outcomes~\cite{barriguinhaVineyardYieldEstimation2021,heFruitYieldPrediction2022,laurentReviewIssuesMethods2021}.
+Other analyses, look at life cycles or cost benefit analyses, reflecting a specific management decision's outcome~\cite{abbottAWRIAssessingEnvironmental2016,beauchetInterannualVariabilityEnvironmental2019,ferraraLifeCycleAssessment2018} or their cost benefits by utilising different scenarios to assess specific practices~\cite{fincoCombiningPrecisionViticulture2022a,singhBibliometricReviewUse2022}.
+It has been noted that some studies have relied too heavily on single points of engagement when relaying the idea of sustainability and vineyard outcomes, especially when the inclusion of more variables or data could have better described a situation as a whole~\cite{baianoOverviewSustainabilityWine2021,santiago-brownSustainabilityAssessmentWineGrape2015}.
+The changing nature of sustainability also calls for measures to be more than simple snapshots of data~\cite{baianoOverviewSustainabilityWine2021,montalvo-falconSustainabilityResearchWine2023,santiago-brownSustainabilityAssessmentWineGrape2015}.
+Given recent studies there is still growing interest from the winegrowing industry in further studies, with several recent reviews calling for more research to demonstrate the potential trade-offs between different operational approaches across different climates and regions~\cite{baianoOverviewSustainabilityWine2021,marianiSustainableWinegrowingCurrent2015,montalvo-falconSustainabilityResearchWine2023}.
 
-A dilemma exists across agriculture through shared fundamental considerations of resource use such as water and fuel, and the resultant yield and crop value that are produced \cite{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016, zhuEffectsNitrogenLevel2017}.
+A dilemma exists across agriculture through shared fundamental considerations of resource use such as water and fuel, and the resultant yield and crop value that are produced~\cite{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016, zhuEffectsNitrogenLevel2017}.
 The average price of grapes for wine production is driven through its integration within the wine industry.
-An important connection between grapes and their price is the grape's perceived quality, which initially defines a wine's potential through the grape's chemical makeup \cite{blackTerpenoidsTheirRole2015,schreierFlavorCompositionWines1979}.
-Grapes of higher perceived quality or grapes from well known regions are likely to have higher prices \cite{wineaustraliaNationalVintageReport2021}.
-Grape quality is connected to the market value of winegrapes, with Wine Australia explicitly defining grape quality through the use of discrete price brackets in their annual reports \cite{winemakersfederationofaustraliaNationalVintageReport2018}.
-It is important to note that the generalisation made to reflect quality through using average price assumes a due diligence of those who purchase the grapes \cite{yeggeInfluenceSensoryNonsensory2001}.
+An important connection between grapes and their price is the grape's perceived quality, which initially defines a wine's potential through the grape's chemical makeup~\cite{blackTerpenoidsTheirRole2015,schreierFlavorCompositionWines1979}.
+Grapes of higher perceived quality or grapes from well known regions are likely to have higher prices~\cite{wineaustraliaNationalVintageReport2021}.
+Grape quality is connected to the market value of winegrapes, with Wine Australia explicitly defining grape quality through the use of discrete price brackets in their annual reports~\cite{winemakersfederationofaustraliaNationalVintageReport2018}.
+It is important to note that the generalisation made to reflect quality through using average price assumes a due diligence of those who purchase the grapes~\cite{yeggeInfluenceSensoryNonsensory2001}.
 The economic sustainability of a vineyard is tied to this market culture, driven by the wine industry.
-The consideration of sustainability within viticulture is also subject to environmental and sociodemographic pressures \cite{santiago-brownSustainabilityAssessmentWineGrape2015}.
-In the Australian context, these pressures include biosecurity, climate and international market changes \cite{canadellMultidecadalIncreaseForest2021,longbottomRoleVineyardPractices2015,oliverReviewSoilPhysical2013}.
+The consideration of sustainability within viticulture is also subject to environmental and sociodemographic pressures~\cite{santiago-brownSustainabilityAssessmentWineGrape2015}.
+In the Australian context, these pressures include biosecurity, climate and international market changes~\cite{canadellMultidecadalIncreaseForest2021,longbottomRoleVineyardPractices2015,oliverReviewSoilPhysical2013}.
 \par
-There is an extensive amount of research into the varied effects of factors on grape quality and yield \cite{heFruitYieldPrediction2022,laurentLocalInfluenceClimate2022,liakosMachineLearningAgriculture2018}.
+There is an extensive amount of research into the varied effects of factors on grape quality and yield~\cite{heFruitYieldPrediction2022,laurentLocalInfluenceClimate2022,liakosMachineLearningAgriculture2018}.
 Due to the lack of long-term and in-depth data there is a lack of research on grape sale price and its driving factors.
-Furthermore, individual factors are often studied in isolation with yield and sales price not appearing together \cite{abbalDecisionSupportSystem2016}.
+Furthermore, individual factors are often studied in isolation with yield and sales price not appearing together~\cite{abbalDecisionSupportSystem2016}.
 The lack of consolidated datasets restricts the ability to gain statistical insights at large scales and across multiple regions.
-As a result, broader studies are lacking \cite{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
+As a result, broader studies are lacking~\cite{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}.
 The dataset used for this analysis includes data spanning 10 years from a multitude of vineyards located over a diverse range of Australian winegrowing regions.
@@
 Data used in these analyses were obtained from Sustainable Winegrowing Australia and Wine Australia.
-Sustainable Winegrowing Australia is Australia's national wine industry sustainability program, which aims to facilitate grape  growers and winemakers in demonstrating and improving their sustainability \cite{swaSustainableWingrowingAustralia2022}.
-Wine Australia is an Australian Government statutory authority governed by the Wine Australia Act \cite{attorney-generalsdepartmentWineAustraliaCorporation2010}.
+Sustainable Winegrowing Australia is Australia's national wine industry sustainability program, which aims to facilitate grape  growers and winemakers in demonstrating and improving their sustainability~\cite{swaSustainableWingrowingAustralia2022}.
+Wine Australia is an Australian Government statutory authority governed by the Wine Australia Act~\cite{attorney-generalsdepartmentWineAustraliaCorporation2010}.
 Data collected by Wine Australia are publicly available.
 The response (predicted) variables in this analysis were yield, defined as the total tonnes of grapes harvested, and average sale price of grapes in Australian dollars per tonne.
-Both response variables were examined as totals and also relative to area harvested (see Table \ref{tab:tab1}).
+Both response variables were examined as totals and also relative to area harvested (see Table~\ref{tab:tab1}).
 Values were compared in this manner to observe how economies of scale affect the use of resources.
@@
 Data recorded by Sustainable Winegrowing Australia were entered manually by winegrowers voluntarily using a web based interface.
-For each model, vineyards were only included if they recorded all the variables used for the corresponding model (see Table \ref{tab:tab1}).
+For each model, vineyards were only included if they recorded all the variables used for the corresponding model (see Table~\ref{tab:tab1}).
 Each vineyard had at least recorded region, harvest year, yield and area harvested.
@@
 To enable direct comparisons between fuels, fuel use was converted to tonnes of carbon dioxide equivalent and collectively referenced to as emissions.
-All variables were continuous except for harvest year and region, which were categorical variables (Table \ref{tab:tab1}).
+All variables were continuous except for harvest year and region, which were categorical variables (Table~\ref{tab:tab1}).
 \par
@@
 The dataset was split into two distinct subsets that could be used across the different models.
-The first subset contained all vineyards and was used for two models (Model 1 and Model 2, see Table \ref{tab:tab1}).
-The second subset contained vineyards which either recorded a value for average price of sale per tonne through Sustainable Winegrowing Australia, or were within a region with an average price of sale recorded by Wine Australia; this subset was used for three further models (Models 3, 4 and 5, see Table \ref{tab:tab1}).
-These subsets meant that the data would be limited to samples which had recorded values for the response variables (see Table \ref{tab:tab1}), as every sample/vineyard had a recorded value for yield but not average price of sale per tonne.
+The first subset contained all vineyards and was used for two models (Model 1 and Model 2, see Table~\ref{tab:tab1}).
+The second subset contained vineyards which either recorded a value for average price of sale per tonne through Sustainable Winegrowing Australia, or were within a region with an average price of sale recorded by Wine Australia; this subset was used for three further models (Models 3, 4 and 5, see Table~\ref{tab:tab1}).
+These subsets meant that the data would be limited to samples which had recorded values for the response variables (see Table~\ref{tab:tab1}), as every sample/vineyard had a recorded value for yield but not average price of sale per tonne.
 \par
-The first subset of data (used for Model 1 and Model 2, see Table \ref{tab:tab1}) contained 5298 samples spanning the period from 2012 to 2022, covering 55 GI Regions and 1261 discrete vineyards.
+The first subset of data (used for Model 1 and Model 2, see Table~\ref{tab:tab1}) contained 5298 samples spanning the period from 2012 to 2022, covering 55 GI Regions and 1261 discrete vineyards.
 \par
-The second subset of data (used for Model 3, Model 4 and Model 5, see Table \ref{tab:tab1}) contained 2878 samples spanning the period from 2015 to 2022, covering 51 GI Regions and 944 separate vineyards.
+The second subset of data (used for Model 3, Model 4 and Model 5, see Table~\ref{tab:tab1}) contained 2878 samples spanning the period from 2015 to 2022, covering 51 GI Regions and 944 separate vineyards.
 Average price of sale per tonne was extracted from both Wine Australia (1842 values) and Sustainable Winegrowing Australia (remaining 1036 values).
@@
 \par
-Data preprocessing was conducted prior to analysis using the Python programming language \cite{g.vanrossumPythonTutorialTechnical1995}.
+Data preprocessing was conducted prior to analysis using the Python programming language~\cite{g.vanrossumPythonTutorialTechnical1995}.
 Preprocessing included the conversion of fuel to scope one emissions and prior calculations for all continuous variables which included logarithmic transformations, centring and scaling by standard deviation.
-We converted multiple emission sources into scope one emissions using the equation given from the Australian National Greenhouse Accounts Factors \cite{agdeeNationalGreenhouseAccounts2021}.
+We converted multiple emission sources into scope one emissions using the equation given from the Australian National Greenhouse Accounts Factors~\cite{agdeeNationalGreenhouseAccounts2021}.
 The calculation conducted using
@@
 \end{equation}
-where emissions was the the product of the quantity of fuel in litres, $Q$, a prescribed Energy Content, $EC$, and emission factors (as given by the Australian National Greenhouse Accounts factors) of scope one, $EF1$, and scope three, $EF3$, to tonnes of Carbon Dioxide Emission equivalent, $tCO2e$ \cite{departmentofclimatechangeenergytheenvironmentandwaterAustralianNationalGreenhouse2022}.
+where emissions was the the product of the quantity of fuel in litres, $Q$, a prescribed Energy Content, $EC$, and emission factors (as given by the Australian National Greenhouse Accounts factors) of scope one, $EF1$, and scope three, $EF3$, to tonnes of Carbon Dioxide Emission equivalent, $tCO2e$~\cite{departmentofclimatechangeenergytheenvironmentandwaterAustralianNationalGreenhouse2022}.
 \par
 \begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{my_map.pdf}}
+  % \resizebox{\textwidth}{!}{
+  % \includegraphics{fig1.pdf}}
   \caption{Map of each GI Regions' (outlined by Wine Australia) average income for a vineyard of that region (average grape sale price per tonne $\times$ total tonnes yielded).}\label{fig:map}
@@
 \par
-Differences in vineyard locations were captured through the use of Geographical Indicator Regions defined by Wine Australia \cite{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
-Although vineyards generally differed in income, there were pronounced differences between regions as shown in Figure \ref{fig:map}.
-These differences were likely due to the site of a vineyard being important as it predetermines several physical parameters such as climate, geology and soil, making location a widely considered key determinant of grape yield and average sale price \cite{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
-Each GI Region has its own unique mixture of climatic and geophysical properties that describes a unique winegrowing region within Australia and is a protected trademark under the Wine Australia act \cite{attorney-generalsdepartmentWineAustraliaCorporation2010}.
+Differences in vineyard locations were captured through the use of Geographical Indicator Regions defined by Wine Australia~\cite{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}.
+Although vineyards generally differed in income, there were pronounced differences between regions as shown in Figure~\ref{fig:map}.
+These differences were likely due to the site of a vineyard being important as it predetermines several physical parameters such as climate, geology and soil, making location a widely considered key determinant of grape yield and average sale price~\cite{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}.
+Each GI Region has its own unique mixture of climatic and geophysical properties that describes a unique winegrowing region within Australia and is a protected trademark under the Wine Australia act~\cite{attorney-generalsdepartmentWineAustraliaCorporation2010}.
 Both Wine Australia and Sustainable Winegrowing Australia used the same GI Region categorical variable format to describe location.
 \par
- The climatic properties of each GI Region were summarised by using predefined classifications as per the \cite{sustainablewinegrowingaustraliaSustainableWinegrowingAustralia2021} user manual.
+ The climatic properties of each GI Region were summarised by using predefined classifications as per the~\cite{sustainablewinegrowingaustraliaSustainableWinegrowingAustralia2021} user manual.
 The user manual describes climates by rainfall and temperature, creating supersets of Regions of similar climatic properties.
@@
 \par
-Five general linear models (GLMs) were created (see Table \ref{tab:tab1}).
+Five general linear models (GLMs) were created (see Table~\ref{tab:tab1}).
 GLMs were chosen as they offer the ability to produce statistical models that are explicit in the relationships between predictors and response variables.
@@
 Model fit was measured in terms of $R^2$ and adjusted $R^2$ as well as F statistics, and t-tests were used to determine if predictors significantly contributed to the models when accounting for other variables, showing which specific years and areas contributed significantly.
-Both the PPCC and GLM analyses were created using the R statistical programming language \cite{rcoreteamLanguageEnvironmentStatistical2021} with the Caret package \cite{kuhnBuildingPredictiveModels2008}.
+Both the PPCC and GLM analyses were created using the R statistical programming language~\cite{rcoreteamLanguageEnvironmentStatistical2021} with the Caret package~\cite{kuhnBuildingPredictiveModels2008}.
 \par
@@
 \subsection*{Summary Statistics}\label{sec:exp_anal}
- Table \ref{tab:summary} shows the summary statistics of each variable in its original units.
-The range of these values shows the level of difference between some vineyards, with operations differing by orders of magnitude in size, yield and average price of sale (See Table \ref{tab:tab1}).
+ Table~\ref{tab:summary} shows the summary statistics of each variable in its original units.
+The range of these values shows the level of difference between some vineyards, with operations differing by orders of magnitude in size, yield and average price of sale (See Table~\ref{tab:tab1}).
 \par
@@
 
-PPCCs values for the transformed, centred and scaled variables are shown in Table \ref{tab:pearson_correlation}.
+PPCCs values for the transformed, centred and scaled variables are shown in Table~\ref{tab:pearson_correlation}.
 The process of centring and scaling variables changes them so that their mean is centred on 0 and all values are scaled to the variables standard deviation, such that 1 is equal to the variables standard deviation.
@@
 The three exceptions were: scope one emissions in Model 3 (p=0.22) and Model 4 (p=0.0.39), and the interaction between area harvested and water used in model 2 (p=0.22).
-Note that scope one emissions was included in all models to directly compare the response variables as ratios of vineyard size to raw values and because it was strongly correlated to the response variable in every model (except model 5) especially for Models 1 and 4 (Table \ref{tab:pearson_correlation}).
+Note that scope one emissions was included in all models to directly compare the response variables as ratios of vineyard size to raw values and because it was strongly correlated to the response variable in every model (except model 5) especially for Models 1 and 4 (Table~\ref{tab:pearson_correlation}).
 %A full list of regression coefficients 95\% CIs and p-values for each of the four models is provided in the appendix. % You gotta add these to the appendix in a more readable fashion
@@
 
-Coefficients related to continuous variables are summarised in Table \ref{tab:contvar}.
+Coefficients related to continuous variables are summarised in Table~\ref{tab:contvar}.
 In Model 1 all coefficients except for the intercept significantly contributed to the model (p $<$ 0.05), and in Model 2 all coefficients were statistically significant (p $<$ 0.05).
@@
 
-The regression coefficients for the Year variable of each model is depicted in Figure \ref{fig:yearly}.
+The regression coefficients for the Year variable of each model is depicted in Figure~\ref{fig:yearly}.
 For each model, the first year for a model's was used as the baseline: 2012 for Models 1 and 2, and 2015 for Models 3, 4 and 5.
@@
 However, the coefficients varied substantially over the years, and were not significantly different from zero in some years.
-Models 4 and 5 are very similar in Figure \ref{fig:yearly}, indicating that the exclusion of area does not greatly affect the contribution from yearly influence.
+Models 4 and 5 are very similar in Figure~\ref{fig:yearly}, indicating that the exclusion of area does not greatly affect the contribution from yearly influence.
 Models 4 and 5 have the most prominent trends, showing an increase in yearly effects over time, with Model 3 also increasing from 2016 to 2018, but plateauing afterwards.
@@
 \begin{figure}
-  \includegraphics{yearly_plots.pdf}
+  % \includegraphics{fig2.pdf}
   \caption{Model Coefficient values for Year, with standard error bars.
@@
 %difference between the continuous variables coefficients is the change from positive to negative values.
-This change occurs between the Models for Yield (Model 1 and 2) and the Models for value (Models 3 and 4); where all but the coefficient for area harvested had the opposite sign (see Table \ref{tab:contvar}).
+This change occurs between the Models for Yield (Model 1 and 2) and the Models for value (Models 3 and 4); where all but the coefficient for area harvested had the opposite sign (see Table~\ref{tab:contvar}).
 These models also differ in an order of magnitude when looking at resource use, with the coefficients for yield being smaller than those for value.
@@
 
-Regional differences by temperature and rainfall are summarised in Figure \ref{fig:spider}.
+Regional differences by temperature and rainfall are summarised in Figure~\ref{fig:spider}.
 The most notable difference is between vineyards within 'Hot' and 'Very Dry' regions (warm inland regions), where high average sale prices are historically low, and yield is high.
@@
 The warmer and drier regions tend to also have larger vineyards.
-These regional differences are further shown in Figure \ref{fig:yield_vs_value_area}, as a ratio of vineyard; where there is a pronounced difference in 'Hot' and 'Very Dry' regions producing more per area with lower average sale prices per area than other regions.
+These regional differences are further shown in Figure~\ref{fig:yield_vs_value_area}, as a ratio of vineyard; where there is a pronounced difference in 'Hot' and 'Very Dry' regions producing more per area with lower average sale prices per area than other regions.
 \par
 \begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{starplots.png}}
+  % \resizebox{\textwidth}{!}{
+  % \includegraphics{fig3.pdf}}
   \caption{Radar plot of climatic profile's resource use, yield and average sale price.
@@
 
-Figure \ref{fig:yield_vs_value_area} further shows the emphasis that 'Hot' areas have on high yields, with low average sale price compared with other regions.
+Figure~\ref{fig:yield_vs_value_area} further shows the emphasis that 'Hot' areas have on high yields, with low average sale price compared with other regions.
 Scaling average price and yield by area shows a strong negative trend, trading quantity for higher sales prices.
@@
 \begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{yield_verse_value_by_area.pdf}}
+  % \resizebox{\textwidth}{!}{
+  % \includegraphics{fig4.pdf}}
     \caption{Scatter plot of vineyard yield against the average sale price as ratios to area harvested.
@@
 
-Table \ref{tab:kfold} shows the validation results of each of the models.
+Table~\ref{tab:kfold} shows the validation results of each of the models.
 The $R^2$ measures of fit show similar results to the initial models, with a slight decrease as expected, indicating that the models are robust and consistent.
@@
 \begin{table}[]
-  \label{tab:kfold}
-  \caption{Model validation using k-fold cross validation, for 10 folds repeated 100 times.}
+  \caption{Model validation using k-fold cross validation, for 10 folds repeated 100 times.}\label{tab:kfold}
   % \resizebox{\textwidth}{!}{
@@
 The vertical integration of winegrowing within the wine industry ties winegrowers to secondary and tertiary industries, such as wine production, packaging, transport and sales.
-This results in unique issues and considerations for each vineyard, where on-the-ground decisions are influenced by other wine industry's choices, such as the use of sustainable practices in vineyards as a requirement for sale in overseas markets; notably these interactions can be further complicated by some winegrowers being completely integrated into a wine company, while others are not \cite{knightFirmResourcesDevelopment2019}.
+This results in unique issues and considerations for each vineyard, where on-the-ground decisions are influenced by other wine industry's choices, such as the use of sustainable practices in vineyards as a requirement for sale in overseas markets; notably these interactions can be further complicated by some winegrowers being completely integrated into a wine company, while others are not~\cite{knightFirmResourcesDevelopment2019}.
 Incorporating decisions into the model could help describe the contributing factors to regional differences beyond resource consumption, motivating the call for more granular data and more sophisticated modelling.
@@
 There are many on-the-ground decisions that influence both sales price and yield.
-The decision to prioritise average sale price over quantity, is governed by complex physical and social forces, for example international market demands, disease pressures and natural disasters \cite{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}, with many of these occurrences being highlighted throughout the reports from Wine Australia over the past decade \cite{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
-However, the changes in the coefficients (see Figure \ref{fig:yearly}) are not reflective of many known occurrences, such as the 2020 bush fires, which had higher values for coefficients than prior years; During the 2020 bush fires 40,000 tonnes of grapes were lost across 18 different wine regions due to bush fires and smoke taint.
-In comparison to countrywide pressures such as drought, this damage made up only 3\% of the total amount of grapes for that year; although acknowledged as a considerable loss on an individual basis, it was deemed to be only a minor national concern by Wine Australia when compared to other environmental pressures such as drought \cite{wineaustraliaNationalVintageReport2020}
+The decision to prioritise average sale price over quantity, is governed by complex physical and social forces, for example international market demands, disease pressures and natural disasters~\cite{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}, with many of these occurrences being highlighted throughout the reports from Wine Australia over the past decade~\cite{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
+However, the changes in the coefficients (see Figure~\ref{fig:yearly}) are not reflective of many known occurrences, such as the 2020 bush fires, which had higher values for coefficients than prior years; During the 2020 bush fires 40,000 tonnes of grapes were lost across 18 different wine regions due to bush fires and smoke taint.
+In comparison to countrywide pressures such as drought, this damage made up only 3\% of the total amount of grapes for that year; although acknowledged as a considerable loss on an individual basis, it was deemed to be only a minor national concern by Wine Australia when compared to other environmental pressures such as drought~\cite{wineaustraliaNationalVintageReport2020}
 \par
 Climatic pressures are an important consideration for growers, especially those in warmer and drier regions.
-The Wine Australia reports also show that warm inland regions have seen a decline in profit over the past decade, whereas regions with lower average sales prices did not \cite{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
+The Wine Australia reports also show that warm inland regions have seen a decline in profit over the past decade, whereas regions with lower average sales prices did not~\cite{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}.
 Vineyards in warm inland regions also tend to be larger, making up for lower sale prices with larger yields.
@@
 However, the increasing climatic pressures mixed with the requirement for larger volumes of water, make the sustainability of some vineyards come into question.
-Furthermore, intensive farming in general is known to jeopardise the sustainability of an operation through the degradation of soil and waterways \cite{capelloEffectsTractorPasses2019,linHydropedologySynergisticIntegration2012,pisciottaGroundwaterNitrateRisk2015}.
+Furthermore, intensive farming in general is known to jeopardise the sustainability of an operation through the degradation of soil and waterways~\cite{capelloEffectsTractorPasses2019,linHydropedologySynergisticIntegration2012,pisciottaGroundwaterNitrateRisk2015}.
 There are established methods that can help to mitigate these effects, such as the use of cover crops, midrow crop rotation and efficient irrigation.
@@
 Some regions appeared to produce grapes of lower average sale price at scale whilst others achieved higher priced grapes in lower volumes.
-This empirical finding is consistent with Wine Australia's annual reports, which shows that some GI regions are known for producing large amounts of lower grade (low value per tonne) grapes \cite{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
+This empirical finding is consistent with Wine Australia's annual reports, which shows that some GI regions are known for producing large amounts of lower grade (low value per tonne) grapes~\cite{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}.
 Comparatively, other regions only produce grapes of higher sales price but in smaller quantities.
@@
 Due to regional differences, different strategies are  employed, such as some regions targeting mass production over higher sales price.
-This is most notable when grouping regions by climate, especially when considering GI Regions in the 'Hot Very Dry' climate (see Figure \ref{fig:yield_vs_value_area}).
-Figure \ref{fig:spider} also shows that comparatively 'Warm' and 'Dry' Regions manage their resources incredibly efficiently having generally larger areas using similar resources to those in 'Cool' regions but having areas comparable to regions in 'Hot' climates.
-The coefficients for Model 4 also show a greater benefit of resource use per area when producing grapes at higher average sale prices, showing higher resource use per area reflective of higher average sales price (see Table \ref{tab:summary}).
+This is most notable when grouping regions by climate, especially when considering GI Regions in the 'Hot Very Dry' climate (see Figure~\ref{fig:yield_vs_value_area}).
+Figure~\ref{fig:spider} also shows that comparatively 'Warm' and 'Dry' Regions manage their resources incredibly efficiently having generally larger areas using similar resources to those in 'Cool' regions but having areas comparable to regions in 'Hot' climates.
+The coefficients for Model 4 also show a greater benefit of resource use per area when producing grapes at higher average sale prices, showing higher resource use per area reflective of higher average sales price (see Table~\ref{tab:summary}).
 %In the preliminary analyses alternative models found that without the direct incorporation of GI Region or year, predictions greatly under performed.
 The effect of climate in the models was never as significant as the more granular GI regions, and always led to less accurate models.
-Although not chosen over GI region, climate was considered to be a large determinant of the ability to produce larger quantities of grapes, as well as a determinant in grape sale price \cite{agostaRegionalClimateVariability2012}.
-The more granular GI Region likely explained a broader mix of geographical phenomenon, such as soil, geology and access to water resources \cite{abbalDecisionSupportSystem2016,carmonaUseParticipatoryObjectOriented2011}.
+Although not chosen over GI region, climate was considered to be a large determinant of the ability to produce larger quantities of grapes, as well as a determinant in grape sale price~\cite{agostaRegionalClimateVariability2012}.
+The more granular GI Region likely explained a broader mix of geographical phenomenon, such as soil, geology and access to water resources~\cite{abbalDecisionSupportSystem2016,carmonaUseParticipatoryObjectOriented2011}.
 The interaction between year and GI Region likely accounted for events such as bushfires, which would be impactful, but only at a local level, both in time and space.
@@
 These differences could be caused by the use of alternative sustainable practices in the field.
-Moreover, while there is evidence to suggest that environmentally sustainable practices can reduce costs, and increase efficiency whilst improving the quality of grapes; more research is needed to link these benefits across different regions and climates \cite{baianoOverviewSustainabilityWine2021,marianiSustainableWinegrowingCurrent2015,montalvo-falconSustainabilityResearchWine2023}.
+Moreover, while there is evidence to suggest that environmentally sustainable practices can reduce costs, and increase efficiency whilst improving the quality of grapes; more research is needed to link these benefits across different regions and climates~\cite{baianoOverviewSustainabilityWine2021,marianiSustainableWinegrowingCurrent2015,montalvo-falconSustainabilityResearchWine2023}.
 \section*{Conclusion}
@@
 \begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{model1_qq.pdf}}
+  % \resizebox{\textwidth}{!}{
+  % \includegraphics{s1.pdf}}
   \caption{QQ-plot of Model 1.}
@@
 \begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{model1_rvf.pdf}}
+  % \resizebox{\textwidth}{!}{
+  % \includegraphics{s2.pdf}}
   \caption{Residuals vs fitted values for Model 1.}
@@
 \begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{model2_qq.pdf}}
+  % \resizebox{\textwidth}{!}{
+  % \includegraphics{s3.pdf}}
   \caption{QQ-plot of Model 2.}
@@
 \begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{model2_rvf.pdf}}
+  % \resizebox{\textwidth}{!}{
+  % \includegraphics{s4.pdf}}
   \caption{Residuals vs fitted values for Model 2.}
@@
 \begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{model3_qq.pdf}}
+  % \resizebox{\textwidth}{!}{
+  % \includegraphics{s5.pdf}}
   \caption{QQ-plot of Model 3.}
@@
 \begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{model3_rvf.pdf}}
+  % \resizebox{\textwidth}{!}{
+  % \includegraphics{s6.pdf}}
   \caption{Residuals vs fitted values for Model 3.}
@@
 \begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{model4_qq.pdf}}
+  % \resizebox{\textwidth}{!}{
+  % \includegraphics{s7.pdf}}
   \caption{QQ-plot of Model 4.}
@@
 \begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{model4_rvf.pdf}}
+  % \resizebox{\textwidth}{!}{
+  % \includegraphics{s8.pdf}}
   \caption{Residuals vs fitted values for Model 4.}
@@
 \begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{model5_qq.pdf}}
+  % \resizebox{\textwidth}{!}{
+  % \includegraphics{s9.pdf}}
   \caption{QQ-plot of Model 5.}
@@
 \begin{figure}
-  \resizebox{\textwidth}{!}{
-  \includegraphics{model5_rvf.pdf}}
+  % \resizebox{\textwidth}{!}{
+  % \includegraphics{s10.pdf}}
   \caption{Residuals vs fitted values for Model 5.}

### 2024-11-30 Manuscript.tex
KM Edits RE rejoinder

% Template for PLoS
% Version 3.6 Aug 2022
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimise problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file 
% and leave only the final text of your manuscript. 
% PLOS recommends using latexdiff to track changes during the review, which will help maintain a clean text file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info, or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that no packages listed in the template may be deleted.
%
% Please do not include colours or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
%, please use "Fig" instead of "Figure" for figure citations.
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or coloured text (cell background colour/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjust width environment as illustrated in the example table in the text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex.
%
% For inline equations, please include all portions of an equation in the math environment. For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanised font is desired).
%
% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
%
% Please add line breaks to long display equations when possible to fit the size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscripts or subscripts outside brackets/braces, please group using {}. For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
%
% Do not use \cal for calligraphic font. Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, valid for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package is used to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage[nopatch=eqnum]{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% colour can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{csvsimple}
\usepackage{longtable}
\usepackage{adjustbox} % For fitting table to page width
\usepackage{underscore} % For handling underscores in CSV files without interference in other parts

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS-provided BibTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}

%% Include all macros below

% \newcommand{\lorem}{{\bf LOREM}}
% \newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

% Original draft packages
\usepackage{array}
\usepackage{amssymb}
\usepackage{adjustbox}
\usepackage{tabulary, ragged2e}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{csvsimple}
\usepackage{longtable}

\begin{document}

\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{The influence of resource use on yield versus sale price trade-off in Australian vineyards}
% \title{The influence of resource use on yield versus sale price trade-off in Australian vineyards}
% Please use "sentence case" for titles and headings (capitalise only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Bryce Boyd\textsuperscript{1,2,3,*},
Kate Helmstedt\textsuperscript{1},
Mardi Longbottom\textsuperscript{2, \dag},
Kerrie Mengersen\textsuperscript{1}
% {\textpilcrow}
\\
\bigskip
\textbf{1} Affiliation Faculty of Mathematical Sciences, Queensland University of Technology, Brisbane, QLD, Australia
\\
\textbf{2} Affiliation Australian Wine Research Institute, Adelaide, SA, Australia
\\
\textbf{3} Affiliation Food Agility CRC, Sydney, NSW, Australia
\\
\bigskip

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
% 
% Remove or comment out the author notes below if they aren't used.
%

% % Primary Equal Contribution Note
% \Yinyang These authors contributed equally to this work.

% % Additional Equal Contribution Note
% % Use this double-dagger symbol for special authorship notes, such as senior authorship.
% \ddag These authors also contributed equally to this work.

% % Current address notes
% \textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
% % \textcurrency b Insert second current address 
% % \textcurrency c Insert third current address

% % Deceased author note
% \dag Deceased

% % Group/Consortium Author Note
% \textpilcrow Membership list can be found in the Acknowledgments section.

% % Use the asterisk to denote the corresponding authorship and provide the email address in the note below.
* Bryce.polley@awri.com.au
\dag Mardi.longbottom@awri.com.au

\end{flushleft}

\section*{Abstract}
% The word limit needs to be checked.

Strategies for achieving sustainability in the winegrowing industry require balancing resource investment against the economic outcomes of resultant yields and sale price of the produce. Although there has been much research into forecasting outcomes and resource use, little has been done to illustrate their effects on one another and the consequential economic outcomes. This analysis uses statistical models to observe relationships between resource use, yield, and sale price. The dataset used for this analysis includes data collected for the past 10 years from 1261 vineyards located over a diverse range of Australian winegrowing regions. Yield and sale price were evaluated regarding resource use factors, such as water use and Greenhouse Gas (GHG) emissions. The analysis confirmed a strong relationship between area and resource use, with the overall area of a vineyard and its access to resources significantly determining the upper limit of yield. However, the area was also negatively related to the average sale price of grapes; we find that higher average sale prices were connected to high resource inputs per area rather than to the overall expenditure of resources. Regional and temporal effects on vineyard yield and average sales price were also identified. The analysis highlighted the importance of considering a vineyard's business goal, region, external pressures, and economies of scale when pursuing higher yields verse higher average sales prices.

% \begin{highlights}
% \item Found that higher resource inputs per area are associated with higher sale prices rather than overall resource expenditure.
% \item Identified significant differences in yield and sale prices across regions, highlighting the influence of localised conditions on vineyard profitability.
% \item Developed baseline models demonstrating the critical relationship between vineyard area, resource usage, and economic outcomes.
% \item Analysis of long-term data offering a comprehensive view of industry trends highlighting resource use trends and economic outcomes.
% \end{highlights}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                main text             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\linenumbers

\section*{Introduction}

The focus on sustainability in agronomic industries has changed how enterprises do business. The diverse nature of vineyard operations results in complicated interactions with both environmental and economic outcomes. The interactions between different factors can put an environmentally advantageous decision at odds with an economic one, and conversely, financial decisions can be detrimental to the environment~\cite{beaumelleBiodiversityConservationEcosystem2023}. As a result, many studies have included more diverse data and scales. However, studies tend to focus on specific aspects, such as reducing chemical use or improving soil health, often focusing on singular relationships or variable outcomes~\cite{barriguinhaVineyardYieldEstimation2021,heFruitYieldPrediction2022,laurentReviewIssuesMethods2021}. Other analyses, look at life cycles or cost-benefit analyses, reflecting a specific management decision's outcome~\cite{abbottAWRIAssessingEnvironmental2016,beauchetInterannualVariabilityEnvironmental2019,ferraraLifeCycleAssessment2018} or their cost benefits by utilising different scenarios to assess specific practices~\cite{fincoCombiningPrecisionViticulture2022a,singhBibliometricReviewUse2022}. It has been noted that some studies have relied too heavily on single points of engagement when relaying the idea of sustainability and vineyard outcomes, especially when the inclusion of more variables or data could have better described a situation as a whole~\cite{baianoOverviewSustainabilityWine2021,santiago-brownSustainabilityAssessmentWineGrape2015}. The changing nature of sustainability also calls for measures to be more than simple snapshots of data~\cite{baianoOverviewSustainabilityWine2021,montalvo-falconSustainabilityResearchWine2023,santiago-brownSustainabilityAssessmentWineGrape2015}. Given recent studies there is still growing interest from the winegrowing industry in further studies, with several recent reviews calling for more research to demonstrate the potential trade-offs between different operational approaches across various climates and regions~\cite{baianoOverviewSustainabilityWine2021,marianiSustainableWinegrowingCurrent2015,montalvo-falconSustainabilityResearchWine2023}.

A dilemma exists across agriculture through shared fundamental considerations of resource use such as water and fuel, and the resultant yield and crop value that are produced~\cite{hemmingCherryTomatoProduction2020,kawasakiQualityMattersMore2016, zhuEffectsNitrogenLevel2017}. The average price of grapes for wine production is driven by its integration within the wine industry. An essential connection between grapes and their price is the grape's perceived quality, which initially defines a wine's potential through the grape's chemical makeup~\cite{blackTerpenoidsTheirRole2015,schreierFlavorCompositionWines1979}. Grapes of higher perceived quality or grapes from well-known regions are likely to have higher prices~\cite{wineaustraliaNationalVintageReport2021}. Grape quality is connected to the market value of winegrapes, with Wine Australia explicitly defining grape quality through the use of discrete price brackets in their annual reports~\cite{winemakersfederationofaustraliaNationalVintageReport2018}. It is important to note that the generalisation made to reflect quality through average price assumes the due diligence of those purchasing the grapes~\cite{yeggeInfluenceSensoryNonsensory2001}. The economic sustainability of a vineyard is tied to this market culture, driven by the wine industry. The consideration of sustainability within viticulture is also subject to environmental and sociodemographic pressures~\cite{santiago-brownSustainabilityAssessmentWineGrape2015}. In the Australian context, these pressures include biosecurity, climate and international market changes~\cite{canadellMultidecadalIncreaseForest2021,longbottomRoleVineyardPractices2015,oliverReviewSoilPhysical2013}.
\par
There is an extensive amount of research into the varied effects of factors on grape quality and yield~\cite{heFruitYieldPrediction2022,laurentLocalInfluenceClimate2022,liakosMachineLearningAgriculture2018}. Due to the lack of long-term and in-depth data there is a lack of research on grape sale price and its driving factors. Furthermore, individual factors are often studied in isolation, with yield and sales price not appearing together~\cite{abbalDecisionSupportSystem2016}. The lack of consolidated datasets restricts gaining statistical insights at large scales and across multiple regions. As a result, broader studies are lacking~\cite{keithjonesAustralianWineIndustry2002,knightFirmResourcesDevelopment2019}. The dataset used for this analysis includes data spanning 10 years from a multitude of vineyards located over a diverse range of Australian winegrowing regions. We use this dataset to describe the relationship between the resources used, such as water and fuel use, with the output yield and average sale price of the resultant product, taking into account the size and location of the vineyard. In a practical sense, this aim is a baseline for comparison: given a vineyard within Australia, one could estimate the comparative efficiency of the trade-off between invested resources, yield and average sale price. This is the first time such a trade-off has been explicitly confirmed across varying regions, scales and climates in the Australian winegrowing industry.
\par
We aim to contribute to understanding resource-use efficiency in the Australian wine industry by providing and analysing these baselines using this unique set of regional and temporal data on vineyard resource allocation, grape yield, and sales price. Our analyses reveal that resource input per area, rather than overall expenditure, is a more accurate predictor of high sales prices, especially in regions with specific climate and size characteristics. We further aim to link this effect to regional attributes, where climatic factors interact with operational scale. These contributions will help to establish a baseline for vineyard managers and policymakers to assess resource trade-offs, ultimately supporting more sustainable and economically viable decision-making in the face of climate and market pressures.

\section*{Methods}
\subsection*{Data}

Data used in these analyses were obtained from Sustainable Winegrowing Australia and Wine Australia. Sustainable Winegrowing Australia is Australia's national wine industry sustainability program, which aims to facilitate grape growers and winemakers in demonstrating and improving their sustainability~\cite{swaSustainableWingrowingAustralia2022}. Wine Australia is an Australian Government statutory authority governed by the Wine Australia Act~\cite{attorney-generalsdepartmentWineAustraliaCorporation2010}. Data collected by Wine Australia are publicly available. This analysis's response (predicted) variables were yield, defined as the total tonnes of grapes harvested and the average sale price of grapes in Australian dollars per tonne. Both response variables were examined as totals and also relative to the area harvested (see Table~\ref{tab:tab1}). Values were compared in this manner to observe how economies of scale affect the use of resources.

\begin{table}[!ht]
\caption{Summary of models; their predictors, covariates and variable interactions.}\label{tab:tab1}
  \resizebox{\textwidth}{!}{\begin{tabular}{@{}ccccc@{}}
    \toprule
    & \textbf{Response} & \textbf{Predictors} & \textbf{Covariates} & \textbf{Interactions} \\ \midrule
    \textbf{Model 1} & Yield & \begin{tabular}[c]{@{}c@{}}Water Used\\ scope one Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & N/A \\
    \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
    \textbf{Model 2} & ${ \textrm{Yield}}\over{ \textrm{Area Harvested}}$ & \begin{tabular}[c]{@{}c@{}}Water Used\\ scope one Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * scope one Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular} \\
    \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
    \textbf{Model 3} & {$\textrm{Yield} {\times} \textrm{Average Sale Price}$} & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope One Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & N/A\\
    \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
    \textbf{Model 4} & Average Sale Price & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope One Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested\\ Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}}Area Harvested * Scope One Emissions\\ Area Harvested * Water Use\\ Year * Region\end{tabular}
        \\
        \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
     \textbf{Model 5} &  Average Sale Price & \begin{tabular}[c]{@{}c@{}}Water Used\\ Scope One Emissions\end{tabular} & \begin{tabular}[c]{@{}c@{}} Year\\ GI Region\end{tabular} & \begin{tabular}[c]{@{}c@{}} Year * Region\end{tabular}
        \end{tabular}}
\end{table}

\par
% Edit ID: 2
Data recorded by Sustainable Winegrowing Australia were entered manually by winegrowers voluntarily using a web-based interface. Data was recorded after the Australian crush (February to April), with entries due for the season by the 31$^{st}$ of August each year. The SWA user manual outlines details surrounding the web interface and its requirements ~\cite{sustainablewinegrowingaustraliaSustainableWinegrowingAustralia2021}. Vineyards volunteered data as part of Sustainable Winegrowing Australia initiatives, which included collaboration, workshops and certification of sustainable practices using third-party auditors. Vineyards were only included for each model if they recorded all the variables used for the corresponding model (see Table~\ref{tab:tab1}). Each vineyard had at least recorded region, harvest year, yield and area harvested. Other variables used but not present for every vineyard were average sale price, water used and fuel used (diesel, petrol, biodiesel and LPG). Fuel use was converted to equivalent tonnes of carbon dioxide and collectively referenced as emissions to enable direct comparisons between fuels. All variables were continuous except for harvest year and region, which were categorical variables (Table~\ref{tab:tab1}).
\par
% Edit ID 3
As data from Sustainable Wine Australia were voluntarily given, missing values were improved using regional average prices from the Wine Australia (previously the Winemakers Federation of Australia) data. Data from Wine Australia were collected via phone surveys and included total tonnes purchased, the average cost per tonne, and yearly change in price for region and grape varietal; the data is publicly available through the Wine Australia Annual reports~\cite{wineaustraliaNationalVintageReport2019, wineaustraliaNationalVintageReport2020, wineaustraliaNationalVintageReport2021, wineaustraliaNationalVintageReport2022, winemakersfederationofaustraliaNationalVintageReport2012, winemakersfederationofaustraliaNationalVintageReport2013, winemakersfederationofaustraliaNationalVintageReport2014, winemakersfederationofaustraliaNationalVintageReport2015, winemakersfederationofaustraliaNationalVintageReport2016, winemakersfederationofaustraliaNationalVintageReport2017, winemakersfederationofaustraliaNationalVintageReport2018}.
\par
The dataset was split into two subsets that could be used across the different models. The first subset contained all vineyards and was used for two models (Model 1 and Model 2, see Table~\ref{tab:tab1}). The second subset contained vineyards that either recorded a value for an average price of sale per tonne through Sustainable Winegrowing Australia or were within a region with an average price of sale recorded by Wine Australia; this subset was used for three further models (Models 3, 4 and 5, see Table~\ref{tab:tab1}). These subsets meant that the data would be limited to samples with recorded values for the response variables (see Table~\ref{tab:tab1}), as every sample/vineyard had a recorded value for yield but not the average sale price per tonne.
\par
The first subset of data (used for Model 1 and Model 2, see Table~\ref{tab:tab1}) contained 5298 samples spanning 2012 to 2022, covering 55 GI Regions and 1261 discrete vineyards.
\par
The second subset of data (used for Model 3, Model 4 and Model 5, see Table~\ref{tab:tab1}) contained 2878 samples spanning the period from 2015 to 2022, covering 51 GI Regions and 944 separate vineyards. The average sale price per tonne was extracted from Wine Australia (1842 values) and Sustainable Winegrowing Australia (remaining 1036 values).
\par
Additional variables were considered for analysis but were excluded due to being either underreported or had insignificant contributions to model accuracies. Variables explored but not used due to low reporting values included fertiliser (kg), electricity (Kw/h), and scope two emissions (TCO$^2$ equivalent). Variables were considered but ultimately removed due to a lack of significant contributions to models, including using renewable energy, contractors, and pressures such as frost, fire, and disease.
\par
Data preprocessing was conducted before analysis using the Python programming language~\cite{g.vanrossumPythonTutorialTechnical1995}. Preprocessing included the conversion of fuel to scope one emission and prior calculations for all continuous variables, including logarithmic transformations, centring and scaling by standard deviation. 
We converted multiple emission sources into scope one emissions using the equation from the Australian National Greenhouse Accounts Factors~\cite{agdeeNationalGreenhouseAccounts2021}. The calculation was conducted using
\begin{equation}
\label{(1)}
    tCO_{2}e={{Q \times EC \times EF1 + EF3}\over{1000}},
\end{equation}
where emissions were the product of the quantity of fuel in litres, $Q$, a prescribed Energy Content, $EC$, and emission factors (as given by the Australian National Greenhouse Accounts factors) of scope one, $EF1$, and scope three, $EF3$, to tonnes of Carbon Dioxide Emission equivalent, $tCO2e$~\cite{departmentofclimatechangeenergytheenvironmentandwaterAustralianNationalGreenhouse2022}.
\par

\par
Differences in vineyard locations were captured through the use of Geographical Indicator Regions defined by Wine Australia~\cite{hallidayAustralianWineEncyclopedia2009,oliverReviewSoilPhysical2013,soarClimateDriversRed2008}. Although vineyards generally differed in income, there were pronounced differences between regions as shown in Fig~\ref{fig:map}. These differences were likely due to the site of a vineyard being important as it predetermines several physical parameters such as climate, geology and soil, making location a widely considered key determinant of grape yield and average sale price~\cite{abbalDecisionSupportSystem2016,agostaRegionalClimateVariability2012,fragaMultivariateClusteringViticultural2017}. Each GI Region has its unique mixture of climatic and geophysical properties that describes a unique winegrowing region within Australia and is a protected trademark under the Wine Australia act~\cite{attorney-generalsdepartmentWineAustraliaCorporation2010}. Wine Australia and Sustainable Winegrowing Australia used the same GI Region categorical variable format to describe location.

\begin{figure}[htb]
  % \resizebox{\textwidth}{!}{
  % \includegraphics{fig1.pdf}}
  \caption{Map of each GI Regions' (outlined by Wine Australia) average income for a vineyard of that region (average grape sale price per tonne $\times$ total tonnes yielded).}\label{fig:map}
\end{figure}%%

\par
% Edit ID 4
 Each GI Region's climatic properties were summarised using predefined classifications per the~\cite{sustainablewinegrowingaustraliaSustainableWinegrowingAustralia2021} user manual. The user manual describes climates by rainfall and temperature, creating supersets of regions of similar climatic properties. The climatic groups illustrated similarities and differences in sets larger than GI Regions.
These classifications were literal descriptors used by industry members to discern between different weather types within winegrowing regions. However, the climate descriptors for Australian GI regions were originally conceived by~\cite{gladstones_climate_2010} using temperature-based indices. A key metric employed was the Mean January Temperature (MJT), which indicates heat accumulation during the growing season. Regions were categorised based on their MJT values into cool, intermediate, warm, and hot climates. This classification was used to aid viticulturists in selecting grape varieties best suited to the thermal conditions of each region. Further in-depth summarisations of Australian wine region climates can be found in greater detail using Wine Australia's climate atlas~\cite{remenyi_australias_2019}. Other climatic descriptors were explored, such as the  Köppen climate classification; however, this index did not offer enough granularity to highlight any trends or clustering within the data.
\par
\subsection*{Analysis}
% Edit ID 17
General Linear Models (GLMs) were chosen as the primary analytical approach due to their ability to quantify and interpret relationships between continuous predictor variables (such as water use and emissions) and the continuous response variables of interest (yield, yield per hectare, and average sale price). GLMs offer the flexibility to examine both main effects and interaction terms~\cite{fahrmeir2013regression}, which is essential for exploring the nuanced interplay between resource use, vineyard size, and regional factors. Many prior studies have used GLMs to model vineyard relationships for response variables such as yield~\cite{SalinasRuiz2023, Baptiste2021}. Additionally, GLMs allow us to incorporate categorical variables (Year and GI Region) to account for temporal and spatial heterogeneity, making it possible to observe differences in vineyard performance across diverse Australian winegrowing regions over time. The interpretability of GLM coefficients makes this approach attractive for practical applications in vineyard management~\cite{kutner2005applied}, providing insights into the magnitude and direction of each variable's influence on yield and sale price outcomes (given the other variables in the model).
\par
Pairwise Pearson Correlation Coefficients (PPCC) were calculated to assess the potential existence of linear relationships between the input and predicted variables. Both the magnitude of the value and the associated confidence intervals were evaluated to determine if a coefficient was indicative of a strong relationship. P-values reflected the significance of a given correlation coefficient, with statistical significance being declared when the corresponding P-value was lower than 0.05. PPCC analyses were undertaken for data on the original scale and data as a logarithmic transform. Transforming data before calculating the coefficients changes several things. The logarithmic transform of the data alters the interpretation of the coefficients to percentage change; a coefficient will indicate the change in percentage of one variable compared to the other; scaling by standard deviation also changes this interpretation to be a percentage of that variable's standard deviation. When considering the logarithmically transformed variables, a coefficient of 1 would indicate that the change of one variable by one percentage of its standard deviation would correlate to the other variable changing by one percentage of its standard deviation. The importance of this is the dimensionless nature of these relationships and that it can be translated directly to any vineyard's case with a well-known distribution.
\par
Five general linear models (GLMs) were created to explore the interactions between predictors (see Table~\ref{tab:tab1}) and allow for easily comparable differences in the influence and magnitude of relationships. Model fit was measured using $R^2$ and adjusted $R^2$ and F statistics. T-tests were used to determine if predictors significantly contributed to the models when accounting for other variables, showing which specific years and areas contributed significantly. Both the PPCC and GLM analyses were created using the R statistical programming language~\cite{rcoreteamLanguageEnvironmentStatistical2021} with the Caret package~\cite{kuhnBuildingPredictiveModels2008}. 
\par
Various alternate methods were explored, including splines, hierarchical regression, general additive, and generalised linear models. These alternative approaches were not used as final models because they offered little further insights or improvements in accuracy.
\par
\subsection*{Model validation}
Models were validated using K-fold cross-validation. This was performed by removing a subset of data from the sample used to train models and then predicting those variables to determine how sensitive the model is to changes in the sample data. For this analysis, each model was validated using 10 folds, repeated 100 times.
\par
\section*{Results}
\subsection*{Summary statistics}\label{sec:exp_anal}
 Table~\ref{tab:summary} shows the summary statistics of each variable in its original units. The range of these values shows the level of difference between some vineyards, with operations differing by orders of magnitude in size, yield and average price of sale (See Table~\ref{tab:tab1}). PPCC values for the transformed, centred and scaled variables are shown in Table~\ref{tab:pearson_correlation}. The process of centring and scaling variables changes them so that their mean is centred on 0, and all values are scaled to the variable's standard deviation, such that 1 is equal to the variable's standard deviation. All correlations were found to be statistically significant (p $<$ 2.200E-16), except for Average Sale Price, all variables were positively correlated. Water use, area harvested, and emissions being positively correlated to yield, it can be considered that more resources and area are likely to lead to greater yields. The negative correlations between Average Sale Price and Yield, Water Use, Area and Scope One Emissions indicated that Area Harvested and Fuel Use were not the determining factors for average sale price. The negative correlations are associative, not causal relationships (i.e. using more water does not cause lower sale prices).
\begin{table}[!ht]
  \caption{Summary statistics of each continuous variable.}\label{tab:summary}
      % \resizebox{\textwidth}{!}{
        \begin{tabular}{@{}cllll@{}}
          \toprule
          \textbf{Variable} & \textbf{Mean} & \textbf{\begin{tabular}[c]{@{}l@{}}Standard\\ Deviation\end{tabular}} & \textbf{Minimum} & \textbf{Maximum} \\ \midrule
          {Yield (tonnes)} & 7.76E+02 & 2.18E+03 & 1.00E+00 & 7.23E+04 \\
          \multicolumn{1}{l}{} &  &  &  &  \\
          {Area Harvested (ha)} & 6.67E+01 & 1.34E+02 & 7.00E-02 & 2.44E+03 \\
          \multicolumn{1}{l}{} &  &  &  &  \\
          {Water Used (ML)} & 7.47E+06 & 5.65E+08 & 1.00E+00 & 4.27E+10 \\
          \multicolumn{1}{l}{} &  &  &  &  \\
          {Scope One Emissions ($tCO_{2}e$)} & 4.17E+04 & 8.57E+04 & 6.76E+00 & 2.11E+06 \\
          \multicolumn{1}{l}{} &  &  &  &  \\
          $\mbox{Yield (tonnes)}\over \mbox{Area harvested (ha)}$ & 1.01E+01 & 8.13E+00 & 4.00E-02 & 8.63E+01 \\
          \multicolumn{1}{l}{} &  &  &  &  \\
          {Average Sale Price (AUD/tonne)} & 1.48E+03 & 9.22E+02 & 1.60E+02 & 2.60E+04 \\
          \multicolumn{1}{l}{} &  &  &  &  \\
          $\mbox{Average Sale Price (AUD/tonne)} \over \mbox{Area Harvested (ha)}$ & 1.35E+02 & 5.71E+02 & 1.75E-01 & 2.98E+04 \\ \bottomrule
          \end{tabular}
  % }
  \end{table}

\par
\begin{table}[!ht]
  \caption{Pairwise Pearson correlation coefficients for logarithmically transformed values.}\label{tab:pearson_correlation}
      \resizebox{\textwidth}{!}{
      \begin{tabular}{@{}cccccccc@{}}
          \toprule
          \textbf{} & \multicolumn{1}{c}{\textbf{Yield}} & \multicolumn{1}{c}{\textbf{Area Harvested}} & \multicolumn{1}{c}{\textbf{Water Used}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Scope One\\ Emissions\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Yield by\\ Area\end{tabular}}} & \multicolumn{1}{c}{\textbf{Average Price}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Average Price\\ by Area\end{tabular}}} \\ \midrule
          Yield & 1.00 & 0.88 & 0.82 & 0.76 & 0.96 & -0.46 & -0.88 \\
          Area Harvested & 0.88 & 1.00 & 0.78 & 0.83 & 0.73 & -0.19 & -0.81 \\
          Water Used & 0.82 & 0.78 & 1.00 & 0.67 & 0.76 & -0.49 & -0.82 \\
          \begin{tabular}[c]{@{}c@{}}Scope One\\ Emissions\end{tabular} & 0.76 & 0.83 & 0.67 & 1.00 & 0.65 & -0.16 & -0.67 \\
          \begin{tabular}[c]{@{}c@{}}Yield by\\ Area\end{tabular} & 0.96 & 0.73 & 0.76 & 0.65 & 1.00 & -0.54 & -0.84 \\
          Average Price & -0.46 & -0.19 & -0.49 & -0.16 & -0.54 & 1.00 & 0.72 \\
          \begin{tabular}[c]{@{}c@{}}Average Price\\ by Area\end{tabular} & -0.88 & -0.81 & -0.82 & -0.67 & -0.84 & 0.72 & 1.00 \\ \bottomrule
          \end{tabular}}
\end{table}

\subsection*{General linear models}

Each model had a high $R^2$ value, indicating that most of the variance within the data was described by the models (see Table~\ref{tab:modelsummary}) and statistical significance of F-tests (p $<$ 2.200E-16). Aside from 3 variables, all regression parameters were statistically significant from zero; F-tests across each model's variables were (p $<$ 0.05). The three exceptions were scope one emissions in Model 3 (p=0.22) and Model 4 (p=0.39) and the interaction between the area harvested and water used in Model 2 (p=0.22). Note that scope one emissions were included in all models to directly compare the response variables as ratios of vineyard size to raw values and because it was strongly correlated to the response variable in every model (except model 5), especially for Models 1 and 4 (Table~\ref{tab:pearson_correlation}). 
%A full list of regression coefficients 95\% CIs and p-values for each of the four models is provided in the appendix. % You gotta add these to the appendix in a more readable fashion
\par
\begin{table}[!ht]
  \caption{Summary of models; their performance, F-statistics and Residual error.}\label{tab:modelsummary}
      \resizebox{\textwidth}{!}{
        \begin{tabular}{@{}cccccccc@{}}
          \toprule
           & $\mathbf{R^2}$ & \textbf{\begin{tabular}[c]{@{}c@{}} Adjusted\\ $\mathbf{R^2}$ \end{tabular}} & \textbf{F-Statistic} & \textbf{P-Value} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual\\ Standard Error\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Sum\\ of Squares\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Mean\\ of Squares\end{tabular}} \\ \midrule
          \textbf{\begin{tabular}[c]{@{}c@{}}Model 1\\ \end{tabular}} & 0.9072 & 0.9061 & 775.3 & 2.200e-16 & 0.3065 & 491.3 & 0.1 \\
          \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
          \textbf{\begin{tabular}[c]{@{}c@{}}Model 2\\ \end{tabular}} & 0.8291 & 0.8141 & 55.07 & 2.200e-16 & 0.4312
           & 905.03 & 0.19 \\
          \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
          \textbf{\begin{tabular}[c]{@{}c@{}}Model 3\\ \end{tabular}} &0.9753 &0.9748 & 1885 & 2.200e-16 & 0.1589 & 71.11 & 0.03 \\
          \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
          \textbf{\begin{tabular}[c]{@{}c@{}}Model 4\\ \end{tabular}} & 0.9091 & 0.9006 & 106.1 & 2.200e-16 & 0.3153  & 261.41 & 0.10 \\
          \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
          \textbf{\begin{tabular}[c]{@{}c@{}}Model 5\\ \end{tabular}} & 0.9089 & 0.9004 & 107.2 & 2.200e-16 & 0.3155 & 262.04 & 0.10 \\ \bottomrule
          \end{tabular}
        }
\end{table}

Table~\ref{tab:contvar} summarises coefficients related to continuous variables. In Model 1, all coefficients except for the intercept significantly contributed to the model (p $<$ 0.05), and in Model 2, all coefficients were statistically significant (p $<$ 0.05). However, scope one emissions did not significantly contribute to models 3, 4 and 5. Model 4 only had statistically significant contributions from the intercept and water use. Although the coefficient for water use was statistically significant for each model, it did not have the highest value; instead, the area harvested, being an order of magnitude greater, dominated the models. Model 5 achieved a similar $R^2$ to Model 4 without area harvested, having stronger influences from water use and scope one emission.

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage{multirow}

% Intercept
% Area Harvested
% Water Used
% Scope One Emissions

% Area Harvested ** Scope One Emissions
% Area Harvested ** Water Used

\begin{table}[!ht]
  \caption{Summary of each Model's coefficients for continuous variables.}\label{tab:contvar}
  \resizebox{\textwidth}{!}{
  \begin{tabular}{@{}cccccccc@{}}
  \toprule
  \textbf{} & \textbf{} & \textbf{Intercept} & \textbf{Area Harvested} & \textbf{Water Used} & \textbf{\begin{tabular}[c]{@{}c@{}}Scope One\\ Emissions\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Area Harvested \\ ** \\ Scope One Emissions\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Area Harvested \\ ** \\ Water Used\end{tabular}} \\ \midrule
  \multirow{3}{*}{Model 1} 
   & Coefficient & -3.32E-02 & 7.42E-01 & 8.66E-02 & 6.73E-02 &  &  \\
   & Std Error & 1.96E-02 & 1.00E-02 & 8.90E-03 & 8.00E-03 &  &  \\
   & $Pr(>|t|)$ & 9.09e-02 & $< 2e-16$ & $< 2e-16$ & $< 2e-16$ &  &  \\
   &  &  &  &  &  &  &  \\
  \multirow{3}{*}{Model 2} 
  & Coefficient & 1.70E-01 & 5.77E-01 & 1.08E-01 & 8.50E-02 & -4.97E-02 & -5.35E-02 \\
   & Std Error & 5.91E-02 & 1.48E-02 & 1.31E-02 & 1.17E-02 & 8.10E-03 & 8.40E-03 \\
   & $Pr(>|t|)$ & 4.15e-03 & $< 2e-16$ & 2.09e-16 & 5.41e-13 & 9.10e-10 & 2.66e-10 \\
   &  &  &  &  &  &  &  \\
  \multirow{3}{*}{Model 3} 
   & Coefficient & 1.81E-02 & 9.71E-01 & -2.31E-02 & -7.00E-03 &  &  \\
   & Std Error & 1.30E-02 & 7.20E-03 & 6.90E-03 & 5.70E-03 &  &  \\
   & $Pr(>|t|)$ & 1.66e-01 & $< 2e-16$ & 8.55e-04 & 2.22e-01  &  &  \\
   &  &  &  &  &  &  &  \\
  \multirow{3}{*}{Model 4} 
  & Coefficient & 1.45E-01 & 2.40E-03 & -4.66E-02 & -1.70E-02 & 1.15E-02 & 1.40E-03 \\
   & Std Error & 5.28E-02 & 1.50E-02 & 1.43E-02 & 1.18E-02 & 7.90E-03 & 8.30E-03 \\
   & $Pr(>|t|)$ & 6.10e-03 & 8.76e-01 & 1.16e-03 & 1.49e-01 & 1.47e-01 & 8.63e-01 \\
   &  &  &  &  &  &  &  \\
  \multirow{3}{*}{Model 5} 
   & Coefficient & 1.52E-01 &  & -4.04E-02 & -1.71E-02 &  &  \\
   & Std Error & 5.27E-02 &  & 1.13E-02 & 9.70E-03 &  & \\
   & $Pr(>|t|)$ & 4.01e-03 &  & 3.53e-04 & 7.97e-02 &  &  \\ \cmidrule(l){2-8} 
  \end{tabular}
  }
  \end{table}

The regression coefficients for the Year variable of each model are depicted in Fig~\ref{fig:yearly}. For each model, the first year of data was used for the model's baseline: 2012 for Models 1 and 2 and 2015 for Models 3, 4 and 5. The Adelaide Hills is used as the regional baseline for each of the models. And, for the interaction between year and region, the first year and the Adelaide Hills is used as the baseline. Region and year contributed more than the other variables in some but not all cases. However, the coefficients varied substantially over the years and were not significantly different from zero in some years. Models 4 and 5 are very similar in Fig~\ref{fig:yearly}, indicating that the exclusion of area does not significantly affect the contribution from annual influence. Models 4 and 5 have the most prominent trends, showing an increase in yearly effects over time, with Model 3 also increasing from 2016 to 2018 but plateauing afterwards. Models 1 and 2 do not show a clear trend but drop during 2017 and 2018 after rising in the first 3 years.
\par
\begin{figure}[!ht]
  % \includegraphics{fig2.pdf}
  \caption{Model Coefficient values for each year with standard error bars. Each coefficient represents the influence that each year had on the model outcomes (see Table 1 for model outcomes and their predictors). Each year's influence is compared to the first year of the model.}t\label{fig:yearly}
  \end{figure}

% Search for where you mention quality and where you mention value, as this does become misleading!!!
%
% This is partly because of the intercept
%
% It depends on what else is in the model.
%
% These direct comparisons are 
%
% The correlation between input vars need to be discussed earlier.
%
%difference between the continuous variables coefficients is the change from positive to negative values. This change occurs between the Models for Yield (Models 1 and 2), and the value models (Models 3 and 4), where all but the coefficient for area harvested had the opposite sign (see Table~\ref{tab:contvar}). These models also differ in an order of magnitude when looking at resource use, with the coefficients for yield being smaller than those for value. 
%
% The resource input is different by an order of magnitude between value and yield
%
% Area has the most decisive influence
%
% The models as a ratio of the area shows opposite trends - this could indicate that smaller vineyards tend toward more valuable crops and are more significant to crops of greater yield. 
%

Fig~\ref{fig:spider} summarises regional differences by temperature and rainfall. The most notable difference is between vineyards within 'Hot' and 'Very Dry' regions (warm inland regions), where high average sale prices are historically low, and yield is high. Water Use changes dramatically between these regions as well, with water being a driving force for yield but not necessarily average sale price. The warmer and drier regions also tend to have larger vineyards. These regional differences are further shown in Fig~\ref{fig:yield_vs_value_area}, as a ratio of vineyard area, where there is a pronounced difference in 'Hot' and 'Very Dry' regions producing more per area with lower average sale prices per area than other regions.
\par
\begin{figure}[!ht]
  % \resizebox{\textwidth}{!}{
  % \includegraphics{fig3.pdf}}
  \caption{Radar plot of climatic profile's resource use, yield and average sale price. Each point shows the overall contribution of a variable to the vineyard's outcomes for that climate. The left reflects vineyards in different climatic temperatures. The right reflects vineyards in different rainfall climates.}\label{fig:spider}
\end{figure}

Fig~\ref{fig:yield_vs_value_area} further shows the emphasis that 'Hot' regions have on high yields, with low average sale prices compared with other regions. Scaling average price and yield by area shows a strong negative trend, trading quantity for higher sales prices.
\par
\begin{figure}[!ht]
  % \resizebox{\textwidth}{!}{
  % \includegraphics{fig4.pdf}}
    \caption{Scatter plot of vineyard yield against the average sale price as ratios to area harvested. The axes are in standard deviations as absolute values with points coloured by climate.}\label{fig:yield_vs_value_area}
\end{figure}

Table~\ref{tab:kfold} shows the validation results of each of the models. The $R^2$ measures of fit show similar results to the initial models, with a slight decrease as expected, indicating that the models are robust and consistent.

\par
\begin{table}[!ht]
  \caption{Model validation using k-fold cross validation, for ten folds repeated 100 times.}\label{tab:kfold}
  % \resizebox{\textwidth}{!}{
  \begin{tabular}{@{}cccc@{}}
    \toprule
    \textbf{} & \textbf{\begin{tabular}[c]{@{}c@{}}Residual Mean\\ Squared Error\end{tabular}} & \textbf{R2} & \textbf{\begin{tabular}[c]{@{}c@{}}Mean Average \\ Error\end{tabular}} \\ \midrule
    \textbf{Model 1} & .309 & .905 & .2165 \\
    \textbf{Model 2} & .457 & .7921 & .313 \\
    \textbf{Model 3} & .165 & .972 & .101 \\
    \textbf{Model 4} & .348 & .878 & .182 \\ 
    \textbf{Model 5} & .348 & .878 & .183 \\ \bottomrule
    \end{tabular}
  % }
  \end{table}
% Edit Id 13
Fig~\ref{fig:cluster} highlights two of the most prominent distinct clusters of vineyards based on climate characteristics, specifically comparing warm/hot inland regions and cooler coastal regions. The clustering reveals that warmer inland areas tend to have higher yields per hectare but at a lower profit per area, whereas cooler coastal regions show lower yields but higher profit margins per hectare. This contrast emphasizes the interplay between climate conditions and economic outcomes.

\begin{figure}[!ht]
  % \resizebox{\textwidth}{!}{
  % \includegraphics{fig5.pdf}}
    \caption{Scatter plot of vineyard yield against the average sale price as ratios to area harvested. This figure includes only vineyards from cool/damp and Hot/Dry climates, plotted against each other to aid in highlighting the division of these clusters. The axes are in standard deviations as absolute values with points coloured by climate.}\label{fig:cluster}
\end{figure}
\par

%
%
% At least summary statistics of the baseline model.
%
%

\section*{Discussion}

There was an expected strong relationship between size and resource use, with the overall area of a vineyard and its access to resources significantly determining the upper limit of potential yield. However, size was also inversely related to the potential average sale price. Higher average sales prices were also associated with high resource inputs per area instead of the overall expenditure of resources. Vineyard yields and sales prices changed considerably by region and year. Even given regional and yearly changes, there was a strong connection between smaller vineyards and higher sales prices. This could have been due to more attention available when managing smaller properties. Additionally, the strong negative correlation between vineyard size and average sale price could be due to regionally differing vineyard sizes, with warmer inland regions tending to have much larger vineyard sizes.
% Edit ID 7
\par
% Edit ID 9 
The difference between regions can also be due to varied reasons, such as climate; where warmer regions result in a shortening of the ripening period and the risk that if harvest occurs in a duration of high temperature, there could be a negative impact on wine quality~\cite{salazar_parra_effects_2010, duchene_grapevine_2005, jones_climate_2000} and yield~\cite{mira_de_orduna_climate_2010, iglesias_regional_2012}. Climate change might move the north and south latitude boundaries of areas suitable for good quality wines~\cite{schultz_climate_2010}. It could even lead to improvements in fruit production and quality improvements in some areas~\cite{olesen_consequences_2002}. However, other areas may be negatively affected by high temperatures and water stress due to reduced water availability. There is significant room to investigate the impact of different climate scenarios on these other regions and further the various consequences of climate change, specifically for water in the Australian context.
\par
The findings from this study reveal notable differences in yield and profitability between vineyard clusters based on climate conditions, specifically between the warm/hot regions and cooler/damper regions, with warmer drier regions tending to be further inland in Australia, compared to the more coastal cooler and damper regions. Warmer areas are characterized by longer growing seasons and higher heat accumulation, and are generally associated with higher yields per hectare (as shown by the results in Figs~\ref{fig:yield_vs_value_area} and~\ref{fig:cluster}). However, these conditions often entail trade-offs in grape quality, as accelerated ripening due to higher temperatures can reduce the complexity of flavours—a key determinant of premium wine quality~\cite{jones2005climate, gladstones2011wine}. Consequently, grapes from warmer regions tend to be lower-priced, with an emphasis on yield rather than quality~\cite{schamel2003wine}. In contrast, cooler coastal regions, which benefit from moderate temperatures and oceanic cooling, produce lower yields per hectare but command higher prices due to the slower ripening process that fosters more nuanced flavours in the grapes~\cite{gladstones2011wine, mira_de_orduna_climate_2010}.
\par
% Edit ID 13
These climate-based clusters suggest distinct strategies for vineyard management that could enhance both economic outcomes and sustainability. Vineyards in warm/hot inland regions may benefit from water-efficient practices, such as mulching~\cite{rossertWoodchipsPotatoChips2022} to maximize yield while conserving resources, thereby sustaining profitability through volume despite lower price points per tonne. Conversely, vineyards in cool/damp coastal regions could focus on quality-enhancing practices, such as canopy managements~\cite{delpuechAdaptingCoverCrop2018}, to optimize profit potential per hectare, reflecting the market demand for premium-quality grapes. Future research could build upon these findings by refining cluster characteristics with additional factors, such as vineyard size and soil type, to offer more targeted recommendations. Such an approach may further clarify the influence of specific environmental and operational factors on vineyard productivity and profitability, providing valuable insights for region-specific, sustainable vineyard management.
\par
Given its F-statistics, the lack of significance of scope one emissions and its contribution to models could indicate that other vineyard activities requiring fuel are not leading factors for a vineyard's average sale price. The relationship between yield, value, and area was not simply about efficiently producing the most grapes. It is possible that the relationship of scope one emission between yield and sale price was closely tied to a vineyard's area due to requiring more fuel to address more issues over greater distances. It is difficult to directly discern the connections between scope one emissions, as fuel can be used for a broad category of activities. As~\cite{rugani2013carbon} highlights, the interpretation of emissions is uncertain and inexplicit when comparing viticulture practices. Many factors, especially size, can directly or indirectly influence the dynamics of carbon emissions in a vineyard's life cycle. Notably, scope one emissions account for only a small portion of the total energy input in vineyards, as other significant inputs, such as fertilizers, pesticides, and electricity, collectively contribute to a substantial share of the carbon footprint in grape production~\cite{point2012life}.
\par
There are important considerations unique to winegrowing compared to other agricultural industries. The vertical integration of winegrowing within the wine industry ties winegrowers to secondary and tertiary industries, such as wine production, packaging, transport and sales. This results in unique issues and considerations for each vineyard, where on-the-ground decisions are influenced by other wine industry choices, such as the use of sustainable practices in vineyards as a requirement for sale in overseas markets; notably, these interactions can be further complicated by some winegrowers being wholly integrated into a wine company, while others are not~\cite{knightFirmResourcesDevelopment2019}. Incorporating decisions into the model could help describe the factors contributing to regional differences beyond resource consumption, motivating the call for more granular data and more sophisticated modelling.
\par
Many on-the-ground decisions influence both sales price and yield. The decision to prioritise average sale price over quantity, is governed by complex physical and social forces, for example international market demands, disease pressures and natural disasters~\cite{abadCoverCropsViticulture2021,cortezUsingDataMining2009,hallWithinseasonTemporalVariation2011,i.goodwinManagingSoilWater2009,kasimatiPredictingGrapeSugar2022,oliverReviewSoilPhysical2013,srivastavaNondestructiveSensingMethods2018}, with many of these occurrences being highlighted throughout the reports from Wine Australia over the past decade~\cite{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2021,wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}. However, the changes in the coefficients (see Fig~\ref{fig:yearly}) are not reflective of many known occurrences, such as the 2020 bush fires, which had higher values for coefficients than prior years; During the 2020 bush fires 40,000 tonnes of grapes were lost across 18 different wine regions due to bush fires and smoke taint. Compared to countrywide pressures such as drought, this damage made up only 3\% of the total amount of grapes for that year; although acknowledged as a considerable loss on an individual basis, it was deemed to be only a minor national concern by Wine Australia when compared to other environmental pressures such as drought~\cite{wineaustraliaNationalVintageReport2020}
\par
Climatic pressures are an important consideration for growers, especially those in warmer and drier regions. The Wine Australia reports also show that warm inland regions have seen a decline in profit over the past decade, whereas regions with lower average sales prices did not~\cite{wineaustraliaNationalVintageReport2019,wineaustraliaNationalVintageReport2020,wineaustraliaNationalVintageReport2021,winemakersfederationofaustraliaNationalVintageReport2013,winemakersfederationofaustraliaNationalVintageReport2014,winemakersfederationofaustraliaNationalVintageReport2015,winemakersfederationofaustraliaNationalVintageReport2016,winemakersfederationofaustraliaNationalVintageReport2017,winemakersfederationofaustraliaNationalVintageReport2018}. Vineyards in warm inland regions also tend to be larger, making up for lower sale prices with larger yields. Considering the negative correlation of average price to area, economies of scale become essential for this strategy. Given the large quantities of grapes that can be produced by some vineyards, even at low margins, there is the potential to be profitable. However, the increasing climatic pressures, mixed with the requirement for larger volumes of water, make the sustainability of some vineyards come into question. Furthermore, intensive farming, in general, is known to jeopardise the sustainability of an operation through the degradation of soil and waterways~\cite{capelloEffectsTractorPasses2019,linHydropedologySynergisticIntegration2012,pisciottaGroundwaterNitrateRisk2015}. There are established methods that can help to mitigate these effects, such as the use of cover crops, midrow crop rotation and efficient irrigation.
\par
Some regions appeared to produce grapes at lower average sale prices at scale, while others produced higher in lower volumes. This empirical finding is consistent with Wine Australia's annual reports, which shows that some GI regions are known for producing large amounts of lower grade (low value per tonne) grapes~\cite{wineaustraliaNationalVintageReport2022,winemakersfederationofaustraliaNationalVintageReport2017}. Comparatively, other regions only produce grapes of higher sales price but in smaller quantities. The difference in pricing per tonne between the lowest and highest regional average sales prices was almost a hundred times, showing that region had a profound influence. Some regions also had a mixture of high and low average sales prices and yield, showing regional variability in pricing, which may be explained by the varieties produced. A further possibility is the existence of regional upper limits on the potential sales price or that there are diminishing returns in some regions when pursuing higher sales prices or quantity; however, these types of relationships may be obfuscated by knowledgeable winegrowers who avoid such pitfalls.
\par
Due to regional differences, different strategies are employed, such as some regions targeting mass production over higher sales prices. This is most notable when grouping regions by climate, especially when considering GI Regions in the 'Hot Very Dry' climate (see Fig~\ref{fig:yield_vs_value_area}). Fig~\ref{fig:spider} also shows that comparatively, 'Warm' and 'Dry' Regions manage their resources incredibly efficiently, having generally larger areas using similar resources to those in 'Cool' regions but having areas comparable to regions in 'Hot' climates. The coefficients for Model 4 also show a more significant benefit of resource use per area when producing grapes at higher average sale prices, showing higher resource use per area reflective of the higher average sales price (see Table~\ref{tab:summary}).
%In the preliminary analyses, alternative models found that predictions significantly underperformed without the direct incorporation of GI RegioYear year. The effect of climate in the models was never as significant as the more granular GI regions and always led to less accurate models. 
Although not chosen over the GI region, climate was considered a significant determinant of the ability to produce larger quantities of grapes and a determinant in grape sale price~\cite{agostaRegionalClimateVariability2012}. The more granular GI Region likely explained a broader mix of geographical phenomena, such as soil, geology and access to water resources~\cite{abbalDecisionSupportSystem2016,carmonaUseParticipatoryObjectOriented2011}. The interaction between the YYearand the GI Region likely accounted for events such as bushfires, which would be impactful, but only locally, both in time and space.
\par
% Edit ID 19
For winegrowers, understanding the inverse relationship between vineyard size and average sale price suggests that smaller vineyards may benefit from focusing on high-value, lower-yield practices, which allow for a more targeted approach to quality management. In larger vineyards, where economies of scale may drive profitability, strategies that optimise resource use efficiency per hectare become crucial, especially in regions facing water scarcity. Here, larger vineyards could benefit from adopting further sustainable and climate adaptive practices such as mulching~\cite{rossertWoodchipsPotatoChips2022}, vineyard canopy and mid-row management strategies~\cite{delpuechAdaptingCoverCrop2018}, and soil nutrient leaching and organic management strategies~\cite{goslingLongtermChangesSoil2005, abadCoverCropsViticulture2021, moonenFunctionalBiodiversityAgroecosystem2008} to help aid in water, fuel and fertiliser reductions. Given the climatic pressures in warmer inland regions, especially those with high water requirements, these practices may also mitigate the risk of crop failure due to drought or heat stress, enhancing both resilience and economic viability. Additionally, international market trends increasingly favour sustainably sourced wine, suggesting that supporting vineyards in sustainable water and fuel management could enhance export opportunities, meeting both environmental goals and market expectations.
\par
We identified two main limitations to our linear modelling. First, models 1 and 2 over-predicting yield may have been due to preventative measures brought on by regional pressures such as fire, frost and disease. More fuel and water were likely used to prevent these issues from spreading within a region, thus disproportionately affecting some vineyards compared to others locally. This type of maintenance is not well captured in the models, especially when some regions, especially those in warmer regions, are not as prone to disease as cooler climates and could potentially have lower fuel and water use per hectare. This could create a discrepancy in vineyards that utilised preventative measures in wetter regions, as opposed to those that did not, thus expending less fuel and energy but risking disease. When reviewing the differences between regions, it is essential to consider that vineyards in 'Hot, Very Dry' regions can be hundreds of times the size in other regions. This limitation could be overcome by incorporating the profitability of vineyards and comparing the financial success of working at different operational scales.
\par
The second limitation was the lack of further explanatory variables to help link models to causal effects. Variables such as using renewable energy, contractors, and the occurrence of disease, fire and frost were explored initially to capture the discrepancies between similar vineyards that produced different yields and crop values. However, none of these variables was significantly correlated with the response variables and did not add to model accuracy, even when considered as interactions. Allowance for nonlinear relationships, specifically through splines, resulted in more normally distributed residuals but at a drastically reduced overall accuracy when comparing $R^2$ and Residual Square Error. Attempts to fully explain minor variations were always overshadowed by the dramatic differences in regional trends. More data for each region would also be beneficial, allowing for better comparisons between regions.
\par
Using other models, such as random forests and decision trees, alongside more variables and data may help uncover the reasons for under or overestimation. The use of alternative sustainable practices in the field could cause these differences. Moreover, while there is evidence to suggest that environmentally sustainable practices can reduce costs and increase efficiency whilst improving the quality of grapes, more research is needed to link these benefits across different regions and climates~\cite{baianoOverviewSustainabilityWine2021,marianiSustainableWinegrowingCurrent2015,montalvo-falconSustainabilityResearchWine2023}.
\section*{Conclusion}

This study delved into the relationships between resource use, grape sales price and yield. The findings underscore the multifaceted nature of vineyard management, where the interplay of size, resource allocation, climate, and regional influences collectively shape both the expected sale price and the quantity of grape yields. The average sales price of grapes was not solely tied to the overall expenditure of resources but rather to the efficient allocation of resources per area. This emphasises that factors beyond sheer scale contribute significantly to the final sale price of grapes produced. Moreover, regional and yearly variations substantially affected vineyard outputs, impacting sales price and quantity. Building on these insights, future research could link these types of interpretable models to other studies through the use of machine learning techniques and causal inference to improve accuracy and link sustainable practice would greatly strive toward the creation of decision support tools to further inform winegrowers of sustainable options and their efficacy.
% Edit ID 7
% Edit ID 20

\nolinenumbers

% \bibliographystyle{plos2015}
\bibliography{references}

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as regular sections
 \appendix

\section*{Appendix}

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{figure}{0} % Reset figure counter for appendix
\setcounter{table}{0}  % Reset table counter for appendix

\par

\begin{table}[htb]
    \caption{P-values for the Pearson correlation coefficients of the non-transformed water used variable with other variables. This table highlights the statistical significance of correlations in the dataset.}\label{tab:waterppc}
  % \resizebox{\textwidth}{!}{
  \begin{tabular}{lr}
    \toprule1
      \textbf{Variable }                                          & \textbf{Water Used} \\ \\ \midrule
      Yield                                              & 7.538E-01  \\
      Area                                               & 6.981E-01  \\
      Scope One Emissions                                & 8.883E-01  \\
      $\textrm{Yield}\over\textrm{Area}$                     & 6.836E-01  \\
      Average Price Per Tonne                            & 5.600E-02  \\
      ${\textrm{Average Price per tonne}\over\textrm{Area}}$ & 1.522E-01 \\ \bottomrule
    \end{tabular}
  \end{table}

\begin{figure}[htb]
  % \resizebox{\textwidth}{!}{
  % \includegraphics{s1.pdf}}
  \caption{Quantile-Quantile (QQ) plot of Model 1 residuals, showing how closely the residuals align with a normal distribution. Deviations from the line indicate departures from normality.}
\end{figure}

\begin{figure}[htb]
  % \resizebox{\textwidth}{!}{
  % \includegraphics{s2.pdf}}
  \caption{Residuals vs. fitted values plot for Model 1, illustrating the linearity and homoscedasticity of the residuals.}
\end{figure}

\begin{figure}[htb]
  % \resizebox{\textwidth}{!}{
  % \includegraphics{s3.pdf}}
  \caption{Quantile-Quantile (QQ) plot of Model 2 residuals, showing how closely the residuals align with a normal distribution. Deviations from the line indicate departures from normality}
\end{figure}

\begin{figure}[htb]
  % \resizebox{\textwidth}{!}{
  % \includegraphics{s4.pdf}}
  \caption{Residuals vs. fitted values plot for Model 2, illustrating the linearity and homoscedasticity of the residuals.}
\end{figure}

\begin{figure}[htb]
  % \resizebox{\textwidth}{!}{
  % \includegraphics{s5.pdf}}
  \caption{Quantile-Quantile (QQ) plot of Model 3 residuals, showing how closely the residuals align with a normal distribution. Deviations from the line indicate departures from normality.}
\end{figure}

\begin{figure}[htb]
  % \resizebox{\textwidth}{!}{
  % \includegraphics{s6.pdf}}
  \caption{Residuals vs. fitted values plot for Model 3, illustrating the linearity and homoscedasticity of the residuals.}
\end{figure}

\begin{figure}[htb]
  % \resizebox{\textwidth}{!}{
  % \includegraphics{s7.pdf}}
  \caption{Quantile-Quantile (QQ) plot of Model 4 residuals, showing how closely the residuals align with a normal distribution. Deviations from the line indicate departures from normality.}
\end{figure}

\begin{figure}[htb]
  % \resizebox{\textwidth}{!}{
  % \includegraphics{s8.pdf}}
  \caption{Residuals vs. fitted values plot for Model 4, illustrating the linearity and homoscedasticity of the residuals.}
\end{figure}
\begin{figure}[htb]
  % \resizebox{\textwidth}{!}{
  % \includegraphics{s9.pdf}}
  \caption{Quantile-Quantile (QQ) plot of Model 5 residuals, showing how closely the residuals align with a normal distribution. Deviations from the line indicate departures from normality.}
\end{figure}

\begin{figure}[htb]
  % \resizebox{\textwidth}{!}{
  % \includegraphics{s10.pdf}}
  \caption{Residuals vs. fitted values plot for Model 5, illustrating the linearity and homoscedasticity of the residuals.}
\end{figure}
\par

% \begin{longtable}{p{4cm} r r r r} % Adjust column width for Variable
% \caption{P Va} \\ % Add your caption here
% \hline
% \textbf{Variable} & \textbf{Estimate} & \textbf{Std. Error} & \textbf{t value} & \textbf{Pr(>|t|)} \\
% \hline
% \csvreader[
%     head to column names,
%     late after line=\\,
%     late after last line=\\\hline
% ]{model1summary.csv}{}%
% { \csvcoli & \csvcolii & \csvcoliii & \csvcoliv & \csvcolv }
% \end{longtable}

% \begin{longtable}{p{4cm} r r r r} % Adjust column width for Variable
% \hline
% \textbf{Variable} & \textbf{Estimate} & \textbf{Std. Error} & \textbf{t value} & \textbf{Pr(>|t|)} \\
% \hline
% \csvreader[
%     head to column names,
%     late after line=\\,
%     late after last line=\\\hline
% ]{model2summary.csv}{}%
% { \csvcoli & \csvcolii & \csvcoliii & \csvcoliv & \csvcolv }
% \end{longtable}

% \begin{longtable}{p{4cm} r r r r} % Adjust column width for Variable
% \hline
% \textbf{Variable} & \textbf{Estimate} & \textbf{Std. Error} & \textbf{t value} & \textbf{Pr(>|t|)} \\
% \hline
% \csvreader[
%     head to column names,
%     late after line=\\,
%     late after last line=\\\hline
% ]{model3summary.csv}{}%
% { \csvcoli & \csvcolii & \csvcoliii & \csvcoliv & \csvcolv }
% \end{longtable}

% \begin{longtable}{p{4cm} r r r r} % Adjust column width for Variable
% \hline
% \textbf{Variable} & \textbf{Estimate} & \textbf{Std. Error} & \textbf{t value} & \textbf{Pr(>|t|)} \\
% \hline
% \csvreader[
%     head to column names,
%     late after line=\\,
%     late after last line=\\\hline
% ]{model4summary.csv}{}%
% { \csvcoli & \csvcolii & \csvcoliii & \csvcoliv & \csvcolv }
% \end{longtable}

% \subsection*{Complete list of coefficients for Model 1.}
% \csvautolongtable[]{model1_coef.csv}

%   \subsection*{Complete list of coefficients for Model 2.}
% \csvautolongtable[]{model2_coef.csv}

%   \subsection*{Complete list of coefficients for Model 3.}
% \csvautolongtable[]{model3_coef.csv}

%   \subsection*{Complete list of coefficients for Model 4.}
% \csvautolongtable[]{model4_coef.csv}

%   \subsection*{Complete list of coefficients for Model 5.}
% \csvautolongtable[]{model5_coefs.csv}

%
%% \section*{}
%% \label{}

%% If you have a bib database file and want BibTeX to generate the
%% bib items, please use
%%

%% else use the following coding to input the bib items directly into the
%% TeX file.

\end{document}
